### 2023-02-21 README.Rmd
more basic rr files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "ROF Camera Trap Data Analysis - Preliminary Report"
```

# rofcamtrap

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/master?urlpath=rstudio)

This repository contains the data and code for our preliminary report.

<!-- This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

### How to cite

Please cite this compendium as:

> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### Licenses

TBD

<!-- **Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2023-02-21 README.Rmd
fix binder

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/main?urlpath=rstudio)
 
@@
 You can download the compendium as a zip from from this URL:
-[master.zip](/archive/master.zip). After unzipping:
+[master.zip](/archive/main.zip). After unzipping:
 - open the `.Rproj` file in RStudio

### 2023-02-22 scripts/1_preprocessing.R
add one script

@@ -1,0 +1,2 @@
+## For functions related to pre-processing camera trap data
+

### 2023-02-22 scripts/1_preprocess_annotations.R
add firsts few scripts

@@ -1,0 +1,57 @@
+# Note: This runs on the csvs created from the label studio json outputs by
+# the post process function in the python utility mdtools
+
+# Folder with label studio annotation outputs
+label_studio_folder <- "analysis/data/raw_data/label_studio_outputs/"
+
+# Annotations -------------------------------------------------------------
+
+# List all the files with the proper file name format
+file_list <- list.files(label_studio_folder, recursive = F,
+                        full.names = T, pattern = "output.csv")
+
+# Create an empty dataframe to be grown
+annotations <- data.frame()
+
+for (file in file_list){
+
+  # print(file)
+
+  outls <- readr::read_csv(file, col_types = readr::cols(
+    id = readr::col_character(),
+    type = readr::col_character(),
+    origin = readr::col_character(),
+    to_name = readr::col_character(),
+    variable = readr::col_character(),
+    original_width = readr::col_double(),
+    original_height = readr::col_double(),
+    value_x = readr::col_double(),
+    value_y = readr::col_double(),
+    value_width = readr::col_double(),
+    value_height = readr::col_double(),
+    source_file = readr::col_character(),
+    tag = readr::col_character(),
+    value_text = readr::col_character()
+  ))
+
+  annotations <- dplyr::bind_rows(annotations, outls)
+
+}
+
+saveRDS(annotations, "analysis/data/derived_data/annotations.rds")
+
+# -------------------------------------------------------------------------
+
+# TODO dead code
+
+# anns <- readRDS("data/anns.rds")
+# Check for duplicates (should have been caught by python code)
+# test <- anns %>% dplyr::filter(variable == "species" |
+#                                  variable == "other_species",
+#                                tag != "Human",
+#                                tag != "Vehicle")  %>%
+#   dplyr::group_by(id, type, origin, to_name, original_width, original_height,
+#                   value_x, value_y, value_width, value_height, source_file,
+#                   value_text, variable) %>%
+#   dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
+#   dplyr::filter(n > 1L)

### 2023-02-22 scripts/1_preprocessing.R

@@ -1,2 +1,0 @@
-## For functions related to pre-processing camera trap data
-

### 2023-02-22 scripts/2_preprocess_detections.R

@@ -1,0 +1,27 @@
+# Note: This takes the COCO formatted detections derived from the python tool
+# mdtools, and outputs a table
+
+# Folder with Megadetector detections in coco format
+detections_folder <- "/media/vlucet/TrailCamST/TrailCamStorage/"
+
+# List all the files with the proper file name format
+file_list <- list.files(detections_folder, recursive = F,
+                        full.names = T, pattern = "output_coco.json")
+
+# Detections --------------------------------------------------------------
+
+# Create an empty dataframe to be grown
+detections <- data.frame()
+
+for (file in file_list){
+
+  # print(file)
+
+  outls <- jsonlite::fromJSON(file, flatten = T)$annotations %>%
+    janitor::clean_names()
+
+  detections <- rbind(detections, outls)
+
+}
+
+saveRDS(detections, "analysis/data/derived_data/detections.rds")

### 2023-02-22 analysis/paper/paper.qmd
commit qmd file, leave out the docx for now

---
title: "Ring of Fire Camera Detections Report"
author:
  - Valentin Lucet:
      correspondence: "yes"
      email: valentin.lucet@gmail.com
      orcid: "0000-0003-0268-818X"
      institute: wlu
  - Frances Stewart:
      institute: wlu
      orcid: "0000-0001-9344-8346"
institute:
  - wlu:
      name: Wilfrid Laurier University
      # address: 23 Science Street, Eureka, Mississippi, USA
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/landscape-ecology.csl" # Insert path for the bib-style
abstract: |
  Abstract TBW
# keywords: |
#   keyword 1; keyword 2; keyword 3
# highlights: |
#   These are the highlights. 
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
```

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored.

With the following code you can access and display values from the yml header above.
Keywords: `r rmarkdown::metadata$keywords`
Highlights: `r rmarkdown::metadata$highlights`

Here is a citation [@Marwick2017] -->

<!-- The actual document text starts here: -->

## Summary

TBW

## Introduction

- What this report is about
- Recycle the other introductions
- Importance of monitoring
- Data collection challenge
- New methods to the rescue
- Data challenges remain
- AI plan

## Methods

### Sampling protocol

- Recycle the info from proposal

### Image detection

```{r}
#| label: load-data
#| echo: false
#| 
if (!file.exists("analysis/data/derived_data/annotations.rds")){
  source("scripts/1_preprocess_annotations.R")
} else {
  anns <- readRDS("analysis/data/derived_data/annotations.rds")
}

if (!file.exists("analysis/data/derived_data/detections.rds")){
  source("scripts/2_preprocess_detections.R")
} else {
  dets <- readRDS("analysis/data/derived_data/detections.rds")
}
```

Images were retrieved from all cameras using a SD card reader and copied onto a 10 TB harddrive od brand LaCie. The images amounted to 1.8TB of memory on disk, for a total of 2,788,375 images. Images were stored with the following directory structure:

- Plot id (e.g. P128)
- Site id (e.g. P128-1)
- Camera id (e.g. CFS-10)
- DCIM folder (created by the camera)
- Overflow folder (e.g. 100RECNX, each such folder contains up to 10,000 images)

The images were processed through the animal detection model megadetector, version 5.0, using a lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete. The model processes each image and produces “detections” (bounding boxes) at a  variable confidence score.

About 46.22 % of the images, or 1,289,022 images, were found by the model to possibly have either a person, animal, or vehicle. The model did not produce detections for 1,499,353 images, or 53.77 % of the total.
This resulted in a total of 2,889,195 detections. The totals for each category are: 1,691,444 animals (or 58.54 %); 989,112 persons (or 34.23 %); and 208,639 vehicles (or 7.22 %).

Using a cutt-off value for confidence threshold of 0.1, this left us with a total of 333,502 detections to sort through (see Figure 1 below for an histogram of the detection scores, grouped by detection types). The totals for each category are: 189,982 animals (or 56.97 %); 119,709 persons (or 35.89 %); and 23,811 vehicles (or 7.13 %).

- Add explanation about image bursts and upload of bursts including empties

### Tagging platform selection

We had in mind the following criterias when selecting the proper tool to tag our images and bounding boxes:

- Deployable online. This was to allow for for multiple taggers
- Flexible tagging interface. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
- Allowing to tag multiple bounding boxes per image. Most platforms only allow for tagging the entire image, and not for portions of it (e.g. bounding boxes of detections produced by megadetector). This is necessary for when we will crop our images to train our neural network model. 
- Optionally, open source, to allow for maximum reproducibility of the process.

We looked at platforms not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners for which the task of tagging is a common step, known as labeling. We found two platforms that matched all those criterias: LabelStudio and CVAT. After testing both platforms, LabelStudio was chosen as it was simpler to implement. The Community edition of LabelStudio was deployed on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and a NGINX reverse proxy, which is a standard approach for custom hosting of websites.

### Tagging progress

Out of the 189,982 animal detections, 65,338 have been processed on the platform (34.39 %) for a total of 54,677 images (1.96 % of the total amount of images, or 4.24 % of the set of images with at least one detection, or 19.44 % of the total amount of images uploaded to the platform). This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and 82,165 images, (2.94 % of the total amount of images, or 29.21 % of the total amount of images uploaded to the platform).
Out of the 65,338 detections, 30,621 were false positives (46,87 %), and 33,237 were true positives (50.86 %), to which 1,480 false negatives were added (manually added bounding boxes, 2.26 %).

- Percentage of predictions changed (no human)

### Issues encountered

- Issues with vegetation falsely detected.

## Preliminary results

### Species ID summary

TBW

### Ordination

TBW

<!-- Here's some example analysis code: -->

```{r}
#| label: get-data
#| eval: false
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

Figure @fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

## Discussion

### Choice of methods

TBW

### Future plans

TBW

## Conclusion

## Acknowledgements

- Thank the volunteers for help with tagging

<!-- The following line inserts a page break  -->

\newpage

## References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

<!-- ### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
 -->

### 2023-02-22 scripts/2_preprocess_detections.R
not use the pipe

@@ -17,10 +17,9 @@ for (file in file_list){
 
   # print(file)
 
-  outls <- jsonlite::fromJSON(file, flatten = T)$annotations %>%
-    janitor::clean_names()
+  outls <- jsonlite::fromJSON(file, flatten = T)$annotations
 
-  detections <- rbind(detections, outls)
+  detections <- rbind(detections, janitor::clean_names(outls))
 
 }
 

### 2023-02-22 analysis/paper/paper.qmd
update

@@
 
-Images were retrieved from all cameras using a SD card reader and copied onto a 10 TB harddrive od brand LaCie.
-The images amounted to 1.8TB of memory on disk, for a total of 2,788,375 images.
+Images were retrieved from all cameras using a SD card reader and copied onto a 10 TB hard drive of brand LaCie.
+The images amounted to 1.8TB of memory on disk, for a total of __`r length(unique(dets$image_id))`__ images.
 Images were stored with the following directory structure:
@@
 The model did not produce detections for 1,499,353 images, or 53.77 % of the total.
+
 This resulted in a total of 2,889,195 detections. The totals for each category are: 1,691,444 animals (or 58.54 %); 989,112 persons (or 34.23 %); and 208,639 vehicles (or 7.22 %).
@@
 This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and 82,165 images, (2.94 % of the total amount of images, or 29.21 % of the total amount of images uploaded to the platform).
+
 Out of the 65,338 detections, 30,621 were false positives (46,87 %), and 33,237 were true positives (50.86 %), to which 1,480 false negatives were added (manually added bounding boxes, 2.26 %).

### 2023-02-22 analysis/paper/paper.qmd
incremental changes

@@
 knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
+options(scipen = 1, digits = 2)
 ```
@@
 #| echo: false
-#| 
+
 if (!file.exists("analysis/data/derived_data/annotations.rds")){
@@
 }
+
+# Compute total amount of images
+nb_image <- length(unique(dets$image_id))
+nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
+
+# Hardcoded check
+stopifnot(nb_image == 2788375)
 ```
@@
 Images were retrieved from all cameras using a SD card reader and copied onto a 10 TB hard drive of brand LaCie.
-The images amounted to 1.8TB of memory on disk, for a total of __`r length(unique(dets$image_id))`__ images.
+The images amounted to 1.8TB of memory on disk, for a total of __`r nb_image_fmt`__ images.
 Images were stored with the following directory structure:
@@
 
-The images were processed through the animal detection model megadetector, version 5.0, using a lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
+The images were processed through the animal detection model MegaDetector, version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
 The model processes each image and produces “detections” (bounding boxes) at a  variable confidence score.
 
-About 46.22 % of the images, or 1,289,022 images, were found by the model to possibly have either a person, animal, or vehicle.
-The model did not produce detections for 1,499,353 images, or 53.77 % of the total.
+```{r}
+#| label: basics-stats-1
+#| echo: false
 
-This resulted in a total of 2,889,195 detections. The totals for each category are: 1,691,444 animals (or 58.54 %); 989,112 persons (or 34.23 %); and 208,639 vehicles (or 7.22 %).
+# Select and distinct down to image-based infos
+dets_images <- dets |>  dplyr::select(image_id, max_confidence, isempty) |> 
+  dplyr::distinct()
 
-Using a cutt-off value for confidence threshold of 0.1, this left us with a total of 333,502 detections to sort through (see Figure 1 below for an histogram of the detection scores, grouped by detection types).
-The totals for each category are: 189,982 animals (or 56.97 %); 119,709 persons (or 35.89 %); and 23,811 vehicles (or 7.13 %).
+# Breakdown of which images triggered a MD detection
+dets_tb <- dets_images$isempty |> table()
+dets_tb_sum <- sum(dets_tb)
+dets_tb_percent <- dets_tb/dets_tb_sum*100
+
+dets_tb_fmt <- prettyNum(dets_tb, big.mark = ",")
+```
+
+About __`r dets_tb_percent[1]` %__ of the images, or __`r dets_tb_fmt[1]`__ images, were found by the model to possibly have either a person, animal, or vehicle.
+The model did not produce detections for __`r dets_tb_fmt[2]`__ images, or __`r dets_tb_percent[2]` %__ of the total.
+
+```{r}
+#| label: basics-stats-2
+#| echo: false
+dets_0 <- dets |> dplyr::filter(confidence > 0)
+dets_0.1 <- dets |> dplyr::filter(confidence > 0.1)
+
+# Breakdown of detections by categories (animal, person, vehicles)
+dets_tb_cat_unfil <- dets_0 |> dplyr::pull(category_id) |> table()
+dets_tb_cat_unfil_sum <- sum(dets_tb_cat_unfil)
+dets_tb_cat_unfil_percent <- dets_tb_cat_unfil/dets_tb_cat_unfil_sum*100
+
+dets_tb_cat_unfil_fmt <- prettyNum(dets_tb_cat_unfil, big.mark = ",")
+dets_tb_cat_unfil_sum_fmt <- prettyNum(dets_tb_cat_unfil_sum, big.mark = ",")
+
+# Breakdown of detections above 0.1 by categories (animal, persons, vehicles)
+dets_tb_cat <- dets_0.1 |> dplyr::pull(category_id) |> table()
+dets_tb_cat_sum <- sum(dets_tb_cat)
+dets_tb_cat_percent <- dets_tb_cat/sum(dets_tb_cat)*100
+
+dets_tb_cat_fmt <- prettyNum(dets_tb_cat, big.mark = ",")
+dets_tb_cat_sum_fmt <- prettyNum(dets_tb_cat_sum, big.mark = ",")
+```
+
+This resulted in a total of __`r dets_tb_cat_unfil_sum_fmt`__ detections.
+The totals for each category are: __`r dets_tb_cat_unfil_fmt[1]`__ animals (or __`r dets_tb_cat_unfil_percent[1]` %__); __`r dets_tb_cat_unfil_fmt[2]`__ persons (or __`r dets_tb_cat_unfil_percent[2]` %__); and __`r dets_tb_cat_unfil_fmt[3]`__ vehicles (or __`r dets_tb_cat_unfil_percent[3]` %__).
+
+Using a cutt-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+The totals for each category are: __`r dets_tb_cat_fmt[1]`__ animals (or __`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (or __`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (or __`r dets_tb_cat_percent[3]` %__).
+
+```{r}
+#| label: fig-detections-histogram
+#| echo: false
+#| fig-cap: "Histogram of detection confidence scores."
+
+# Histogram of confidence scores above 0.1
+dets_0.1 |> 
+  dplyr::mutate(category_id = as.character(category_id)) |> 
+  dplyr::mutate(Type = dplyr::case_match(category_id,
+                                         "0" ~ "Empty",
+                                         "1" ~ "Animal",
+                                         "2" ~ "Person",
+                                         "3" ~ "Vehicle")) |> 
+  ggplot2::ggplot() +
+  ggplot2::theme_bw() +
+  ggplot2::geom_histogram(ggplot2::aes(x=confidence, group=Type, fill=Type),
+                          bins = 100) +
+  ggplot2::scale_fill_viridis_d() +
+  ggplot2::xlab("Confidence Score") +
+  ggplot2::ylab("Count")
+```
 
@@
 
-We had in mind the following criterias when selecting the proper tool to tag our images and bounding boxes:
+We had in mind the following criteria when selecting the proper tool to tag our images and bounding boxes:
 
@@
 Most platforms only allow for tagging the entire image, and not for portions of it (e.g.
-bounding boxes of detections produced by megadetector).
+bounding boxes of detections produced by MegaDetector).
 This is necessary for when we will crop our images to train our neural network model.
@@
 We looked at platforms not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners for which the task of tagging is a common step, known as labeling.
-We found two platforms that matched all those criterias: LabelStudio and CVAT.
+We found two platforms that matched all those criteria: LabelStudio and CVAT.
 After testing both platforms, LabelStudio was chosen as it was simpler to implement.
@@
 ### Tagging progress
+
+```{r}
+#| label: basics-stats-3
+#| echo: false
+
+# Number of uploaded images
+uploaded <- dets |> dplyr::filter(confidence > 0.1) |> dplyr::pull(image_id) |> 
+  unique() |> length()
+
+uploaded_fmt <- prettyNum(uploaded, big.mark = ",")
+
+# Check total number of images
+stopifnot(dets_tb_sum == length(unique(dets$image_id)))
+```
 
@@
 TBW
-
-<!-- Here's some example analysis code: -->
-
-```{r}
-#| label: get-data
-#| eval: false
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
-```
-
-```{r}
-#| label: fig-demo-plot
-#| fig-cap: "A plot of random numbers"
-plot(rnorm(10))
-```
-
-Figure @fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`
-
-```{r}
-#| label: demo-inline-code
-#| echo: false
-x <- round(pi, 2)
-```
 

### 2023-02-22 scripts/1_preprocess_annotations.R
remove deadcode

@@ -39,19 +39,3 @@ for (file in file_list){
 }
 
 saveRDS(annotations, "analysis/data/derived_data/annotations.rds")
-
-# -------------------------------------------------------------------------
-
-# TODO dead code
-
-# anns <- readRDS("data/anns.rds")
-# Check for duplicates (should have been caught by python code)
-# test <- anns %>% dplyr::filter(variable == "species" |
-#                                  variable == "other_species",
-#                                tag != "Human",
-#                                tag != "Vehicle")  %>%
-#   dplyr::group_by(id, type, origin, to_name, original_width, original_height,
-#                   value_x, value_y, value_width, value_height, source_file,
-#                   value_text, variable) %>%
-#   dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
-#   dplyr::filter(n > 1L)

### 2023-02-23 scripts/1_preprocess_annotations.R
add wide format

@@ -39,3 +39,20 @@ for (file in file_list){
 }
 
 saveRDS(annotations, "analysis/data/derived_data/annotations.rds")
+
+# Wide format -------------------------------------------------------------
+
+annotations_wide <- annotations |>
+  dplyr::select(-type) |>
+  dplyr::mutate(manual = stringr::str_detect(id, "_M")) |>
+  dplyr::mutate(image_id = stringr::str_split(id, "_")) |>
+  dplyr::mutate(image_id =
+                  unlist(purrr::map(image_id,
+                                    ~paste0(.x[1:5], collapse = "_")))) |>
+  tidyr::pivot_wider(names_from = "variable", values_from = "tag") |>
+  dplyr::filter(is.na(value_text)) |>
+  dplyr::mutate (species = ifelse(species == 'Other [See more species]',
+                                  other_species, species)) |>
+  dplyr::select(-other_species)
+
+saveRDS(annotations_wide, "analysis/data/derived_data/annotations_wide.rds")

### 2023-02-23 analysis/paper/paper.qmd
updates

@@
   - Valentin Lucet:
-      correspondence: "yes"
+      # correspondence: "yes"
       email: valentin.lucet@gmail.com
@@
   docx:
-    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
+    reference-doc: "../templates/template_edited.docx" # Insert path for the DOCX file
 execute:
@@
 
+\newpage
+
 ## Introduction
@@
 
-if (!file.exists("analysis/data/derived_data/annotations.rds")){
+if (!file.exists("analysis/data/derived_data/annotations.rds") |
+    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
   source("scripts/1_preprocess_annotations.R")
@@
   anns <- readRDS("analysis/data/derived_data/annotations.rds")
+  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
 }
@@
 dets_tb_fmt <- prettyNum(dets_tb, big.mark = ",")
+
+# Check total number of images
+stopifnot(dets_tb_sum == length(unique(dets$image_id)))
 ```
@@
 
-Using a cutt-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+Using a cut-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
 The totals for each category are: __`r dets_tb_cat_fmt[1]`__ animals (or __`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (or __`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (or __`r dets_tb_cat_percent[3]` %__).
@@
 
-- Deployable online. This was to allow for for multiple taggers
-- Flexible tagging interface. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
-- Allowing to tag multiple bounding boxes per image.
+- *Deployable online*. This was to allow for for multiple taggers
+- *Flexible tagging interface*. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
+- *Allowing to tag multiple bounding boxes per image*.
 Most platforms only allow for tagging the entire image, and not for portions of it (e.g.
@@
 This is necessary for when we will crop our images to train our neural network model.
-- Optionally, open source, to allow for maximum reproducibility of the process.
+- *Optionally, open source,* to allow for maximum reproducibility of the process.
 
@@
 
-# Check total number of images
-stopifnot(dets_tb_sum == length(unique(dets$image_id)))
+# Number of images viewed
+viewed <- length(unique(anns_wide$image_id))
+viewed_fmt <- prettyNum(viewed, big.mark = ",")
+
+# Percentage of image viewed (in the total)
+viewed_pct <- viewed/length(unique(dets$image_id))*100
+
+# Percentage of image viewed from what was uploaded to the platform
+viewed_pct_up <- length(unique(anns_wide$image_id))/uploaded*100
+
+# Filter for non empty images
+anns_wide_noempty <- anns_wide |>
+  dplyr::filter(!is.na(to_name))
+
+# Number of processed annotations
+processed <- nrow(anns_wide_noempty)
+processed_fmt <- prettyNum(processed, big.mark = ",")
+
+# Percentage of animal annotations processed
+processed_pct <- processed/dets_tb_cat[1]*100
+
+# Number of nonempty images viewed
+processed_img <- length(unique(anns_wide_noempty$image_id))
+processed_img_fmt <- prettyNum(processed_img, big.mark = ",")
+
+# Percentage of nonempty images viewed (in the total)
+processed_img_pct <- processed_img/length(unique(dets$image_id))*100
+
+# Percentage of nonempty images viewed (in the detected images > 0.1)
+processed_img_up_pct <- processed_img/uploaded*100
 ```
 
-Out of the 189,982 animal detections, 65,338 have been processed on the platform (34.39 %) for a total of 54,677 images (1.96 % of the total amount of images, or 4.24 % of the set of images with at least one detection, or 19.44 % of the total amount of images uploaded to the platform).
-This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and 82,165 images, (2.94 % of the total amount of images, or 29.21 % of the total amount of images uploaded to the platform).
+Out of the __`r dets_tb_cat_fmt[1]`__ animal detections, __`r processed_fmt`__ have been processed on the platform (or __`r processed_pct` %__) for a total of __`r processed_img_fmt`__ images (i.e.
+__`r processed_img_pct` %__ of the total amount of images, or __`r processed_img_up_pct` %__ of the total amount of images uploaded to the platform).
+This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and __`r viewed_fmt`__, (i.e.
+__`r viewed_pct` %__ of the total amount of images, or __`r viewed_pct_up` %__ of the total amount of images uploaded to the platform).
 
-Out of the 65,338 detections, 30,621 were false positives (46,87 %), and 33,237 were true positives (50.86 %), to which 1,480 false negatives were added (manually added bounding boxes, 2.26 %).
+```{r}
+#| label: mans
+#| echo: false
 
-- Percentage of predictions changed (no human)
+# Deleted predictions
+dels <- anns_wide_noempty |> dplyr::filter(is.na(species))
+dels_nb <- nrow(dels)
+dels_nb_fmt <- prettyNum(dels_nb, big.mark = ",")
+dels_nb_pct <- dels_nb/nrow(anns_wide_noempty)*100
+
+# Should be no manuals here
+stopifnot((anns_wide_noempty |> dplyr::filter(is.na(species)) 
+           |> dplyr::filter(manual)) == 0)
+
+# Actual annotations
+true_anns <- anns_wide_noempty |> dplyr::filter(!is.na(species))
+true_anns_nb <- nrow(true_anns)
+# true_anns_nb_fmt <- prettyNum(true_anns_nb, big.mark = ",")
+# true_anns_nb_pct <- true_anns_nb/processed*100
+
+# Non manuals 
+non_mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
+  dplyr::filter(!manual)
+non_mans_nb <- nrow(non_mans)
+non_mans_nb_fmt <- prettyNum(non_mans_nb, big.mark = ",")
+non_mans_nb_pct <- non_mans_nb/nrow(anns_wide_noempty)*100
+
+# Mans
+mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
+  dplyr::filter(manual)
+mans_nb <- nrow(mans)
+mans_nb_fmt <- prettyNum(mans_nb, big.mark = ",")
+mans_nb_pct <- mans_nb/nrow(anns_wide_noempty)*100
+
+# adjusted
+orig_tb <- anns_wide_noempty$origin |> table()
+orig_tb_pct <- orig_tb/sum(orig_tb)*100
+orig_tb_fmt <- prettyNum(orig_tb, big.mark = ",")
+```
+
+Out of the __`r processed_fmt`__ detections, __`r dels_nb_fmt`__ were false positives (or __`r dels_nb_pct` %__), and __`r non_mans_nb_fmt`__ were true positives (or __`r non_mans_nb_pct` %__), to which __`r mans_nb_fmt`__ false negatives were added (manually added bounding boxes, __`r mans_nb_pct` %__).
+Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted,
 
 ### Issues encountered
+
+```{r}
+#| label: quality-control
+#| eval: false
+#| echo: false
+
+# Only annotations for animals
+anns_wide_animals <- anns_wide_noempty %>% 
+  filter(species != 'Human') %>% 
+  filter(species != 'Vehicle')
+
+# Unknowns and bird breakdown
+anns_wide_animals %>% filter(species == "Unknown") %>% pull(id) %>% 
+  str_sub(1,4) %>% table() %>% sort()
+anns_wide_animals %>% filter(str_detect(species,"Bird")) %>% pull(id) %>% 
+  str_sub(1,4) %>% table() %>% sort()
+
+# Join
+anns_wide_animals_joined <- anns_wide_animals %>% 
+  right_join(info_images, by = "image_id")
+```
 
@@
 
+```{r}
+#| label: species
+#| eval: false
+
+# Summary by species
+
+anns_wide_animals_table <- anns_wide_animals %>% 
+  mutate(camera_id = anns_wide_animals$id %>% str_split("_") %>% 
+           map(~.x[[2]]) %>% unlist()) %>% 
+  dplyr::select(camera_id, species) %>% 
+  filter(species != "Unknown") %>%
+  table() %>% 
+  `>`(0) %>% 
+  `+`(0) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("camera_id") %>% 
+  arrange(camera_id)
+anns_wide_animals_table_tb <- anns_wide_animals_table %>% 
+  dplyr::select(-camera_id)
+rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
+
+```
+
 ### Ordination
@@
 TBW
+
+```{r}
+#| label: spatial
+#| eval: false
+
+RoF_CameraDeploymentLog_WLU <- 
+  read_excel("data/raw/WLU_Cameras_Log_2023.xlsx") %>% 
+  janitor::clean_names()
+
+RoF_CameraDeploymentLog_WLU_old <- 
+  read_excel("data/raw/RoF_CameraDeploymentLog_WLU.xlsx",
+             sheet = "Sheet1") %>% 
+  janitor::clean_names()
+
+cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
+  filter(!is.na(latitude)) %>% 
+  filter(!is.na(longitude)) %>% 
+  st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326")
+plot(cams_sf$geometry)
+
+lu <- raster("data/raw/CAN_LC_2015_CAL.tif")
+
+lu_dat <- read_csv("data/raw/lu.txt") %>% 
+  mutate(cats = as.factor(code))
+
+cams_sf_buffers <- cams_sf %>% 
+  st_transform(st_crs(lu)) %>% 
+  st_buffer(dist = 1000)
+
+extr <- exact_extract(lu, cams_sf_buffers)
+
+extr_list <- extr %>% 
+  lapply(FUN = function(x){
+    the_tab <- table(x$value)
+    props <- the_tab/sum(the_tab)
+    cats <- names(the_tab)
+    df <- tibble(cats, props)
+  })
+names(extr_list) <- cams_sf_buffers$camera_id
+extr_table <- bind_rows(extr_list, .id = "camera_id") %>% 
+  mutate(props = as.double(props)) %>% 
+  left_join(lu_dat, by="cats") %>% 
+  dplyr::select(camera_id,category_code, props) %>% 
+  # mutate(cats = paste0("cat_", cats)) %>% 
+  pivot_wider(names_from = "category_code", values_from = "props") %>% 
+  as.data.frame()
+extr_table[is.na(extr_table)] <- 0
+
+extr_table_sub <- extr_table %>% 
+  filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
+  arrange(camera_id)
+extr_table_sub_tb <- extr_table_sub %>% 
+  dplyr::select(-camera_id)
+rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
+
+# -------------------------------------------------------------------------
+
+# Ordination
+all(rownames(anns_wide_animals_table_tb)==rownames(extr_table_sub_tb))
+# plot(rda(anns_wide_animals_table_tb,extr_table_sub_tb))
+
+test_rda <- rda(anns_wide_animals_table_tb ~ .,
+                extr_table_sub_tb)
+
+sp_scores <- scores(test_rda,
+                    choices = 1:2,
+                    display = "sp"
+)
+
+plot(test_rda,
+     display = c("sp", "lc", "cn")
+)
+```
 

### 2023-02-23 analysis/paper/paper.qmd
update

@@
 
-- Add explanation about image bursts and upload of bursts including empties
+Finally, it is important to note that not only pictures with detections were uploaded to the platform.
+Because camera traps produce images in bursts, we included all the pictures from a given bursts if any of the images in the bursts include a detection with certainty above the threshold.
 

### 2023-02-27 analysis/paper/paper.qmd
progress on qmd

@@
 options(scipen = 1, digits = 2)
+library(magrittr)
 ```
@@
 
-### Sampling protocol
+### Camera deployment
 
 - Recycle the info from proposal
+
+```{r}
+#| label: spatial
+#| echo: false
+
+RoF_CameraDeploymentLog_WLU <- 
+  readxl::read_excel(paste0("analysis/data/raw_data/camera_deployments/",
+                            "WLU_Cameras_Log_2023.xlsx")) %>% 
+  janitor::clean_names()
+
+cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
+  dplyr::filter(!is.na(latitude)) %>% 
+  dplyr::filter(!is.na(longitude)) %>% 
+  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
+  dplyr::mutate(is_retrieved = is.na(date_retrieved))
+
+# can <- rnaturalearth::ne_countries(country = "Canada", returnclass = "sf") %>% 
+#   sf::st_crop(cams_sf)
+# 
+# can %>% 
+#   ggplot2::ggplot() +
+#   ggplot2::theme_bw() +
+#   ggplot2::geom_sf() +
+#   ggplot2::geom_sf(data=cams_sf)
+
+cams_sf %>%
+  ggplot2::ggplot() +
+  ggplot2::theme_bw() +
+  ggplot2::geom_sf(ggplot2::aes(color = is_retrieved))
+```
 
@@
 stopifnot(nb_image == 2788375)
+
+exif <- readr::read_csv("analysis/data/raw_data/exif/exif_combined.csv") %>% 
+  dplyr::select(-SourceFile) %>% 
+  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) %>% 
+  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) %>% 
+  dplyr::mutate(image_id = stringr::str_sub(image_id, 6)) %>% 
+  janitor::clean_names()
 ```
@@
 # Select and distinct down to image-based infos
-dets_images <- dets |>  dplyr::select(image_id, max_confidence, isempty) |> 
+dets_images <- dets %>%  dplyr::select(image_id, max_confidence, isempty) %>% 
   dplyr::distinct()
@@
 # Breakdown of which images triggered a MD detection
-dets_tb <- dets_images$isempty |> table()
+dets_tb <- dets_images$isempty %>% table()
 dets_tb_sum <- sum(dets_tb)
@@
 #| echo: false
-dets_0 <- dets |> dplyr::filter(confidence > 0)
-dets_0.1 <- dets |> dplyr::filter(confidence > 0.1)
+dets_0 <- dets %>% dplyr::filter(confidence > 0)
+dets_0.1 <- dets %>% dplyr::filter(confidence > 0.1)
 
 # Breakdown of detections by categories (animal, person, vehicles)
-dets_tb_cat_unfil <- dets_0 |> dplyr::pull(category_id) |> table()
+dets_tb_cat_unfil <- dets_0 %>% dplyr::pull(category_id) %>% table()
 dets_tb_cat_unfil_sum <- sum(dets_tb_cat_unfil)
@@
 # Breakdown of detections above 0.1 by categories (animal, persons, vehicles)
-dets_tb_cat <- dets_0.1 |> dplyr::pull(category_id) |> table()
+dets_tb_cat <- dets_0.1 %>% dplyr::pull(category_id) %>% table()
 dets_tb_cat_sum <- sum(dets_tb_cat)
@@
 # Histogram of confidence scores above 0.1
-dets_0.1 |> 
-  dplyr::mutate(category_id = as.character(category_id)) |> 
+dets_0.1 %>% 
+  dplyr::mutate(category_id = as.character(category_id)) %>% 
   dplyr::mutate(Type = dplyr::case_match(category_id,
@@
                                          "2" ~ "Person",
-                                         "3" ~ "Vehicle")) |> 
+                                         "3" ~ "Vehicle")) %>% 
   ggplot2::ggplot() +
@@
 # Number of uploaded images
-uploaded <- dets |> dplyr::filter(confidence > 0.1) |> dplyr::pull(image_id) |> 
-  unique() |> length()
+uploaded <- dets %>% dplyr::filter(confidence > 0.1) %>% dplyr::pull(image_id) %>% 
+  unique() %>% length()
 
@@
 # Filter for non empty images
-anns_wide_noempty <- anns_wide |>
+anns_wide_noempty <- anns_wide %>%
   dplyr::filter(!is.na(to_name))
@@
 # Deleted predictions
-dels <- anns_wide_noempty |> dplyr::filter(is.na(species))
+dels <- anns_wide_noempty %>% dplyr::filter(is.na(species))
 dels_nb <- nrow(dels)
@@
 # Should be no manuals here
-stopifnot((anns_wide_noempty |> dplyr::filter(is.na(species)) 
-           |> dplyr::filter(manual)) == 0)
+stopifnot((anns_wide_noempty %>% dplyr::filter(is.na(species)) 
+           %>% dplyr::filter(manual)) == 0)
 
 # Actual annotations
-true_anns <- anns_wide_noempty |> dplyr::filter(!is.na(species))
+true_anns <- anns_wide_noempty %>% dplyr::filter(!is.na(species))
 true_anns_nb <- nrow(true_anns)
@@
 # Non manuals 
-non_mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
+non_mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
   dplyr::filter(!manual)
@@
 # Mans
-mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
+mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
   dplyr::filter(manual)
@@
 # adjusted
-orig_tb <- anns_wide_noempty$origin |> table()
+orig_tb <- anns_wide_noempty$origin %>% table()
 orig_tb_pct <- orig_tb/sum(orig_tb)*100
@@
 #| label: quality-control
-#| eval: false
 #| echo: false
@@
 anns_wide_animals <- anns_wide_noempty %>% 
-  filter(species != 'Human') %>% 
-  filter(species != 'Vehicle')
+  dplyr::filter(species != 'Human') %>% 
+  dplyr::filter(species != 'Vehicle') %>% 
+  dplyr::mutate(camera_id = id %>% stringr::str_split("_") %>% 
+                  purrr::map(~.x[[2]]) %>% unlist())
 
 # Unknowns and bird breakdown
-anns_wide_animals %>% filter(species == "Unknown") %>% pull(id) %>% 
-  str_sub(1,4) %>% table() %>% sort()
-anns_wide_animals %>% filter(str_detect(species,"Bird")) %>% pull(id) %>% 
-  str_sub(1,4) %>% table() %>% sort()
+anns_wide_animals %>% dplyr::filter(species == "Unknown") %>% 
+  dplyr::pull(id) %>% 
+  stringr::str_sub(1,4) %>% table() %>% sort() 
+anns_wide_animals %>% dplyr::filter(stringr::str_detect(species,"Bird")) %>% 
+  dplyr::pull(id) %>% 
+  stringr::str_sub(1,4) %>% table() %>% sort()
 
@@
 anns_wide_animals_joined <- anns_wide_animals %>% 
-  right_join(info_images, by = "image_id")
+  dplyr::right_join(dets_images, by = "image_id")
 ```
 
-- Issues with vegetation falsely detected.
+The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation.
+
+### Analysis
+
+We present a summary of the species identified and an ordination of presence absence data, using proportions of land cover classes observed in a radius of 1 km around each camera location.
+We used the 2015 Land Cover of Canada dataset.
 
@@
 #| label: species
-#| eval: false
+#| echo: false
 
-# Summary by species
+# Summary by speciesanns_wide_noempty$origin %>% table()
 
 anns_wide_animals_table <- anns_wide_animals %>% 
-  mutate(camera_id = anns_wide_animals$id %>% str_split("_") %>% 
-           map(~.x[[2]]) %>% unlist()) %>% 
   dplyr::select(camera_id, species) %>% 
-  filter(species != "Unknown") %>%
+  dplyr::filter(species != "Unknown") %>%
   table() %>% 
@@
   as.data.frame() %>% 
-  rownames_to_column("camera_id") %>% 
-  arrange(camera_id)
+  tibble::rownames_to_column("camera_id") %>% 
+  dplyr::arrange(camera_id)
 anns_wide_animals_table_tb <- anns_wide_animals_table %>% 
@@
 rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
-
 ```
@@
 ```{r}
-#| label: spatial
-#| eval: false
+#| label: extract
+#| echo: false
 
-RoF_CameraDeploymentLog_WLU <- 
-  read_excel("data/raw/WLU_Cameras_Log_2023.xlsx") %>% 
-  janitor::clean_names()
+lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
+lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
 
-RoF_CameraDeploymentLog_WLU_old <- 
-  read_excel("data/raw/RoF_CameraDeploymentLog_WLU.xlsx",
-             sheet = "Sheet1") %>% 
-  janitor::clean_names()
+lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
+  dplyr::mutate(cats = as.factor(code))
 
-cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
-  filter(!is.na(latitude)) %>% 
-  filter(!is.na(longitude)) %>% 
-  st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326")
-plot(cams_sf$geometry)
+cams_sf_buffers_16 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_16)) %>% 
+  sf::st_buffer(dist = 1000)
 
-lu <- raster("data/raw/CAN_LC_2015_CAL.tif")
+cams_sf_buffers_17 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_17)) %>% 
+  sf::st_buffer(dist = 1000)
 
-lu_dat <- read_csv("data/raw/lu.txt") %>% 
-  mutate(cats = as.factor(code))
+extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16)
+extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17)
 
-cams_sf_buffers <- cams_sf %>% 
-  st_transform(st_crs(lu)) %>% 
-  st_buffer(dist = 1000)
+ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
+ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
 
-extr <- exact_extract(lu, cams_sf_buffers)
+stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
+
+extr_16[ind_16] <- extr_17[ind_16]
+
+extr <- extr_16
 
@@
     cats <- names(the_tab)
-    df <- tibble(cats, props)
+    df <- tibble::tibble(cats, props)
   })
-names(extr_list) <- cams_sf_buffers$camera_id
-extr_table <- bind_rows(extr_list, .id = "camera_id") %>% 
-  mutate(props = as.double(props)) %>% 
-  left_join(lu_dat, by="cats") %>% 
+names(extr_list) <- cams_sf_buffers_16$camera_id
+
+extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
+  dplyr::mutate(props = as.double(props)) %>% 
+  dplyr::left_join(lu_dat, by="cats") %>% 
   dplyr::select(camera_id,category_code, props) %>% 
-  # mutate(cats = paste0("cat_", cats)) %>% 
-  pivot_wider(names_from = "category_code", values_from = "props") %>% 
+  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
   as.data.frame()
@@
 extr_table_sub <- extr_table %>% 
-  filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
-  arrange(camera_id)
+  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
+  dplyr::arrange(camera_id)
 extr_table_sub_tb <- extr_table_sub %>% 
@@
 
-# -------------------------------------------------------------------------
+# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
+# raster::plot(lu_cropped)
+```
+
+```{r}
+#| label: ordination
+#| echo: false
 
 # Ordination
-all(rownames(anns_wide_animals_table_tb)==rownames(extr_table_sub_tb))
-# plot(rda(anns_wide_animals_table_tb,extr_table_sub_tb))
+stopifnot(all(rownames(anns_wide_animals_table_tb)==rownames(extr_table_sub_tb)))
 
-test_rda <- rda(anns_wide_animals_table_tb ~ .,
+test_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
                        extr_table_sub_tb)
-
-sp_scores <- scores(test_rda,
-                    choices = 1:2,
-                    display = "sp"
-)
 
@@
 )
+```
+
+```{r}
+#| label: ordination
+#| echo: false
+
+source("scripts/3_ordination.R")
+
+custom_rda_plot(test_rda)
+```
+
+### Events definition
+
+```{r}
+#| label: ordination
+#| echo: false
+
+anns_wide_animals_joined <- anns_wide_animals %>% 
+  dplyr::left_join(exif, by = c("source_file", "image_id"))
+
+events_tb <- anns_wide_animals_joined %>% 
+  dplyr::filter(species != "Unknown") %>%
+  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
+  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
+                                               tz = "Canada/Eastern"),
+                year = lubridate::year(date_time),
+                month = lubridate::month(date_time),
+                day = lubridate::day(date_time),
+                ymd = paste(year, month, day, sep = "_")) %>% 
+  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
+  # dplyr::select(camera_id, image_id, species, ymd) %>% 
+  dplyr::select(plot_id, species, ymd) %>% 
+  # dplyr::group_by(image_id, ind, ymd) %>% 
+  # dplyr::summarise(ind_count=dplyr::n())
+  dplyr::distinct() %>% 
+  dplyr::group_by(plot_id, species) %>%
+  dplyr::summarise(day_count=dplyr::n()) %>% 
+  dplyr::ungroup() %>% 
+  # dplyr::group_by(camera_id) %>% 
+  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
+  dplyr::arrange(plot_id) %>%
+  tibble::column_to_rownames("plot_id")
+
+events_tb[is.na(events_tb)] <- 0
+events_tb <- events_tb[,colSums(events_tb)>50]
+tmp_col <- ncol(events_tb)
+
+# events <- events/rowSums(events)
+events <- events_tb %>% 
+  tibble::rownames_to_column("plot_id")
+
+```
+
+### RDA based on events
+
+```{r}
+events_tb_cam <- anns_wide_animals_joined %>% 
+  dplyr::filter(species != "Unknown") %>%
+  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
+  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
+                                               tz = "Canada/Eastern"),
+                year = lubridate::year(date_time),
+                month = lubridate::month(date_time),
+                day = lubridate::day(date_time),
+                ymd = paste(year, month, day, sep = "_")) %>% 
+  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
+  # dplyr::select(camera_id, image_id, species, ymd) %>% 
+  dplyr::select(camera_id, species, ymd) %>% 
+  # dplyr::group_by(image_id, ind, ymd) %>% 
+  # dplyr::summarise(ind_count=dplyr::n())
+  dplyr::distinct() %>% 
+  dplyr::group_by(camera_id, species) %>%
+  dplyr::summarise(day_count=dplyr::n()) %>% 
+  dplyr::ungroup() %>% 
+  # dplyr::group_by(camera_id) %>% 
+  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
+  dplyr::arrange(camera_id) %>%
+  tibble::column_to_rownames("camera_id")
+
+events_tb_cam[is.na(events_tb_cam)] <- 0
+# events_tb_cam <- events_tb_cam[,colSums(events_tb_cam)>50]
+# tmp_col <- ncol(events_tb_cam)
+
+events_rda <- vegan::rda(vegan::decostand(events_tb_cam, "hellinger") ~ .,
+                         extr_table_sub_tb)
+
+custom_rda_plot(events_rda, sp_scale = 1.5)
+```
+
+```{r}
+coords <- cams_sf %>% 
+  sf::st_coordinates() %>% 
+  as.data.frame()
+cams_sf_with_props <- dplyr::bind_cols(cams_sf, coords) %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::select(plot_id,X,Y) %>% 
+  dplyr::group_by(plot_id) %>% 
+  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
+  dplyr::distinct() %>% 
+  dplyr::ungroup() %>% 
+  dplyr::left_join(events, by = "plot_id")
+
+ggplot2::ggplot() +
+  scatterpie::geom_scatterpie(data = cams_sf_with_props,
+                              ggplot2::aes(x=X, y=Y, group=plot_id),
+                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
+  ggplot2::coord_equal() #+ 
+  #ggplot2::theme(legend.position = "none")
+```
+
+```{r}
+zones <- 
+  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp") %>% 
+  sf::st_transform(crs = sf::st_crs(cams_sf))
+
+plot(zones$geometry)
+```
+
+```{r}
+zones_filter <- zones %>%
+  sf::st_filter(cams_sf)
+plot(zones_filter)
+```
+
+```{r}
+ggplot2::ggplot(zones_filter) +
+  ggplot2::theme_bw() +
+  ggplot2::geom_sf()
+```
+
+```{r}
+ggplot2::ggplot(zones_filter) +
+  ggplot2::theme_bw() +
+  ggplot2::geom_sf() +
+  ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
+                   ggplot2::aes(col = is_retrieved)) +
+  ggplot2::coord_sf(xlim = c(-89, -84),
+                    ylim = c(50, 55))
+```
+
+```{r}
+pols <- sf::st_coordinates(zones_filter) %>% 
+  as.data.frame()
+
+regions_labels <- zones_filter %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::bind_cols(data.frame(x = c(-85, -85, -87.5),
+                              y = c(54, 52, 52.7)))
+
+base <- ggplot2::ggplot() +
+  ggplot2::theme_bw()
+
+for (pol in 1:3){
+  base <- base + 
+    ggplot2::geom_path(data = dplyr::filter(pols, L3 == pol),
+                       ggplot2::aes(x=X, y=Y),
+                       linewidth = 0.2)
+}
+
+base +
+  ggplot2::coord_equal(xlim = c(-88, -84.5),
+                       ylim = c(50, 54)) +
+  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION), 
+                      data = regions_labels) +
+  scatterpie::geom_scatterpie(data = cams_sf_with_props,
+                              ggplot2::aes(x=X, y=Y, group=plot_id),
+                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
+                              sorted_by_radius = TRUE,
+                              legend_name = "Species",
+                              pie_scale = 1.5)
+```
+
+```{r}
+set.seed(123)
+long <- rnorm(50, sd=100)
+lat <- rnorm(50, sd=50)
+d <- data.frame(long=long, lat=lat)
+d <- with(d, d[abs(long) < 150 & abs(lat) < 70,])
+n <- nrow(d)
+d$region <- factor(1:n)
+d$A <- abs(rnorm(n, sd=1))
+d$B <- abs(rnorm(n, sd=2))
+d$C <- abs(rnorm(n, sd=3))
+d$D <- abs(rnorm(n, sd=4))
+d[1, 4:7] <- d[1, 4:7] * 3
+head(d)
+ggplot2::ggplot() + scatterpie::geom_scatterpie(ggplot2::aes(x=long, y=lat, group=region), data=d,
+                           cols=LETTERS[1:4]) + ggplot2::coord_equal()
 ```
@@
 ```
+
+```{r}
+# -------------------------------------------------------------------------
+
+# # Create ordered dataframe, and calculate time interval between images.
+# x1 <- x %>%
+#   # Sometimes VNA sneaks in here
+#   mutate(number_individuals = as.numeric(ifelse(number_individuals == "VNA", 1, number_individuals))) %>%
+#   # Amalgamate tags of same species in same image; currently broken into two separate rows
+#   group_by(location, {{datetime_col}}, common_name) %>%
+#   mutate(number_individuals = sum(number_individuals)) %>%
+#   distinct(location, {{datetime_col}}, common_name, number_individuals, .keep_all = TRUE) %>%
+#   ungroup() %>%
+#   # Order the dataframe
+#   arrange(project, location, {{datetime_col}}, common_name) %>%
+#   group_by(project, location, common_name) %>%
+#   # Calculate the time difference between subsequent images
+#   mutate(interval = int_length({{datetime_col}} %--% lag({{datetime_col}}))) %>%
+#   # Is this considered a new detection?
+#   mutate(new_detection = ifelse(is.na(interval) | abs(interval) >= threshold, TRUE, FALSE)) %>%
+#   ungroup() %>%
+#   # Number independent detections
+#   mutate(detection = c(1, cumsum(new_detection[-1]) + 1))
+# 
+# # Summarise detections
+# x2 <- x1 %>%
+#   group_by(detection, project, location, common_name, scientific_name) %>%
+#   summarise(start_time = min({{datetime_col}}),
+#             end_time = max({{datetime_col}}),
+#             total_duration_seconds = int_length(start_time %--% end_time),
+#             n_images = n(),
+#             avg_animals = mean(number_individuals),
+#             max_animals = max(number_individuals))
+```
+
 -->

### 2023-02-27 analysis/paper/paper.qmd
updates

@@
 
-events_rda <- vegan::rda(vegan::decostand(events_tb_cam, "hellinger") ~ .,
+events_tb_cam_hell <- vegan::decostand(events_tb_cam, "hellinger")
+
+events_rda <- vegan::rda( events_tb_cam_hell ~ .,
                          extr_table_sub_tb)
 
+# vegan::RsquareAdj(events_rda)
+
 custom_rda_plot(events_rda, sp_scale = 1.5)
+```
+
+### Clustering
+
+```{r}
+clust <- hclust(cluster::daisy(events_tb_cam_hell, "gower"), 
+                method = "ward.D2")
+plot(clust)
+```
+
+```{r}
+clust_grp <- cutree(clust, k=5)
+custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_grp)
 ```

### 2023-02-27 scripts/3_ordination.R

@@ -1,0 +1,82 @@
+custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA) {
+
+  mult <- # 1.7
+    vegan::ordiArrowMul(vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2))
+
+  sc_si <- vegan::scores(the_rda, display="sites", choices=c(1,2), scaling=2) %>%
+    as.data.frame() %>%
+    tibble::rownames_to_column("name")
+  sc_sp <- (vegan::scores(the_rda, display="species", choices=c(1,2), scaling=2) %>%
+              as.data.frame() * sp_scale) %>%
+    tibble::rownames_to_column("name") %>%
+    dplyr::mutate(highlight =ifelse((abs(RDA1) > 0.4 | abs(RDA2) > 0.4),
+                                    name, ""))
+  sc_bp <- (vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2) %>%
+              as.data.frame() * mult) %>%
+    tibble::rownames_to_column("name")
+
+  base <-  ggplot2::ggplot() +
+    ggplot2::theme_bw() +
+    ggplot2::coord_fixed(xlim = c(-1.5, 1.5)) +
+    ggplot2::geom_hline(yintercept=0, lty = 3) +
+    ggplot2::geom_vline(xintercept=0, lty = 3) +
+
+    ggplot2::geom_point(data = sc_sp,
+                        ggplot2::aes(x=RDA1, y=RDA2),
+                        color = "coral4",
+                        pch = 3) +
+
+    ggplot2::geom_segment(data = sc_bp, color = "steelblue4",
+                          arrow = ggplot2::arrow(length = ggplot2::unit(0.010, "npc")),
+                          ggplot2::aes(x=0, y=0,
+                                       xend=RDA1, yend=RDA2)) +
+
+    ggrepel::geom_text_repel(data = sc_bp,
+                             ggplot2::aes(x=RDA1, y=RDA2, label = name),
+                             family = "Poppins",
+                             size = 3,
+                             min.segment.length = 10,
+                             seed = 5,
+                             direction = "both",
+                             # arrow = arrow(length = unit(0.010, "npc")),
+                             # nudge_x = .15,
+                             # nudge_y = .5,
+                             color = "steelblue4") +
+    ggrepel::geom_text_repel(data = sc_sp,
+                             ggplot2::aes(x=RDA1, y=RDA2, label = highlight),
+                             family = "Poppins",
+                             size = 3,
[…]

### 2023-02-28 analysis/paper/paper.qmd
updates

@@
   sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
-  dplyr::mutate(is_retrieved = is.na(date_retrieved))
+  dplyr::mutate(is_retrieved = !is.na(date_retrieved))
 
@@
 ```{r}
-#| label: ordination
+#| label: ordination-base
 #| echo: false
@@
 ```{r}
-#| label: ordination
+#| label: ordination-custom
 #| echo: false
@@
 ```{r}
-#| label: ordination
+#| label: events
 #| echo: false
@@
 ```{r}
-events_tb_cam <- anns_wide_animals_joined %>% 
+#| label: ordination-events
+#| echo: false
+
+anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
   dplyr::filter(species != "Unknown") %>%
@@
                 day = lubridate::day(date_time),
-                ymd = paste(year, month, day, sep = "_")) %>% 
+                ymd = paste(year, month, day, sep = "_"))
+
+events_tb_cam <- anns_wide_animals_joined_mut %>% 
   # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
@@
 ```{r}
-clust <- hclust(cluster::daisy(events_tb_cam_hell, "gower"), 
-                method = "ward.D2")
+#| label: cluster-base
+#| echo: false
+
+events_gower <- cluster::daisy(events_tb_cam_hell, "gower")
+clust <- hclust(events_gower, method = "ward.D2")
 plot(clust)
@@
 ```{r}
+#| label: rda-clust
+#| echo: false
+
 clust_grp <- cutree(clust, k=5)
@@
 ```{r}
+#| label: test-pie
+#| echo: false
+
 coords <- cams_sf %>% 
@@
   as.data.frame()
-cams_sf_with_props <- dplyr::bind_cols(cams_sf, coords) %>% 
+
+cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)
+
+cams_sf_with_props <- cams_sf_coords %>% 
   sf::st_drop_geometry() %>% 
@@
 ```{r}
+#| label: ecoregions
+#| echo: false
+
 zones <- 
@@
 ```{r}
+#| label: ecoregions-basic
+#| echo: false
+
 zones_filter <- zones %>%
@@
 ```{r}
+#| label: ecoregions-ggplot
+#| echo: false
+
 ggplot2::ggplot(zones_filter) +
@@
 ```{r}
+#| label: ecoregions-ggplot-retrieved
+#| echo: false
+
 ggplot2::ggplot(zones_filter) +
@@
   ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
-                   ggplot2::aes(col = is_retrieved)) +
+                   ggplot2::aes(col=is_retrieved)) +
   ggplot2::coord_sf(xlim = c(-89, -84),
@@
 ```{r}
+#| label: map-pies
+#| echo: false
+
 pols <- sf::st_coordinates(zones_filter) %>% 
@@
   base <- base + 
-    ggplot2::geom_path(data = dplyr::filter(pols, L3 == pol),
+    ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
                           ggplot2::aes(x=X, y=Y),
-                       linewidth = 0.2)
+                          col = 1, fill = "grey90")
 }
@@
                               legend_name = "Species",
-                              pie_scale = 1.5)
+                              pie_scale = 1.5) +
+  ggplot2::labs(x="Lat", y="Long")
 ```
@@
 ```{r}
-set.seed(123)
-long <- rnorm(50, sd=100)
-lat <- rnorm(50, sd=50)
-d <- data.frame(long=long, lat=lat)
-d <- with(d, d[abs(long) < 150 & abs(lat) < 70,])
-n <- nrow(d)
-d$region <- factor(1:n)
-d$A <- abs(rnorm(n, sd=1))
-d$B <- abs(rnorm(n, sd=2))
-d$C <- abs(rnorm(n, sd=3))
-d$D <- abs(rnorm(n, sd=4))
-d[1, 4:7] <- d[1, 4:7] * 3
-head(d)
-ggplot2::ggplot() + scatterpie::geom_scatterpie(ggplot2::aes(x=long, y=lat, group=region), data=d,
-                           cols=LETTERS[1:4]) + ggplot2::coord_equal()
+#| label: map-groups
+#| echo: false
+
+grps_df <- data.frame(camera_id = names(clust_grp),
+                      grp = as.factor(clust_grp))
+
+cams_sf_clust <- cams_sf %>% 
+  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
+  dplyr::left_join(grps_df, by = "camera_id") %>% 
+  dplyr::filter(!(is.na(grp)))
+
+ggplot2::ggplot(zones_filter) +
+  ggplot2::theme_bw() +
+  ggplot2::geom_sf() +
+  ggplot2::geom_sf(data = cams_sf_clust, cex = 1, 
+                   ggplot2::aes(col = grp)) +
+  ggplot2::coord_sf(xlim = c(-89, -84),
+                    ylim = c(50, 55))
+```
+
+```{r}
+#| label: clust-geo
+#| echo: false
+
+# Spatially constrained clust
+
+events_gower_mod <- events_gower
+class(events_gower_mod) <- "dist"
+
+tree <- ClustGeo::hclustgeo(events_gower_mod)
+# plot(tree)
+# rect.hclust(tree ,k = 5, border = c(4,5,3,2,1))
+# legend("topright", legend = paste("cluster",1:5), 
+#        fill=1:5,bty= "n", border = "white")
+
+dist_geo <- cams_sf_clust %>% 
+  dplyr::arrange(camera_id) %>% 
+  sf::st_distance() %>% 
+  as.dist()
+
+range.alpha <- seq(0,1,0.1)
+K <- 5
+cr <- ClustGeo::choicealpha(events_gower_mod, dist_geo, range.alpha, 
+  K, graph = F)
+
+clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
+                                 dist_geo, alpha = 1)
+clust_geo_grp <- cutree(clust_geo, k=K)
+
+grps_geo_df <- data.frame(camera_id = names(clust_geo_grp),
+                          grp = as.factor(clust_geo_grp))
+
+cams_sf_geo_clust <- cams_sf %>% 
+  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
+  dplyr::left_join(grps_geo_df, by = "camera_id") %>% 
+  dplyr::filter(!(is.na(grp)))
+
+plot(clust_geo)
+```
+
+```{r}
+#| label: clust-geo-plot
+#| echo: false
+
+ggplot2::ggplot(zones_filter) +
+  ggplot2::theme_bw() +
+  ggplot2::geom_sf() +
+  ggplot2::geom_sf(data = cams_sf_geo_clust, cex = 1, 
+                   ggplot2::aes(col = grp)) +
+  ggplot2::coord_sf(xlim = c(-89, -84),
+                    ylim = c(50, 55))
+```
+
+```{r}
+#| label: rda-clust-geo
+#| echo: false
+
+custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
+```
+
+### Density heatmaps
+
+```{r}
+#| label: join-sf
+#| echo: false
+
+cams_sf_sp <- anns_wide_animals_joined_mut %>% 
+  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
+  dplyr::filter(species %in% colnames(events[,-1]))
+
+cams_sf_sp_sum <- cams_sf_sp %>% 
+  dplyr::group_by(camera_id, X, Y, species, month) %>% 
+  dplyr::summarise(count=dplyr::n()) %>% 
+  dplyr::ungroup()
+# stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])
+
+# cams_sf_sp %>% 
+cams_sf_sp_sum %>% 
+  # dplyr::filter(species %in% colnames(events[,-1])) %>% 
+  dplyr::filter(species == "Caribou") %>% 
+  # dplyr::filter(species == "Caribou", month == 8) %>% 
+  ggplot2::ggplot() +
+  # ggplot2::geom_hex(ggplot2::aes(x=X, y=Y), bins = 70) +
+  # ggplot2::stat_density_2d(ggplot2::aes(fill = ..level..), geom = "polygon", colour="white") +
+  # ggplot2::geom_density_2d(ggplot2::aes(x=X, y=Y)) +
+  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
+  #                          geom = "raster", contour = T) +
+  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
+  #                          geom = "raster", contour = F) +
+  # ggplot2::geom_tile(ggplot2::aes(x=X, y=Y, fill = count),interpolate=TRUE) +
+  ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size = count)) +
+  # ggplot2::scale_fill_continuous(type = "viridis") +
+  # ggplot2::scale_fill_distiller(palette=4, direction=1) +
+  # ggplot2::theme_bw() + 
+  ggplot2::facet_wrap(~month) #+
+  # ggplot2::scale_x_continuous(expand = c(0, 0)) +
+  # ggplot2::scale_y_continuous(expand = c(0, 0)) # +
+  # ggplot2::theme(
+  #   legend.position='none'
+  # )
+```
+
+```{r}
+ggplot2::ggplot(ggplot2::faithfuld, ggplot2::aes(waiting, eruptions)) +
+ ggplot2::geom_raster(ggplot2::aes(fill = density))
 ```

### 2023-02-28 scripts/3_ordination.R

@@ -65,9 +65,11 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA) {
       dplyr::left_join(grps, by = "name")
 
     base <- base +
-      ggplot2::geom_text(data = sc_si,
-                         ggplot2::aes(x=RDA1, y=RDA2, label=name, color=grp),
-                         size = 1.8)
+      ggplot2::geom_point(data = sc_si,
+                         ggplot2::aes(x=RDA1, y=RDA2, alpha=0.5, color=grp),
+                         size = 1) +
+      ggplot2::stat_ellipse(data = sc_si,
+                            ggplot2::aes(x=RDA1, y=RDA2, color=grp))
 
   } else {
     base <- base +

### 2023-03-02 analysis/paper/paper.qmd
updates

@@
   sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
-  dplyr::mutate(is_retrieved = !is.na(date_retrieved))
+  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))
 
@@
 #   ggplot2::geom_sf(data=cams_sf)
-
-cams_sf %>%
-  ggplot2::ggplot() +
-  ggplot2::theme_bw() +
-  ggplot2::geom_sf(ggplot2::aes(color = is_retrieved))
 ```
@@
 
-# Join
+# Joins
+
 anns_wide_animals_joined <- anns_wide_animals %>% 
-  dplyr::right_join(dets_images, by = "image_id")
+  dplyr::left_join(exif, by = c("source_file", "image_id"))
+
+anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
+  dplyr::filter(species != "Unknown") %>%
+  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
+  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
+                                               tz = "Canada/Eastern"),
+                year = lubridate::year(date_time),
+                month = lubridate::month(date_time, label = TRUE),
+                day = lubridate::day(date_time),
+                ymd = paste(year, month, day, sep = "_"))
+
+# anns_wide_animals_joined_dets <- anns_wide_animals %>% 
+#   dplyr::right_join(dets_images, by = "image_id")
 ```
@@
 
-extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16)
-extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17)
+extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
+                                        progress = FALSE)
+extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
+                                        progress = FALSE)
 
@@
 ```{r}
-#| label: ordination-base
+#| label: RDA-presence-absence
 #| echo: false
+
+source("scripts/3_ordination.R")
 
@@
 
-test_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
+pa_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
                        extr_table_sub_tb)
 
-plot(test_rda,
-     display = c("sp", "lc", "cn")
-)
-```
-
-```{r}
-#| label: ordination-custom
-#| echo: false
-
-source("scripts/3_ordination.R")
-
-custom_rda_plot(test_rda)
+custom_rda_plot(pa_rda)
 ```
@@
 
-anns_wide_animals_joined <- anns_wide_animals %>% 
-  dplyr::left_join(exif, by = c("source_file", "image_id"))
-
-events_tb <- anns_wide_animals_joined %>% 
-  dplyr::filter(species != "Unknown") %>%
-  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
-  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
-                                               tz = "Canada/Eastern"),
-                year = lubridate::year(date_time),
-                month = lubridate::month(date_time),
-                day = lubridate::day(date_time),
-                ymd = paste(year, month, day, sep = "_")) %>% 
+events_plot_days <- anns_wide_animals_joined_mut %>% 
   # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
@@
   # dplyr::summarise(ind_count=dplyr::n())
-  dplyr::distinct() %>% 
+  dplyr::distinct()
+
+events_plot <- events_plot_days %>% 
   dplyr::group_by(plot_id, species) %>%
   dplyr::summarise(day_count=dplyr::n()) %>% 
-  dplyr::ungroup() %>% 
+  dplyr::ungroup()
+
+events_plot_tb <- events_plot %>% 
   # dplyr::group_by(camera_id) %>% 
@@
 
-events_tb[is.na(events_tb)] <- 0
-events_tb <- events_tb[,colSums(events_tb)>50]
-tmp_col <- ncol(events_tb)
+events_plot_tb[is.na(events_plot_tb)] <- 0
+events_plot_tb_filt <- events_plot_tb[,colSums(events_plot_tb)>50]
+tmp_col <- ncol(events_plot_tb_filt)
 
 # events <- events/rowSums(events)
-events <- events_tb %>% 
+events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
   tibble::rownames_to_column("plot_id")
-
 ```
@@
 ```{r}
-#| label: ordination-events
+#| label: RDA-events
 #| echo: false
 
-anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
-  dplyr::filter(species != "Unknown") %>%
-  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
-  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
-                                               tz = "Canada/Eastern"),
-                year = lubridate::year(date_time),
-                month = lubridate::month(date_time),
-                day = lubridate::day(date_time),
-                ymd = paste(year, month, day, sep = "_"))
-
-events_tb_cam <- anns_wide_animals_joined_mut %>% 
+events_cam_days <- anns_wide_animals_joined_mut %>% 
   # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
   # dplyr::select(camera_id, image_id, species, ymd) %>% 
-  dplyr::select(camera_id, species, ymd) %>% 
+  dplyr::select(camera_id, species, month, ymd) %>% 
   # dplyr::group_by(image_id, ind, ymd) %>% 
   # dplyr::summarise(ind_count=dplyr::n())
-  dplyr::distinct() %>% 
+  dplyr::distinct()
+
+events_cam <- events_cam_days %>% 
   dplyr::group_by(camera_id, species) %>%
   dplyr::summarise(day_count=dplyr::n()) %>% 
-  dplyr::ungroup() %>% 
+  dplyr::ungroup()
+
+events_tb_cam <- events_cam %>% 
   # dplyr::group_by(camera_id) %>% 
@@
 events_tb_cam[is.na(events_tb_cam)] <- 0
-# events_tb_cam <- events_tb_cam[,colSums(events_tb_cam)>50]
+# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
 # tmp_col <- ncol(events_tb_cam)
@@
 
-events_rda <- vegan::rda( events_tb_cam_hell ~ .,
+events_rda <- vegan::rda(events_tb_cam_hell ~ .,
                          extr_table_sub_tb)
@@
 # vegan::RsquareAdj(events_rda)
-
-custom_rda_plot(events_rda, sp_scale = 1.5)
+# custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
+#                 thres_plot = 0.3)
 ```
@@
 clust <- hclust(events_gower, method = "ward.D2")
-plot(clust)
+# plot(clust)
 ```
@@
 clust_grp <- cutree(clust, k=5)
-custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_grp) 
+custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
+                thres_plot = 0.3, clust = clust_grp)
 ```
@@
   dplyr::ungroup() %>% 
-  dplyr::left_join(events, by = "plot_id")
+  dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")
 
-ggplot2::ggplot() +
-  scatterpie::geom_scatterpie(data = cams_sf_with_props,
-                              ggplot2::aes(x=X, y=Y, group=plot_id),
-                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
-  ggplot2::coord_equal() #+ 
+# ggplot2::ggplot() +
+#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
+#                               ggplot2::aes(x=X, y=Y, group=plot_id),
+#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
+#   ggplot2::coord_equal() #+ 
 #ggplot2::theme(legend.position = "none")
@@
 zones <- 
-  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp") %>% 
+  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
+              quiet=TRUE) %>% 
   sf::st_transform(crs = sf::st_crs(cams_sf))
 
-plot(zones$geometry)
+# plot(zones$geometry)
 ```
@@
   sf::st_filter(cams_sf)
-plot(zones_filter)
-```
-
-```{r}
-#| label: ecoregions-ggplot
-#| echo: false
-
-ggplot2::ggplot(zones_filter) +
-  ggplot2::theme_bw() +
-  ggplot2::geom_sf()
+# plot(zones_filter)
 ```
@@
 
-ggplot2::ggplot(zones_filter) +
+ggplot2::ggplot() +
   ggplot2::theme_bw() +
-  ggplot2::geom_sf() +
-  ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
+  ggplot2::geom_sf(data = zones_filter) +
+  ggplot2::geom_sf(data = sf::st_jitter(cams_sf, 0.05), cex = 0.5, 
                    ggplot2::aes(col=is_retrieved)) +
   ggplot2::coord_sf(xlim = c(-89, -84),
-                    ylim = c(50, 55))
+                    ylim = c(50, 55)) +
+  ggplot2::labs(col = "Camera \nRetrieved")
 ```
@@
   ggplot2::geom_sf() +
-  ggplot2::geom_sf(data = cams_sf_clust, cex = 1, 
+  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_clust, 0.05), cex = 1, 
                    ggplot2::aes(col = grp)) +
@@
 clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
-                                 dist_geo, alpha = 1)
+                                 dist_geo, alpha = 0.2)
 clust_geo_grp <- cutree(clust_geo, k=K)
@@
 
-plot(clust_geo)
+# plot(clust_geo)
 ```
@@
   ggplot2::geom_sf() +
-  ggplot2::geom_sf(data = cams_sf_geo_clust, cex = 1, 
+  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_geo_clust, 0.05), cex = 1, 
                    ggplot2::aes(col = grp)) +
@@
 
-custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
+# custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
 ```
@@
 
-cams_sf_sp <- anns_wide_animals_joined_mut %>% 
-  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
-  dplyr::filter(species %in% colnames(events[,-1]))
+# cams_sf_sp <- anns_wide_animals_joined_mut %>% 
+#   dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
+#   dplyr::filter(species %in% colnames(events[,-1]))
+
+cams_sf_sp <- events_cam_days %>% 
+  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) %>% 
+  dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))
 
 cams_sf_sp_sum <- cams_sf_sp %>% 
-  dplyr::group_by(camera_id, X, Y, species, month) %>% 
+  dplyr::group_by(camera_id, X, Y, geometry, species, month) %>% 
   dplyr::summarise(count=dplyr::n()) %>% 
-  dplyr::ungroup()
+  dplyr::ungroup() # %>% 
+  # dplyr::group_by(camera_id, X, Y, species) %>% 
+  # dplyr::mutate(count=count/sum(count)) %>% 
+  # dplyr::ungroup() 
 # stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])
-
-# cams_sf_sp %>% 
-cams_sf_sp_sum %>% 
-  # dplyr::filter(species %in% colnames(events[,-1])) %>% 
-  dplyr::filter(species == "Caribou") %>% 
-  # dplyr::filter(species == "Caribou", month == 8) %>% 
-  ggplot2::ggplot() +
-  # ggplot2::geom_hex(ggplot2::aes(x=X, y=Y), bins = 70) +
-  # ggplot2::stat_density_2d(ggplot2::aes(fill = ..level..), geom = "polygon", colour="white") +
-  # ggplot2::geom_density_2d(ggplot2::aes(x=X, y=Y)) +
-  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
-  #                          geom = "raster", contour = T) +
-  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
-  #                          geom = "raster", contour = F) +
-  # ggplot2::geom_tile(ggplot2::aes(x=X, y=Y, fill = count),interpolate=TRUE) +
-  ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size = count)) +
-  # ggplot2::scale_fill_continuous(type = "viridis") +
-  # ggplot2::scale_fill_distiller(palette=4, direction=1) +
-  # ggplot2::theme_bw() + 
-  ggplot2::facet_wrap(~month) #+
-  # ggplot2::scale_x_continuous(expand = c(0, 0)) +
-  # ggplot2::scale_y_continuous(expand = c(0, 0)) # +
-  # ggplot2::theme(
-  #   legend.position='none'
-  # )
 ```
@@
 ```{r}
-ggplot2::ggplot(ggplot2::faithfuld, ggplot2::aes(waiting, eruptions)) +
- ggplot2::geom_raster(ggplot2::aes(fill = density))
+cams_sf_sp %>%
+  dplyr::filter(species == "Caribou", month == "Jun") %>%
+  ggplot2::ggplot(ggplot2::aes(X, Y)) +
+  ggplot2::geom_density_2d_filled(binwidth = 0.05) +
+  ggplot2::coord_equal(xlim = c(-89, -84),
+                       ylim = c(50, 55)) +
+  ggplot2::theme(
+    legend.position='none'
+  )
+```
+
+```{r}
+cams_sf_sp %>%
+  dplyr::filter(species == "Caribou") %>%
+  ggplot2::ggplot(ggplot2::aes(X, Y)) +
+  ggplot2::geom_density_2d_filled(breaks = c(seq(0,1,by=0.1),5, 10, 20, 50,100),
+                                  n=20) +
+  ggplot2::facet_wrap(~month) +
+  ggplot2::coord_equal(xlim = c(-89, -84),
+                       ylim = c(50, 55)) +
+  ggplot2::theme(
+    legend.position='none'
+  )
+```
+
+```{r}
+# cams_sf_sp_sum %>%
+#   # dplyr::filter(species == "Caribou") %>%
+#   ggplot2::ggplot() +
+#   ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size=count, col =species)) +
+#   ggplot2::facet_wrap(~month) +
+#   ggplot2::coord_equal(xlim = c(-89, -84),
+#                        ylim = c(50, 55))
+
+for (sp in unique(cams_sf_sp_sum$species)){
+
+  dat <- sf::st_as_sf(cams_sf_sp_sum) %>% 
+    dplyr::filter(species == sp)
+
+  gg <- ggplot2::ggplot(zones_filter) +
+    ggplot2::theme_bw() +
+    ggplot2::geom_sf() +
+    ggplot2::geom_sf(data = dat,
+                     ggplot2::aes(size = count)) +
+    ggplot2::coord_sf(xlim = c(-89, -84),
+                      ylim = c(50, 54)) +
+    ggplot2::facet_wrap(~month) +
+    ggplot2::labs(size="Days \ndetected")
+
+  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,".png"))
+}
+```
+
+```{r}
+# cams_sf_sp_sum %>%
+#   dplyr::filter(species == "Caribou") %>%
+#   ggplot2::ggplot() +
+#   ggplot2::geom_raster(ggplot2::aes(x=X, y=Y, z=count)) +
+#   ggplot2::facet_wrap(~month) +
+#   ggplot2::coord_equal(xlim = c(-89, -84),
+#                        ylim = c(50, 55))
 ```
@@
 
+```{r}
+#| label: ecoregions-ggplot
+#| echo: false
+
+# ggplot2::ggplot(zones_filter) +
+#   ggplot2::theme_bw() +
+#   ggplot2::geom_sf()
+```
 -->

### 2023-03-02 scripts/3_ordination.R

@@ -1,7 +1,11 @@
-custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA) {
+custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA,
+                            thres_plot=0.4,
+                            scale_x_sp=0.8, scale_y_sp=0.6,
+                            scale_x_bp=0.1, scale_y_bp=0.1) {
 
   mult <- # 1.7
-    vegan::ordiArrowMul(vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2))
+    vegan::ordiArrowMul(vegan::scores(the_rda, display="bp", choices=c(1, 2),
+                                      scaling=2))
 
   sc_si <- vegan::scores(the_rda, display="sites", choices=c(1,2), scaling=2) %>%
     as.data.frame() %>%
@@ -9,7 +13,7 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA) {
   sc_sp <- (vegan::scores(the_rda, display="species", choices=c(1,2), scaling=2) %>%
               as.data.frame() * sp_scale) %>%
     tibble::rownames_to_column("name") %>%
-    dplyr::mutate(highlight =ifelse((abs(RDA1) > 0.4 | abs(RDA2) > 0.4),
+    dplyr::mutate(highlight =ifelse((abs(RDA1) > thres_plot | abs(RDA2) > thres_plot),
                                     name, ""))
   sc_bp <- (vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2) %>%
               as.data.frame() * mult) %>%
@@ -27,21 +31,25 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA) {
                         pch = 3) +
 
     ggplot2::geom_segment(data = sc_bp, color = "steelblue4",
-                          arrow = ggplot2::arrow(length = ggplot2::unit(0.010, "npc")),
+                          arrow = ggplot2::arrow(
+                            length = ggplot2::unit(0.010, "npc")),
                           ggplot2::aes(x=0, y=0,
                                        xend=RDA1, yend=RDA2)) +
 
     ggrepel::geom_text_repel(data = sc_bp,
                              ggplot2::aes(x=RDA1, y=RDA2, label = name),
                              family = "Poppins",
-                             size = 3,
+                             size = 2.5,
                              min.segment.length = 10,
-                             seed = 5,
-                             direction = "both",
+                             seed = 77,
+                             # direction = "both",
[…]

### 2023-03-02 analysis/paper/paper.qmd
updates

@@
 ---
-title: "Ring of Fire Camera Detections Report"
+title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across Ontario’s Far North"
 author:
@@
   docx:
-    reference-doc: "../templates/template_edited.docx" # Insert path for the DOCX file
+    reference-doc: "../templates/template_modified.docx" # Insert path for the DOCX file
 execute:
@@
 abstract: |
-  Abstract TBW
+  Northern Ontario holds some of the world’s last intact wild places, but it is under pressure from resource extraction and climate change.
+Our understanding of biodiversity measures like species occurrence, abundance, diversity, and density for the region are lacking, yet needed for strategic decision-making about the impacts of resource extraction on ecosystem services and biodiversity.
+The use of non-invasive survey methods, such as camera trap arrays, has significantly changed our ability to quantify wildlife across large and remote regions.
+However, these methods generate a large amount of data and require tools to speed up the identification of species.
+Machine learning methods, such as deep neural networks, are a promising tool to overcome this data bottle-neck.
+Yet, variables like strong seasonality, changes in vegetation type, and species characteristics can strongly impact their efficacy.
+The extent to which these methods perform for northern regions is not well established.
+Using an array of cameras deployed in an heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
+We've recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
+
 # keywords: |
@@
 
-<!-- The actual document text starts here: -->
+<!-- The actual document text starts here:
 
@@
 
-\newpage
+\newpage -->
 
@@
 
-- What this report is about
-- Recycle the other introductions
-- Importance of monitoring
-- Data collection challenge
-- New methods to the rescue
-- Data challenges remain
-- AI plan
+The James Bay Lowlands in Northern Ontario hold some of the world’s last intact wild places but it is under pressure from resource extraction and global warming.
+Our understanding of basic biodiversity measures like species occurrence, abundance, diversity, and density for the region are still rudimental but needed to weigh ecological-economical decision tradeoffs.
+Traditionally, quantifying and monitoring biodiversity in northern regions required expensive and time-consuming field surveys.
+But today, the use of non-intrusive methods, such as arrays of 100s of camera traps, has significantly changed field practices for reliably quantifying wildlife across large and remote regions.
+However, these methods generate a large amount of data and require tools to speed up the identification, and processing, of image datasets.
+
+Machine learning methods, such as deep neural networks, are now in use to automate the process of species identification and help in that endeavour.
+We aim to evaluate the viability of deep neural networks for camera trap surveys currently taking place in Northern Ontario’s heterogeneous landscapes, including the Hudson Bay Lowlands and Ontario Shield ecozones, as variables like weather, seasonality, vegetation type, and species characteristics can impact the performance of these methods.
+There are currently no available machine learning models for identifying species from images in northern regions and developing one would be beneficial as a monitoring and modeling tool for northern boreal ecosystems.
+
+The extent to which these methods perform for northern regions is not well established.
+Using an array of cameras deployed in an heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
+We've recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
 
@@
 
-- Recycle the info from proposal
+The camera deployment protocol follows a spatially balanced hierarchical design that provides an even distribution of sample sites across biotic and abiotic conditions within the study area.
+The study sites span three ecoregions and two ecozones.
+The network was established by CWS in 2020 and seasonally equipped in a rolling design with 500 autonomous recording units (ARUs) to monitor audible species diversity paired with site-specific on-ground vegetation imagery.
+In 2022, 232 unbaited wildlife camera traps (CTs) were added to a subset of study sites to quantify mammalian occurrence and diversity.
+@fig-deployment shows the location of the and their current status.
+retrieved cameras were retrieved in September 2022.
 
 ```{r}
-#| label: spatial
+#| label: spatial-basics
 #| echo: false
@@
 
-# can <- rnaturalearth::ne_countries(country = "Canada", returnclass = "sf") %>% 
-#   sf::st_crop(cams_sf)
-# 
-# can %>% 
-#   ggplot2::ggplot() +
-#   ggplot2::theme_bw() +
-#   ggplot2::geom_sf() +
-#   ggplot2::geom_sf(data=cams_sf)
+coords <- cams_sf %>% 
+  sf::st_coordinates() %>% 
+  as.data.frame()
+
+cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)
+
+zones <- 
+  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
+              quiet=TRUE, ) %>% 
+  sf::st_transform(crs = sf::st_crs(cams_sf))
+
+zones_filter <- zones %>%
+  sf::st_filter(cams_sf)
+
+pols <- sf::st_coordinates(zones_filter) %>% 
+  as.data.frame()
+
+regions_labels <- zones_filter %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
+                              y = c(55, 52, 52.6)))
+
+map <- ggmap::get_stamenmap( bbox = c(left = -90, bottom = 50, right = -83, top = 56), 
+                             zoom = 9, maptype = "terrain-background")
+```
+
+```{r}
+#| label: fig-deployment
+#| echo: false
+#| fig-cap: "Camera deployment status"
+
+base <- ggmap::ggmap(map) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") +
+  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
+  ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
+
+p <- base +
+  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+                      size = 2.5, data = regions_labels) +
+  ggplot2::geom_point(data = cams_sf_coords, cex = 1, shape=21, col="black",
+                      ggplot2::aes(x=X, y=Y, fill=is_retrieved), 
+                      inherit.aes = FALSE) +
+  ggplot2::labs(fill = "Camera \nRetrieved")
+p
 ```
@@
 
+The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation.
+
+### Analysis
+
 ```{r}
@@
 # Unknowns and bird breakdown
-anns_wide_animals %>% dplyr::filter(species == "Unknown") %>% 
-  dplyr::pull(id) %>% 
-  stringr::str_sub(1,4) %>% table() %>% sort() 
-anns_wide_animals %>% dplyr::filter(stringr::str_detect(species,"Bird")) %>% 
-  dplyr::pull(id) %>% 
-  stringr::str_sub(1,4) %>% table() %>% sort()
+# anns_wide_animals %>% dplyr::filter(species == "Unknown") %>% 
+#   dplyr::pull(id) %>% 
+#   stringr::str_sub(1,4) %>% table() %>% sort() 
+# anns_wide_animals %>% dplyr::filter(stringr::str_detect(species,"Bird")) %>% 
+#   dplyr::pull(id) %>% 
+#   stringr::str_sub(1,4) %>% table() %>% sort()
 
@@
 
-The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation.
+#### Event definition
 
-### Analysis
+- days
 
-We present a summary of the species identified and an ordination of presence absence data, using proportions of land cover classes observed in a radius of 1 km around each camera location.
-We used the 2015 Land Cover of Canada dataset.
+#### Ordination and clustering
+
+- two types of ordination: pa and events
+- two clustering: basics and geo
 
@@
 
-TBW
+- Most common species
 
@@
   # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
-  # dplyr::select(camera_id, image_id, species, ymd) %>% 
   dplyr::select(plot_id, species, ymd) %>% 
-  # dplyr::group_by(image_id, ind, ymd) %>% 
-  # dplyr::summarise(ind_count=dplyr::n())
   dplyr::distinct()
@@
 events_plot_tb <- events_plot %>% 
-  # dplyr::group_by(camera_id) %>% 
-  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
   tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
@@
 
-# events <- events/rowSums(events)
 events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
@@
 events_cam_days <- anns_wide_animals_joined_mut %>% 
-  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
-  # dplyr::select(camera_id, image_id, species, ymd) %>% 
   dplyr::select(camera_id, species, month, ymd) %>% 
-  # dplyr::group_by(image_id, ind, ymd) %>% 
-  # dplyr::summarise(ind_count=dplyr::n())
   dplyr::distinct()
@@
 events_tb_cam <- events_cam %>% 
-  # dplyr::group_by(camera_id) %>% 
-  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
   tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
@@
 
-coords <- cams_sf %>% 
-  sf::st_coordinates() %>% 
-  as.data.frame()
-
-cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)
-
 cams_sf_with_props <- cams_sf_coords %>% 
@@
 ```{r}
-#| label: ecoregions
-#| echo: false
-
-zones <- 
-  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
-              quiet=TRUE) %>% 
-  sf::st_transform(crs = sf::st_crs(cams_sf))
-
-# plot(zones$geometry)
-```
-
-```{r}
-#| label: ecoregions-basic
-#| echo: false
-
-zones_filter <- zones %>%
-  sf::st_filter(cams_sf)
-# plot(zones_filter)
-```
-
-```{r}
-#| label: ecoregions-ggplot-retrieved
-#| echo: false
-
-ggplot2::ggplot() +
-  ggplot2::theme_bw() +
-  ggplot2::geom_sf(data = zones_filter) +
-  ggplot2::geom_sf(data = sf::st_jitter(cams_sf, 0.05), cex = 0.5, 
-                   ggplot2::aes(col=is_retrieved)) +
-  ggplot2::coord_sf(xlim = c(-89, -84),
-                    ylim = c(50, 55)) +
-  ggplot2::labs(col = "Camera \nRetrieved")
-```
-
-```{r}
 #| label: map-pies
@@
 
-pols <- sf::st_coordinates(zones_filter) %>% 
-  as.data.frame()
-
-regions_labels <- zones_filter %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::bind_cols(data.frame(x = c(-85, -85, -87.5),
-                              y = c(54, 52, 52.7)))
-
-base <- ggplot2::ggplot() +
+base2 <- ggplot2::ggplot() +
   ggplot2::theme_bw()
-
 for (pol in 1:3){
-  base <- base + 
+  base2 <- base2 + 
     ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
@@
 
-base +
-  ggplot2::coord_equal(xlim = c(-88, -84.5),
+base2 +
+  ggplot2::coord_equal(xlim = c(-88.5, -84),
                        ylim = c(50, 54)) +
-  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION), 
-                      data = regions_labels) +
   scatterpie::geom_scatterpie(data = cams_sf_with_props,
@@
 
-cams_sf_clust <- cams_sf %>% 
+cams_sf_clust <- cams_sf_coords %>% 
   # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
@@
 
-ggplot2::ggplot(zones_filter) +
-  ggplot2::theme_bw() +
-  ggplot2::geom_sf() +
-  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_clust, 0.05), cex = 1, 
-                   ggplot2::aes(col = grp)) +
-  ggplot2::coord_sf(xlim = c(-89, -84),
-                    ylim = c(50, 55))
+base_cut <- base +
+  ggplot2::coord_map(xlim = c(-89, -84),
+                     ylim = c(50, 54.5))
+
+base_cut +
+  ggplot2::geom_point(data = cams_sf_clust, cex = 1, shape=21, col="black",
+                      ggplot2::aes(x=X, y=Y, fill=grp), 
+                      inherit.aes = FALSE) +
+  ggplot2::labs(fill="Group")
 ```
@@
 
-cams_sf_geo_clust <- cams_sf %>% 
+cams_sf_geo_clust <- cams_sf_coords %>% 
   # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
@@
 
-ggplot2::ggplot(zones_filter) +
-  ggplot2::theme_bw() +
-  ggplot2::geom_sf() +
-  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_geo_clust, 0.05), cex = 1, 
-                   ggplot2::aes(col = grp)) +
-  ggplot2::coord_sf(xlim = c(-89, -84),
-                    ylim = c(50, 55))
+base_cut +
+  ggplot2::geom_point(data = cams_sf_geo_clust, cex = 1, shape=21, col="black",
+                      ggplot2::aes(x=X, y=Y, fill=grp), 
+                      inherit.aes = FALSE) +
+  ggplot2::labs(fill="Group")
 ```
@@
 
-### Density heatmaps
+### Density contourmaps
 
@@
 ```{r}
-cams_sf_sp %>%
-  dplyr::filter(species == "Caribou", month == "Jun") %>%
-  ggplot2::ggplot(ggplot2::aes(X, Y)) +
-  ggplot2::geom_density_2d_filled(binwidth = 0.05) +
-  ggplot2::coord_equal(xlim = c(-89, -84),
-                       ylim = c(50, 55)) +
-  ggplot2::theme(
-    legend.position='none'
-  )
+#| label: heatmap
+#| echo: false
+
+cams_sf_sp_filt <- cams_sf_sp %>% 
+  dplyr::filter(species == "Caribou", month == "Jun")
+
+base_cut +
+  ggplot2::geom_density2d_filled(
+    data = cams_sf_sp_filt, ggplot2::aes(X, Y),
+    binwidth = 0.05, alpha = 0.5)
 ```
@@
 ```{r}
-cams_sf_sp %>%
-  dplyr::filter(species == "Caribou") %>%
-  ggplot2::ggplot(ggplot2::aes(X, Y)) +
-  ggplot2::geom_density_2d_filled(breaks = c(seq(0,1,by=0.1),5, 10, 20, 50,100),
-                                  n=20) +
+#| label: heatmap-facets
+#| echo: false
+
+cams_sf_sp_caribou <- cams_sf_sp %>%
+  dplyr::filter(species == "Caribou")
+
+base_cut +
+  ggplot2::geom_density2d_filled(
+    data = cams_sf_sp_caribou, ggplot2::aes(X, Y),
+    alpha = 0.5,breaks = c(seq(0.001,1,by=0.1),5, 10, 20, 50,100), n=20) +
   ggplot2::facet_wrap(~month) +
@@
 ```{r}
+#| label: bubbleplots
+#| echo: false
+
 # cams_sf_sp_sum %>%
@@
 
-  gg <- ggplot2::ggplot(zones_filter) +
-    ggplot2::theme_bw() +
-    ggplot2::geom_sf() +
-    ggplot2::geom_sf(data = dat,
-                     ggplot2::aes(size = count)) +
-    ggplot2::coord_sf(xlim = c(-89, -84),
-                      ylim = c(50, 54)) +
+  gg <- base_cut +
+    ggplot2::geom_point(data = dat, col="black",
+                        ggplot2::aes(x=X, y=Y, size = count), 
+                        inherit.aes = FALSE) +
     ggplot2::facet_wrap(~month) +
@@
 
-  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,".png"))
+  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_bplot.png"))
 }
@@
 
+![alt text](../figures/Caribou_bplot.png "Caribou")
 ```{r}
+#| label: heatmaps
+#| echo: false
+
+for (sp in unique(cams_sf_sp_sum$species)){
+
+  cams_sf_sp_spe <- cams_sf_sp %>%
+    dplyr::filter(species == sp)
+
+  gg <- base_cut +
+    ggplot2::geom_density2d_filled(
+      data = cams_sf_sp_spe, ggplot2::aes(X, Y),
+      alpha = 0.5,breaks = c(seq(0.001,1,by=0.1),5, 10, 20, 50,100), n=20) +
+    ggplot2::facet_wrap(~month) +
+    ggplot2::coord_equal(xlim = c(-89, -84),
+                         ylim = c(50, 55)) +
+    ggplot2::theme(
+      legend.position='none'
+    )
+
+  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_heatmap.png"))
+}
+```
+![alt text](../figures/Caribou_heatmap.png "Caribou")
+
+```{r}
+#| label: plot-tests
+#| echo: false
+
 # cams_sf_sp_sum %>%
@@
 
-- Thank the volunteers for help with tagging
+- Thank the volunteers for help with tagging.
 
 <!-- The following line inserts a page break  -->
-
+<!--
 \newpage
@@
 
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
+The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  
+-->
 
@@
 
-<!-- ### Colophon
+ ### Colophon
 
@@
 
+<!-- 
 ```{r}

### 2023-03-03 analysis/paper/paper.qmd
update qmd

@@
 ---
-title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across Ontario’s Far North"
+title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North"
 author:
@@
 knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
+backup_options <- options()
 options(scipen = 1, digits = 2)
@@
 
-We looked at platforms not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners for which the task of tagging is a common step, known as labeling.
+We looked at platforms not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners for which the task of tagging is a common step, often referred to as labeling.
 We found two platforms that matched all those criteria: LabelStudio and CVAT.
@@
 
-### Issues encountered
-
-The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation.
-
-### Analysis
-
 ```{r}
@@
 
-#### Event definition
+<!-- ### Issues encountered
 
-- days
+The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation. -->
+
+### Analysis
+
+#### Species detections
+
+We analysed the results of the tagging process under two main forms: presence absence data for each camera; and relative species abundance measured as the number of days, for a given month, that a given species was detected by a given camera.
 
@@
 
-- two types of ordination: pa and events
-- two clustering: basics and geo
+We analysed the relative abundance data using well known numerical ecology community analysis methods (Legendre and Legendre 2013).
+As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
+We used Ward clustering on the relative abundances matrix.
+We extracted land cover information in a 1-km radius around each camera site, using the Far North Land Cover data set from Ontario's Ministry of Natural Resources and Forestry.
+We computed the relative abundances of land cover classes and used these as constraining variable in a redundancy analysis, a method well suited for multidimensional data.
 
@@
 
-- Most common species
+@fig-map-pies show relative species abundances at plot sites, for species detected more than 50 days in the season.
 
@@
 #| echo: false
-
-# Summary by speciesanns_wide_noempty$origin %>% table()
-
 anns_wide_animals_table <- anns_wide_animals %>% 
@@
 rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
+```
+
+```{r}
+#| label: events-Plots
+#| echo: false
+
+events_plot_days <- anns_wide_animals_joined_mut %>% 
+  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
+  dplyr::select(plot_id, species, ymd) %>% 
+  dplyr::distinct()
+
+events_plot <- events_plot_days %>% 
+  dplyr::group_by(plot_id, species) %>%
+  dplyr::summarise(day_count=dplyr::n()) %>% 
+  dplyr::ungroup()
+
+events_plot_tb <- events_plot %>% 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
+  dplyr::arrange(plot_id) %>%
+  tibble::column_to_rownames("plot_id")
+
+events_plot_tb[is.na(events_plot_tb)] <- 0
+events_plot_tb_filt <- events_plot_tb[,colSums(events_plot_tb)>50]
+tmp_col <- ncol(events_plot_tb_filt)
+
+events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
+  tibble::rownames_to_column("plot_id")
+```
+
+```{r}
+#| label: events-Cams
+#| echo: false
+
+events_cam_days <- anns_wide_animals_joined_mut %>% 
+  dplyr::select(camera_id, species, month, ymd) %>% 
+  dplyr::distinct()
+
+events_cam <- events_cam_days %>% 
+  dplyr::group_by(camera_id, species) %>%
+  dplyr::summarise(day_count=dplyr::n()) %>% 
+  dplyr::ungroup()
+
+events_tb_cam <- events_cam %>% 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
+  dplyr::arrange(camera_id) %>%
+  tibble::column_to_rownames("camera_id")
+
+events_tb_cam[is.na(events_tb_cam)] <- 0
+# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
+# tmp_col <- ncol(events_tb_cam)
+```
+
+```{r}
+#| label: fig-map-pies
+#| echo: false
+#| fig-cap: "Relative species abundances at plot sites (only species species detected more than 50 days in the season are shown)"
+
+cams_sf_with_props <- cams_sf_coords %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::select(plot_id,X,Y) %>% 
+  dplyr::group_by(plot_id) %>% 
+  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
+  dplyr::distinct() %>% 
+  dplyr::ungroup() %>% 
+  dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")
+
+# ggplot2::ggplot() +
+#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
+#                               ggplot2::aes(x=X, y=Y, group=plot_id),
+#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
+#   ggplot2::coord_equal() #+ 
+#ggplot2::theme(legend.position = "none")
+
+base2 <- ggplot2::ggplot() +
+  ggplot2::theme_bw()
+for (pol in 1:3){
+  base2 <- base2 + 
+    ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
+                          ggplot2::aes(x=X, y=Y),
+                          col = 1, fill = "grey90")
+}
+
+base2 +
+  ggplot2::coord_equal(xlim = c(-88.5, -84),
+                       ylim = c(50, 54)) +
+  scatterpie::geom_scatterpie(data = cams_sf_with_props,
+                              ggplot2::aes(x=X, y=Y, group=plot_id),
+                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
+                              sorted_by_radius = TRUE,
+                              legend_name = "Species",
+                              pie_scale = 1.5) +
+  ggplot2::labs(x="Lat", y="Long")
 ```
@@
 
-### Events definition
-
-```{r}
-#| label: events
-#| echo: false
-
-events_plot_days <- anns_wide_animals_joined_mut %>% 
-  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
-  dplyr::select(plot_id, species, ymd) %>% 
-  dplyr::distinct()
-
-events_plot <- events_plot_days %>% 
-  dplyr::group_by(plot_id, species) %>%
-  dplyr::summarise(day_count=dplyr::n()) %>% 
-  dplyr::ungroup()
-
-events_plot_tb <- events_plot %>% 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
-  dplyr::arrange(plot_id) %>%
-  tibble::column_to_rownames("plot_id")
-
-events_plot_tb[is.na(events_plot_tb)] <- 0
-events_plot_tb_filt <- events_plot_tb[,colSums(events_plot_tb)>50]
-tmp_col <- ncol(events_plot_tb_filt)
-
-events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
-  tibble::rownames_to_column("plot_id")
-```
-
 ### RDA based on events
@@
 #| echo: false
-
-events_cam_days <- anns_wide_animals_joined_mut %>% 
-  dplyr::select(camera_id, species, month, ymd) %>% 
-  dplyr::distinct()
-
-events_cam <- events_cam_days %>% 
-  dplyr::group_by(camera_id, species) %>%
-  dplyr::summarise(day_count=dplyr::n()) %>% 
-  dplyr::ungroup()
-
-events_tb_cam <- events_cam %>% 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
-  dplyr::arrange(camera_id) %>%
-  tibble::column_to_rownames("camera_id")
-
-events_tb_cam[is.na(events_tb_cam)] <- 0
-# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
-# tmp_col <- ncol(events_tb_cam)
-
 events_tb_cam_hell <- vegan::decostand(events_tb_cam, "hellinger")
@@
                 thres_plot = 0.3, clust = clust_grp)
-```
-
-```{r}
-#| label: test-pie
-#| echo: false
-
-cams_sf_with_props <- cams_sf_coords %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::select(plot_id,X,Y) %>% 
-  dplyr::group_by(plot_id) %>% 
-  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
-  dplyr::distinct() %>% 
-  dplyr::ungroup() %>% 
-  dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")
-
-# ggplot2::ggplot() +
-#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
-#                               ggplot2::aes(x=X, y=Y, group=plot_id),
-#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
-#   ggplot2::coord_equal() #+ 
-#ggplot2::theme(legend.position = "none")
-```
-
-```{r}
-#| label: map-pies
-#| echo: false
-
-base2 <- ggplot2::ggplot() +
-  ggplot2::theme_bw()
-for (pol in 1:3){
-  base2 <- base2 + 
-    ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
-                          ggplot2::aes(x=X, y=Y),
-                          col = 1, fill = "grey90")
-}
-
-base2 +
-  ggplot2::coord_equal(xlim = c(-88.5, -84),
-                       ylim = c(50, 54)) +
-  scatterpie::geom_scatterpie(data = cams_sf_with_props,
-                              ggplot2::aes(x=X, y=Y, group=plot_id),
-                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
-                              sorted_by_radius = TRUE,
-                              legend_name = "Species",
-                              pie_scale = 1.5) +
-  ggplot2::labs(x="Lat", y="Long")
 ```
@@
 #| echo: false
+#| eval: false
+
+options(backup_options)
 
@@
   dplyr::filter(species == "Caribou", month == "Jun")
+
+cams_sf_sp_sum_filt <- cams_sf_sp_sum %>% 
+  dplyr::filter(species == "Caribou", month == "Jun")
+
+mod <- mgcv::gam(count~s(X,Y, bs=""), data = cams_sf_sp_sum_filt)
+pred_grid <- 
+  tidyr::expand_grid(X=seq(-89, -84, by=0.1),
+                     Y=seq(50, 55, by=0.1))
+preds <- pred_grid %>% 
+  dplyr::mutate(count=predict(mod, pred_grid))
+
+rast <- raster::rasterFromXYZ(preds)
+raster::plot(rast)
+rast[rast<0] <- NA
+raster::plot(rast)
 
@@
     binwidth = 0.05, alpha = 0.5)
+
+base_cut +
+  ggplot2::geom_tile(
+    data = dplyr::filter(preds, count>=0), ggplot2::aes(X, Y, fill=count),
+    binwidth = 0.05, alpha = 0.5, inherit.aes = FALSE) +
+  ggplot2::scale_fill_viridis_c()
 ```
@@
 #| label: heatmap-facets
-#| echo: false
+#| eval: false
 
@@
 
-## Discussion
-
-### Choice of methods
-
-TBW
-
-### Future plans
-
-TBW
-
-## Conclusion
-
 ## Acknowledgements
 
-- Thank the volunteers for help with tagging.
+We thank the student volunteers at Wilfrid Laurier University for their help with tagging: [TODO: list volunteers here].
 

### 2023-03-03 analysis/paper/paper.qmd
improvements to text

@@
 Here is a citation [@Marwick2017] -->
-
-<!-- The actual document text starts here:
-
-## Summary
-
-TBW
-
-\newpage -->
 
@@
 
+<!-- ### Issues encountered
+
+The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation. -->
+
+### Analysis
+
+#### Species detections
+
+We analysed the results of the tagging process under two main forms: presence absence data for each camera; and relative species abundance measured as the number of days, for a given month, that a given species was detected by a given camera.
+
 ```{r}
@@
 
-<!-- ### Issues encountered
-
-The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation. -->
-
-### Analysis
-
-#### Species detections
-
-We analysed the results of the tagging process under two main forms: presence absence data for each camera; and relative species abundance measured as the number of days, for a given month, that a given species was detected by a given camera.
-
 #### Ordination and clustering
@@
 
+```{r}
+#| label: extract
+#| echo: false
+
+lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
+lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
+
+lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
+  dplyr::mutate(cats = as.factor(code))
+
+cams_sf_buffers_16 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_16)) %>% 
+  sf::st_buffer(dist = 1000)
+
+cams_sf_buffers_17 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_17)) %>% 
+  sf::st_buffer(dist = 1000)
+
+extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
+                                        progress = FALSE)
+extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
+                                        progress = FALSE)
+
+ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
+ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
+
+stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
+
+extr_16[ind_16] <- extr_17[ind_16]
+
+extr <- extr_16
+
+extr_list <- extr %>% 
+  lapply(FUN = function(x){
+    the_tab <- table(x$value)
+    props <- the_tab/sum(the_tab)
+    cats <- names(the_tab)
+    df <- tibble::tibble(cats, props)
+  })
+names(extr_list) <- cams_sf_buffers_16$camera_id
+
+extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
+  dplyr::mutate(props = as.double(props)) %>% 
+  dplyr::left_join(lu_dat, by="cats") %>% 
+  dplyr::select(camera_id,category_code, props) %>% 
+  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
+  as.data.frame()
+extr_table[is.na(extr_table)] <- 0
+
+extr_table_sub <- extr_table %>% 
+  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
+  dplyr::arrange(camera_id)
+extr_table_sub_tb <- extr_table_sub %>% 
+  dplyr::select(-camera_id)
+rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
+
+# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
+# raster::plot(lu_cropped)
+```
+
 ## Preliminary results
 
-### Species ID summary
+### Community data
 
@@
 
-### Ordination
+### Ordination - presence-absence
 
-TBW
+@fig-rda-pa shows the triplot for the RDA on presence-absence data.
 
 ```{r}
-#| label: extract
+#| label: fig-rda-pa
 #| echo: false
-
-lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
-lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
-
-lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
-  dplyr::mutate(cats = as.factor(code))
-
-cams_sf_buffers_16 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_16)) %>% 
-  sf::st_buffer(dist = 1000)
-
-cams_sf_buffers_17 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_17)) %>% 
-  sf::st_buffer(dist = 1000)
-
-extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
-                                        progress = FALSE)
-extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
-                                        progress = FALSE)
-
-ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
-ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
-
-stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
-
-extr_16[ind_16] <- extr_17[ind_16]
-
-extr <- extr_16
-
-extr_list <- extr %>% 
-  lapply(FUN = function(x){
-    the_tab <- table(x$value)
-    props <- the_tab/sum(the_tab)
-    cats <- names(the_tab)
-    df <- tibble::tibble(cats, props)
-  })
-names(extr_list) <- cams_sf_buffers_16$camera_id
-
-extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
-  dplyr::mutate(props = as.double(props)) %>% 
-  dplyr::left_join(lu_dat, by="cats") %>% 
-  dplyr::select(camera_id,category_code, props) %>% 
-  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
-  as.data.frame()
-extr_table[is.na(extr_table)] <- 0
-
-extr_table_sub <- extr_table %>% 
-  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
-  dplyr::arrange(camera_id)
-extr_table_sub_tb <- extr_table_sub %>% 
-  dplyr::select(-camera_id)
-rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
-
-# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
-# raster::plot(lu_cropped)
-```
-
-```{r}
-#| label: RDA-presence-absence
-#| echo: false
+#| fig-cap: "Triplot of RDA on presence-absence data"
 
@@
 # Ordination
-stopifnot(all(rownames(anns_wide_animals_table_tb)==rownames(extr_table_sub_tb)))
+stopifnot(all(rownames(anns_wide_animals_table_tb)==
+                rownames(extr_table_sub_tb)))
 
@@
 
-### RDA based on events
+### Ordination - relative abundances
+
+@fig-rda-events shows the triplot for the RDA on relative abundance data.
 
 ```{r}
-#| label: RDA-events
+#| label: fig-rda-events
 #| echo: false
+#| fig-cap: "Triplot of RDA on relative abundance data"
+
 events_tb_cam_hell <- vegan::decostand(events_tb_cam, "hellinger")
@@
 # vegan::RsquareAdj(events_rda)
-# custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
-#                 thres_plot = 0.3)
+custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
+                thres_plot = 0.3)
 ```
@@
 
+@fig-rda-events-clust shows the triplot for the RDA on relative abundance data,  with Ward clusters as ellipses.
+
 ```{r}
-#| label: rda-clust
+#| label: fig-rda-events-clust
 #| echo: false
+#| fig-cap: "Triplot of RDA on relative abundance data, with Ward clusters as ellipses"
 
@@
 
+@fig-map-clust shows a map of clustered camera sites.
+
 ```{r}
-#| label: map-groups
+#| label: fig-map-clust
 #| echo: false
+#| fig-cap: "Map of clustered camera sites"
 
@@
 ```
+### Spatial clustering
 
@@
 
+@fig-map-clust-geo shows a map of geographically constrained clustered camera sites.
+
 ```{r}
-#| label: clust-geo-plot
+#| label: fig-map-clust-geo
 #| echo: false
+#| fig-cap: "Map of geographically constrained clustered camera sites"
 
@@
 
-### Density contourmaps
+### Species heatmaps and bubbleplots
 
@@
 #| label: heatmap-facets
+#| echo: false
 #| eval: false
@@
 ![alt text](../figures/Caribou_bplot.png "Caribou")
+
 ```{r}
@@
 ![alt text](../figures/Caribou_heatmap.png "Caribou")
-
-```{r}
-#| label: plot-tests
-#| echo: false
-
-# cams_sf_sp_sum %>%
-#   dplyr::filter(species == "Caribou") %>%
-#   ggplot2::ggplot() +
-#   ggplot2::geom_raster(ggplot2::aes(x=X, y=Y, z=count)) +
-#   ggplot2::facet_wrap(~month) +
-#   ggplot2::coord_equal(xlim = c(-89, -84),
-#                        ylim = c(50, 55))
-```
 

### 2023-03-21 analysis/paper/paper.qmd
move code
update qmd

@@
 
+# zones <- 
+#   sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
+#               quiet=TRUE, ) %>% 
+#   sf::st_transform(crs = sf::st_crs(cams_sf))
+
 zones <- 
-  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
+  sf::st_read("analysis/data/raw_data/ecozones/EcoZone/EcoZone.shp",
               quiet=TRUE, ) %>% 
@@
 
+# regions_labels <- zones_filter %>% 
+#   sf::st_drop_geometry() %>% 
+# dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
+#                             y = c(55, 52, 52.6)))
+
 regions_labels <- zones_filter %>% 
   sf::st_drop_geometry() %>% 
-  dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
-                              y = c(55, 52, 52.6)))
+  dplyr::bind_cols(data.frame(x = c(-84.5, -88.5),
+                              y = c(52, 52.6)))
 
@@
 p <- base +
-  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+  #                     size = 2.5, data = regions_labels) +
+  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
                       size = 2.5, data = regions_labels) +
@@
                       ggplot2::aes(x=X, y=Y, fill=is_retrieved), 
+                      inherit.aes = FALSE) +
+  ggplot2::labs(fill = "Camera \nRetrieved")
+p
+```
+
+```{r}
+#| label: fig-deployment-poster
+#| eval: false
+
+base <- ggmap::ggmap(map) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") +
+  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
+  ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
+
+p <- base +
+  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+  #                     size = 2.5, data = regions_labels) +
+  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
+                      size = 2.5, data = regions_labels) +
+  ggplot2::geom_point(data = cams_sf_coords %>% 
+                        dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
+                      col="black", fill = "red",
+                      ggplot2::aes(x=X, y=Y), 
                       inherit.aes = FALSE) +
@@
 #| fig-cap: "Histogram of detection confidence scores."
+#| fig.width: 8
+#| fig.height: 4.5
 
@@
 
-We had in mind the following criteria when selecting the proper tool to tag our images and bounding boxes:
+We had the following criteria in mind when selecting an adequate tool to tag our images and bounding boxes:
 
@@
 ```
-
-#### Ordination and clustering
-
-We analysed the relative abundance data using well known numerical ecology community analysis methods (Legendre and Legendre 2013).
-As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
-We used Ward clustering on the relative abundances matrix.
-We extracted land cover information in a 1-km radius around each camera site, using the Far North Land Cover data set from Ontario's Ministry of Natural Resources and Forestry.
-We computed the relative abundances of land cover classes and used these as constraining variable in a redundancy analysis, a method well suited for multidimensional data.
-
-```{r}
-#| label: extract
-#| echo: false
-
-lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
-lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
-
-lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
-  dplyr::mutate(cats = as.factor(code))
-
-cams_sf_buffers_16 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_16)) %>% 
-  sf::st_buffer(dist = 1000)
-
-cams_sf_buffers_17 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_17)) %>% 
-  sf::st_buffer(dist = 1000)
-
-extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
-                                        progress = FALSE)
-extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
-                                        progress = FALSE)
-
-ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
-ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
-
-stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
-
-extr_16[ind_16] <- extr_17[ind_16]
-
-extr <- extr_16
-
-extr_list <- extr %>% 
-  lapply(FUN = function(x){
-    the_tab <- table(x$value)
-    props <- the_tab/sum(the_tab)
-    cats <- names(the_tab)
-    df <- tibble::tibble(cats, props)
-  })
-names(extr_list) <- cams_sf_buffers_16$camera_id
-
-extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
-  dplyr::mutate(props = as.double(props)) %>% 
-  dplyr::left_join(lu_dat, by="cats") %>% 
-  dplyr::select(camera_id,category_code, props) %>% 
-  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
-  as.data.frame()
-extr_table[is.na(extr_table)] <- 0
-
-extr_table_sub <- extr_table %>% 
-  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
-  dplyr::arrange(camera_id)
-extr_table_sub_tb <- extr_table_sub %>% 
-  dplyr::select(-camera_id)
-rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
-
-# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
-# raster::plot(lu_cropped)
-```
-
-## Preliminary results
-
-### Community data
-
-@fig-map-pies show relative species abundances at plot sites, for species detected more than 50 days in the season.
 
@@
 
+#### Ordination and clustering
+
+We analysed the relative abundance data using well known numerical ecology community analysis methods (Legendre and Legendre 2013).
+As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
+We used Ward clustering on the relative abundances matrix.
+We extracted land cover information in a 1-km radius around each camera site, using the Far North Land Cover data set from Ontario's Ministry of Natural Resources and Forestry.
+We computed the frequencies of land cover classes and used these as constraining variable in a redundancy analysis, a method well suited for multidimensional data.
+
+```{r}
+#| label: extract
+#| echo: false
+
+lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
+lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
+
+lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
+  dplyr::mutate(cats = as.factor(code))
+
+cams_sf_buffers_16 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_16)) %>% 
+  sf::st_buffer(dist = 1000)
+
+cams_sf_buffers_17 <- cams_sf %>% 
+  sf::st_transform(sf::st_crs(lu_17)) %>% 
+  sf::st_buffer(dist = 1000)
+
+extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
+                                        progress = FALSE)
+extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
+                                        progress = FALSE)
+
+ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
+ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
+
+stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
+
+extr_16[ind_16] <- extr_17[ind_16]
+
+extr <- extr_16
+
+extr_list <- extr %>% 
+  lapply(FUN = function(x){
+    the_tab <- table(x$value)
+    props <- the_tab/sum(the_tab)
+    cats <- names(the_tab)
+    df <- tibble::tibble(cats, props)
+  })
+names(extr_list) <- cams_sf_buffers_16$camera_id
+
+extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
+  dplyr::mutate(props = as.double(props)) %>% 
+  dplyr::left_join(lu_dat, by="cats") %>% 
+  dplyr::filter(!is.na(category_code)) %>% 
+  dplyr::select(camera_id,category_code, props) %>% 
+  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
+  as.data.frame()
+extr_table[is.na(extr_table)] <- 0
+
+# Combine the categories
+extr_table_simple <- extr_table %>% 
+  dplyr::mutate(WAT = WAT + XWAT,
+                MixTRE = MixTRE + DecTRE,
+                Dist  = TrOrSHr + NSWood, 
+                SWA = ThSWA + ConSWA, 
+                Anth = URB + MIN + BED) %>% 
+  dplyr::select(-XWAT,
+                -DecTRE,
+                -TrOrSHr, -NSWood, 
+                -ThSWA, - ConSWA, 
+                -URB, -MIN, -BED)
+
+extr_table_sub <- extr_table_simple %>% 
+  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
+  dplyr::arrange(camera_id)
+extr_table_sub_tb <- extr_table_sub %>% 
+  dplyr::select(-camera_id)
+rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
+extr_table_sub_tb <- extr_table_sub_tb[, colSums(extr_table_sub_tb) != 0]
+
+# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
+# raster::plot(lu_cropped)
+```
+
+```{r}
+sum(unlist(lapply(extr_list, function(x) sum(x$cats == "NA"))))
+sum(unlist(lapply(extr_list, function(x) is.na(x$cats))))
+#  [1] "1"   "2"   "8"   "9"   "12"  "15"  "16"  "17"  "18"  "11"  "13"  "14"  "19"  "20"  "23"  "157" "247" "21"  "22"
+
+pairs(extr_table_sub_tb, col = "dodgerblue")
+round(cor(extr_table_sub_tb), 2)
+```
+
+## Preliminary results
+
+### Community data
+
+Using the number of days detected as a measure of relative abundance, the most common species (detected more than 10 days in the season) are listed in Table 1 (the full table is available in the appendix at the end of this report).
+Note the large amount detections in the "bird" and "duck" species categories which could be further distinguished into specific species.
+
 ```{r}
@@
 ```{r}
+#| label: tab-species-table
+#| echo: false
+
+sp_tab <- dplyr::rename(.data = events_tb_cam %>% colSums() %>% 
+                          sort(decreasing = TRUE) %>% as.data.frame(), 
+                        days_detected=`.`) %>% 
+  tibble::rownames_to_column("species")
+
+sp_tab %>%
+  dplyr::filter(days_detected >= 10) %>%
+  gt::gt() %>%
+  gt::tab_header(
+    title = "Relative abundances for species above 10 detection days",
+    subtitle = "Computed in number of days detected in the season"
+  ) %>%
+  gt::fmt_number(
+    decimals = 0,
+    columns = days_detected,
+    suffixing = TRUE
+  ) %>% 
+  gt::cols_label(
+    days_detected ="Days",
+    species = "Species",
+  ) %>% 
+  gt::opt_stylize(style = 6, color = "gray")
+
+```
+
+As for the spatial distribution of those relative abundances, @fig-map-pies show relative species abundances across the landscape, for species detected more than 50 days in the season (note that the here the detections have grouped by plot number for better visibility, with a plot containing up to 4 cameras) .
+There are no clear spatial patterns, but some tendencies can be noted: Moose detections are predominant in the northwestern of the sampling array, while caribou and sandhill cranes detections share the most of the detections for the rest of the array.
+
+```{r}
 #| label: fig-map-pies
@@
 #| fig-cap: "Relative species abundances at plot sites (only species species detected more than 50 days in the season are shown)"
+#| fig.width: 8
+#| fig.height: 6
 
@@
 
-base2 <- ggplot2::ggplot() +
-  ggplot2::theme_bw()
-for (pol in 1:3){
-  base2 <- base2 + 
-    ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
-                          ggplot2::aes(x=X, y=Y),
-                          col = 1, fill = "grey90")
-}
-
-base2 +
+# base2 <- ggplot2::ggplot() +
+#   ggplot2::theme_bw()
+# for (pol in 1:2){
+#   base2 <- base2 + 
+#     ggplot2::geom_polygon(data = dplyr::filter(pols, L2 == pol),
+#                           ggplot2::aes(x=X, y=Y),
+#                           col = 1, fill = "grey90")
+# }
+# 
+# base2 +
+#   ggplot2::coord_equal(xlim = c(-88.5, -84),
+#                        ylim = c(50, 54)) +
+#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
+#                               ggplot2::aes(x=X, y=Y, group=plot_id),
+#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
+#                               sorted_by_radius = TRUE,
+#                               legend_name = "Species",
+#                               pie_scale = 1.5) +
+#   ggplot2::labs(x="Lat", y="Long")
+
+base +
   ggplot2::coord_equal(xlim = c(-88.5, -84),
@@
 
-@fig-rda-pa shows the triplot for the RDA on presence-absence data.
-
-```{r}
-#| label: fig-rda-pa
-#| echo: false
-#| fig-cap: "Triplot of RDA on presence-absence data"
+```{r}
+#| label: rda-pa
+#| echo: false
 
@@
 
+anns_wide_animals_table_tb_ch <- vegan::decostand(anns_wide_animals_table_tb, 
+                                                  "hellinger")
+
 pa_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
                      extr_table_sub_tb)
+```
+
+@fig-rda-pa shows the triplot for the RDA on presence-absence data, which yielded a low $R^2$ of `r vegan::RsquareAdj(pa_rda)[2]`, but which is still relevant for exploring species associations.
+Land cover might not be able to explain community structure, but it can still show which species are found together and around what kind of land cover class.
+Here, we see that caribou and sandhill cranes are associated with open and treed wetlands (bogs and fens) while black bear and moose are found together in more forested areas.
+Canada geese are, as is to be expected, associated with water while smaller species like squirrels and martens score away from the larger mammals in a mix of habitats.
+Note that a full legend of the land cover abbreviations are to be found in the appendix of this report.
+
+```{r}
+#| label: fig-rda-pa
+#| echo: false
+#| fig-cap: "Triplot of RDA on presence-absence data"
 
@@
 ```
+```{r}
+# Jaccard pcoa
+
+anns_wide_animals_table_tb_jac <- vegan::vegdist(anns_wide_animals_table_tb, 
+                                                 "jac", binary = TRUE)
+
+pa_pcoa <- ape::pcoa(anns_wide_animals_table_tb_jac)
+
+biplot(pa_pcoa, anns_wide_animals_table_tb)
+```
 
@@
 
-@fig-rda-events shows the triplot for the RDA on relative abundance data.
+```{r}
+#| label: rda-events
+#| echo: false
+#| fig-cap: "Triplot of RDA on relative abundance data"
+
+events_tb_cam_hell <- vegan::decostand(events_tb_cam, "hellinger")
+
+events_rda <- vegan::rda(events_tb_cam_hell ~ .,
+                         extr_table_sub_tb)
+```
+
+@fig-rda-events shows the triplot for the RDA on relative abundance data, which  yielded a slightly larger $R^2$ of `r vegan::RsquareAdj(events_rda)[2]`, but reduces the axis scores of the rarer species.
+We observe similar patterns for caribou, crane, moose, black bear and marten.
 
@@
 # plot(clust)
-```
-
+# vegan::cascadeKM(events_gower,2, 5) %>% plot()
+```
+
+Clustering allows us to find pattern in the ordination data.
 @fig-rda-events-clust shows the triplot for the RDA on relative abundance data, with Ward clusters as ellipses.
+We used the cascading K-means function `cascadeKM` in the `vegan` package to determine the number of relevant clusters to show (in this case, 2 clusters were found to be most relevant.
+This divides our ordination plot into 2 clusters along the first RDA axis.
+One cluster for sites primarily found with caribou and cranes, and a second cluster for the rest of the cameras and species.
 
@@
 
-clust_grp <- cutree(clust, k=5)
+clust_grp <- cutree(clust, k=2)
 custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
-                thres_plot = 0.3, clust = clust_grp)
-```
-
-@fig-map-clust shows a map of clustered camera sites.
+                thres_plot = 0.2716, clust = clust_grp)
+```
+
+These clusters can be visualized on a map, which @fig-map-clust shows. There is a slight pattern in which more sites in group 1 are found in the northern part of the array.
 
@@
 range.alpha <- seq(0,1,0.1)
-K <- 5
+K <- 2
 cr <- ClustGeo::choicealpha(events_gower_mod, dist_geo, range.alpha, 
@@
 
-@fig-map-clust-geo shows a map of geographically constrained clustered camera sites.
+Spatial clustering allows to add distance between sites as a soft constraint in the clustering algorithm.
+The extent to which these distances influence the result in balance with the community data can be adjusted.
+We used a weak effect of `r 0.2` using the `hclustgeo` function in the `ClustGeo` package.
+@fig-map-clust-geo shows the resulting map of geographically constrained clustered camera sites.
+The North/South pattern observed below is therein reinforced: camera sites in the northern part of the sampling area are moose likely to detect cranes, caribous and mooses.
 
@@
 ```{r}
-#| label: rda-clust-geo
-#| echo: false
-
-# custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
+#| label: fig-rda-clust-geo
+#| echo: false
+#| eval: false
+#| fig-cap: "Triplot of RDA on relative abundance data, with geographical Ward clusters as ellipses"
+
+custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
 ```
@@
 
-![alt text](../figures/Caribou_bplot.png "Caribou")
+In order to visualize relative abundances through time and space, we can use bubbleplots. The following plots are for caribou and sandhill crane.
+
+![Caribou](../figures/Caribou_bplot.png)
+![Sandhill crane](../figures/Sandhill crane_bplot.png)
+<!-- ![Moose](../figures/Moose_bplot.png) -->
+
+Alternatively, we can visualize those results as heatmaps. Here is an example for caribou.
 
@@
 #| label: heatmaps
+#| eval: false
 #| echo: false
@@
 ```
-![alt text](../figures/Caribou_heatmap.png "Caribou")
+
+![Caribou](../figures/Caribou_heatmap.png "Caribou")
+
+## Next steps
+
+Using the tagging information we have produced so far, we were able to show the general patterns that are starting to emerge about our data.
+The next steps for this project are to complete the tagging process and identify all the detections tagged as unknowns, bird, duck and mustelid species.
+We will also further explore the community patterns and possibly model abundances using unmarked catch and recapture models.
+Finally, as mentioned in the introduction, the next major step in this project will be tp use the bounding boxes information we have generated to train an image classifier.
 
@@
 We thank the student volunteers at Wilfrid Laurier University for their help with tagging: [TODO: list volunteers here].
+
+\newpage
+
+## Appendix
+
+```{r}
+#| label: tab-species-table-full
+#| echo: false
+
+sp_tab %>%
+  gt::gt() %>%
+  gt::tab_header(
+    title = "Relative species abundances",
+    subtitle = "Computed in number of days detected in the season"
+  ) %>%
+  gt::fmt_number(
+    decimals = 0,
+    columns = days_detected,
+    suffixing = TRUE
+  ) %>% 
+  gt::cols_label(
+    days_detected ="Days",
+    species = "Species",
+  ) %>% 
+  gt::opt_stylize(style = 6, color = "gray")
+```
+
+```{r}
+#| label: tab-landcover
+#| echo: false
+
+lu_dat %>% 
+  dplyr::select(category_code, label) %>% 
+  dplyr::filter(category_code != "ND") %>% 
+  dplyr::arrange(category_code) %>% 
+  gt::gt() %>%
+  gt::tab_header(
+    title = "Land cover classes abbreviations",
+    subtitle = "Adapted from the ontario Far North land cover dataset"
+  ) %>%
+  gt::cols_label(
+    label ="Land cover class",
+    category_code = "Abbreviations",
+  ) %>% 
+  gt::opt_stylize(style = 1, color = "gray")
+```
 

### 2023-04-14 analysis/paper/paper.qmd
progress

@@
 title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North"
+subtitle: "Prepared for: Canadian Wildlife Service, Canadian Forest Service, and Wilfrid Laurier University"
 author:
@@
       institute: wlu
+  - Samantha McFarlane:
+      institute: eccc
+  - Jennifer Baltzer:
+      institute: wlu
+  - Cheryl A. Johnson:   
+      institute: eccc
+  - Mathieu Leblond:
+      institute: eccc
+  - Eric Neilson:
+      institute: nrcan
+  - Philip Weibe:
+      institute: nrcan
   - Frances Stewart:
@@
       orcid: "0000-0001-9344-8346"
+  - Josie Hughes:
+      institute: eccc
 institute:
@@
       name: Wilfrid Laurier University
-      # address: 23 Science Street, Eureka, Mississippi, USA
+  - eccc:
+      name: Environment & Climate Change Canada
+  - nrcan:
+      name: Natural Resources Canada
 title-block-published: "Last updated"  
@@
 The extent to which these methods perform for northern regions is not well established.
-Using an array of cameras deployed in an heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
-We've recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
+Using an array of cameras deployed across an heterogeneous sampling area straddling two Northern Ontario ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
+We’ve recently developed an open-source workflow for camera trap image identification, access, and storage.
+We continue to use this workflow for ongoing image identification while investigating requirements to develop a machine learning image classifier.
 
@@
 backup_options <- options()
-options(scipen = 1, digits = 2)
+options(scipen = 1, digits = 12)
 library(magrittr)
@@
 
-The James Bay Lowlands in Northern Ontario hold some of the world’s last intact wild places but it is under pressure from resource extraction and global warming.
+Northern Ontario holds some of the world’s last intact wild places, but it is under pressure from resource extraction and global warming.
 Our understanding of basic biodiversity measures like species occurrence, abundance, diversity, and density for the region are still rudimental but needed to weigh ecological-economical decision tradeoffs.
 Traditionally, quantifying and monitoring biodiversity in northern regions required expensive and time-consuming field surveys.
-But today, the use of non-intrusive methods, such as arrays of 100s of camera traps, has significantly changed field practices for reliably quantifying wildlife across large and remote regions.
+Today, the use of non-intrusive methods, such as arrays of 100s of camera traps, has significantly changed field practices for reliably quantifying wildlife across large and remote regions.
 However, these methods generate a large amount of data and require tools to speed up the identification, and processing, of image datasets.
 
-Machine learning methods, such as deep neural networks, are now in use to automate the process of species identification and help in that endeavour.
+Machine learning methods, such as deep neural networks, are now in use to automate the process of species identification and help in that endeavor.
 We aim to evaluate the viability of deep neural networks for camera trap surveys currently taking place in Northern Ontario’s heterogeneous landscapes, including the Hudson Bay Lowlands and Ontario Shield ecozones, as variables like weather, seasonality, vegetation type, and species characteristics can impact the performance of these methods.
+The extent to which these methods perform for northern regions is not well established.
 There are currently no available machine learning models for identifying species from images in northern regions and developing one would be beneficial as a monitoring and modeling tool for northern boreal ecosystems.
 
-The extent to which these methods perform for northern regions is not well established.
 Using an array of cameras deployed in an heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
-We've recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
+We’ve recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
 
@@
 
-### Camera deployment
-
-The camera deployment protocol follows a spatially balanced hierarchical design that provides an even distribution of sample sites across biotic and abiotic conditions within the study area.
-The study sites span three ecoregions and two ecozones.
+### Camera deployment and setup
+
+The camera deployment protocol follows a spatially balanced hierarchical design that provides an even distribution of sample sites across biotic and abiotic conditions within the study area (@vanDamBates2018).
+The study sites span three ecoregions (Northern Taiga, James Bay, Big Trout Lake) and two ecozones (Hudson Bay Lowlands, Ontario Shield).
 The network was established by CWS in 2020 and seasonally equipped in a rolling design with 500 autonomous recording units (ARUs) to monitor audible species diversity paired with site-specific on-ground vegetation imagery.
-In 2022, 232 unbaited wildlife camera traps (CTs) were added to a subset of study sites to quantify mammalian occurrence and diversity.
-@fig-deployment shows the location of the and their current status.
-retrieved cameras were retrieved in September 2022.
+In March through June of 2022 193 unbaited and unlured wildlife camera traps ([Reconyx Hyper Fire II](https://reconyx.com/product/hyperfire-2-covert-ir-camera)) were added to this study design.
+All cameras were all programmed to take bursts of 5 images (one per second) at each trigger of their motion sensor.
+
+Briefly, 5 ARUs are deployed per site, with 100 m between ARUs.
+Up to 4 cameras were paired with ARUs at each of these sites, with at least one of these cameras placed at a random ARU site and one other prioritized to focus on a game trail or open area to increase detection probability (in total, we have 43 plots with 4 cameras, 5 plots with 3 cameras and 3 plots with 2 cameras). @fig-deployment shows the location of the clustered cameras and their current retrieval status.
+170 cameras were retrieved in September 2022, totaling a deployment spanning 5-7 months.
+5 additional cameras were only retrieved in March 2023 due to logistical and inaccessibility issues.
+There remains 18 cameras in the field to be retrieved in 2023.
+
+The work conducted under this project focuses on quantifying and developing an open-source workflow to quantifying the wildlife species detected in those images that have been retrieved to date.
 
@@
   dplyr::filter(!is.na(longitude)) %>% 
-  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
+  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326", remove=F) %>% 
   dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))
@@
 #| echo: false
-#| fig-cap: "Camera deployment status"
+#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
 
@@
 #| eval: false
-
-base <- ggmap::ggmap(map) +
-  ggplot2::labs(x = "Longitude", y = "Latitude") +
-  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
-  ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
-
-p <- base +
-  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+#| echo: false
+
+# base <- ggmap::ggmap(map) +
+#   ggplot2::labs(x = "Longitude", y = "Latitude") +
+#   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
+#   ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
+# 
+# p <- base +
+#   # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+#   #                     size = 2.5, data = regions_labels) +
+#   ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
 #                       size = 2.5, data = regions_labels) +
-  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
-                      size = 2.5, data = regions_labels) +
-  ggplot2::geom_point(data = cams_sf_coords %>% 
-                        dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
-                      col="black", fill = "red",
-                      ggplot2::aes(x=X, y=Y), 
-                      inherit.aes = FALSE) +
-  ggplot2::labs(fill = "Camera \nRetrieved")
-p
+#   ggplot2::geom_point(data = cams_sf_coords %>% 
+#                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
+#                       col="black", fill = "red",
+#                       ggplot2::aes(x=X, y=Y), 
+#                       inherit.aes = FALSE) +
+#   ggplot2::labs(fill = "Camera \nRetrieved")
+# p
 ```
@@
 
-Images were retrieved from all cameras using a SD card reader and copied onto a 10 TB hard drive of brand LaCie.
-The images amounted to 1.8TB of memory on disk, for a total of __`r nb_image_fmt`__ images.
+Images were retrieved from all cameras using a 32 GB SD card and copied onto a 10 TB hard drive (LaCie d2 Professional).
+The images amounted to 1.8TB of memory on this disk, for a total of __`r nb_image_fmt`__ images.
 Images were stored with the following directory structure:
@@
 - Site id (e.g. P128-1)
-- Camera id (e.g. CFS-10)
+- Camera id (e.g. CFS-10, with the first 3 letters corresponding to the group owning the camera: CFS for Canadian Forest Service, WLU for Wilfrid Laurier University)
 - DCIM folder (created by the camera)
-- Overflow folder (e.g. 100RECNX, each such folder contains up to 10,000 images)
-
-The images were processed through the animal detection model MegaDetector, version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
+- Overflow folder (e.g. 100RECNX, 101RECNX ; each such folder contains up to 10,000 images)
+
+The images were processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
 The model processes each image and produces “detections” (bounding boxes) at a  variable confidence score.
@@
 
-About __`r dets_tb_percent[1]` %__ of the images, or __`r dets_tb_fmt[1]`__ images, were found by the model to possibly have either a person, animal, or vehicle.
-The model did not produce detections for __`r dets_tb_fmt[2]`__ images, or __`r dets_tb_percent[2]` %__ of the total.
+At this first step into the process, no filtering was applied beyond the model predictions which classifies images as either likely to contain an object of relevance or empty.
+About __`r dets_tb_percent[1]` %__ of the images, (__`r dets_tb_fmt[1]`__) images, were found by the model to possibly contain either a person, animal, or vehicle.
+The model did not produce detections for __`r dets_tb_fmt[2]`__ images, (__`r dets_tb_percent[2]` %__) of the total.
 
@@
 
-This resulted in a total of __`r dets_tb_cat_unfil_sum_fmt`__ detections.
+Because the model can produce more than one detection per image, this resulted in a total of __`r dets_tb_cat_unfil_sum_fmt`__ detections.
 The totals for each category are: __`r dets_tb_cat_unfil_fmt[1]`__ animals (or __`r dets_tb_cat_unfil_percent[1]` %__); __`r dets_tb_cat_unfil_fmt[2]`__ persons (or __`r dets_tb_cat_unfil_percent[2]` %__); and __`r dets_tb_cat_unfil_fmt[3]`__ vehicles (or __`r dets_tb_cat_unfil_percent[3]` %__).
 
-Using a cut-off value for confidence threshold of 0.1, this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
-The totals for each category are: __`r dets_tb_cat_fmt[1]`__ animals (or __`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (or __`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (or __`r dets_tb_cat_percent[3]` %__).
+Megadetector is an AI model which produces confidence scores with each of its predictive detections.
+With no filtering, the overwhelming majority of the detections are of very low confidence (85% of detections are below a score of 0.1).
+This explains the high number of "person" and "vehicle" detections which are of course unrealistic.
+It is recommended to filter low probability detections, and cut-off values of 0.1 or 0.2 are usually suggested in the Megadetector documentation.
+
+Using a cut-off value for the detection confidence of 0.1 (,i.e.
+discarding all detections with confidence below 0.1), this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process  for each category are:  __`r dets_tb_cat_fmt[1]`__ animals (__`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (__`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (__`r dets_tb_cat_percent[3]` %__).
 
@@
 #| echo: false
-#| fig-cap: "Histogram of detection confidence scores."
+#| fig-cap: "Histogram of detection confidence scores produced for each bounding box predicted by Megadetector, showing a bimodal distribution.
+The colors represent the three categories of object identified by Megadetector: animals, persons and vehicles.
+Note that the hsitogram is truncated at our selected confidence threshold of 0.1."
 #| fig.width: 8
@@
   ggplot2::xlab("Confidence Score") +
-  ggplot2::ylab("Count")
-```
-
-Finally, it is important to note that not only pictures with detections were uploaded to the platform.
-Because camera traps produce images in bursts, we included all the pictures from a given bursts if any of the images in the bursts include a detection with certainty above the threshold.
+  ggplot2::ylab("Count") +
+  ggplot2::geom_vline(ggplot2::aes(xintercept = 0.1), lty=2, col = "firebrick")
+```
+
+Finally, it is important to note that not only pictures with detections were uploaded to the LabelStudio platform.
+As stated before, camera traps produce images in bursts (in our case, 5 images per trigger).
+It is possible for Megadetector to miss a detection within a burst, or to produce variable confidence scores within that burst, which could lead to missing true detections if not all pictures from a given burst are filtered in.
+Therefore, we included all the images from a given burst if any of the images in the burst include a detection with certainty above the threshold, including potentially empty images from that burst.
 
@@
 
-- *Deployable online*. This was to allow for for multiple taggers
+- *Deployable online*. This was to allow for multiple people to help process the images.
 - *Flexible tagging interface*. Few platforms allow for a full customization of the tagging interface, down to the smallest details.
-- *Allowing to tag multiple bounding boxes per image*.
+- *Ability to tag multiple bounding boxes per image*.
 Most platforms only allow for tagging the entire image, and not for portions of it (e.g.
@@
 
-We looked at platforms not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners for which the task of tagging is a common step, often referred to as labeling.
-We found two platforms that matched all those criteria: LabelStudio and CVAT.
-After testing both platforms, LabelStudio was chosen as it was simpler to implement.
-The Community edition of LabelStudio was deployed on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and a NGINX reverse proxy, which is a standard approach for custom hosting of websites.
+To guide our choice, we used these criteria to informally compare 51 different platforms and tools currently available for processing and managing camera trap images.
+We first looked at platforms traditionally used for tagging camera trap images, but found them often lacking in two main aspects: flexibility of the tagging interface, and ability to manipulate bounding box information.
+
+These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners.
+Indeed, machine learning datasets outside can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features.
+Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.
+
+We found two platforms that matched all of our criteria: LabelStudio (@LabelStudio) and CVAT (@boris_sekachev_2020_4009388).
+After testing both platforms, LabelStudio was chosen because the data format it uses is more easily compatible with Megadetector's COCO format output, making it easier to setup.
+We deployed the Community edition LabelStudio (version 1.6.0rc5) on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and a NGINX reverse proxy, which is a standard approach for custom hosting of websites.
 
@@
 
+saveRDS(anns_wide_noempty, "analysis/data/derived_data/annotations_wide_noempty.rds")
+
 # Number of processed annotations
@@
 
-Out of the __`r dets_tb_cat_fmt[1]`__ animal detections, __`r processed_fmt`__ have been processed on the platform (or __`r processed_pct` %__) for a total of __`r processed_img_fmt`__ images (i.e.
+Out of the __`r dets_tb_cat_fmt[1]`__ animal detections, __`r processed_fmt`__ have been processed to date using our platform (or __`r processed_pct` %__) for a total of __`r processed_img_fmt`__ images (i.e.
 __`r processed_img_pct` %__ of the total amount of images, or __`r processed_img_up_pct` %__ of the total amount of images uploaded to the platform).
-This does not include the empty images also viewed as part of a set, bringing that amount to 92,826 “detections” and __`r viewed_fmt`__, (i.e.
+
+As stated above, all images of a given burst are viewed if at least one image in the burst contains a detections with confidence above the threshold.
+This include potentially empty images.
+If we add the empty images viewed as part of a burst, it brings the total amount to __`r viewed_fmt`__ images, (i.e.
 __`r viewed_pct` %__ of the total amount of images, or __`r viewed_pct_up` %__ of the total amount of images uploaded to the platform).
@@
 
-Out of the __`r processed_fmt`__ detections, __`r dels_nb_fmt`__ were false positives (or __`r dels_nb_pct` %__), and __`r non_mans_nb_fmt`__ were true positives (or __`r non_mans_nb_pct` %__), to which __`r mans_nb_fmt`__ false negatives were added (manually added bounding boxes, __`r mans_nb_pct` %__).
-Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted,
-
-<!-- ### Issues encountered
-
-The main issue we encountered was a problem in the performances of MegaDetector which caused a lot of false positives, as the model was often tricked by vegetation. -->
+Out of the __`r processed_fmt`__ detections, __`r dels_nb_fmt`__ were false positives (or __`r dels_nb_pct` %__), where Megadetector identified something, but there was nothing.
+In addition, __`r non_mans_nb_fmt`__ were true positives (or __`r non_mans_nb_pct` %__), to which __`r mans_nb_fmt`__ false negatives were added (manually added bounding boxes, __`r mans_nb_pct` %__).
+Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted.
+The below summarises information for the tagging progress.
+
+```{r}
+#| label: tab-summary-table
+#| echo: false
+
+sum_tab <- tibble::tribble(
+  ~Status,                ~Images,           ~Detections,
+  "Total",                nb_image_fmt,      dets_tb_cat_unfil_sum_fmt,
+  "With detections",      dets_tb_fmt[1],    "N/A",
+  "Empty",                dets_tb_fmt[2],    "N/A",
+  "Filtered (> 0.1)",     uploaded_fmt,      dets_tb_cat_sum_fmt,
+  "Processed",            processed_img_fmt, processed_fmt
+)
+
+knitr::kable(sum_tab,
+             caption = "Summary table on tagging progress")
+```
+
+The relatively high rate of false positives is attributable to a combination of factors.
+First, the high seasonality of Ontario's far north makes it so that it is difficult to predict whether a camera that was set up in the winter won't be overgrown by vegetation in the spring.
+Locations where vegetation has grown in front of the sensor, combined with windy events, creates a perfect storm for false cameras triggers when vegetation is moving in front of the sensors.
+In addition, Megadetector is a likely to mistake a certain proportion of those vegetation pictures for actual animals.
+We are currently exploring avenues to tackle this roadblock by identifying repeating bounding images and by using a model to sort false from true positives.
+
+#### Reproducibility
+
+We are committed to producing a reproducible workflow.
+All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap).
+The two software tools used in our workflow are also available at online repositories:
+
+- [**mdtools**](https://github.com/StewartWILDlab/mdtools): a python-based command line tool for converting Megadetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
+
+- [**mdscripts**](https://github.com/StewartWILDlab/mdscripts): a bash-based command line tool for running Megadetector and mdtools.
 
@@
 
-We analysed the results of the tagging process under two main forms: presence absence data for each camera; and relative species abundance measured as the number of days, for a given month, that a given species was detected by a given camera.
+We summarized our existing species data to date for analysis in two main ways:   (1) species presence/absence data for each camera and, (2) relative species abundance, measured as the number of days, within each month that each  species was detected at a camera site.
 
@@
 ```{r}
-sum(unlist(lapply(extr_list, function(x) sum(x$cats == "NA"))))
-sum(unlist(lapply(extr_list, function(x) is.na(x$cats))))
-#  [1] "1"   "2"   "8"   "9"   "12"  "15"  "16"  "17"  "18"  "11"  "13"  "14"  "19"  "20"  "23"  "157" "247" "21"  "22"
-
-pairs(extr_table_sub_tb, col = "dodgerblue")
-round(cor(extr_table_sub_tb), 2)
-```
-
-## Preliminary results
-
-### Community data
-
-Using the number of days detected as a measure of relative abundance, the most common species (detected more than 10 days in the season) are listed in Table 1 (the full table is available in the appendix at the end of this report).
-Note the large amount detections in the "bird" and "duck" species categories which could be further distinguished into specific species.
+#| label: extract-cats
+#| echo: false
+#| eval: false
+
+# sum(unlist(lapply(extr_list, function(x) sum(x$cats == "NA"))))
+# sum(unlist(lapply(extr_list, function(x) is.na(x$cats))))
+# #  [1] "1"   "2"   "8"   "9"   "12"  "15"  "16"  "17"  "18"  "11"  "13"  "14"  "19"  "20"  "23"  "157" "247" "21"  "22"
+# 
+# pairs(extr_table_sub_tb, col = "dodgerblue")
+# round(cor(extr_table_sub_tb), 2)
+```
+
+## Data exploration and preliminary results
+
+### Wildlife community compposition
+
+We obtained repeat detections of at least 31 mammal and bird species across 193 sites and 5-7 months of detections The most common species detected (detected more than 10 days in the season) are listed in Table 1 (the full table is available in the appendix at the end of this report).
+Note the large amount detections in the “bird” and “duck” species categories which could be further distinguished into specific species.
 
@@
 
+# sci_names <- taxize::comm2sci(sp_tab$species)
+# saveRDS(sci_names, "analysis/data/derived_data/sci_names.rds")
+sci_names <- readRDS("analysis/data/derived_data/sci_names.rds") %>% 
+  unlist() %>% tibble::as_tibble(rownames = NA) %>% 
+  tibble::rownames_to_column() %>% 
+  dplyr::rename(Scientific_name = value,
+                species = rowname) %>% 
+  tibble::add_row(Scientific_name = "Ursus americanus", species="Black bear") %>% 
+  tibble::add_row(Scientific_name = "Perisoreus canadensis", species="Canada Jay")
+
 sp_tab %>%
+  dplyr::left_join(sci_names, by="species") %>% 
+  dplyr::relocate(Scientific_name, .after=species) %>% 
   dplyr::filter(days_detected >= 10) %>%
@@
   gt::tab_header(
-    title = "Relative abundances for species above 10 detection days",
+    title = "Relative abundances for species with more than 10 detection days",
     subtitle = "Computed in number of days detected in the season"
@@
     species = "Species",
+    Scientific_name = "Scientific name"
   ) %>% 
+  gt::tab_style(style = list(
+      gt::cell_text(style="italic")),
+    locations = gt::cells_body(columns = Scientific_name)) %>% 
   gt::opt_stylize(style = 6, color = "gray")
@@
 
-As for the spatial distribution of those relative abundances, @fig-map-pies show relative species abundances across the landscape, for species detected more than 50 days in the season (note that the here the detections have grouped by plot number for better visibility, with a plot containing up to 4 cameras) .
-There are no clear spatial patterns, but some tendencies can be noted: Moose detections are predominant in the northwestern of the sampling array, while caribou and sandhill cranes detections share the most of the detections for the rest of the array.
+As for the spatial distribution of those relative abundances, @fig-map-pies shows relative species abundances across the landscape, for species detected more than 50 days in the season (this number is roughly equal to the standard deviation of the distribution of detection days, which has a mean of 32 days).
+Note that for this figure, the detections have grouped by plot number for better visibility, with a plot containing up to 4 cameras (see deployment methods).
+There are no clear spatial patterns, but some generalities can be noted: Moose detections are predominant in the northwestern and southeastern zones of the sampling array, while caribou and sandhill crane detections share the most of the detections for the central part of the array.
 
@@
 #| echo: false
-#| fig-cap: "Relative species abundances at plot sites (only species species detected more than 50 days in the season are shown)"
+#| fig-cap: "Pie chart map of relative species abundances across our study area Detections have grouped by plot for better visibility, with a plot containing up to 4 cameras.
+Only species detected more than 50 days in the season are shown for clarity."
 #| fig.width: 8
@@
 
+regions_labels_bis <- zones_filter %>% 
+  sf::st_drop_geometry() %>% 
+  dplyr::bind_cols(data.frame(x = c(-85, -88),
+                              y = c(52, 52.6)))
+
 base +
@@
                               pie_scale = 1.5) +
-  ggplot2::labs(x="Lat", y="Long")
+  ggplot2::labs(x="Lat", y="Long") +
+    ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
+                      size = 2.5, data = regions_labels_bis)
 ```
@@
 
+To start exploring general species-habitat associations we completed an redundancy analysis (RDA, using the `vegan` R package, @vegan).
+RDAs allow to study the relationships between two matrices of data, and is often used in community studies, using a constraining matrix of environmental variables for each site and a corresponding matrix of species observations (either presence absence or relative abundances).
+Ordinations also combine very well with clustering analyses to produce insights about associations between sites.
+
+We used the Ontario Far North Land Cover dataset to produce proportions of land covers at each site (@hogg2014far, a full legend of the land cover abbreviations are to be found in the appendix of this report).
 @fig-rda-pa shows the triplot for the RDA on presence-absence data, which yielded a low $R^2$ of `r vegan::RsquareAdj(pa_rda)[2]`, but which is still relevant for exploring species associations.
 Land cover might not be able to explain community structure, but it can still show which species are found together and around what kind of land cover class.
-Here, we see that caribou and sandhill cranes are associated with open and treed wetlands (bogs and fens) while black bear and moose are found together in more forested areas.
-Canada geese are, as is to be expected, associated with water while smaller species like squirrels and martens score away from the larger mammals in a mix of habitats.
-Note that a full legend of the land cover abbreviations are to be found in the appendix of this report.
+Here, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
+Canada geese are, as is to be expected, associated with water (WAT) while smaller species like squirrels and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
 
@@
 #| echo: false
-#| fig-cap: "Triplot of RDA on presence-absence data"
+#| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected in March through September of 2022 reveals that different species associate with different land cover classes.
+For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
 
@@
 ```
-```{r}
-# Jaccard pcoa
-
-anns_wide_animals_table_tb_jac <- vegan::vegdist(anns_wide_animals_table_tb, 
-                                                 "jac", binary = TRUE)
-
-pa_pcoa <- ape::pcoa(anns_wide_animals_table_tb_jac)
-
-biplot(pa_pcoa, anns_wide_animals_table_tb)
+
+```{r}
+#| label: jaccard-pcoa
+#| echo: false
+#| eval: false
+
+# anns_wide_animals_table_tb_jac <- vegan::vegdist(anns_wide_animals_table_tb, 
+#                                                  "jac", binary = TRUE)
+# 
+# pa_pcoa <- ape::pcoa(anns_wide_animals_table_tb_jac)
+# 
+# biplot(pa_pcoa, anns_wide_animals_table_tb)
 ```
@@
 
-@fig-rda-events shows the triplot for the RDA on relative abundance data, which  yielded a slightly larger $R^2$ of `r vegan::RsquareAdj(events_rda)[2]`, but reduces the axis scores of the rarer species.
+To continue exploring abundance-habitat relationships as a proxy for species relative habitat use, we also conducted an RDA using relative abundances instead of presence absence.
+@fig-rda-events shows the triplot for the RDA on relative abundance data.
+This analysis yielded a slightly larger $R^2$ of `r vegan::RsquareAdj(events_rda)[2]`, than the analysis in @fig-rda-pa presenting presence-absence data, but reduces the axis scores of the rarer species.
 We observe similar patterns for caribou, crane, moose, black bear and marten.
@@
 #| echo: false
-#| fig-cap: "Triplot of RDA on relative abundance data"
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data."
 
@@
 # plot(clust)
-# vegan::cascadeKM(events_gower,2, 5) %>% plot()
-```
-
-Clustering allows us to find pattern in the ordination data.
-@fig-rda-events-clust shows the triplot for the RDA on relative abundance data, with Ward clusters as ellipses.
-We used the cascading K-means function `cascadeKM` in the `vegan` package to determine the number of relevant clusters to show (in this case, 2 clusters were found to be most relevant.
-This divides our ordination plot into 2 clusters along the first RDA axis.
-One cluster for sites primarily found with caribou and cranes, and a second cluster for the rest of the cameras and species.
+# km <- vegan::cascadeKM(events_gower, 2, 5)
+```
+
+To start inferring patters from the above ordination analyses we used a clustering algorithm on the relative abundance data presented in @fig-rda-events.
+@fig-rda-events-clust shows the triplot for the RDA on relative abundance data, with Ward clusters as ellipses (@Legendre2012).
+We divided our ordination plot into 2 clusters along the first RDA axis.
+One cluster for sites primarily found with caribou and cranes, and a second cluster for the rest of the cameras and species.This clustering generally reflects the spatial patterns in species abundance that we observed in @fig-map-pies.
+
+<!-- We used the cascading K-means function `cascadeKM` in the `vegan` package (@vegan) to determine the number of relevant clusters to show.
+We used the Calinski criteria (@Legendre2012) to decide on the appropriate number of cluster.
+This criteria is // We found that the optimal number of clusters was 2.
+-->
 
@@
 #| echo: false
-#| fig-cap: "Triplot of RDA on relative abundance data, with Ward clusters as ellipses"
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
+Two clusters determined using Ward clustering are shown as ellipses."
 
@@
 
-These clusters can be visualized on a map, which @fig-map-clust shows. There is a slight pattern in which more sites in group 1 are found in the northern part of the array.
+These clusters can be visualized on a map, which @fig-map-clust shows. There is no apparent discernible spatial pattern.
 
@@
 #| echo: false
-#| fig-cap: "Map of clustered camera sites"
+#| fig-cap: "Map of clustered camera sites. The two clusters were determined using the Ward clustering method."
 
@@
 ```
+
 ### Spatial clustering
@@
 
-range.alpha <- seq(0,1,0.1)
+# range.alpha <- seq(0,0.2,0.01)
 K <- 2
-cr <- ClustGeo::choicealpha(events_gower_mod, dist_geo, range.alpha, 
-                            K, graph = F)
+
+# cr <- ClustGeo::choicealpha(events_gower_mod, dist_geo, range.alpha, 
+#                             K, graph = F)
+# plot(cr)
 
 clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
-                                 dist_geo, alpha = 0.2)
+                                 dist_geo, alpha = 0.1)
 clust_geo_grp <- cutree(clust_geo, k=K)
@@
 Spatial clustering allows to add distance between sites as a soft constraint in the clustering algorithm.
-The extent to which these distances influence the result in balance with the community data can be adjusted.
-We used a weak effect of `r 0.2` using the `hclustgeo` function in the `ClustGeo` package.
+The extent to which these distances influence the result in balance with the community data can be adjusted, by setting the strength of the mixing parameter alpha.
+The `ClustGeo` package (@Chavent2018) allows to test different values of that parameter to determine optimal mixing of the clustering forces, with the function `choicealpha`.
+
+We found that the optimal value for alpha be `r 0.1`, and performed mixed clustering using the `hclustgeo` function.
 @fig-map-clust-geo shows the resulting map of geographically constrained clustered camera sites.
-The North/South pattern observed below is therein reinforced: camera sites in the northern part of the sampling area are moose likely to detect cranes, caribous and mooses.
+This reveals a North/South pattern in the spatial clustering of sites.
 
@@
 #| echo: false
-#| fig-cap: "Map of geographically constrained clustered camera sites"
+#| fig-cap: "Map of geographically constrained clustered camera sites, using geographically constrained clustering with a mixing parameter of 0.1."
 
@@
 
+We can plot those geographically constrained clusters on top of the abundance-based RDA of @fig-rda-events.
+@fig-rda-clust-geo shows that those geographic clusters do not map well on top of the ordination, with one cluster seeming to represent a subset of a larger trend experienced in the larger part of the data.
+
 ```{r}
@@
 #| echo: false
-#| eval: false
-#| fig-cap: "Triplot of RDA on relative abundance data, with geographical Ward clusters as ellipses"
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
+Two clusters determined using geographically constrained clustering with a mixing parameter of 0.1."
 
@@
 cams_sf_sp <- events_cam_days %>% 
-  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) %>% 
-  dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))
+  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) # %>% 
+  # dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))
 
@@
 
-mod <- mgcv::gam(count~s(X,Y, bs=""), data = cams_sf_sp_sum_filt)
+mod <- mgcv::gam(count~s(X,Y), data = cams_sf_sp_sum_filt)
 pred_grid <- 
@@
 
-rast <- raster::rasterFromXYZ(preds)
-raster::plot(rast)
-rast[rast<0] <- NA
-raster::plot(rast)
+# rast <- raster::rasterFromXYZ(preds)
+# raster::plot(rast)
+# rast[rast<0] <- NA
+# raster::plot(rast)
 
@@
 
-base_cut +
-  ggplot2::geom_tile(
-    data = dplyr::filter(preds, count>=0), ggplot2::aes(X, Y, fill=count),
-    binwidth = 0.05, alpha = 0.5, inherit.aes = FALSE) +
-  ggplot2::scale_fill_viridis_c()
+# base_cut +
+#   ggplot2::geom_tile(
+#     data = dplyr::filter(preds, count>=0), ggplot2::aes(X, Y, fill=count),
+#     binwidth = 0.05, alpha = 0.5, inherit.aes = FALSE) +
+#   ggplot2::scale_fill_viridis_c()
 ```
@@
     ggplot2::facet_wrap(~month) +
-    ggplot2::labs(size="Days \ndetected")
+    ggplot2::labs(size="Days \ndetected",
+                  title=sp)
 
@@
 
-In order to visualize relative abundances through time and space, we can use bubbleplots. The following plots are for caribou and sandhill crane.
+In order to visualize relative abundances through time and space, we can use bubbleplots.
+The following plots are for caribou, sandhill crane, moose, and wolf.
+They represent days detected at each camera site per month.
 
@@
 ![Sandhill crane](../figures/Sandhill crane_bplot.png)
-<!-- ![Moose](../figures/Moose_bplot.png) -->
-
-Alternatively, we can visualize those results as heatmaps. Here is an example for caribou.
+![Moose](../figures/Moose_bplot.png)
+
+![Wolf](../figures/Wolf_bplot.png)
+
+As more of the camera images are labeled and our dataset grows, we can transform the above bubble plots into heat maps of species presence/absence or relative abundance.
+Below is an example using the existing caribou data to date.
+It is important to recognize that not all cameras had been deployed until June, leaving April and Ma.
+y maps with large portions of the northern study area where cameras were not yet present.
 
@@
 
-![Caribou](../figures/Caribou_heatmap.png "Caribou")
+<!-- ![Caribou](../figures/Caribou_heatmap.png "Caribou") -->
+
+### Take Home messages
+
+We can derive a few take home messages from our initial exploration of the data:
+
+- Using an object detection model allowed us to filter out large quantities of empty images, but performed somewhat poorly in certain conditions (namely, windy sites with drastic changes in vegetation).
+
+- According to our analysis, species tend to associate with land use changes classes that generally corresponds to their expected habitat, such as wetlands for caribou and cranes, or forested areas for moose.
+
+- Clustering reveals two large groups among sites, but spatial patterns are difficult to tease apart at this stage.
 
@@
 
-Using the tagging information we have produced so far, we were able to show the general patterns that are starting to emerge about our data.
+Our existing image identification workflow and tagging process has allowed us to produce an open and reproducible data set of spatially explicit wildlife species counts in Ontario’s Far North.
+Using this data set we have started to explore spatial patterns of species richness, occurrence, and abundance.
+All of this work is conducted through opensource platforms, and data exploration and analysis code hosted as a [GitHub repository](https://github.com/StewartWILDlab/rofcamtrap) for ease of access and reproducibilty.
+The information necessay to reproduce this report is in the colophon in the appendix.
+
+Over the coming months we will continue to process the retrieved data and anticipate adding to this data set over the summer, as data from ## cameras has yet to be retrieved.
 The next steps for this project are to complete the tagging process and identify all the detections tagged as unknowns, bird, duck and mustelid species.
-We will also further explore the community patterns and possibly model abundances using unmarked catch and recapture models.
-Finally, as mentioned in the introduction, the next major step in this project will be tp use the bounding boxes information we have generated to train an image classifier.
+We will also further explore the spatial community composition patterns at each camera site.
+Given a high enough data density for certain species future possibilities for this project include modelling species occupancy (@MacKenzie2002) and spatially explicit density estimation using spatially explicit capture-recapture models (@Chandler2013).
+
+After data processing and exploration is complete the next major step in this project will be to use the bounding boxes information we have generated to train an image classifier.
+This classifier will be based on Megaclassifier (of type EfficientNet architecture, @EfficientNet), an classifier already trained to pair with Megadetector's output.
+We will retrain this model to fit our taxa, possibly leveraging mentoring opportunities offered by the CV4E workshop ("Computer Vision for Ecology"), if availability allows.
 
@@
 
-We thank the student volunteers at Wilfrid Laurier University for their help with tagging: [TODO: list volunteers here].
+We thank the student volunteers at Wilfrid Laurier University for their help with tagging: Teea Curlew, Bridget Matthews, Meghna Pal, Rafay Siraj, and Dietrich Westberg.
+We also thank members of the WILDlab for feedback throughout this process: Claudia Haas, Eric Jolin, Charlotte Rentmeister, and Sheyda Zand.
+This work was supported by a G&C to Frances Stewart.
+
+\newpage
+
+## References
+
+<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files): -->
+
+::: {#refs}
+:::
 
@@
   gt::tab_header(
-    title = "Land cover classes abbreviations",
+    title = "Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
     subtitle = "Adapted from the ontario Far North land cover dataset"
@@
 <!-- The following line inserts a page break  -->
-<!--
-\newpage
-
-## References
-
-The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  
--->
-
-::: {#refs}
-:::
 

### 2023-04-14 analysis/paper/paper.qmd
try remove spce
fixes

@@
     reference-doc: "../templates/template_modified.docx" # Insert path for the DOCX file
+  pdf:
+    toc: true
+    number-sections: true
+    colorlinks: true
 execute:
@@
 
-![Caribou](../figures/Caribou_bplot.png)
-![Sandhill crane](../figures/Sandhill crane_bplot.png)
-![Moose](../figures/Moose_bplot.png)
-
-![Wolf](../figures/Wolf_bplot.png)
+![](../figures/Caribou_bplot.png)
+![](../figures/Sandhill crane_bplot.png)
+![](../figures/Moose_bplot.png)
+
+![](../figures/Wolf_bplot.png)
 
@@
 
-<!-- ![Caribou](../figures/Caribou_heatmap.png "Caribou") -->
-
 ### Take Home messages
@@
 This work was supported by a G&C to Frances Stewart.
-
-\newpage
 

### 2023-04-20 analysis/paper/paper.qmd
ready to submit

@@
 format:
-  docx:
-    reference-doc: "../templates/template_modified.docx" # Insert path for the DOCX file
+  # docx:
+  #   reference-doc: "../templates/template_modified.docx" # Insert path for the DOCX file
   pdf:
@@
     colorlinks: true
+    extra_dependencies: ["float"]
 execute:
@@
 knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
+knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
 backup_options <- options()
-options(scipen = 1, digits = 12)
+options(scipen = 1, digits = 2)
 library(magrittr)
@@
 Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted.
-The below summarises information for the tagging progress.
+The table below summarises information for the tagging progress.
 
@@
 
-We obtained repeat detections of at least 31 mammal and bird species across 193 sites and 5-7 months of detections The most common species detected (detected more than 10 days in the season) are listed in Table 1 (the full table is available in the appendix at the end of this report).
+We obtained repeat detections of at least 31 mammal and bird species across 193 sites and 5-7 months of detections The most common species detected (detected more than 10 days in the season) are listed in Table 2 below (the full table is available in the appendix at the end of this report).
 Note the large amount detections in the “bird” and “duck” species categories which could be further distinguished into specific species.
@@
   gt::tab_header(
-    title = "Relative abundances for species with more than 10 detection days",
+    title = "Table 2: Relative abundances for species with more than 10 detection days",
     subtitle = "Computed in number of days detected in the season"
@@
 ![](../figures/Caribou_bplot.png)
+\vspace{1pt}
+
 ![](../figures/Sandhill crane_bplot.png)
+\vspace{1pt}
+
 ![](../figures/Moose_bplot.png)
+\vspace{1pt}
 
 ![](../figures/Wolf_bplot.png)
+\vspace{1pt}
 
@@
 
-Over the coming months we will continue to process the retrieved data and anticipate adding to this data set over the summer, as data from ## cameras has yet to be retrieved.
+Over the coming months we will continue to process the retrieved data and anticipate adding to this data set over the summer, as data from 18 cameras has yet to be retrieved.
 The next steps for this project are to complete the tagging process and identify all the detections tagged as unknowns, bird, duck and mustelid species.
@@
   gt::tab_header(
-    title = "Relative species abundances",
+    title = "Table 3: Relative species abundances",
     subtitle = "Computed in number of days detected in the season"
@@
   gt::tab_header(
-    title = "Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
+    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
     subtitle = "Adapted from the ontario Far North land cover dataset"
@@
 # what commit is this file at? 
-if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
+if ("git2r" %in% installed.packages() & 
+    git2r::in_repository(path = ".")) {
+  git2r::repository(here::here())  
+}
 ```

### 2023-04-20 scripts/3_ordination.R

@@ -38,7 +38,6 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA,
 
     ggrepel::geom_text_repel(data = sc_bp,
                              ggplot2::aes(x=RDA1, y=RDA2, label = name),
-                             family = "Poppins",
                              size = 2.5,
                              min.segment.length = 10,
                              seed = 77,
@@ -52,7 +51,6 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA,
 
     ggrepel::geom_text_repel(data = sc_sp,
                              ggplot2::aes(x=RDA1, y=RDA2, label = highlight),
-                             family = "Poppins",
                              size = 3,
                              # min.segment.length = 0,
                              seed = 77,

### 2023-08-21 analysis/paper/paper.qmd
updates

@@
     !file.exists("analysis/data/derived_data/annotations_wide.rds")){
-  source("scripts/1_preprocess_annotations.R")
+  # source("scripts/1_preprocess_annotations.R")
 } else {
-  anns <- readRDS("analysis/data/derived_data/annotations.rds")
-  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
+  # anns <- readRDS("../analysis/data/derived_data/annotations.rds")
+  # anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
+  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide_treated.rds")
 }
@@
 if (!file.exists("analysis/data/derived_data/detections.rds")){
-  source("scripts/2_preprocess_detections.R")
+  # source("scripts/2_preprocess_detections.R")
 } else {
@@
 #   dplyr::right_join(dets_images, by = "image_id")
+```
+
+```{r}
+# library(forcats)
+# anns_wide_animals_joined_mut %>% 
+#   ggplot(aes(x=date_time, y=species)) + geom_point()
+# anns_wide_animals_joined_mut %>% 
+#   dplyr::filter(species == "Fisher") %>% View()
+
+camera_days <- exif %>% dplyr::group_by(maker_notes_user_label) %>% 
+  dplyr::summarise(camera_first_time=min(exif_date_time_original)) %>% 
+  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
+                                               tz = "Canada/Eastern"),
+                year = lubridate::year(date_time),
+                month = lubridate::month(date_time, label = TRUE),
+                day = lubridate::day(date_time),
+                ymd = paste(year, month, day, sep = "_")) %>% 
+  dplyr::select(camera_id = maker_notes_user_label, 
+                camera_first_day = date_time)
+
+camera_regions <- cams_sf %>% sf::st_join(zones) %>% 
+  sf::st_drop_geometry() %>% dplyr::select(camera_id, ECOZONE_NA)
+
+bp <- anns_wide_animals_joined_mut %>%
+
+  dplyr::filter(species != "Fisher") %>% 
+  dplyr::left_join(camera_days, by = "camera_id") %>%
+
+  dplyr::select(species, camera_id, 
+                date_time, camera_first_day, ymd) %>%
+
+  dplyr::group_by(species, camera_id) %>% 
+  dplyr::summarise(first_day = min(date_time),
+                   camera_first_day = min(camera_first_day),
+                   camera_events = length(unique(ymd))) %>% 
+  dplyr::ungroup() %>%
+
+  dplyr::left_join(camera_regions, by = "camera_id") %>%
+
+  dplyr::group_by(species) %>% 
+  dplyr::mutate(total_events = sum(camera_events),
+                n_cameras = dplyr::n()) %>% 
+  dplyr::ungroup() %>%
+
+  # dplyr::filter(n_cameras > 4) %>%
+  dplyr::filter(n_cameras > 3) %>%
+
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%
+
+  ggplot(aes(x = diff_time, 
+             y = forcats::fct_reorder(species, total_events),
+             color = ECOZONE_NA)) +
+  geom_boxplot() +
+
+  scale_x_continuous()+
+  labs(x = "Detection time (in days)",
+       y = "Species",
+       color = "Ecozone") + 
+  theme_bw()
+ggsave(plot = bp, filename = "analysis/figures/bp.png", width = 10, height = 15)
+
+pts <- anns_wide_animals_joined_mut %>%
+
+  dplyr::filter(species != "Fisher") %>% 
+  dplyr::left_join(camera_days, by = "camera_id") %>%
+
+  dplyr::select(species, camera_id, 
+                date_time, camera_first_day, ymd) %>%
+
+  dplyr::group_by(species, camera_id) %>% 
+  dplyr::summarise(first_day = min(date_time),
+                   camera_first_day = min(camera_first_day),
+                   camera_events = length(unique(ymd))) %>% 
+  dplyr::ungroup() %>%
+
+  dplyr::left_join(camera_regions, by = "camera_id") %>%
+
+  dplyr::group_by(species) %>% 
+  dplyr::mutate(total_events = sum(camera_events),
+                n_cameras = dplyr::n()) %>% 
+  dplyr::ungroup() %>%
+
+  # dplyr::filter(n_cameras > 4) %>%
+  dplyr::filter(n_cameras > 3) %>%
+
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA)) %>%
+
+  ggplot(aes(x = diff_time, 
+             y = forcats::fct_reorder(species, total_events),
+             # size = camera_events, 
+             color = ECOZONE_NA)) +
+
+  geom_point(aes(x = diff_time, 
+                  y = forcats::fct_reorder(species, total_events),
+                  size = camera_events, 
+                  color = ECOZONE_NA,
+                  group = ECOZONE_NA)) +
+
+  scale_x_continuous()+
+  labs(x = "Detection time (in days)",
+       y = "Species",
+       color = "Ecozone",
+       size = "Day counts") + 
+  theme_bw()
+ggsave(plot = pts, filename = "analysis/figures/pts.png", width = 10, height = 15)
+pts
+```
+
+```{r}
+for_vic <- anns_wide_animals_joined_mut %>% 
+  dplyr::select(id,species,date_time, camera_id) %>% 
+  dplyr::left_join(cams_sf %>% dplyr::select(camera_id, plot_id) %>% sf::st_drop_geometry(),
+                   by = "camera_id")
+tot_camera <- for_vic$camera_id %>% unique() %>% length()
+hit_rates <- for_vic %>% dplyr::select(species, camera_id) %>% 
+  dplyr::group_by(species) %>% dplyr::summarise(n_cameras=length(unique(camera_id))) %>% 
+  dplyr::ungroup() %>% dplyr::mutate(rate = n_cameras/tot_camera*100) %>% 
+  dplyr::arrange(dplyr::desc(rate))
+# det_history_two_weeks <-
+
+readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
+readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
 ```

### 2023-08-21 scripts/1_preprocess_annotations.R

@@ -1,23 +1,19 @@
 # Note: This runs on the csvs created from the label studio json outputs by
 # the post process function in the python utility mdtools
 
+# for FILE in *
+#   mdtools postprocess --write-csv $FILE
+# end
+
+library(magrittr)
+
 # Folder with label studio annotation outputs
 label_studio_folder <- "analysis/data/raw_data/label_studio_outputs/"
 
 # Annotations -------------------------------------------------------------
 
-# List all the files with the proper file name format
-file_list <- list.files(label_studio_folder, recursive = F,
-                        full.names = T, pattern = "output.csv")
-
-# Create an empty dataframe to be grown
-annotations <- data.frame()
-
-for (file in file_list){
-
-  # print(file)
-
-  outls <- readr::read_csv(file, col_types = readr::cols(
+read_ls_file <- function(the_file){
+  readr::read_csv(the_file, col_types = readr::cols(
     id = readr::col_character(),
     type = readr::col_character(),
     origin = readr::col_character(),
@@ -34,12 +30,54 @@ for (file in file_list){
     tag = readr::col_character(),
     value_text = readr::col_character()
   ))
+}
 
-  annotations <- dplyr::bind_rows(annotations, outls)
+widen <- function(df) {
+  df %>%
+    dplyr::select(-type) |>
+    dplyr::mutate(manual = stringr::str_detect(id, "_M")) |>
+    dplyr::mutate(image_id = stringr::str_split(id, "_")) |>
+    dplyr::mutate(image_id =
+                    unlist(purrr::map(image_id,
+                                      ~paste0(.x[1:5], collapse = "_")))) |>
+    tidyr::pivot_wider(names_from = "variable", values_from = "tag") |>
+    dplyr::filter(is.na(value_text)) |>
+    dplyr::mutate (species = ifelse(species == 'Other [See more species]',
+                                    other_species, species)) # |>
+    # dplyr::select(-other_species)
+}
+
+# List all the files with the proper file name format
+file_list <- list.files(label_studio_folder, recursive = F,
+                        full.names = T, pattern = "output.csv")
+
+# Create an empty dataframe to be grown
+annotations_no_dup <- data.frame()
+annotations_dup <- data.frame()
+
+# separate into dups and not dups. process not dups normally and
[…]

### 2023-08-21 scripts/4_cropping.R

@@ -197,12 +197,12 @@ trees_test <- testing(trees_split)
 
 tree_rec <- recipe(is_TP~size+mean_HUE+id, data = trees_train) %>%
   update_role(id, new_role = "id variable") # %>%
-  # step_other(species, caretaker, threshold = 0.01) %>%
-  # step_other(site_info, threshold = 0.005) %>%
-  # step_dummy(all_nominal(), -all_outcomes()) %>%
-  # step_date(date, features = c("year")) %>%
-  # step_rm(date) %>%
-  # step_downsample(legal_status)
+# step_other(species, caretaker, threshold = 0.01) %>%
+# step_other(site_info, threshold = 0.005) %>%
+# step_dummy(all_nominal(), -all_outcomes()) %>%
+# step_date(date, features = c("year")) %>%
+# step_rm(date) %>%
+# step_downsample(legal_status)
 
 tree_prep <- prep(tree_rec)
 juiced <- juice(tree_prep)
@@ -311,4 +311,111 @@ final_res %>%
   geom_point(size = 0.5, alpha = 0.5) +
   labs(color = NULL) +
   scale_color_manual(values = c("gray80", "darkred"))
+
+# -------------------------------------------------------------------------
+
+# read in old and new 58
+library(magrittr)
+
+old <- readr::read_csv("old_58.csv") %>%
+  dplyr::mutate(from = "old")
+# new <- readr::read_csv("new_58.csv") %>%
+#   dplyr::mutate(from = "new")
+
+# row_sums <- nrow(old) + nrow(new)
+
+# full <- old %>% dplyr::full_join(new)
+
+old <- old %>%
+  dplyr::mutate(value_rectanglelabels =
+                  stringr::str_remove(value_rectanglelabels,
+                                      stringr::fixed("['"))) %>%
+  dplyr::mutate(value_rectanglelabels =
+                  stringr::str_remove(value_rectanglelabels,
+                                      stringr::fixed("']"))) |>
+  dplyr::select(-type) |>
+  dplyr::mutate(manual = stringr::str_detect(id, "_M")) |>
+  dplyr::mutate(image_id = stringr::str_split(id, "_")) |>
+  dplyr::mutate(image_id =
+                  unlist(purrr::map(image_id,
+                                    ~paste0(.x[1:5], collapse = "_")))) |>
+  tidyr::pivot_wider(names_from = "variable", values_from = "tag") |>
[…]

### 2023-10-17 analysis/paper/paper.qmd
update paper

@@
 
-```{r}
+```{r eval=FALSE}
 # library(forcats)
+library(ggplot2)
 # anns_wide_animals_joined_mut %>% 
@@
 
-```{r}
+```{r eval=FALSE}
 for_vic <- anns_wide_animals_joined_mut %>% 

### 2023-10-17 scripts/4_cropping.R
add cropping script

@@ -1,13 +1,96 @@
 library(magrittr)
+library(magick)
 
 # -------------------------------------------------------------------------
 
 anns_wide_noempty <- readRDS(
   "analysis/data/derived_data/annotations_wide_noempty.rds") %>%
   dplyr::mutate(crop_name = NA)
+# anns_wide <- readRDS(
+#   "analysis/data/derived_data/annotations_wide.rds") %>%
+#   dplyr::mutate(crop_name = NA)
 
 out_folder <- "/media/vlucet/TrailCamST/Cropped/"
 base_folder <- "/media/vlucet/TrailCamST/TrailCamStorage/"
+
+# -------------------------------------------------------------------------
+# Cropping for AI
+# -------------------------------------------------------------------------
+
+make_crop_name_label <- function(r, out_folder = "/media/vlucet/TrailCamST/Cropped/all/") {
+  img_name <- paste0(out_folder, stringr::str_replace_all(stringr::str_replace_all(
+    r$species, pattern = c(" "), replacement = "_"),
+    pattern = stringr::fixed("."), replacement = ""), "_", r$id)
+  return(img_name)
+}
+
+for (row in seq_len(nrow(anns_wide_noempty))){
+
+  r <- anns_wide_noempty[row,]
+
+  if (is.na(r$species)) {
+    next
+  } else if(r$species == "Human") {
+    next
+  } else if(r$species == "Vehicle") {
+    next
+  } else{
+
+    img_name <- make_crop_name_label(r)
+    anns_wide_noempty[row,"crop_name"] <- img_name
+
+    if (file.exists(img_name)) {
+
+      next
+
+    } else {
+
+      print(r$id)
+
+      img <- magick::image_read(paste0(base_folder, r$source_file))
+      img_cropped <- image_crop(img, magick::geometry_area(
+        width = (r$value_width/100)*r$original_width,
+        height = (r$value_height/100)*r$original_height,
+        x_off = (r$value_x/100)*r$original_width,
+        y_off = (r$value_y/100)*r$original_height))
+
+      magick::image_write(img_cropped, img_name)
+
+      image_destroy(img)
+      image_destroy(img_cropped)
+      gc()
+
+    }
+
+  }
+
+}
+
+saveRDS(anns_wide_noempty, "analysis/data/derived_data/annotations_wide_noempty_with_crop.rds")
+
[…]

### 2023-10-17 scripts/5_repeats_test.R
add scripts

@@ -1,0 +1,162 @@
+library(magrittr)
+
+# if (!file.exists("analysis/data/derived_data/detections.rds")){
+#   source("scripts/2_preprocess_detections.R")
+# } else {
+#   dets <- readRDS("analysis/data/derived_data/detections.rds")
+# }
+
+# ls_out <- readr::read_csv("out_542011_output.csv", col_types = readr::cols(
+#   id = readr::col_character(),
+#   type = readr::col_character(),
+#   origin = readr::col_character(),
+#   to_name = readr::col_character(),
+#   variable = readr::col_character(),
+#   original_width = readr::col_double(),
+#   original_height = readr::col_double(),
+#   value_x = readr::col_double(),
+#   value_y = readr::col_double(),
+#   value_width = readr::col_double(),
+#   value_height = readr::col_double(),
+#   value_rectanglelabels = readr::col_character(),
+#   source_file = readr::col_character(),
+#   tag = readr::col_character(),
+#   value_text = readr::col_character()
+# ))
+#
+# ls_out <- ls_out %>%
+#   dplyr::select(-type) |>
+#   dplyr::mutate(manual = stringr::str_detect(id, "_M")) |>
+#   dplyr::mutate(image_id = stringr::str_split(id, "_")) |>
+#   dplyr::mutate(image_id =
+#                   unlist(purrr::map(image_id,
+#                                     ~paste0(.x[1:5], collapse = "_")))) |>
+#   tidyr::pivot_wider(names_from = "variable", values_from = "tag") # |>
+#   # dplyr::filter(is.na(value_text)) |>
+#   # dplyr::mutate (species = ifelse(species == 'Other [See more species]',
+#   #                                 other_species, species)) |>
+#   # dplyr::select(-other_species)
+
+# -------------------------------------------------------------------------
+
+ls_out <- jsonlite::fromJSON("/media/vlucet/TrailCamST/P312_out_ls.json",
+                             simplifyVector = T,
+                             flatten = T)
+coco <- jsonlite::fromJSON("/media/vlucet/TrailCamST/TrailCamStorage/P312_output_coco.json",
+                           simplifyVector = T,
+                           flatten = T)
+# ----
+
[…]

### 2023-10-17 scripts/6_phenopix_analysis.R

@@ -1,0 +1,286 @@
+library(phenopix)
+library(raster)
+library(magrittr)
+
+# -------------------------------------------------------------------------
+
+my.path <- structureFolder("~/Documents/Cropped/FP/Animal/")
+img <- brick('~/Documents/Cropped/FP/Animal/REF/animal_P224-1_WLU-147_DCIM_104RECNX_RCNX4369_7.JPG')
+img <- brick('~/Documents/Cropped/FP/Animal/animal_P028-1_WLU-10_DCIM_100RECNX_RCNX0148_6.JPG')
+plotRGB(img)
+
+plot(img[[1]])
+plot(img[[2]])
+plot(img[[3]])
+
+DrawMULTIROI(path_img_ref = '~/Documents/Cropped/FP/Animal/REF/animal_P224-1_WLU-147_DCIM_104RECNX_RCNX4369_7.JPG',
+             nroi = 1, path_ROIs = '~/Documents/Cropped/FP/Animal/REF/')
+
+PrintROI(path_img_ref = '~/Documents/Cropped/FP/Animal/REF/animal_P224-1_WLU-147_DCIM_104RECNX_RCNX4369_7.JPG',
+         path_ROIs = '~/Documents/Cropped/FP/Animal/REF/')
+
+load('~/Documents/Cropped/FP/Animal/REF/roi.data.Rdata')
+
+extractVIs(img.path = '~/Documents/Cropped/FP/Animal/IMG',file.type = ".JPG",
+           date.code = 'yyyymmdd',
+           roi.path = '~/Documents/Cropped/FP/Animal/REF/')
+
+load("VI.data.Rdata")
+VI.data
+
+# -------------------------------------------------------------------------
+
+base <- "/media/vlucet/TrailCamST/Cropped/"
+all_files <- list.files(base, recursive = T)
+
+library(exifr)
+exif_df <- read_exif(file.path(base, all_files), tags = c("filename", "imagesize",  	"DateTimeOriginal"))
+View(exif_df)
+
+saveRDS(exif_df,"exif_df.rds")
+
+exif_df <- readRDS("exif_df.rds")
+
+collect_indices <- function(img, base = "/media/vlucet/TrailCamST/Cropped/"){
+
+  # browser()
+  img_brick <- brick(file.path(base, img))
+
+  if(nlayers(img_brick) == 3) {
+
+    img_split <- unlist(strsplit(img, "/"))
+    status <- img_split[1]
+    category <- img_split[2]
+    file <- img_split[3]
+
[…]

### 2023-10-17 scripts/7_error_tests.R

@@ -1,0 +1,20 @@
+library(magrittr)
+
+old <- readr::read_csv("analysis/data/raw_data/label_studio_outputs/culled_post_test/old_128.csv") %>%
+ dplyr::mutate(from = "old")
+new <- readr::read_csv("analysis/data/raw_data/label_studio_outputs/culled_post_test/new_128.csv") %>%
+ dplyr::mutate(from = "new")
+
+row_sums <- nrow(old) + nrow(new)
+
+# full <- old %>% dplyr::full_join(new)
+
+full <- dplyr::bind_rows(old,new)
+
+nrow(full) == row_sums
+
+length(unique(full$source_file))
+
+full %>% janitor::get_dupes(id) %>% View()
+
+full %>% dplyr::filter(id %in% old$id[which(old$id %in% new$id)]) %>% View()

### 2023-10-18 R/count_files.R
fuse other repo

@@ -1,0 +1,46 @@
+# V. LUCET 2022
+# For Sam, to count survey pictures
+
+# Change path
+folder <- "/media/vlucet/TrailCamST/TrailCamStorage"
+
+# -------------------------------------------------------------------------
+
+# FUNCTIONS
+# Taken from
+# https://www.r-bloggers.com/2017/04/a-python-like-walk-function-for-r/
+
+checkname <- function(drname, filelist, arg) {
+  if (length(filelist)>0){
+    arg$tot <- arg$tot + sum(sapply(filelist, pattern = arg$flname, FUN = grepl))
+  }
+  arg
+}
+
+walk <- function(currdir, f, arg) {
+
+  # "leave trail of bread crumbs"
+  savetop <- getwd()
+  setwd(currdir)
+
+  print(currdir)
+
+  fls <- list.files()
+  arg <- f(currdir, fls, arg)
+
+  # subdirectories of this directory
+  dirs <- list.dirs(recursive=FALSE)
+
+  for (d in dirs) arg <- walk(d,f,arg)
+
+  setwd(savetop) # go back to calling directory
+  return(arg)
+}
+
+countinst <- function(startdir, flname) {
+  walk(startdir, checkname,
+       list(flname=flname, tot=0))
+}
+
+ret <- countinst(folder, ".JPG")
+print(ret$tot)

### 2023-10-18 R/utils.R

@@ -1,0 +1,53 @@
+
+#'@export
+summarise_folder <- function(){
+  return(NULL)
+}
+
+
+folder <- "/media/vlucet/TrailCamST/TrailCamStorage"
+
+folders <- list.dirs(folder, recursive = FALSE)
+
+list_fun <- function(the_dir){
+  list.files(the_dir, pattern="JPG$", recursive = TRUE)
+}
+
+file_list <- lapply(X = folders, FUN = list_fun)
+
+
+# -------------------------------------------------------------------------
+
+checkname <- function(drname, filelist, arg) {
+  if (length(filelist)>0){
+    arg$tot <- arg$tot + sum(sapply(filelist, pattern = arg$flname, FUN = grepl))
+  }
+  arg
+}
+
+walk <- function(currdir, f, arg) {
+
+  # "leave trail of bread crumbs"
+  savetop <- getwd()
+  setwd(currdir)
+
+  print(currdir)
+
+  fls <- list.files()
+  arg <- f(currdir, fls, arg)
+
+  # subdirectories of this directory
+  dirs <- list.dirs(recursive=FALSE)
+
+  for (d in dirs) arg <- walk(d,f,arg)
+
+  setwd(savetop) # go back to calling directory
+  return(arg)
+}
+
+countinst <- function(startdir, flname) {
+  walk(startdir, checkname,
+       list(flname=flname, tot=0))
+}
+
+countinst(folder, ".JPG")

### 2023-10-18 docker/build.sh
add build script reminder

@@ -1,0 +1,3 @@
+#!/usr/bin/bash
+
+sudo docker build -t testmd .

### 2023-10-23 dockerfiles/build.sh
update build.sh

@@ -1,3 +1,3 @@
 #!/usr/bin/bash
 
-sudo docker build -t testmd .
+sudo docker build -t rofcamtrap dockerfiles

### 2023-10-23 scripts/bash/camtrap.sh
add bash scripts from old repo

@@ -1,0 +1,385 @@
+#!/usr/bin/bash
+
+# ------------------------------------------------------------------
+
+## CLI UTILITY FOR Camera trap workflow
+## WARNING: Make sure to activate conda env first
+
+# ------------------------------------------------------------------
+
+# Declare variables first
+
+STORAGE_DIR=""
+BASE_FOLDER=""
+
+# ------------------------------------------------------------------
+
+print_main_usage(){
+cat <<EOM
+    Usage: camtrap.sh [-s storage_dir] [-b base_folder] <subcommand>
+
+    CLI UTILITY FOR Camera trap workflow
+    WARNING: Make sure to activate conda env first
+
+    Options:
+      -s    Set the storage directory for camera trap images.
+      -b    Set the base folder for the workflow.
+
+    Subcommands:
+      all      Run the entire camera trap workflow.
+      prep     Run the preparation steps for the workflow.
+      md       Run the megadetector (MD) step.
+      convert  Run the converter to LS step.
+      repeat   Run the repeat detection and conversion step.
+
+    Examples:
+      camtrap.sh -s /path/to/storage -b /path/to/base_folder all
+      camtrap.sh -s /path/to/storage -b /path/to/base_folder prep
+      camtrap.sh -s /path/to/storage -b /path/to/base_folder md
+      camtrap.sh -s /path/to/storage -b /path/to/base_folder convert
+      camtrap.sh -s /path/to/storage -b /path/to/base_folder repeat
+EOM
+}
+
+# ------------------------------------------------------------------
+
+export_default_vars(){
+
+    # STORAGE_DIR="/media/vlucet/TrailCamST/TrailCamStorage"
+    # STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
+    # STORAGE_DIR="/media/vlucet/My Passport/Images"
+    # BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
+    
+    MD_FOLDER="$BASE_FOLDER/cameratraps"
[…]

### 2023-10-23 scripts/bash/move_files_second_retrieval.sh

@@ -1,0 +1,31 @@
+#!/usr/bin/bash
+
+# New storage dir
+STORAGE_DIR="/media/vlucet/My Passport/Images"
+
+# Start with empty array
+DIRS=()
+
+# Find all folders
+echo "Finding all folders"
+for FILE in "$STORAGE_DIR"/*; do
+    echo "$FILE"
+    [[ -d "$FILE" ]] && DIRS+=("$FILE")
+done
+
+echo "      Directory list:"
+printf "      %s\n" "${DIRS[@]}"
+echo ""
+
+for DIR in "${DIRS[@]}"; do
+
+	echo "$DIR"
+	# mkdir "$DIR/Wildlife"
+
+	for SUBFILE in "$DIR"/*; do
+		# echo "$SUBFILE"
+    	[[ -f "$SUBFILE" ]] && mv "$SUBFILE" "$DIR/Wildlife/$(basename "$SUBFILE")"
+	done
+
+done
+

### 2023-10-23 scripts/bash/old_run.sh

@@ -1,0 +1,159 @@
+#!/usr/bin/bash
+
+## CLI UTILITY FOR MD RUN
+## Make sure to activate conda env first
+
+# VARS
+# STORAGE_DIR="/media/vlucet/TrailCamST/TrailCamStorage"
+STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
+BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
+MD_FOLDER="$BASE_FOLDER/cameratraps"
+MODEL="md_v5a.0.0.pt"
+CHECKPOINT_FREQ=1000
+THRESHOLD_FILTER=0.1
+# THRESHOLD=0.0001
+
+OVERWRITE_MD=true
+
+OVERWRITE_LS=true
+
+OVERWRITE_MD_CSV=true
+OVERWRITE_EXIF_CSV=true
+OVERWRITE_COMBINED=true
+
+OVERWRITE_MD_COMBINED=false
+OVERWRITE_EXIF_COMBINED=false
+
+# Export python path
+export PYTHONPATH="$PYTHONPATH:$MD_FOLDER"
+export PYTHONPATH="$PYTHONPATH:$BASE_FOLDER/ai4eutils"
+export PYTHONPATH="$PYTHONPATH:$BASE_FOLDER/yolov5"
+
+# Start with empty array
+DIRS=()
+
+# Find all folders
+echo "Finding all folders"
+for FILE in $STORAGE_DIR/*; do
+    [[ -d $FILE ]] && DIRS+=("$FILE")
+done
+
+# Print the list
+# echo "Directory list:"
+# printf "%s\n" "${DIRS[@]}"
+
+OLD_DIR=$PWD
+
+for DIR in "${DIRS[@]}"; do # "P072"; do 
+
+    echo "*** RUNNING MD ***"
+    
+    RUN_DIR=$STORAGE_DIR/$(basename $DIR)
+    echo "Running on directory: $RUN_DIR"
+
+    OUTPUT_JSON="$(basename $DIR)_output.json"
+    echo $OUTPUT_JSON
+    
+    CHECKPOINT_PATH="$(basename $DIR)_checkpoint.json"
+    echo $CHECKPOINT_PATH
+
+    if [ -f "$STORAGE_DIR/$OUTPUT_JSON" ] && [ "$OVERWRITE_MD" != true ]; then # if output exist, do nothing
+        
[…]

### 2023-10-23 scripts/bash/test.sh

@@ -1,0 +1,64 @@
+#!/usr/bin/bash
+
+STORAGE_DIR=""
+BASE_FOLDER=""
+
+export_default_vars(){
+    echo "Storage Directory: $STORAGE_DIR"
+    echo "Base Folder: $BASE_FOLDER"
+}
+
+run_prep(){
+    export_default_vars
+    # Other prep steps
+}
+
+run_all(){
+    run_prep
+    # Other steps for run_all
+}
+
+
+shift $((OPTIND - 1))
+
+subcommand=$1; shift
+case "$subcommand" in
+    all )
+        echo "Running 'run_all'"
+        while getopts ":s:b:" opt; do
+		  case ${opt} in
+		    s )
+		      STORAGE_DIR=$OPTARG
+		      ;;
+		    b )
+		      BASE_FOLDER=$OPTARG
+		      ;;
+		    \? )
+		      echo "Invalid option -$OPTARG"
+		      exit 1
+		      ;;
+		  	esac
+		done
+        run_all
+        ;;
+    prep )
+        echo "Running 'run_prep'"
+        run_prep
+        ;;
+    md )
+        run_prep
+        # run_md
+        ;;
+    convert )
+        run_prep
+        # run_convert
+        ;;
+    repeat )
+        run_prep
+        # Other steps for repeat
+        ;;
+    \? )
+        echo "Invalid subcommand: $subcommand"
+        exit 1
+        ;;
+esac

### 2023-10-23 scripts/bash/camtrap.sh
encoding

@@ -49,7 +49,7 @@ export_default_vars(){
     # STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
     # STORAGE_DIR="/media/vlucet/My Passport/Images"
     # BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
-    
+
     MD_FOLDER="$BASE_FOLDER/cameratraps"
     MODEL="md_v5a.0.0.pt"
     CHECKPOINT_FREQ=1000
@@ -137,7 +137,7 @@ run_md(){
 
     OLD_DIR=$PWD
 
-    for DIR in "${DIRS[@]}"; do # "P072"; do 
+    for DIR in "${DIRS[@]}"; do # "P072"; do
 
         echo "*** RUNNING MD ***"
 
@@ -146,12 +146,12 @@ run_md(){
 
         OUTPUT_JSON="$(basename "$DIR")_output.json"
         echo $OUTPUT_JSON
-        
+
         CHECKPOINT_PATH="$(basename "$DIR")_checkpoint.json"
         echo $CHECKPOINT_PATH
 
         if [ -f "$STORAGE_DIR/$OUTPUT_JSON" ] && [ "$OVERWRITE_MD" != true ]; then # if output exist, do nothing
-            
+
             echo "Output file $OUTPUT_JSON exists, moving to the next folder"
 
         elif [ -f "$STORAGE_DIR/$CHECKPOINT_PATH" ]; then # else, if checkpoint exists, use it
@@ -199,7 +199,7 @@ run_convert(){
         echo $OUTPUT_JSON_LS
 
         if [ -f "$STORAGE_DIR/$OUTPUT_JSON_LS" ] && [ "$OVERWRITE_LS" != true ]; then # if output exist, do nothing
-            
+
             echo "Output file $OUTPUT_JSON_LS exists, moving to the next folder"
 
         else
@@ -279,7 +279,7 @@ run_convert_repeat(){
     #     echo $OUTPUT_COCO_REPEAT
 
     #     if [ -f "$STORAGE_DIR/$OUTPUT_COCO_REPEAT" ] && [ "$OVERWRITE_COCO_REPEAT" != true ]; then # if output exist, do nothing
-            
+
     #         echo "Output file $OUTPUT_COCO_REPEAT exists, moving to the next folder"
 
     #     else
@@ -303,7 +303,7 @@ run_convert_repeat(){
         echo $OUTPUT_JSON_LS
 
[…]

### 2023-10-23 scripts/bash/camtrap.sh
keep improving script to run properly

@@ -49,9 +49,9 @@ export_default_vars(){
     # STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
     # STORAGE_DIR="/media/vlucet/My Passport/Images"
     # BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
+    # MODEL="md_v5a.0.0.pt"
 
-    MD_FOLDER="$BASE_FOLDER/cameratraps"
-    MODEL="md_v5a.0.0.pt"
+    MD_FOLDER="$BASE_FOLDER/MegaDetector"
     CHECKPOINT_FREQ=1000
     THRESHOLD_FILTER=0.1
     # THRESHOLD=0.0001
@@ -59,7 +59,7 @@ export_default_vars(){
     IOU_THRESHOLD=0.85
     NDIR_LEVEL=1
 
-    OVERWRITE_MD=false
+    OVERWRITE_MD=true
     OVERWRITE_LS=true
     OVERWRITE_MD_CSV=true
     OVERWRITE_EXIF_CSV=true
@@ -157,7 +157,7 @@ run_md(){
         elif [ -f "$STORAGE_DIR/$CHECKPOINT_PATH" ]; then # else, if checkpoint exists, use it
 
             python $MD_FOLDER/detection/run_detector_batch.py \
-                $MD_FOLDER/$MODEL "$RUN_DIR" "$STORAGE_DIR/$OUTPUT_JSON" \
+                $MODEL_PATH "$RUN_DIR" "$STORAGE_DIR/$OUTPUT_JSON" \
                 --output_relative_filenames --recursive \
                 --checkpoint_frequency $CHECKPOINT_FREQ \
                 --checkpoint_path "$STORAGE_DIR/$CHECKPOINT_PATH" \
@@ -168,7 +168,7 @@ run_md(){
 
         else # else, start new run
             python $MD_FOLDER/detection/run_detector_batch.py \
-                $MD_FOLDER/$MODEL "$RUN_DIR" "$STORAGE_DIR/$OUTPUT_JSON" \
+                $MODEL_PATH "$RUN_DIR" "$STORAGE_DIR/$OUTPUT_JSON" \
                 --output_relative_filenames --recursive \
                 --include_max_conf \
                 --checkpoint_frequency $CHECKPOINT_FREQ \
@@ -320,13 +320,16 @@ run_convert_repeat(){
 
[…]

### 2023-10-23 scripts/bash/camtrap.sh
updatEs

@@ -11,6 +11,8 @@
 
 STORAGE_DIR=""
 BASE_FOLDER=""
+INPUT_DIR=""
+OUTPUT_DIR=""
 
 # ------------------------------------------------------------------
 
@@ -24,6 +26,9 @@ cat <<EOM
     Options:
       -s    Set the storage directory for camera trap images.
       -b    Set the base folder for the workflow.
+      -m    Set the path to model file to use.
+      -o    Set the output folder.
+      -h    Prints usage.
 
     Subcommands:
       all      Run the entire camera trap workflow.
@@ -45,19 +50,12 @@ EOM
 
 export_default_vars(){
 
-    # STORAGE_DIR="/media/vlucet/TrailCamST/TrailCamStorage"
-    # STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
-    # STORAGE_DIR="/media/vlucet/My Passport/Images"
-    # BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
-    # MODEL="md_v5a.0.0.pt"
-
     MD_FOLDER="$BASE_FOLDER/MegaDetector"
     CHECKPOINT_FREQ=1000
     THRESHOLD_FILTER=0.1
-    # THRESHOLD=0.0001
 
     IOU_THRESHOLD=0.85
-    NDIR_LEVEL=1
+    NDIR_LEVEL=0
 
     OVERWRITE_MD=true
     OVERWRITE_LS=true
@@ -74,6 +72,13 @@ export_default_vars(){
     export PYTHONPATH="$PYTHONPATH:$MD_FOLDER"
     export PYTHONPATH="$PYTHONPATH:$BASE_FOLDER/ai4eutils"
     export PYTHONPATH="$PYTHONPATH:$BASE_FOLDER/yolov5"
+
+    # STORAGE_DIR="/media/vlucet/TrailCamST/TrailCamStorage"
+    # STORAGE_DIR="/home/vlucet/Documents/WILDLab/mdtools/tests/test_images/"
+    # STORAGE_DIR="/media/vlucet/My Passport/Images"
+    # BASE_FOLDER="/home/vlucet/Documents/WILDLab/repos/MDtest/git"
+    # MODEL="md_v5a.0.0.pt"
+    # THRESHOLD=0.0001
 }
 
 print_vars(){
@@ -83,7 +88,7 @@ print_vars(){
[…]

### 2023-10-23 README.Rmd
update
add readme info
progress

@@
 
-# rofcamtrap
+# ROF Camera Trap Data Analysis Research Compedium
 
@@
 
-This repository contains the data and code for our preliminary report.
+Prerequisites:
+ - docker
+ - nvidia-container-toolkit
 
-<!-- This repository contains the data and code for our paper:
+## Main Workflow
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
+1. At the terminal, clone this repository.
 
-Our pre-print is online here:
+```{bash eval=FALSE}
+git clone https://github.com/StewartWILDlab/rofcamtrap
+```
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->
+2. Build the docker image of the computing environment, by running the build script.
 
-### How to cite
+```{bash eval=FALSE}
+rofcamtrap/dockerfiles/build.sh
+```
 
-Please cite this compendium as:
+3.
+Run docker image, with the proper volumes.
+We currently have two separate storage volumes, for each of the two camera retrievals that took place.
+We also make sure to hook the `rofcamtrap` folder as a volume.
 
-> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`).
-_Compendium of R code and data for `r Title`_.
-Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/xxx/xxx>
+```{bash eval=FALSE}
+# -e DISABLE_AUTH=true --shm-size 50G
+sudo docker run \
+  -v "$(pwd):/workspace/rofcamtrap" \
+  -v "/media/vlucet/TrailCamST/TrailCamStorage:/workspace/storage/TrailCamStorage" \
+  -v "/media/vlucet/My Passport/Images:/workspace/storage/my_passport_images" \
+  --gpus all \
+  -it rofcamtrap
+```
 
-## Contents
+4. Running mega detector on images using the bash script.
 
-The **analysis** directory contains:
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/my_passport_images" \ 
+  -m "/workspace/models/md_v5a.0.0.pt" \
+  md
+```
+
+5. Run the repeat detector using the bash script, and remove all instances of trues positives.
+
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/my_passport_images" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs" \
+  repeat-detect
+```
+
+6. Patch MD's output with the repeat detect results.
+
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/my_passport_images" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs" \
+  -o "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats" \
+  repeat-remove
+```
+
+## Classifier training workflow
+
+### Species classifier
+
+### False detections classifier
+
+## LabelStudio instance setup
+
+## Labelme? WildTrax?
+
+## Reports
+
+The **reports** directory currently contains the `GnC_report` folder wich is structured as such:
 
@@
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
+  - [:file\_folder: supplementary-materials](/analysis/supplementary-materials):
     Supplementary materials including notes and other documents
@@
 By participating in this project you agree to abide by its terms.
+
+### How to cite
+
+Please cite this compendium as:
+
+> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`).
+_Compendium of R code and data for `r Title`_.
+Accessed `r format(Sys.Date(), "%d %b %Y")`.
+Online at <https://doi.org/xxx/xxx>
+
+### Notes
+
+```
+for FILE in project*
+    mdtools postprocess --write-csv $FILE
+end
+```
+
+<!-- This repository contains the data and code for our paper:
+
+> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>
+
+Our pre-print is online here:
+
+> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->

### 2023-10-23 R/process_md_outputs.R
add fucntion

@@ -1,0 +1,6 @@
+
+#' @export
+read_md_json <- function(md_json_file){
+  stopifnot(file.exists(md_json_file))
+  jsonlite::read_json(md_json_file, simplifyVector=F)
+}

### 2023-10-30 README.Rmd
updatEs

@@
 
-4. Running mega detector on images using the bash script.
+4. Activate ENV, then run mega detector on images using the bash script.
 
 ```{bash eval=FALSE}
+mamba activate cameratraps-detector
 rofcamtrap/scripts/bash/camtrap.sh \
@@
   repeat-remove
+```
+
+7. Optionally, write out the visualizations of the detections.
+
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/my_passport_images" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/" \
+  -o "/workspace/rofcamtrap/1_MegaDetector/2_visualize/" \
+  viz
+```
+
+8. Convert to coco and ls [and 3rd format?].
+
+```{bash eval=FALSE}
+mamba deactivate 
+cd mdtools
+poetry shell
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/my_passport_images" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/" \
+  -o "/workspace/rofcamtrap/2_LabelStudio/0_inputs/" \
+  repeat-convert
 ```

### 2023-10-30 dockerfiles/build.sh
improvements

@@ -1,3 +1,3 @@
 #!/usr/bin/bash
 
-sudo docker build -t rofcamtrap dockerfiles
+docker build -t rofcamtrap dockerfiles

### 2023-10-30 scripts/bash/camtrap.sh

@@ -27,6 +27,7 @@ cat <<EOM
       -s    Set the storage directory for camera trap images.
       -b    Set the base folder for the workflow.
       -m    Set the path to model file to use.
+      -i    Set the input folder
       -o    Set the output folder.
       -h    Prints usage.
 
@@ -34,6 +35,7 @@ cat <<EOM
       all      Run the entire camera trap workflow.
       prep     Run the preparation steps for the workflow.
       md       Run the megadetector (MD) step.
+      viz      Run the visualization step.
       convert  Run the converter to LS step.
       repeat   Run the repeat detection and conversion step.
 
@@ -62,12 +64,12 @@ export_default_vars(){
     OVERWRITE_MD_CSV=true
     OVERWRITE_EXIF_CSV=true
 
-    OVERWRITE_COCO_REPEAT=false
-    OVERWRITE_LS_REPEAT=false
+    OVERWRITE_COCO_REPEAT=true
+    OVERWRITE_LS_REPEAT=true
 
     OVERWRITE_COMBINED=true
-    OVERWRITE_MD_COMBINED=false
-    OVERWRITE_EXIF_COMBINED=false
+    OVERWRITE_MD_COMBINED=true
+    OVERWRITE_EXIF_COMBINED=true
 
     export PYTHONPATH="$PYTHONPATH:$MD_FOLDER"
     export PYTHONPATH="$PYTHONPATH:$BASE_FOLDER/ai4eutils"
@@ -112,16 +114,16 @@ crawl_dirs(){
     # Find all folders
     echo "Finding all folders"
     for FILE in "$STORAGE_DIR"/*; do
-        echo "$FILE"
+        # echo "$FILE"
         if [[ "$FILE" == *"repeat"* ]];then
             continue
         fi
         [[ -d "$FILE" ]] && DIRS+=("$FILE")
     done
 
-    echo "      Directory list:"
-    printf "      %s\n" "${DIRS[@]}"
-    echo ""
+    # echo "      Directory list:"
+    # printf "      %s\n" "${DIRS[@]}"
+    # echo ""
 
     export DIRS
 }
@@ -237,9 +239,7 @@ run_remove_repeat(){
 
 run_convert_repeat(){
 
-    OLD_DIR=$PWD
-
-    for DIR in "${DIRS[@]}"; do
[…]

### 2023-10-31 dockerfiles/build.sh
add apptainer

@@ -1,3 +1,7 @@
 #!/usr/bin/bash
 
 docker build -t rofcamtrap dockerfiles
+docker save rofcamtrap -o rofcamtrap.tar
+# docker image rm rofcamtrap
+apptainer build --fakeroot rofcamtrap.sif docker-archive://rofcamtrap.tar
+rm rofcamtrap.tar

### 2023-10-31 scripts/bash/test.sh
remove useless file

@@ -1,64 +1,0 @@
-#!/usr/bin/bash
-
-STORAGE_DIR=""
-BASE_FOLDER=""
-
-export_default_vars(){
-    echo "Storage Directory: $STORAGE_DIR"
-    echo "Base Folder: $BASE_FOLDER"
-}
-
-run_prep(){
-    export_default_vars
-    # Other prep steps
-}
-
-run_all(){
-    run_prep
-    # Other steps for run_all
-}
-
-
-shift $((OPTIND - 1))
-
-subcommand=$1; shift
-case "$subcommand" in
-    all )
-        echo "Running 'run_all'"
-        while getopts ":s:b:" opt; do
-		  case ${opt} in
-		    s )
-		      STORAGE_DIR=$OPTARG
-		      ;;
-		    b )
-		      BASE_FOLDER=$OPTARG
-		      ;;
-		    \? )
-		      echo "Invalid option -$OPTARG"
-		      exit 1
-		      ;;
-		  	esac
-		done
-        run_all
-        ;;
-    prep )
-        echo "Running 'run_prep'"
-        run_prep
-        ;;
-    md )
-        run_prep
-        # run_md
-        ;;
-    convert )
-        run_prep
-        # run_convert
-        ;;
-    repeat )
-        run_prep
-        # Other steps for repeat
-        ;;
-    \? )
-        echo "Invalid subcommand: $subcommand"
-        exit 1
-        ;;
-esac

### 2023-10-31 scripts/bash/classify_species.sh
add scripts

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+# export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/InternImage"
+# export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/InternImage/classification/ops_dcnv3"
+# export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/metaformer"
+
+export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
+
+python3 ./scripts/python/species_classifier/classify_species.py
+# python3 ./scripts/classsify_lightning.py

### 2023-10-31 scripts/python/species_classifier/classify_species.py

@@ -1,0 +1,326 @@
+#############################################################
+
+import matplotlib.pyplot as plt
+import pandas as pd
+from PIL import Image
+from tqdm import tqdm
+from sklearn.model_selection import train_test_split
+
+import torch
+from torch.utils.data import Dataset
+from torchvision import transforms
+
+from torch.utils.data import DataLoader
+
+from torch import nn
+import torchvision.models as models
+
+# from metaformer.models.MetaFG import * 
+# from metaformer.models.MetaFG_meta import *
+
+do_train = True
+do_predict = True
+batch_size = 32
+epochs = 1
+random_state = 777
+fit_split = 0.25
+eval_split = 0.25 
+
+print(torch.cuda.is_available())
+device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
+# device = "cpu"
+print(device)
+
+#############################################################
+
+dat = pd.read_csv("3_Classifiers/1_species_classifier/labels.csv", index_col = "id")
+limit = 3
+dat_info = dat.iloc[:, :limit]
+dat_labels = dat.iloc[:, limit:]
+
+print(dat.head())
+print(dat_info.head())
+
+all_x = dat_info.iloc[:,0:1]
+all_y = dat_labels
+
+print(list(all_x.columns))
+print(list(all_y.columns))
+
+species_labels = (all_y.columns.unique().tolist())
+print(species_labels)
+number_of_categories = len(species_labels)
+
+# print(all_y.sum().divide(all_y.shape[0]).sort_values(ascending=False))
+
+# x_fit, x_predict, y_fit, y_predict = train_test_split(
+#     all_x, all_y, stratify=all_y, test_size=fit_split, random_state=random_state
+# )
+# print(x_fit.shape, x_predict.shape, y_fit.shape, y_predict.shape)
+
+# x_train, x_eval, y_train, y_eval = train_test_split(
+#     x_fit, y_fit, stratify=y_fit, test_size=eval_split, random_state=random_state
+# )
+# print(x_train.shape, y_train.shape, x_eval.shape, y_eval.shape)
+
+# split_pcts_fit = pd.DataFrame(
+#     {
+#         "fit": y_fit.idxmax(axis=1).value_counts(normalize=True),
+#         "predict": y_predict.idxmax(axis=1).value_counts(normalize=True),
+#     }
+# )
+# print("Species percentages by split")
+# print((split_pcts_fit.fillna(0) * 100).astype(int))
+
[…]

### 2023-10-31 README.Rmd
updates

@@
 
-2. Build the docker image of the computing environment, by running the build script.
+2. Build the docker image of the computing environment, by running the build script. This will also build the apptainer image needed on HPC.
 
@@
 ## Classifier training workflow
+
+On beluga, we use the apptainer iamge instead.
+
+```{bash eval=FALSE}
+. /workspace/conda/etc/profile.d/conda.sh 
+. /workspace/conda/etc/profile.d/mamba.sh
+PATH="$PATH:$HOME/.local/bin"
+mamba activate cameratraps-detector
+rofcamtrap/scripts/bash/camtrap.sh -b "/workspace/git" -s "/workspace/storage/my_passport_images" -m "/workspace/models/md_v5a.0.0.pt" md
+```
 

### 2023-10-31 hcp/test_job.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+#SBATCH --account=rrg-fstewart
+#SBATCH --gpus-per-node=1         # Number of GPU(s) per node
+#SBATCH --cpus-per-task=1         # CPU cores/threads
+#SBATCH --mem=4GB               # memory per node
+#SBATCH --time=0-03:00
+
+module load apptainer
+
+apptainer run --nv -c rofcamtrap.sif

### 2023-10-31 scripts/python/species_classifier/classify_species.py

@@ -240,17 +240,17 @@ if do_train:
     plt.ylabel("Loss")
     plt.legend(loc=0)
     # plt.show()
-    plt.savefig('loss.png')
+    plt.savefig('3_Classifiers/1_species_classifier/loss.png')
 
     print("TRAINING DONE")
 
-    torch.save(model, "model.pth")
+    torch.save(model, "3_Classifiers/1_species_classifier/model.pth")
 
     print("MODEL SAVED")
 
 #############################################################
 
-loaded_model = torch.load("model.pth")
+loaded_model = torch.load("3_Classifiers/1_species_classifier/model.pth")
 
 eval_dataset = ImagesDataset(x_eval, y_eval, device=device)
 eval_dataloader = DataLoader(eval_dataset, batch_size=batch_size)
@@ -287,7 +287,7 @@ if do_predict:
 
     eval_preds_df = pd.concat(preds_collector)
 
-    eval_preds_df.to_csv("predictions.csv")
+    eval_preds_df.to_csv("3_Classifiers/1_species_classifier/predictions.csv")
 
 # eval_preds_df = pd.read_csv("predictions.csv", index_col=0)
 

### 2023-10-31 README.Rmd
add code

--- 2023-10-30 README.Rmd
@@
 ## Classifier training workflow
+
+On beluga, we use the apptainer iamge instead.
+
+```{bash eval=FALSE}
+. /workspace/conda/etc/profile.d/conda.sh 
+. /workspace/conda/etc/profile.d/mamba.sh
+PATH="$PATH:$HOME/.local/bin"
+mamba activate cameratraps-detector
+rofcamtrap/scripts/bash/camtrap.sh -b "/workspace/git" -s "/workspace/storage/my_passport_images" -m "/workspace/models/md_v5a.0.0.pt" md
+
+$ mkdir -p /scratch/$USER/apptainer/{cache,tmp}
+$ export APPTAINER_CACHEDIR="/scratch/$USER/apptainer/cache"
+$ export APPTAINER_TMPDIR="/scratch/$USER/apptainer/tmp"
+
+salloc --time=00:05:00 --mem=4G --ntasks=1 --gpus-per-task=1 --cpus-per-task=1 --account=rrg-fstewart
+
+apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap"  -B "/media/vlucet/TrailCamST/TrailCamStorage:/workspace/storage/TrailCamStorage"  -B "/media/vlucet/My Passport/Images:/workspace/storage/my_passport_images" rofcamtrap.sif
+
+apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif
+```
 

### 2023-10-31 scripts/python/species_classifier/classify_species.py

@@ -20,8 +20,10 @@ import torchvision.models as models
 
 do_train = True
 do_predict = True
+replace_path = True
+path_replacement = "/workspace/project/data/images/"
 batch_size = 32
-epochs = 1
+epochs = 100
 random_state = 777
 fit_split = 0.25
 eval_split = 0.25 
@@ -109,7 +111,10 @@ class ImagesDataset(Dataset):
         self.device = device
 
     def __getitem__(self, index):
-        image = Image.open(self.data.iloc[index]["filepath"]).convert("RGB")
+        image_path = self.data.iloc[index]["filepath"]
+        if replace_path:
+          image_path = path_replacement + os.path.basename(image_path)
+        image = Image.open(image_path).convert("RGB")
         image = self.transform(image).to(self.device)
         image_id = self.data.index[index]
         # if we don't have labels (e.g. for test set) just return the image and image id

### 2023-10-31 scripts/python/species_classifier/classify_species.py
troubleshoot resnet

@@ -137,7 +137,9 @@ train_dataloader = DataLoader(train_dataset, batch_size=batch_size)
 
 #############################################################
 
-model = models.resnet50('IMAGENET1K_V2')
+model = models.resnet50()
+state_dict = torch.load("models/resnet/pretrained/resnet50-11ad3fa6.pth")
+model.load_state_dict(state_dict)
 for param in model.parameters():
     param.requires_grad = False
 print(model)

### 2023-10-31 scripts/python/species_classifier/classify_species.py
fix osd import

@@ -15,6 +15,8 @@ from torch.utils.data import DataLoader
 from torch import nn
 import torchvision.models as models
 
+import os
+
 # from metaformer.models.MetaFG import * 
 # from metaformer.models.MetaFG_meta import *
 

### 2023-10-31 hcp/test_job.sh
update

@@ -3,8 +3,9 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=4GB               # memory per node
-#SBATCH --time=0-03:00
+#SBATCH --time=00:05:00
 
 module load apptainer
 
-apptainer run --nv -c rofcamtrap.sif
+apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
+  "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif

### 2023-10-31 scripts/bash/classify_species.sh

@@ -6,5 +6,9 @@
 
 export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
 
+. /workspace/conda/etc/profile.d/conda.sh
+. /workspace/conda/etc/profile.d/mamba.sh
+PATH="$PATH:$HOME/.local/bin"
+mamba activate cameratraps-detector
+
 python3 ./scripts/python/species_classifier/classify_species.py
-# python3 ./scripts/classsify_lightning.py

### 2023-10-31 hcp/test_job.sh
add cd

@@ -7,5 +7,7 @@
 
 module load apptainer
 
+cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
+
 apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
   "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif

### 2023-10-31 dockerfiles/build.sh
comment out

@@ -4,4 +4,4 @@ docker build -t rofcamtrap dockerfiles
 docker save rofcamtrap -o rofcamtrap.tar
 # docker image rm rofcamtrap
 apptainer build --fakeroot rofcamtrap.sif docker-archive://rofcamtrap.tar
-rm rofcamtrap.tar
+# rm rofcamtrap.tar

### 2023-10-31 hcp/test_job.sh
update job

@@ -10,4 +10,4 @@ module load apptainer
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
 apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
-  "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif
+  "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
change to exec

@@ -9,5 +9,5 @@ module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
-apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
+apptainer exec --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
   "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
try run instead

@@ -9,5 +9,5 @@ module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
-apptainer exec --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
+apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
   "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
back to shell

@@ -9,5 +9,4 @@ module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
-apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B \
-  "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh
+apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
try again

@@ -8,5 +8,5 @@
 module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
-
-apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ./scripts/bash/classify_species.sh
+ls
+apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ls && ./scripts/bash/classify_species.sh

### 2023-10-31 scripts/bash/classify_species.sh
change to source

@@ -6,8 +6,8 @@
 
 export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
 
-. /workspace/conda/etc/profile.d/conda.sh
-. /workspace/conda/etc/profile.d/mamba.sh
+source /workspace/conda/etc/profile.d/conda.sh
+source /workspace/conda/etc/profile.d/mamba.sh
 PATH="$PATH:$HOME/.local/bin"
 mamba activate cameratraps-detector
 

### 2023-10-31 hcp/test_job.sh
run again

@@ -9,4 +9,4 @@ module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 ls
-apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ls && ./scripts/bash/classify_species.sh
+apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ls && ./scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
try again

@@ -8,5 +8,5 @@
 module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
-ls
-apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif cd /workspace/rofcamtrap && ls && ./scripts/bash/classify_species.sh
+
+apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif ./scripts/bash/classify_species.sh

### 2023-10-31 scripts/bash/classify_species.sh

@@ -6,6 +6,12 @@
 
 export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
 
+ls
+
+cd /workspace/rofcamtrap
+
+ls
+
 source /workspace/conda/etc/profile.d/conda.sh
 source /workspace/conda/etc/profile.d/mamba.sh
 PATH="$PATH:$HOME/.local/bin"

### 2023-10-31 hcp/test_job.sh
cahnge time

@@ -3,7 +3,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=4GB               # memory per node
-#SBATCH --time=00:05:00
+#SBATCH --time=00:03:00
 
 module load apptainer
 

### 2023-10-31 hcp/test_job.sh
cahnge again

@@ -9,4 +9,4 @@ module load apptainer
 
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
-apptainer run --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif ./scripts/bash/classify_species.sh
+apptainer exec --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif /workspace/rofcamtrap/scripts/bash/classify_species.sh

### 2023-10-31 hcp/test_job.sh
update

@@ -3,7 +3,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=4GB               # memory per node
-#SBATCH --time=00:03:00
+#SBATCH --time=00:15:00
 
 module load apptainer
 

### 2023-10-31 scripts/bash/classify_species.sh

@@ -6,12 +6,7 @@
 
 export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
 
-ls
-
 cd /workspace/rofcamtrap
-
-ls
-
 source /workspace/conda/etc/profile.d/conda.sh
 source /workspace/conda/etc/profile.d/mamba.sh
 PATH="$PATH:$HOME/.local/bin"

### 2023-10-31 scripts/python/species_classifier/classify_species.py

@@ -25,7 +25,7 @@ do_predict = True
 replace_path = True
 path_replacement = "/workspace/project/data/images/"
 batch_size = 32
-epochs = 100
+epochs = 1
 random_state = 777
 fit_split = 0.25
 eval_split = 0.25 

### 2023-10-31 hcp/test_job.sh
test

@@ -1,9 +1,9 @@
 #!/bin/bash
 #SBATCH --account=rrg-fstewart
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
-#SBATCH --cpus-per-task=1         # CPU cores/threads
-#SBATCH --mem=4GB               # memory per node
-#SBATCH --time=00:15:00
+#SBATCH --cpus-per-node=2         # CPU cores/threads
+#SBATCH --mem=8GB                 # memory per node
+#SBATCH --time=00:30:00
 
 module load apptainer
 

### 2023-10-31 hcp/test_job.sh
update

@@ -1,7 +1,10 @@
 #!/bin/bash
+#SBATCH --mail-user=valentin.lucet@gmail.com
+#SBATCH --mail-type=END,FAIL
 #SBATCH --account=rrg-fstewart
+#SBATCH --ntasks=1
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
-#SBATCH --cpus-per-node=2         # CPU cores/threads
+#SBATCH --cpus-per-task=2         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
 #SBATCH --time=00:30:00
 

### 2023-10-31 hcp/test_job.sh
add comments

@@ -8,8 +8,16 @@
 #SBATCH --mem=8GB                 # memory per node
 #SBATCH --time=00:30:00
 
+# Load the apptainer module, latest is fine
 module load apptainer
 
+# Move into the rofcamtrap folder
 cd /home/vlucet/projects/rrg-fstewart/vlucet/rofcamtrap
 
-apptainer exec --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif /workspace/rofcamtrap/scripts/bash/classify_species.sh
+# Run apptainer
+#   --nv makes nvidia cuda work
+#   -C contains the filesystem, used for simplicity
+#   -B allows to bind volumes, similar to -v for docker
+apptainer exec --nv -C -B "$(pwd):/workspace/rofcamtrap" \
+  -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" \
+  rofcamtrap.sif /workspace/rofcamtrap/scripts/bash/classify_species.sh

### 2023-10-31 scripts/bash/classify_species.sh

@@ -1,15 +1,22 @@
 #!/bin/bash
 
+# This is for putting the models other than resnets on path
 # export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/InternImage"
 # export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/InternImage/classification/ops_dcnv3"
 # export PYTHONPATH="$PYTHONPATH:/media/vlucet/TrailCamST/Cropped/metaformer"
 
-export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
+# Necessary to make torch work properly on laptop
+# export 'PYTORCH_CUDA_ALLOC_CONF=max_split_size_mb:512'
 
+# Necessary only on cluster
+# Move to folder within container,
+#   source the conda and mamba binaries, put bin on path
+#   activate the MD environment in which we also run the classifiers
 cd /workspace/rofcamtrap
 source /workspace/conda/etc/profile.d/conda.sh
 source /workspace/conda/etc/profile.d/mamba.sh
 PATH="$PATH:$HOME/.local/bin"
 mamba activate cameratraps-detector
 
+# Run main script
 python3 ./scripts/python/species_classifier/classify_species.py

### 2023-10-31 hcp/test_job.sh
get ready for big run

@@ -6,7 +6,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=2         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
-#SBATCH --time=00:30:00
+#SBATCH --time=26:00:00
 
 # Load the apptainer module, latest is fine
 module load apptainer

### 2023-10-31 scripts/python/species_classifier/classify_species.py

@@ -25,7 +25,7 @@ do_predict = True
 replace_path = True
 path_replacement = "/workspace/project/data/images/"
 batch_size = 32
-epochs = 1
+epochs = 100
 random_state = 777
 fit_split = 0.25
 eval_split = 0.25 

### 2023-11-01 hcp/test_job.sh
test non freeze

@@ -6,7 +6,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=2         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
-#SBATCH --time=26:00:00
+#SBATCH --time=00:10:00
 
 # Load the apptainer module, latest is fine
 module load apptainer

### 2023-11-01 scripts/python/species_classifier/classify_species.py

@@ -23,7 +23,8 @@ import os
 do_train = True
 do_predict = True
 replace_path = True
-path_replacement = "/workspace/project/data/images/"
+# path_replacement = "/workspace/project/data/images/"
+path_replacement = "/home/vlucet/Documents/WILDLab/all/"
 batch_size = 32
 epochs = 100
 random_state = 777
@@ -143,8 +144,8 @@ model = models.resnet50()
 state_dict = torch.load("models/resnet/pretrained/resnet50-11ad3fa6.pth")
 model.load_state_dict(state_dict)
 for param in model.parameters():
-    param.requires_grad = False
-print(model)
+    param.requires_grad = True
+# print(model)
 model.fc = nn.Sequential(
     nn.Linear(2048, 100),  # dense layer takes a 2048-dim input and outputs 100-dim
     nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity

### 2023-11-09 hcp/test_job.sh
test

@@ -6,7 +6,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
-#SBATCH --time=20:00:00
+#SBATCH --time=15:00:00
 #SBATCH --output slurm-%j.log
 
 # https://unix.stackexchange.com/questions/285690/slurm-custom-standard-output-name

### 2023-11-09 scripts/python/species_classifier/classify_species.py
different optimizer

@@ -31,7 +31,7 @@ if replace_path:
 else:
   the_basepath = None
 
-batch_size = 16
+batch_size = 32
 epochs = 100
 random_state = 777
 fit_split = 0.25
@@ -152,14 +152,17 @@ model.load_state_dict(state_dict)
 for param in model.parameters():
     param.requires_grad = False
 # print(model)
-model.fc = nn.Sequential(
-    nn.Linear(2048, 100),  # dense layer takes a 2048-dim input and outputs 100-dim
-    nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
-    nn.Dropout(0.1),  # common technique to mitigate overfitting
-    nn.Linear(
-        100, number_of_categories
-    ),  # final dense layer outputs 8-dim corresponding to our target classes
-)
+
+if model_name in ["resnet50", "resnet101"]:
+    model.fc = nn.Sequential(
+        nn.Linear(2048, 100),  # dense layer takes a 2048-dim input and outputs 100-dim
+        nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
+        nn.Dropout(0.1),  # common technique to mitigate overfitting
+        nn.Linear(
+            100, number_of_categories
+        ),  # final dense layer outputs 8-dim corresponding to our target classes
+    )
+
 for param in model.fc.parameters():
     param.requires_grad = True
 # print(model)
@@ -171,7 +174,8 @@ model = model.to(device)
 import torch.optim as optim
 
 criterion = nn.CrossEntropyLoss()
-optimizer = optim.SGD(model.parameters(), lr=0.001, momentum=0.9)
+# optimizer = optim.SGD(model.parameters(), lr=0.001, momentum=0.9)
+optimizer = optim.Adam(model.parameters(), lr=0.001)
 
 #############################################################
 
@@ -202,7 +206,8 @@ if do_train:
 
             # 4) compute our gradients
             loss.backward()
-            # update our weights
+            
+            # 5) update our weights
             optimizer.step()
 
     tracking_loss = pd.Series(tracking_loss)
@@ -246,8 +251,10 @@ if do_predict:
     # we aren't updating our weights so no need to calculate gradients
     with torch.no_grad():
         for batch in tqdm(eval_dataloader, total=len(eval_dataloader)):
+          
             # 1) run the forward step
             logits = model.forward(batch["image"])
+            
             # 2) apply softmax so that model outputs are in range [0,1]
             preds = nn.functional.softmax(logits, dim=1)
 

### 2023-11-09 hcp/test_job.sh
test 30 min

@@ -6,7 +6,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
-#SBATCH --time=15:00:00
+#SBATCH --time=00:30:00
 #SBATCH --output slurm-%j.log
 
 # https://unix.stackexchange.com/questions/285690/slurm-custom-standard-output-name

### 2023-11-10 scripts/python/species_classifier/classify_species.py
add val loss

@@ -15,10 +15,14 @@ from torch.utils.data import DataLoader
 from torch import nn
 import torchvision.models as models
 
+from sklearn.metrics import ConfusionMatrixDisplay
+
 import os
 
 # from metaformer.models.MetaFG import * 
 # from metaformer.models.MetaFG_meta import *
+
+#############################################################
 
 do_train = True
 do_predict = True
@@ -34,7 +38,7 @@ else:
 batch_size = 32
 epochs = 100
 random_state = 777
-fit_split = 0.25
+# fit_split = 0.25
 eval_split = 0.25 
 
 model_name = "resnet101"
@@ -49,7 +53,7 @@ print(run_name)
 
 #############################################################
 
-dat = pd.read_csv("3_Classifiers/1_species_classifier/labels.csv", index_col = "id")
+dat = pd.read_csv("3_Classifiers/1_species_classifier/labels.csv", index_col = "id")#.sample(frac=0.1)
 limit = 3
 dat_info = dat.iloc[:, :limit]
 dat_labels = dat.iloc[:, limit:]
@@ -74,10 +78,13 @@ print(x_train.shape, y_train.shape, x_eval.shape, y_eval.shape)
 
 split_pcts_train = pd.DataFrame(
     {
-        "train": y_train.idxmax(axis=1).value_counts(normalize=True),
-        "eval": y_eval.idxmax(axis=1).value_counts(normalize=True),
+        "train": y_train.idxmax(axis=1).value_counts(normalize=False),
+        "eval": y_eval.idxmax(axis=1).value_counts(normalize=False),
+        "train_pct": y_train.idxmax(axis=1).value_counts(normalize=True),
+        "eval_pct": y_eval.idxmax(axis=1).value_counts(normalize=True),
     }
 )
+print(split_pcts_train)
 
 #############################################################
 
@@ -139,6 +146,9 @@ class ImagesDataset(Dataset):
 train_dataset = ImagesDataset(the_basepath, x_train, y_train, device=device, model=model_name)
 train_dataloader = DataLoader(train_dataset, batch_size=batch_size)
 
+eval_dataset = ImagesDataset(the_basepath, x_eval, y_eval, device=device, model=model_name)
+eval_dataloader = DataLoader(eval_dataset, batch_size=batch_size)
+
 #############################################################
 
 if model_name in ["resnet50", "resnet101"]:
@@ -155,13 +165,23 @@ for param in model.parameters():
 
 if model_name in ["resnet50", "resnet101"]:
     model.fc = nn.Sequential(
[…]

### 2023-11-11 hcp/test_job.sh
try again

@@ -6,7 +6,7 @@
 #SBATCH --gpus-per-node=1         # Number of GPU(s) per node
 #SBATCH --cpus-per-task=1         # CPU cores/threads
 #SBATCH --mem=8GB                 # memory per node
-#SBATCH --time=20:00:00
+#SBATCH --time=35:00:00
 #SBATCH --output slurm-%j.log
 
 # https://unix.stackexchange.com/questions/285690/slurm-custom-standard-output-name

### 2023-11-14 scripts/bash/camtrap.sh
addoverwrite checks

@@ -59,10 +59,13 @@ export_default_vars(){
     IOU_THRESHOLD=0.85
     NDIR_LEVEL=0
 
-    OVERWRITE_MD=true
+    OVERWRITE_MD=false
     OVERWRITE_LS=true
     OVERWRITE_MD_CSV=true
     OVERWRITE_EXIF_CSV=true
+    OVERWRITE_VIZ=false
+    OVERWRITE_REPEAT=false
+    OVERWRITE_REMOVE_REPEAT=false
 
     OVERWRITE_COCO_REPEAT=true
     OVERWRITE_LS_REPEAT=true
@@ -199,12 +202,19 @@ run_detect_repeat(){
         OUTPUT_BASE="$(basename $DIR)_repeat"
         echo $OUTPUT_BASE
 
-        python $MD_FOLDER/api/batch_processing/postprocessing/repeat_detection_elimination/find_repeat_detections.py \
-            "$INPUT_DIR/$INPUT_JSON" --imageBase "$RUN_DIR" \
-            --outputBase "$STORAGE_DIR/$OUTPUT_BASE" \
-            --confidenceMin $THRESHOLD_FILTER \
-            --iouThreshold $IOU_THRESHOLD \
-            --nDirLevelsFromLeaf $NDIR_LEVEL
+        if [ -d "$STORAGE_DIR/$OUTPUT_BASE" ] && [ "$OVERWRITE_REPEAT" != true ]; then
+
+            echo "Output folder $OUTPUT_BASE exists, moving to the next folder"
+
+        else
+
+            python $MD_FOLDER/api/batch_processing/postprocessing/repeat_detection_elimination/find_repeat_detections.py \
+                "$INPUT_DIR/$INPUT_JSON" --imageBase "$RUN_DIR" \
+                --outputBase "$STORAGE_DIR/$OUTPUT_BASE" \
+                --confidenceMin $THRESHOLD_FILTER \
+                --iouThreshold $IOU_THRESHOLD \
+                --nDirLevelsFromLeaf $NDIR_LEVEL
+        fi
 
     done
 }
@@ -229,8 +239,16 @@ run_remove_repeat(){
         FILT_DIR=$(ls -td "$STORAGE_DIR/$OUTPUT_BASE"/*/ | head -1)
         echo $FILT_DIR
 
-        python $MD_FOLDER/api/batch_processing/postprocessing/repeat_detection_elimination/remove_repeat_detections.py \
[…]

### 2023-11-14 README.Rmd
update readme
readme
fix pathing

@@
 Prerequisites:
+
   - docker
@@
 # -e DISABLE_AUTH=true --shm-size 50G
-sudo docker run \
+docker run \
   -v "$(pwd):/workspace/rofcamtrap" \
   -v "/media/vlucet/TrailCamST/TrailCamStorage:/workspace/storage/TrailCamStorage" \
-  -v "/media/vlucet/My Passport/Images:/workspace/storage/my_passport_images" \
+  -v "/media/vlucet/TrailCamST/TrailCamStorage_2:/workspace/storage/TrailCamStorage_2" \
   --gpus all \
@@
   -b "/workspace/git" \
-  -s "/workspace/storage/my_passport_images" \ 
+  -s "/workspace/storage/TrailCamStorage_2" \
   -m "/workspace/models/md_v5a.0.0.pt" \
+  -o "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
   md
@@
   -b "/workspace/git" \
-  -s "/workspace/storage/my_passport_images" \
-  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs" \
+  -s "/workspace/storage/TrailCamStorage_2" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
   repeat-detect
@@
   -b "/workspace/git" \
-  -s "/workspace/storage/my_passport_images" \
-  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs" \
-  -o "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats" \
+  -s "/workspace/storage/TrailCamStorage_2" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
+  -o "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
   repeat-remove
@@
   -b "/workspace/git" \
-  -s "/workspace/storage/my_passport_images" \
-  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/" \
-  -o "/workspace/rofcamtrap/1_MegaDetector/2_visualize/" \
+  -s "/workspace/storage/TrailCamStorage_2" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
+  -o "/workspace/rofcamtrap/1_MegaDetector/2_visualize/TrailCamStorage_2" \
   viz
@@
   -b "/workspace/git" \
-  -s "/workspace/storage/my_passport_images" \
-  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/" \
-  -o "/workspace/rofcamtrap/2_LabelStudio/0_inputs/" \
+  -s "/workspace/storage/TrailCamStorage_2" \
+  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
+  -o "/workspace/rofcamtrap/2_LabelStudio/0_inputs/TrailCamStorage_2" \
   repeat-convert
@@
 
-On beluga, we use the apptainer iamge instead.
+On beluga, we use the apptainer image instead.
 
@@
 ### Species classifier
+
+We need to download the models... TBA
 

### 2023-11-14 dockerfiles/build.sh
Add exiftool to dockerfile

@@ -1,6 +1,6 @@
 #!/usr/bin/bash
 
-docker build -t rofcamtrap dockerfiles
+docker build -t rofcamtrap rofcamtrap/dockerfiles
 docker save rofcamtrap -o rofcamtrap.tar
 # docker image rm rofcamtrap
 apptainer build --fakeroot rofcamtrap.sif docker-archive://rofcamtrap.tar

### 2023-11-15 dockerfiles/build.sh
fix build

@@ -1,7 +1,7 @@
 #!/usr/bin/bash
 
 docker build -t rofcamtrap rofcamtrap/dockerfiles
-docker save rofcamtrap -o rofcamtrap.tar
+docker save rofcamtrap -o rofcamtrap/rofcamtrap.tar
 # docker image rm rofcamtrap
-apptainer build --fakeroot rofcamtrap.sif docker-archive://rofcamtrap.tar
+apptainer build --fakeroot rofcamtrap/rofcamtrap.sif docker-archive://rofcamtrap.tar
 # rm rofcamtrap.tar

### 2023-11-15 dockerfiles/build.sh
change commented

@@ -4,4 +4,4 @@ docker build -t rofcamtrap rofcamtrap/dockerfiles
 docker save rofcamtrap -o rofcamtrap/rofcamtrap.tar
 # docker image rm rofcamtrap
 apptainer build --fakeroot rofcamtrap/rofcamtrap.sif docker-archive://rofcamtrap.tar
-# rm rofcamtrap.tar
+# rm rofcamtrap/rofcamtrap.tar

### 2023-11-16 README.Rmd
fix readme

@@
 poetry shell
+cd ../
 rofcamtrap/scripts/bash/camtrap.sh \

### 2023-11-16 scripts/bash/camtrap.sh
update

@@ -59,13 +59,13 @@ export_default_vars(){
     IOU_THRESHOLD=0.85
     NDIR_LEVEL=0
 
-    OVERWRITE_MD=false
+    OVERWRITE_MD=true
     OVERWRITE_LS=true
     OVERWRITE_MD_CSV=true
     OVERWRITE_EXIF_CSV=true
-    OVERWRITE_VIZ=false
-    OVERWRITE_REPEAT=false
-    OVERWRITE_REMOVE_REPEAT=false
+    OVERWRITE_VIZ=true
+    OVERWRITE_REPEAT=true
+    OVERWRITE_REMOVE_REPEAT=true
 
     OVERWRITE_COCO_REPEAT=true
     OVERWRITE_LS_REPEAT=true
@@ -93,7 +93,6 @@ print_vars(){
     echo "      STORAGE_DIR = $STORAGE_DIR"
     echo "      BASE_FOLDER = $BASE_FOLDER"
     echo "      MD_FOLDER = $MD_FOLDER"
-    echo "      MODEL PATH = $MODEL_PATH"
     echo "      CHECKPOINT_FREQ = $CHECKPOINT_FREQ"
     echo "      THRESHOLD_FILTER = $THRESHOLD_FILTER"
     echo "      IOU_THRESHOLD = $IOU_THRESHOLD"

### 2023-11-21 scripts/python/species_classifier/classify_species.py
try new architecture

@@ -165,12 +165,15 @@ for param in model.parameters():
 
 if model_name in ["resnet50", "resnet101"]:
     model.fc = nn.Sequential(
-        # nn.Linear(2048, number_of_categories),  # dense layer takes a 2048-dim input and outputs 100-dim
+        nn.Linear(2048, 1000),  # dense layer takes a 2048-dim input and outputs 100-dim
         nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
         nn.Dropout(0.5),  # common technique to mitigate overfitting
+        nn.Linear(2048, 100),
+        nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
+        nn.Dropout(0.5),
         nn.Linear(
-            2048, number_of_categories
-        ),  # final dense layer outputs 8-dim corresponding to our target classes
+            100, number_of_categories
+        ),  # final dense layer outputs x-dim corresponding to our target classes
     )
     
 # if model_name in ["resnet50", "resnet101"]:

### 2023-11-21 scripts/python/species_classifier/classify_species.py
fix bug

@@ -168,7 +168,7 @@ if model_name in ["resnet50", "resnet101"]:
         nn.Linear(2048, 1000),  # dense layer takes a 2048-dim input and outputs 100-dim
         nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
         nn.Dropout(0.5),  # common technique to mitigate overfitting
-        nn.Linear(2048, 100),
+        nn.Linear(1000, 100),
         nn.ReLU(inplace=True),  # ReLU activation introduces non-linearity
         nn.Dropout(0.5),
         nn.Linear(

### 2023-11-22 README.Rmd
update readme

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/main?urlpath=rstudio)
+<!-- [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/main?urlpath=rstudio) -->
 
@@
 
+9. Crop annotations.
+
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -b "/workspace/git" \
+  -s "/workspace/storage/TrailCamStorage_2" \
+  -i "/workspace/rofcamtrap/2_LabelStudio/0_inputs/TrailCamStorage_2" \
+  crop
+```
+
+## Label studio outputs
+
+1. Enter the container running the label studio app and output ths number of projects
+
+```{bash eval=FALSE}
+docker exec -it label-studio-app-1 bash
+curl -X GET http://localhost:8080/api/projects/ -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d' -o files/outputs/project_counts.json
+```
+
+2. On the instance outside the container, run the Python script to extract projects id
+
+```{python eval=FALSE, python.reticulate = FALSE}
+import json
+
+with open('data/outputs/project_counts.json', 'r') as file:
+  data = json.load(file)
+
+ids = [data['results'][i]['id'] for i in range(len(data['results']))]
+
+with open('data/outputs/project_ids.txt', 'w') as file:
+  for the_id in ids:
+        file.write(str(the_id) + '\n')
+```
+
+3. Back in the container
+
+```{bash eval=FALSE}
+arr=(170 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10); 
+for id in ${arr[@]}; do         
+  # url="http://localhost:8080/api/projects/${id}/export?exportType=JSON&download_all_tasks=true";
+  url="http://localhost:8080/api/projects/${id}/export?exportType=JSON";
+  curl -X GET "$url" -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d'\
+    -o "files/outputs/ls/output_file_$id.json";
+done
+```
+
+4. Check for file numbers
+
+```{bash eval=FALSE}
+ls data/outputs/ls # 59 nov 21
+```
+
+5. Copy to local machine
+
+```{bash eval=FALSE}
+scp -i ssh_key/arbutus_def_fstewart_prod 'ubuntu@206.12.94.17:~/data/outputs/ls/*' rofcamtrap/2_LabelStudio/1_outputs_downloaded/
+```
+
+6. Process those outputs
+
+```{bash eval=FALSE}
+rofcamtrap/scripts/bash/camtrap.sh \
+  -i "/workspace/rofcamtrap/2_LabelStudio/1_outputs_downloaded/" \
+  -o "/workspace/rofcamtrap/2_LabelStudio/2_outputs_processed" \
+  post
+```
+
 ## Classifier training workflow
@@
 
-```
+```{bash eval=FALSE}
 for FILE in project*

### 2023-11-22 scripts/R/postprocess/1_preprocess_annotations.R
progress

@@ -1,14 +1,8 @@
 # Note: This runs on the csvs created from the label studio json outputs by
 # the post process function in the python utility mdtools
 
-# for FILE in *
-#   mdtools postprocess --write-csv $FILE
-# end
-
-library(magrittr)
-
 # Folder with label studio annotation outputs
-label_studio_folder <- "analysis/data/raw_data/label_studio_outputs/"
+label_studio_folder <- "2_LabelStudio/2_outputs_processed/"
 
 # Annotations -------------------------------------------------------------
 
@@ -33,7 +27,7 @@ read_ls_file <- function(the_file){
 }
 
 widen <- function(df) {
-  df %>%
+  df |>
     dplyr::select(-type) |>
     dplyr::mutate(manual = stringr::str_detect(id, "_M")) |>
     dplyr::mutate(image_id = stringr::str_split(id, "_")) |>
@@ -42,8 +36,8 @@ widen <- function(df) {
                                       ~paste0(.x[1:5], collapse = "_")))) |>
     tidyr::pivot_wider(names_from = "variable", values_from = "tag") |>
     dplyr::filter(is.na(value_text)) |>
-    dplyr::mutate (species = ifelse(species == 'Other [See more species]',
-                                    other_species, species)) # |>
+    dplyr::mutate(species = ifelse(species == 'Other [See more species]',
+                                   other_species, species)) # |>
     # dplyr::select(-other_species)
 }
 
@@ -59,15 +53,15 @@ annotations_dup <- data.frame()
 # process dups specifically
 projects <- unlist(lapply(stringr::str_split(basename(file_list), "_"), function(x)x[1]))
 names(file_list) <- projects
-dups_names <- names(which(table(projects)>1))
-non_dups_names <- names(which(!(table(projects)>1)))
+dups_names <- names(which(table(projects) > 1))
+non_dups_names <- names(which(!(table(projects) > 1)))
 
 file_list_no_dups <- file_list[names(file_list) %in% non_dups_names]
 file_list_dups <- file_list[names(file_list) %in% dups_names]
 
 # -------------------------------------------------------------------------
 
-for (file in file_list_no_dups){
+for (file in file_list_no_dups) {
 
   # print(file)
 
[…]

### 2023-11-22 scripts/R/postprocess/2_preprocess_detections.R

@@ -1,19 +1,22 @@
 # Note: This takes the COCO formatted detections derived from the python tool
 # mdtools, and outputs a table
 
+library("purrr")
+
 # Folder with Megadetector detections in coco format
-detections_folder <- "/media/vlucet/TrailCamST/TrailCamStorage/"
+detections_folders <- c("2_LabelStudio/0_inputs/TrailCamStorage",
+                        "2_LabelStudio/0_inputs/TrailCamStorage_2")
 
 # List all the files with the proper file name format
-file_list <- list.files(detections_folder, recursive = F,
-                        full.names = T, pattern = "output_coco.json")
+file_list <- map(detections_folders, list.files, recursive = F,
+                 full.names = T, pattern = "output_coco.json") |> list_c()
 
 # Detections --------------------------------------------------------------
 
 # Create an empty dataframe to be grown
 detections <- data.frame()
 
-for (file in file_list){
+for (file in file_list) {
 
   # print(file)
 
@@ -23,4 +26,4 @@ for (file in file_list){
 
 }
 
-saveRDS(detections, "analysis/data/derived_data/detections.rds")
+saveRDS(detections, "data/detections.rds")

### 2023-11-22 scripts/bash/camtrap.sh

@@ -117,6 +117,9 @@ crawl_dirs(){
         if [[ "$FILE" == *"repeat"* ]];then
             continue
         fi
+        if [[ "$FILE" == *"crop"* ]];then
+            continue
+        fi
         [[ -d "$FILE" ]] && DIRS+=("$FILE")
     done
 
@@ -280,25 +283,11 @@ run_convert_repeat(){
             mdtools convert ls "$INPUT_DIR/$OUTPUT_JSON_REPEAT" "$RUN_DIR" "$OUTPUT_DIR" \
                 -ct $THRESHOLD_FILTER \
                 -ru "data/local-files/?d=$(basename $STORAGE_DIR)/$(basename $DIR)" \
-                --write-ls --repeat --write-csv --write-coco
+                --write-ls --write-csv --write-coco --repeat
         fi
 
     done
 
-    # for DIR in "${DIRS[@]}"; do
-    #     echo "*** RUNNING REPEAT CONVERTER TO COCO ***"
-    #     RUN_DIR=$STORAGE_DIR/$(basename $DIR)
-    #     echo "Running on directory: $RUN_DIR"
-    #     OUTPUT_JSON_REPEAT="$(basename $DIR)_output_norepeats.json"
-    #     echo $OUTPUT_JSON_REPEAT
-    #     OUTPUT_COCO_REPEAT="$(basename $DIR)_output_coco_norepeats.json"
-    #     echo $OUTPUT_COCO_REPEAT
-    #     if [ -f "$STORAGE_DIR/$OUTPUT_COCO_REPEAT" ] && [ "$OVERWRITE_COCO_REPEAT" != true ]; then # if output exist, do nothing
-    #         echo "Output file $OUTPUT_COCO_REPEAT exists, moving to the next folder"
-    #     else
-    #         mdtools convert cct $STORAGE_DIR/$OUTPUT_JSON_REPEAT $RUN_DIR --write-coco --repeat
-    #     fi
-    # done
 }
 
 run_convert(){
@@ -316,15 +305,16 @@ run_convert(){
         OUTPUT_JSON_LS="$(basename $DIR)_output_ls.json"
         echo $OUTPUT_JSON_LS
 
-        if [ -f "$STORAGE_DIR/$OUTPUT_JSON_LS" ] && [ "$OVERWRITE_LS" != true ]; then # if output exist, do nothing
[…]

### 2023-11-22 scripts/R/data_management/folder-structure-2nd-deployment.R
clean up a bit

@@ -1,0 +1,32 @@
+library(readxl)
+library(magrittr)
+library(dplyr)
+library(tidyr)
+
+deps <- read_excel("data-raw/RoF_CameraDeploymentLog.xlsx")
+
+folders <- list.files("/media/vlucet/My Passport/Images/")
+folders
+
+deps_filt <- deps %>%
+  filter(`Camera ID` %in% folders)
+stopifnot(nrow(deps_filt) == length(folders))
+
+for (plot_id in unique(deps_filt$PlotID)) {
+
+  print(plot_id)
+
+  dir.create(paste0("/media/vlucet/My Passport/Images/", plot_id))
+
+  plot_dat <- deps_filt %>%
+    filter(PlotID == plot_id)
+
+  for (site_id in unique(plot_dat$SiteID)) {
+
+    dir.create(paste0("/media/vlucet/My Passport/Images/", plot_id, "/", site_id))
+
+  }
+
+}
+
+deps_filt %>% arrange(`Camera ID`, PlotID, SiteID) %>% View()

### 2023-11-22 scripts/R/rof_survey.R

@@ -1,0 +1,69 @@
+## code to prepare `rof_survey` dataset
+
+# Load packages
+library(readxl)
+library(janitor)
+library(dplyr)
+library(tidyr)
+library(lubridate)
+library(stringr)
+
+# Load both sheets
+rof_survey_summary_raw <-
+  readxl::read_xlsx("data-raw/RoF_CameraDeploymentLog_WLU.xlsx",
+                    sheet = "Summary", col_names = FALSE)
+rof_survey_raw <-
+  readxl::read_xlsx("data-raw/RoF_CameraDeploymentLog_WLU.xlsx",
+                    sheet = "Sheet1")
+
+# Clean up summary
+names(rof_survey_summary_raw) <- c("key", "value")
+rof_survey_summary_clean <- rof_survey_summary_raw %>%
+  dplyr::slice(-3)
+
+
+# Clean up survey
+
+# Needed function
+split_fun <- function(x, i) {
+  if (length(x) <= 1){
+    return(NA)
+  } else {
+    return(x[i])
+  }
+}
+
+assembly_fun <- function(x) {
+  if (length(x) <= 1){
+    return(NA)
+  } else {
+    return(file.path(x[3], x[4], x[5]))
+  }
+}
+
+rof_survey_clean <-
+  rof_survey_raw %>%
+  # Clean the column names
+  janitor::clean_names() %>%
+  # Deal with time columns
+  dplyr::rename(date_deployed_full = date_deployed,
+                date_retrieved_full = date_retrieved) %>%
+  dplyr::mutate(date_deployed_full = lubridate::ymd_hms(date_deployed_full),
+                date_deployed = lubridate::date(date_deployed_full),
+                date_retrieved_full = lubridate::ymd(date_retrieved_full),
+                date_retrieved = lubridate::date(date_retrieved_full),
+                deploy_interval = date_retrieved - date_deployed) %>%
+  # Deal with paths
+  dplyr::rename(photo_storage_network_location_SF =
+                  photo_storage_network_location) %>%
+  dplyr::mutate(photo_storage_network_location_split =
+                  stringr::str_split(photo_storage_network_location_SF,
+                                     pattern = "\\\\"),
+                photo_storage_network_location_root = sapply(photo_storage_network_location_split,
+                                                             FUN = split_fun, 2),
+                photo_storage_network_location_path = sapply(photo_storage_network_location_split,
+                                                             FUN = assembly_fun))
+
[…]

### 2024-01-11 scripts/bash/camtrap.sh
add exif

@@ -401,6 +401,36 @@ run_post(){
 
 # ------------------------------------------------------------------
 
+run_exif(){
+
+    for DIR in "${DIRS[@]}"; do
+
+        echo "*** RUNNING EXIF  ***"
+
+        RUN_DIR=$STORAGE_DIR/$(basename $DIR)
+        echo "Running on directory: $RUN_DIR"
+
+        OUTPUT_JSON="$(basename $DIR)_output.json"
+        echo $OUTPUT_JSON
+
+        OUTPUT_EXIF="$(basename $DIR)_exif.csv"
+        echo $OUTPUT_JSON_LS
+
+        if [ -f "$OUTPUT_DIR/$OUTPUT_EXIF" ] && [ "$OVERWRITE_EXIF_CSV" != true ]; then # if output exist, do nothing
+
+            echo "Output file $OUTPUT_EXIF exists, moving to the next folder"
+
+        else
+
+            mdtools readexif "$INPUT_DIR/$OUTPUT_JSON" "$STORAGE_DIR" "$OUTPUT_DIR" \
+                --write-csv
+        fi
+
+    done
+}
+
+# ------------------------------------------------------------------
+
 while getopts ":s:b:m:i:o:h:" opt; do
   case ${opt} in
     s )
@@ -506,7 +536,16 @@ case "$subcommand" in
 
      post)
         echo "Running postprocessing step"
+        run_prep
         run_post
+
+        shift $((OPTIND -1))
+        ;;
+
+     exif)
+        echo "Running exif step"
+        run_prep
+        run_exif
 
         shift $((OPTIND -1))
         ;;

### 2024-01-11 scripts/bash/camtrap.sh
updates

@@ -263,7 +263,7 @@ run_convert_repeat(){
 
     for DIR in "${DIRS[@]}"; do # "P072"; do # @ 0
 
-        echo "*** RUNNING REPEAT CONVERTER TO LS ***"
+        echo "*** RUNNING REPEAT CONVERTER ***"
 
         RUN_DIR=$STORAGE_DIR/$(basename $DIR)
         echo "Running on directory: $RUN_DIR"
@@ -294,7 +294,7 @@ run_convert(){
 
     for DIR in "${DIRS[@]}"; do
 
-        echo "*** RUNNING CONVERTER TO LS ***"
+        echo "*** RUNNING CONVERTER ***"
 
         RUN_DIR=$STORAGE_DIR/$(basename $DIR)
         echo "Running on directory: $RUN_DIR"
@@ -311,10 +311,10 @@ run_convert(){
 
         else
 
-            mdtools convert cct "$INPUT_DIR/$OUTPUT_JSON" "$RUN_DIR" "$OUTPUT_DIR" \
+            mdtools convert ls "$INPUT_DIR/$OUTPUT_JSON" "$RUN_DIR" "$OUTPUT_DIR" \
                 -ct $THRESHOLD_FILTER \
                 -ru "data/local-files/?d=$(basename $STORAGE_DIR)/$(basename $DIR)" \
-                --write-csv --write-coco
+                --write-ls --write-csv --write-coco
         fi
 
     done

### 2024-01-11 README.Rmd
fix readme

@@
 
-8. Convert to coco and ls [and 3rd format?].
+8. Switch environments
+
+```{bash eval=FALSE}
+mamba deactivate 
+cd mdtools
+poetry shell
+cd ../
+```
+
+9. Convert to coco and ls [and 3rd format?].
 
@@
 
-9. Crop annotations.
+10. Crop annotations.
 
@@
 docker exec -it label-studio-app-1 bash
-curl -X GET http://localhost:8080/api/projects/ -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d' -o files/outputs/project_counts.json
+curl -X GET http://localhost:8080/api/projects/?page_size=1000 -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d' -o files/outputs/project_counts.json
 ```
@@
 ```{bash eval=FALSE}
-arr=(170 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10); 
+arr=(192 191 190 189 188 187 186 185 184 183 182 181 180 179 178 177 176 175 174 173 172 171 170 87 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10);
 for id in ${arr[@]}; do         
@@
 
-4. Check for file numbers
+4. Back outside the container, check for file numbers
 
 ```{bash eval=FALSE}
-ls data/outputs/ls # 59 nov 21
+ls data/outputs/ls # 81 Jan 11 2024
 ```

### 2024-01-12 README.Rmd
fix readme list

@@
 ```{bash eval=FALSE}
-arr=(192 191 190 189 188 187 186 185 184 183 182 181 180 179 178 177 176 175 174 173 172 171 170 87 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10);
+arr=(192 191 190 189 188 187 186 185 184 183 182 181 180 179 178 177 176 175 174 173 172 171 170 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10);
 for id in ${arr[@]}; do         

### 2024-01-12 scripts/R/postprocess/1_preprocess_annotations.R
fix pathing

@@ -203,4 +203,3 @@ annotations_wide <- widen(annotations_no_dup) |>
   dplyr::bind_rows(all_anns) |>
   dplyr::mutate(species = ifelse(species == "None", NA, species))
 
-saveRDS(annotations_wide, "data/annotations_wide_treated.rds")

### 2024-01-18 scripts/R/postprocess/1_preprocess_annotations.R
update annotations processing

@@ -51,10 +51,17 @@ annotations_dup <- data.frame()
 
 # separate into dups and not dups. process not dups normally and
 # process dups specifically
-projects <- unlist(lapply(stringr::str_split(basename(file_list), "_"), function(x)x[1]))
+base_list <- basename(file_list)
+base_list <- stringr::str_replace_all(base_list, "TrailCamStorage_2", "T2")
+base_list <- stringr::str_replace_all(base_list, "TrailCamStorage", "T1")
+
+projects <- unlist(lapply(stringr::str_split(base_list, "_"),
+                          function(x) paste0(x[1:2], collapse = "_")))
 names(file_list) <- projects
 dups_names <- names(which(table(projects) > 1))
 non_dups_names <- names(which(!(table(projects) > 1)))
+
+dups_names
 
 file_list_no_dups <- file_list[names(file_list) %in% non_dups_names]
 file_list_dups <- file_list[names(file_list) %in% dups_names]
@@ -203,3 +210,4 @@ annotations_wide <- widen(annotations_no_dup) |>
   dplyr::bind_rows(all_anns) |>
   dplyr::mutate(species = ifelse(species == "None", NA, species))
 
+saveRDS(annotations_wide, "data/objects/annotations_wide_treated.rds")

### 2024-01-18 reports/GnC_report/paper/paper.qmd
fix up the qmd, but I really need to get back into it using the code Eric is using

--- 2023-04-20 analysis/paper/paper.qmd
@@
 options(scipen = 1, digits = 2)
-library(magrittr)
 ```
@@
 RoF_CameraDeploymentLog_WLU <- 
-  readxl::read_excel(paste0("analysis/data/raw_data/camera_deployments/",
-                            "WLU_Cameras_Log_2023.xlsx")) %>% 
+  readxl::read_excel(paste0("data-raw/tabular/RoF_CameraDeploymentLog.xlsx")) |> 
   janitor::clean_names()
 
-cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
-  dplyr::filter(!is.na(latitude)) %>% 
-  dplyr::filter(!is.na(longitude)) %>% 
-  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326", remove=F) %>% 
+cams_sf <- RoF_CameraDeploymentLog_WLU |> 
+  dplyr::rename(longitude = x, latitude = y) |> 
+  dplyr::filter(!is.na(latitude)) |> 
+  dplyr::filter(!is.na(longitude)) |> 
+  sf::st_as_sf(coords = c("longitude","latitude"), crs = "EPSG:4326", remove = F) |> 
   dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))
 
-coords <- cams_sf %>% 
-  sf::st_coordinates() %>% 
+coords <- cams_sf |> 
+  sf::st_coordinates() |> 
   as.data.frame()
@@
 #   sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
-#               quiet=TRUE, ) %>% 
+#               quiet=TRUE, ) |> 
 #   sf::st_transform(crs = sf::st_crs(cams_sf))
@@
 zones <- 
-  sf::st_read("analysis/data/raw_data/ecozones/EcoZone/EcoZone.shp",
-              quiet=TRUE, ) %>% 
+  sf::st_read("data-raw/spatial/ecozones/EcoZone/EcoZone.shp",
+              quiet = TRUE, ) |> 
   sf::st_transform(crs = sf::st_crs(cams_sf))
 
-zones_filter <- zones %>%
+zones_filter <- zones |>
   sf::st_filter(cams_sf)
 
-pols <- sf::st_coordinates(zones_filter) %>% 
+pols <- sf::st_coordinates(zones_filter) |> 
   as.data.frame()
 
-# regions_labels <- zones_filter %>% 
-#   sf::st_drop_geometry() %>% 
+# regions_labels <- zones_filter |> 
+#   sf::st_drop_geometry() |> 
 # dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
@@
 
-regions_labels <- zones_filter %>% 
-  sf::st_drop_geometry() %>% 
+regions_labels <- zones_filter |> 
+  sf::st_drop_geometry() |> 
   dplyr::bind_cols(data.frame(x = c(-84.5, -88.5),
@@
 
-map <- ggmap::get_stamenmap( bbox = c(left = -90, bottom = 50, right = -83, top = 56), 
-                             zoom = 9, maptype = "terrain-background")
+map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
+                            zoom = 9, maptype = "stamen_terrain")
 ```
@@
   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
-  ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
+  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)
 
@@
   #                     size = 2.5, data = regions_labels) +
-  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
+  ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA), inherit.aes = FALSE,
                       size = 2.5, data = regions_labels) +
-  ggplot2::geom_point(data = cams_sf_coords, cex = 1, shape=21, col="black",
-                      ggplot2::aes(x=X, y=Y, fill=is_retrieved), 
-                      inherit.aes = FALSE) +
-  ggplot2::labs(fill = "Camera \nRetrieved")
+  ggplot2::geom_point(data = cams_sf_coords, #cex = 1, 
+                      shape = 24, col = "black", fill = "red",
+                      ggplot2::aes(x = X, y = Y, #fill=is_retrieved
+                                   ), 
+                      inherit.aes = FALSE) # +
+  # ggplot2::labs(fill = "Camera \nRetrieved")
 p
+ggplot2::ggsave(plot = p, filename = "figures/camera_array.png", 
+                width = 6, height = 8, dpi = 300)
+ggplot2::ggsave(plot = p, filename = "reports/presentations/rofcamtrap-nov2023/images/camera_array.png", 
+                width = 6, height = 8, dpi = 300)
 ```
@@
 #                       size = 2.5, data = regions_labels) +
-#   ggplot2::geom_point(data = cams_sf_coords %>% 
+#   ggplot2::geom_point(data = cams_sf_coords |> 
 #                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
@@
 
-if (!file.exists("analysis/data/derived_data/annotations.rds") |
-    !file.exists("analysis/data/derived_data/annotations_wide.rds")){
-  source("scripts/1_preprocess_annotations.R")
-} else {
-  anns <- readRDS("analysis/data/derived_data/annotations.rds")
-  anns_wide <- readRDS("analysis/data/derived_data/annotations_wide.rds")
-}
-
-if (!file.exists("analysis/data/derived_data/detections.rds")){
-  source("scripts/2_preprocess_detections.R")
-} else {
-  dets <- readRDS("analysis/data/derived_data/detections.rds")
-}
+anns_wide <- readRDS("data/objects/annotations_wide_treated.rds")
+dets <- readRDS("data/objects/detections.rds")
 
@@
 # Hardcoded check
-stopifnot(nb_image == 2788375)
-
-exif <- readr::read_csv("analysis/data/raw_data/exif/exif_combined.csv") %>% 
-  dplyr::select(-SourceFile) %>% 
-  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) %>% 
-  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) %>% 
-  dplyr::mutate(image_id = stringr::str_sub(image_id, 6)) %>% 
-  janitor::clean_names()
+stopifnot(nb_image == 3288178) # 2788375
+
+exif <- readr::read_csv("data/tabular/all_exif.csv") |>
+  janitor::clean_names() |>
+  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
+  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |>
+
+    dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
+  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |>
+
+  dplyr::mutate(source_file = stringr::str_remove(source_file, 
+                                               "/workspace/storage/TrailCamStorage_2/")) |>
+  dplyr::mutate(source_file = stringr::str_remove(source_file, 
+                                               "/workspace/storage/TrailCamStorage/")) |>
+
+  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
+  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |>
+
+  dplyr::mutate(image_id = stringr::str_remove(image_id, 
+                                               "_workspace_storage_TrailCamStorage_2_")) |>
+  dplyr::mutate(image_id = stringr::str_remove(image_id, 
+                                               "_workspace_storage_TrailCamStorage_")) |>
+
+  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
+                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
+                  unlist()) |>
+
+  dplyr::mutate(
+    source_file = 
+      stringr::str_remove(source_file, 
+                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
+  dplyr::mutate(
+    source_file = 
+      stringr::str_remove(source_file, 
+                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
+  # dplyr::mutate(image_id = stringr::str_sub(image_id, category6))
 ```
@@
 # Select and distinct down to image-based infos
-dets_images <- dets %>%  dplyr::select(image_id, max_confidence, isempty) %>% 
+dets_images <- dets |>  dplyr::select(image_id, max_confidence, isempty) |> 
   dplyr::distinct()
@@
 # Breakdown of which images triggered a MD detection
-dets_tb <- dets_images$isempty %>% table()
+dets_tb <- dets_images$isempty |> table()
 dets_tb_sum <- sum(dets_tb)
@@
 #| echo: false
-dets_0 <- dets %>% dplyr::filter(confidence > 0)
-dets_0.1 <- dets %>% dplyr::filter(confidence > 0.1)
+dets_0 <- dets |> dplyr::filter(confidence > 0)
+dets_0.1 <- dets |> dplyr::filter(confidence > 0.1)
 
 # Breakdown of detections by categories (animal, person, vehicles)
-dets_tb_cat_unfil <- dets_0 %>% dplyr::pull(category_id) %>% table()
+dets_tb_cat_unfil <- dets_0 |> dplyr::pull(category_id) |> table()
 dets_tb_cat_unfil_sum <- sum(dets_tb_cat_unfil)
@@
 # Breakdown of detections above 0.1 by categories (animal, persons, vehicles)
-dets_tb_cat <- dets_0.1 %>% dplyr::pull(category_id) %>% table()
+dets_tb_cat <- dets_0.1 |> dplyr::pull(category_id) |> table()
 dets_tb_cat_sum <- sum(dets_tb_cat)
@@
 # Histogram of confidence scores above 0.1
-dets_0.1 %>% 
-  dplyr::mutate(category_id = as.character(category_id)) %>% 
+dets_0.1 |> 
+  dplyr::mutate(category_id = as.character(category_id)) |> 
   dplyr::mutate(Type = dplyr::case_match(category_id,
@@
                                          "2" ~ "Person",
-                                         "3" ~ "Vehicle")) %>% 
+                                         "3" ~ "Vehicle")) |> 
   ggplot2::ggplot() +
   ggplot2::theme_bw() +
-  ggplot2::geom_histogram(ggplot2::aes(x=confidence, group=Type, fill=Type),
+  ggplot2::geom_histogram(ggplot2::aes(x = confidence, group = Type, fill = Type),
                           bins = 100) +
@@
   ggplot2::ylab("Count") +
-  ggplot2::geom_vline(ggplot2::aes(xintercept = 0.1), lty=2, col = "firebrick")
+  ggplot2::geom_vline(ggplot2::aes(xintercept = 0.1), lty = 2, col = "firebrick")
 ```
@@
 # Number of uploaded images
-uploaded <- dets %>% dplyr::filter(confidence > 0.1) %>% dplyr::pull(image_id) %>% 
-  unique() %>% length()
+uploaded <- dets |> dplyr::filter(confidence > 0.1) |> dplyr::pull(image_id) |> 
+  unique() |> length()
 
@@
 # Filter for non empty images
-anns_wide_noempty <- anns_wide %>%
+anns_wide_noempty <- anns_wide |>
   dplyr::filter(!is.na(to_name))
 
-saveRDS(anns_wide_noempty, "analysis/data/derived_data/annotations_wide_noempty.rds")
+saveRDS(anns_wide_noempty, "data/objects/annotations_wide_noempty.rds")
 
@@
 # Deleted predictions
-dels <- anns_wide_noempty %>% dplyr::filter(is.na(species))
+dels <- anns_wide_noempty |> dplyr::filter(is.na(species))
 dels_nb <- nrow(dels)
@@
 # Should be no manuals here
-stopifnot((anns_wide_noempty %>% dplyr::filter(is.na(species)) 
-           %>% dplyr::filter(manual)) == 0)
+stopifnot((anns_wide_noempty |> dplyr::filter(is.na(species)) 
+           |> dplyr::filter(manual)) == 0)
 
 # Actual annotations
-true_anns <- anns_wide_noempty %>% dplyr::filter(!is.na(species))
+true_anns <- anns_wide_noempty |> dplyr::filter(!is.na(species))
 true_anns_nb <- nrow(true_anns)
@@
 # Non manuals 
-non_mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
+non_mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
   dplyr::filter(!manual)
@@
 # Mans
-mans <- anns_wide_noempty %>% dplyr::filter(!is.na(species)) %>% 
+mans <- anns_wide_noempty |> dplyr::filter(!is.na(species)) |> 
   dplyr::filter(manual)
@@
 # adjusted
-orig_tb <- anns_wide_noempty$origin %>% table()
+orig_tb <- anns_wide_noempty$origin |> table()
 orig_tb_pct <- orig_tb/sum(orig_tb)*100
@@
 # Only annotations for animals
-anns_wide_animals <- anns_wide_noempty %>% 
-  dplyr::filter(species != 'Human') %>% 
-  dplyr::filter(species != 'Vehicle') %>% 
-  dplyr::mutate(camera_id = id %>% stringr::str_split("_") %>% 
-                  purrr::map(~.x[[2]]) %>% unlist())
+# TODO !!!! investigate weird id of cut length 1
+anns_wide_animals <- anns_wide_noempty |>
+  dplyr::filter(species != 'Human') |> 
+  dplyr::filter(species != 'Vehicle') |> 
+  dplyr::mutate(id_cut_length = purrr::map_dbl(id , \(x) {
+    x |> stringr::str_split("_") |> purrr::map_dbl(~length(.x))
+  })) |> 
+  dplyr::filter(id_cut_length > 1) |> 
+  dplyr::mutate(camera_id = purrr::map_chr(id , \(x) {
+    x |>
+      stringr::str_split("_") |>
+      purrr::map(~.x[[2]]) |>
+      unlist()
+  }))
 
 # Unknowns and bird breakdown
-# anns_wide_animals %>% dplyr::filter(species == "Unknown") %>% 
-#   dplyr::pull(id) %>% 
-#   stringr::str_sub(1,4) %>% table() %>% sort() 
-# anns_wide_animals %>% dplyr::filter(stringr::str_detect(species,"Bird")) %>% 
-#   dplyr::pull(id) %>% 
-#   stringr::str_sub(1,4) %>% table() %>% sort()
+# anns_wide_animals |> dplyr::filter(species == "Unknown") |> 
+#   dplyr::pull(id) |> 
+#   stringr::str_sub(1,4) |> table() |> sort() 
+# anns_wide_animals |> dplyr::filter(stringr::str_detect(species,"Bird")) |> 
+#   dplyr::pull(id) |> 
+#   stringr::str_sub(1,4) |> table() |> sort()
 
@@
 
-anns_wide_animals_joined <- anns_wide_animals %>% 
+anns_wide_animals_joined <- anns_wide_animals |>
   dplyr::left_join(exif, by = c("source_file", "image_id"))
 
-anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
-  dplyr::filter(species != "Unknown") %>%
-  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
+anns_wide_animals_joined_mut <- anns_wide_animals_joined |> 
+  dplyr::filter(species != "Unknown") |>
+  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) |> 
   dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
@@
 
-# anns_wide_animals_joined_dets <- anns_wide_animals %>% 
+# anns_wide_animals_joined_dets <- anns_wide_animals |> 
 #   dplyr::right_join(dets_images, by = "image_id")
@@
 
+<!-- Other camera plots -->
+
+```{r eval=FALSE}
+# library(forcats)
+library(ggplot2)
+# anns_wide_animals_joined_mut |> 
+#   ggplot(aes(x=date_time, y=species)) + geom_point()
+# anns_wide_animals_joined_mut |> 
+#   dplyr::filter(species == "Fisher") |> View()
+
+camera_days <- exif |> dplyr::group_by(maker_notes_user_label) |> 
+  dplyr::summarise(camera_first_time=min(exif_date_time_original)) |> 
+  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
+                                               tz = "Canada/Eastern"),
+                year = lubridate::year(date_time),
+                month = lubridate::month(date_time, label = TRUE),
+                day = lubridate::day(date_time),
+                ymd = paste(year, month, day, sep = "_")) |> 
+  dplyr::select(camera_id = maker_notes_user_label, 
+                camera_first_day = date_time)
+
+camera_regions <- cams_sf |> sf::st_join(zones) |> 
+  sf::st_drop_geometry() |> dplyr::select(camera_id, ECOZONE_NA)
+
+bp <- anns_wide_animals_joined_mut |>
+
+  dplyr::filter(species != "Fisher") |> 
+  dplyr::left_join(camera_days, by = "camera_id") |>
+
+  dplyr::select(species, camera_id, 
+                date_time, camera_first_day, ymd) |>
+
+  dplyr::group_by(species, camera_id) |> 
+  dplyr::summarise(first_day = min(date_time),
+                   camera_first_day = min(camera_first_day),
+                   camera_events = length(unique(ymd))) |> 
+  dplyr::ungroup() |>
+
+  dplyr::left_join(camera_regions, by = "camera_id") |>
+
+  dplyr::group_by(species) |> 
+  dplyr::mutate(total_events = sum(camera_events),
+                n_cameras = dplyr::n()) |> 
+  dplyr::ungroup() |>
+
+  # dplyr::filter(n_cameras > 4) |>
+  dplyr::filter(n_cameras > 3) |>
+
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA)) |>
+
+  ggplot(aes(x = diff_time, 
+             y = forcats::fct_reorder(species, total_events),
+             color = ECOZONE_NA)) +
+  geom_boxplot() +
+
+  scale_x_continuous()+
+  labs(x = "Detection time (in days)",
+       y = "Species",
+       color = "Ecozone") + 
+  theme_bw()
+
+ggsave(plot = bp, filename = "figures/bp.png", width = 10, height = 15)
+
+pts <- anns_wide_animals_joined_mut |>
+
+  dplyr::filter(species != "Fisher") |> 
+  dplyr::left_join(camera_days, by = "camera_id") |>
+
+  dplyr::select(species, camera_id, 
+                date_time, camera_first_day, ymd) |>
+
+  dplyr::group_by(species, camera_id) |> 
+  dplyr::summarise(first_day = min(date_time),
+                   camera_first_day = min(camera_first_day),
+                   camera_events = length(unique(ymd))) |> 
+  dplyr::ungroup() |>
+
+  dplyr::left_join(camera_regions, by = "camera_id") |>
+
+  dplyr::group_by(species) |> 
+  dplyr::mutate(total_events = sum(camera_events),
+                n_cameras = dplyr::n()) |> 
+  dplyr::ungroup() |>
+
+  # dplyr::filter(n_cameras > 4) |>
+  dplyr::filter(n_cameras > 3) |>
+
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA)) |>
+
+  ggplot(aes(x = diff_time, 
+             y = forcats::fct_reorder(species, total_events),
+             # size = camera_events, 
+             color = ECOZONE_NA)) +
+
+  geom_point(aes(x = diff_time, 
+                  y = forcats::fct_reorder(species, total_events),
+                  size = camera_events, 
+                  color = ECOZONE_NA,
+                  group = ECOZONE_NA)) +
+
+  scale_x_continuous()+
+  labs(x = "Detection time (in days)",
+       y = "Species",
+       color = "Ecozone",
+       size = "Day counts") + 
+  theme_bw()
+ggsave(plot = pts, filename = "figures/pts.png", width = 10, height = 15)
+pts
+```
+
+```{r eval=FALSE}
+# for_vic <- anns_wide_animals_joined_mut |> 
+#   dplyr::select(id,species,date_time, camera_id) |> 
+#   dplyr::left_join(cams_sf |> dplyr::select(camera_id, plot_id) |> sf::st_drop_geometry(),
+#                    by = "camera_id")
+# tot_camera <- for_vic$camera_id |> unique() |> length()
+# hit_rates <- for_vic |> dplyr::select(species, camera_id) |> 
+#   dplyr::group_by(species) |> dplyr::summarise(n_cameras=length(unique(camera_id))) |> 
+#   dplyr::ungroup() |> dplyr::mutate(rate = n_cameras/tot_camera*100) |> 
+#   dplyr::arrange(dplyr::desc(rate))
+# # det_history_two_weeks <- 
+# 
+# readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
+# readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
+```
+
 ```{r}
@@
 #| echo: false
-anns_wide_animals_table <- anns_wide_animals %>% 
-  dplyr::select(camera_id, species) %>% 
-  dplyr::filter(species != "Unknown") %>%
-  table() %>% 
-  `>`(0) %>% 
-  `+`(0) %>% 
-  as.data.frame() %>% 
-  tibble::rownames_to_column("camera_id") %>% 
+anns_wide_animals_table <- anns_wide_animals |>
+  dplyr::select(camera_id, species) |> 
+  dplyr::filter(species != "Unknown") |>
+  table() 
+anns_wide_animals_table <- anns_wide_animals_table > 0
+anns_wide_animals_table <- anns_wide_animals_table + 0
+anns_wide_animals_table <- anns_wide_animals_table |> 
+  as.data.frame() |> 
+  tibble::rownames_to_column("camera_id") |> 
   dplyr::arrange(camera_id)
-anns_wide_animals_table_tb <- anns_wide_animals_table %>% 
+anns_wide_animals_table_tb <- anns_wide_animals_table |> 
   dplyr::select(-camera_id)
@@
 
-lu_16 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
-lu_17 <- raster::raster("analysis/data/raw_data/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
-
-lu_dat <- readr::read_csv("analysis/data/raw_data/land_use/attr_table_northen_ont_lc.txt") %>% 
+lu_16 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
+lu_17 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
+
+lu_dat <- readr::read_csv("data-raw/spatial/land_use/attr_table_northen_ont_lc.txt") |> 
   dplyr::mutate(cats = as.factor(code))
 
-cams_sf_buffers_16 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_16)) %>% 
+cams_sf_buffers_16 <- cams_sf |> 
+  sf::st_transform(sf::st_crs(lu_16)) |> 
   sf::st_buffer(dist = 1000)
 
-cams_sf_buffers_17 <- cams_sf %>% 
-  sf::st_transform(sf::st_crs(lu_17)) %>% 
+cams_sf_buffers_17 <- cams_sf |> 
+  sf::st_transform(sf::st_crs(lu_17)) |> 
   sf::st_buffer(dist = 1000)
@@
 
-ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) %>% unlist())
-ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) %>% unlist())
+ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) |> unlist())
+ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) |> unlist())
 
@@
 
-extr_list <- extr %>% 
+extr_list <- extr |> 
   lapply(FUN = function(x){
@@
 
-extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") %>% 
-  dplyr::mutate(props = as.double(props)) %>% 
-  dplyr::left_join(lu_dat, by="cats") %>% 
-  dplyr::filter(!is.na(category_code)) %>% 
-  dplyr::select(camera_id,category_code, props) %>% 
-  tidyr::pivot_wider(names_from = "category_code", values_from = "props") %>% 
+extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") |> 
+  dplyr::mutate(props = as.double(props)) |> 
+  dplyr::left_join(lu_dat, by="cats") |> 
+  dplyr::filter(!is.na(category_code)) |> 
+  dplyr::select(camera_id,category_code, props) |> 
+  tidyr::pivot_wider(names_from = "category_code", values_from = "props") |> 
   as.data.frame()
@@
 # Combine the categories
-extr_table_simple <- extr_table %>% 
+extr_table_simple <- extr_table |> 
   dplyr::mutate(WAT = WAT + XWAT,
@@
                 SWA = ThSWA + ConSWA, 
-                Anth = URB + MIN + BED) %>% 
+                Anth = URB + MIN + BED) |> 
   dplyr::select(-XWAT,
@@
 
-extr_table_sub <- extr_table_simple %>% 
-  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) %>% 
+extr_table_sub <- extr_table_simple |> 
+  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) |> 
   dplyr::arrange(camera_id)
-extr_table_sub_tb <- extr_table_sub %>% 
+extr_table_sub_tb <- extr_table_sub |> 
   dplyr::select(-camera_id)
@@
 
-events_plot_days <- anns_wide_animals_joined_mut %>% 
-  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
-  dplyr::select(plot_id, species, ymd) %>% 
+events_plot_days <- anns_wide_animals_joined_mut |> 
+  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) |> 
+  dplyr::select(plot_id, species, ymd) |> 
   dplyr::distinct()
 
-events_plot <- events_plot_days %>% 
-  dplyr::group_by(plot_id, species) %>%
-  dplyr::summarise(day_count=dplyr::n()) %>% 
+events_plot <- events_plot_days |> 
+  dplyr::group_by(plot_id, species) |>
+  dplyr::summarise(day_count=dplyr::n()) |> 
   dplyr::ungroup()
 
-events_plot_tb <- events_plot %>% 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
-  dplyr::arrange(plot_id) %>%
+events_plot_tb <- events_plot |> 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
+  dplyr::arrange(plot_id) |>
   tibble::column_to_rownames("plot_id")
@@
 
-events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
+events_plot_tb_filt_with_col <- events_plot_tb_filt |> 
   tibble::rownames_to_column("plot_id")
@@
 
-events_cam_days <- anns_wide_animals_joined_mut %>% 
-  dplyr::select(camera_id, species, month, ymd) %>% 
+events_cam_days <- anns_wide_animals_joined_mut |> 
+  dplyr::select(camera_id, species, month, ymd) |> 
   dplyr::distinct()
 
-events_cam <- events_cam_days %>% 
-  dplyr::group_by(camera_id, species) %>%
-  dplyr::summarise(day_count=dplyr::n()) %>% 
+events_cam <- events_cam_days |> 
+  dplyr::group_by(camera_id, species) |>
+  dplyr::summarise(day_count=dplyr::n()) |> 
   dplyr::ungroup()
 
-events_tb_cam <- events_cam %>% 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
-  dplyr::arrange(camera_id) %>%
+events_tb_cam <- events_cam |> 
+  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
+  dplyr::arrange(camera_id) |>
   tibble::column_to_rownames("camera_id")
@@
 
-sp_tab <- dplyr::rename(.data = events_tb_cam %>% colSums() %>% 
-                          sort(decreasing = TRUE) %>% as.data.frame(), 
-                        days_detected=`.`) %>% 
+sp_tab <- dplyr::rename(.data = events_tb_cam |> colSums() |> 
+                          sort(decreasing = TRUE) |> as.data.frame(), 
+                        days_detected=`sort(colSums(events_tb_cam), decreasing = TRUE)`) |> 
   tibble::rownames_to_column("species")
@@
 # saveRDS(sci_names, "analysis/data/derived_data/sci_names.rds")
-sci_names <- readRDS("analysis/data/derived_data/sci_names.rds") %>% 
-  unlist() %>% tibble::as_tibble(rownames = NA) %>% 
-  tibble::rownames_to_column() %>% 
+sci_names <- readRDS("data/objects/sci_names.rds") |> 
+  unlist() |> tibble::as_tibble(rownames = NA) |> 
+  tibble::rownames_to_column() |> 
   dplyr::rename(Scientific_name = value,
-                species = rowname) %>% 
-  tibble::add_row(Scientific_name = "Ursus americanus", species="Black bear") %>% 
+                species = rowname) |> 
+  tibble::add_row(Scientific_name = "Ursus americanus", species="Black bear") |> 
   tibble::add_row(Scientific_name = "Perisoreus canadensis", species="Canada Jay")
 
-sp_tab %>%
-  dplyr::left_join(sci_names, by="species") %>% 
-  dplyr::relocate(Scientific_name, .after=species) %>% 
-  dplyr::filter(days_detected >= 10) %>%
-  gt::gt() %>%
+sp_tab |>
+  dplyr::left_join(sci_names, by="species") |> 
+  dplyr::relocate(Scientific_name, .after=species) |> 
+  dplyr::filter(days_detected >= 10) |>
+  gt::gt() |>
   gt::tab_header(
@@
     subtitle = "Computed in number of days detected in the season"
-  ) %>%
+  ) |>
   gt::fmt_number(
@@
     suffixing = TRUE
-  ) %>% 
+  ) |> 
   gt::cols_label(
@@
     Scientific_name = "Scientific name"
-  ) %>% 
+  ) |> 
   gt::tab_style(style = list(
       gt::cell_text(style="italic")),
-    locations = gt::cells_body(columns = Scientific_name)) %>% 
+    locations = gt::cells_body(columns = Scientific_name)) |> 
   gt::opt_stylize(style = 6, color = "gray")
@@
 
-cams_sf_with_props <- cams_sf_coords %>% 
-  sf::st_drop_geometry() %>% 
-  dplyr::select(plot_id,X,Y) %>% 
-  dplyr::group_by(plot_id) %>% 
-  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
-  dplyr::distinct() %>% 
-  dplyr::ungroup() %>% 
+cams_sf_with_props <- cams_sf_coords |> 
+  sf::st_drop_geometry() |> 
+  dplyr::select(plot_id,X,Y) |> 
+  dplyr::group_by(plot_id) |> 
+  dplyr::mutate(X=mean(X), Y=mean(Y)) |>
+  dplyr::distinct() |> 
+  dplyr::ungroup() |> 
   dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")
@@
 
-regions_labels_bis <- zones_filter %>% 
-  sf::st_drop_geometry() %>% 
+regions_labels_bis <- zones_filter |> 
+  sf::st_drop_geometry() |> 
   dplyr::bind_cols(data.frame(x = c(-85, -88),
@@
 
-source("scripts/3_ordination.R")
+source("scripts/R/postprocess/3_ordination.R")
 
@@
 For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
-
+library(magrittr)
 custom_rda_plot(pa_rda)
@@
 
-clust_grp <- cutree(clust, k=2)
+clust_grp <- cutree(clust, k=4)
 custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
@@
 
-cams_sf_clust <- cams_sf_coords %>% 
-  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
-  dplyr::left_join(grps_df, by = "camera_id") %>% 
+cams_sf_clust <- cams_sf_coords |> 
+  # dplyr::filter(camera_id %in% names(clust_grp)) |> 
+  dplyr::left_join(grps_df, by = "camera_id") |> 
   dplyr::filter(!(is.na(grp)))
@@
 
-dist_geo <- cams_sf_clust %>% 
-  dplyr::arrange(camera_id) %>% 
-  sf::st_distance() %>% 
+dist_geo <- cams_sf_clust |> 
+  dplyr::arrange(camera_id) |> 
+  sf::st_distance() |> 
   as.dist()
@@
 # range.alpha <- seq(0,0.2,0.01)
-K <- 2
+K <- 4
 
@@
 
-cams_sf_geo_clust <- cams_sf_coords %>% 
-  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
-  dplyr::left_join(grps_geo_df, by = "camera_id") %>% 
+cams_sf_geo_clust <- cams_sf_coords |> 
+  # dplyr::filter(camera_id %in% names(clust_grp)) |> 
+  dplyr::left_join(grps_geo_df, by = "camera_id") |> 
   dplyr::filter(!(is.na(grp)))
@@
 
-# cams_sf_sp <- anns_wide_animals_joined_mut %>% 
-#   dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
+# cams_sf_sp <- anns_wide_animals_joined_mut |> 
+#   dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) |> 
 #   dplyr::filter(species %in% colnames(events[,-1]))
 
-cams_sf_sp <- events_cam_days %>% 
-  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) # %>% 
+cams_sf_sp <- events_cam_days |> 
+  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) # |> 
   # dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))
 
-cams_sf_sp_sum <- cams_sf_sp %>% 
-  dplyr::group_by(camera_id, X, Y, geometry, species, month) %>% 
-  dplyr::summarise(count=dplyr::n()) %>% 
-  dplyr::ungroup() # %>% 
-# dplyr::group_by(camera_id, X, Y, species) %>% 
-# dplyr::mutate(count=count/sum(count)) %>% 
+cams_sf_sp_sum <- cams_sf_sp |> 
+  dplyr::group_by(camera_id, X, Y, geometry, species, month) |> 
+  dplyr::summarise(count=dplyr::n()) |> 
+  dplyr::ungroup() # |> 
+# dplyr::group_by(camera_id, X, Y, species) |> 
+# dplyr::mutate(count=count/sum(count)) |> 
 # dplyr::ungroup() 
@@
 
-```{r}
+```{r, eval = FALSE}
 #| label: heatmap
@@
 
-cams_sf_sp_filt <- cams_sf_sp %>% 
+cams_sf_sp_filt <- cams_sf_sp |> 
   dplyr::filter(species == "Caribou", month == "Jun")
 
-cams_sf_sp_sum_filt <- cams_sf_sp_sum %>% 
+cams_sf_sp_sum_filt <- cams_sf_sp_sum |> 
   dplyr::filter(species == "Caribou", month == "Jun")
@@
                      Y=seq(50, 55, by=0.1))
-preds <- pred_grid %>% 
+preds <- pred_grid |> 
   dplyr::mutate(count=predict(mod, pred_grid))
@@
 
-cams_sf_sp_caribou <- cams_sf_sp %>% 
+cams_sf_sp_caribou <- cams_sf_sp |> 
   dplyr::filter(species == "Caribou")
@@
 
-# cams_sf_sp_sum %>%
-#   # dplyr::filter(species == "Caribou") %>%
+# cams_sf_sp_sum |>
+#   # dplyr::filter(species == "Caribou") |>
 #   ggplot2::ggplot() +
@@
 
-  dat <- sf::st_as_sf(cams_sf_sp_sum) %>% 
+  dat <- sf::st_as_sf(cams_sf_sp_sum) |> 
     dplyr::filter(species == sp)
@@
 
-```{r}
+```{r, eval = FALSE}
 #| label: heatmaps
@@
 
-  cams_sf_sp_spe <- cams_sf_sp %>%
+  cams_sf_sp_spe <- cams_sf_sp |>
     dplyr::filter(species == sp)
@@
 
-sp_tab %>%
-  gt::gt() %>%
+sp_tab |>
+  gt::gt() |>
   gt::tab_header(
@@
     species = "Species",
-  ) %>% 
+  ) |> 
   gt::opt_stylize(style = 6, color = "gray")
@@
 
-lu_dat %>% 
-  dplyr::select(category_code, label) %>% 
-  dplyr::filter(category_code != "ND") %>% 
-  dplyr::arrange(category_code) %>% 
-  gt::gt() %>%
+lu_dat |> 
+  dplyr::select(category_code, label) |> 
+  dplyr::filter(category_code != "ND") |> 
+  dplyr::arrange(category_code) |> 
+  gt::gt() |>
   gt::tab_header(
@@
     subtitle = "Adapted from the ontario Far North land cover dataset"
-  ) %>%
+  ) |>
   gt::cols_label(
@@
     category_code = "Abbreviations",
-  ) %>% 
+  ) |> 
   gt::opt_stylize(style = 1, color = "gray")
@@
 # # Create ordered dataframe, and calculate time interval between images.
-# x1 <- x %>%
+# x1 <- x |>
 #   # Sometimes VNA sneaks in here
-#   mutate(number_individuals = as.numeric(ifelse(number_individuals == "VNA", 1, number_individuals))) %>%
+#   mutate(number_individuals = as.numeric(ifelse(number_individuals == "VNA", 1, number_individuals))) |>
 #   # Amalgamate tags of same species in same image; currently broken into two separate rows
-#   group_by(location, {{datetime_col}}, common_name) %>%
-#   mutate(number_individuals = sum(number_individuals)) %>%
-#   distinct(location, {{datetime_col}}, common_name, number_individuals, .keep_all = TRUE) %>%
-#   ungroup() %>%
+#   group_by(location, {{datetime_col}}, common_name) |>
+#   mutate(number_individuals = sum(number_individuals)) |>
+#   distinct(location, {{datetime_col}}, common_name, number_individuals, .keep_all = TRUE) |>
+#   ungroup() |>
 #   # Order the dataframe
-#   arrange(project, location, {{datetime_col}}, common_name) %>%
-#   group_by(project, location, common_name) %>%
+#   arrange(project, location, {{datetime_col}}, common_name) |>
+#   group_by(project, location, common_name) |>
 #   # Calculate the time difference between subsequent images
-#   mutate(interval = int_length({{datetime_col}} %--% lag({{datetime_col}}))) %>%
+#   mutate(interval = int_length({{datetime_col}} %--% lag({{datetime_col}}))) |>
 #   # Is this considered a new detection?
-#   mutate(new_detection = ifelse(is.na(interval) | abs(interval) >= threshold, TRUE, FALSE)) %>%
-#   ungroup() %>%
+#   mutate(new_detection = ifelse(is.na(interval) | abs(interval) >= threshold, TRUE, FALSE)) |>
+#   ungroup() |>
 #   # Number independent detections
@@
 # # Summarise detections
-# x2 <- x1 %>%
-#   group_by(detection, project, location, common_name, scientific_name) %>%
+# x2 <- x1 |>
+#   group_by(detection, project, location, common_name, scientific_name) |>
 #   summarise(start_time = min({{datetime_col}}),

### 2024-01-28 scripts/bash/move_files.sh
move file

@@ -1,0 +1,33 @@
+#!/bin/bash
+
+# Specify the directory containing your folders
+main_directory="TrailCamStorage_2"
+
+# Loop through each folder in the main directory
+for folder in "$main_directory"/*; do
+    if [ -d "$folder" ]; then
+        echo "Processing files in folder: $folder"
+
+        # Create subdirectories for NonWildlife and Wildlife files
+        mkdir -p "$folder/NonWildlife"
+        mkdir -p "$folder/Wildlife"
+
+        # Loop through each file in the folder
+        for file in "$folder"/*; do
+            if [ -f "$file" ]; then
+                filename=$(basename "$file")
+                
+                # Check if the filename contains "NonWildlife"
+                if [[ "$filename" == *NonWildlife* ]]; then
+                    mv "$file" "$folder/NonWildlife"
+                    echo "Moved $filename to NonWildlife folder"
+                else
+                    mv "$file" "$folder/Wildlife"
+                    echo "Moved $filename to Wildlife folder"
+                fi
+            fi
+        done
+
+        echo "Processing complete for folder: $folder"
+    fi
+done

### 2024-01-28 README.Rmd
add a couple of things

@@
 
+## Processing annotations
+
+1. Process and tabulate the annotations. TODO: first local then containerized
+
+```{r}
+
+```
+
 ## Classifier training workflow
@@
 ### Species classifier
+
+1. Crop all annotations using the functions in the R package. TODO: first local then containerized (code now written for locals)
+
+```{r, eval=FALSE}
+# Read in the annotations
+anns <- readRDS("data/objects/annotations_wide_noempty.rds")
+rofcamtrap::crop_from_annotations(anns, 
+                                  base_dir = "/media/vlucet/TrailCamST1/",
+                                  out_dir = "data/images/cropped/")
+```
 

### 2024-01-28 README.Rmd
updates readme a bit

@@
 
-Prerequisites:
+*Dependencies:*
 
-  - docker
-  - nvidia-container-toolkit
+-   *docker*
+-   *nvidia-container-toolkit*
+
+## Report(s) and website
+
+The **reports** directory currently contains the [:file\_folder: paper](/analysis/paper) directory with the Quarto source document for the GnC report.
+Includes code to reproduce the figures and tables generated by the analysis.
+
+Website is under construction.
+
+## How to run in your browser or download and run locally
+
+This research compendium has been developed using the statistical programming
+language R. To work with the compendium, you will need
+installed on your computer the [R software](https://cloud.r-project.org/)
+itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).
+
+You can download the compendium as a zip from from this URL:
+[master.zip](/archive/main.zip). After unzipping:
+- open the `.Rproj` file in RStudio
+- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
+[DESCRIPTION](/DESCRIPTION) file).
+- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console
+
+### How to cite
+
+Please cite this compendium as:
+
+> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`).
+_Compendium of R code and data for `r Title`_.
+Accessed `r format(Sys.Date(), "%d %b %Y")`.
+Online at <https://doi.org/xxx/xxx>
 
@@
 
-## Reports
-
-The **reports** directory currently contains the `GnC_report` folder wich is structured as such:
-
-  - [:file\_folder: paper](/analysis/paper): Quarto source document
-    for manuscript. Includes code to reproduce the figures and tables
-    generated by the analysis. It also has a rendered version,
-    `paper.docx`, suitable for reading (the code is replaced by figures
-    and tables in this file)
-  - [:file\_folder: data](/analysis/data): Data used in the analysis.
-  - [:file\_folder: figures](/analysis/figures): Plots and other
-    illustrations
-  - [:file\_folder: supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
-
-## How to run in your browser or download and run locally
-
-This research compendium has been developed using the statistical programming
-language R. To work with the compendium, you will need
-installed on your computer the [R software](https://cloud.r-project.org/)
-itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).
-
-You can download the compendium as a zip from from this URL:
-[master.zip](/archive/main.zip). After unzipping:
-- open the `.Rproj` file in RStudio
-- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
-[DESCRIPTION](/DESCRIPTION) file).
-- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console
-
+````{=html}
+<!--
 ### Licenses
@@
 
-<!-- **Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
+**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
 
@@
 
-**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->
+**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
 
@@
 By participating in this project you agree to abide by its terms.
-
-### How to cite
-
-Please cite this compendium as:
-
-> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`).
-_Compendium of R code and data for `r Title`_.
-Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/xxx/xxx>
 
@@
 
-<!-- This repository contains the data and code for our paper:
+This repository contains the data and code for our paper:
 
@@
 
-> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->
+> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> 
+-->
+````

### 2024-01-28 reports/GnC_report/paper/paper.qmd
updates

@@
 author:
-  - Valentin Lucet:
-      # correspondence: "yes"
+  - name: Valentin Lucet
     email: valentin.lucet@gmail.com
-      orcid: "0000-0003-0268-818X"
-      institute: wlu
-  - Samantha McFarlane:
-      institute: eccc
-  - Jennifer Baltzer:
-      institute: wlu
-  - Cheryl A. Johnson:   
-      institute: eccc
-  - Mathieu Leblond:
-      institute: eccc
-  - Eric Neilson:
-      institute: nrcan
-  - Philip Weibe:
-      institute: nrcan
-  - Frances Stewart:
-      institute: wlu
-      orcid: "0000-0001-9344-8346"
-  - Josie Hughes:
-      institute: eccc
-institute:
-  - wlu:
+    orcid: 0000-0003-0268-818X
+    affiliation: 
+      - ref: wlu
+  - name: Samantha McFarlane
+    affiliation:
+      - ref: eccc
+  - name: Jennifer Baltzer
+    affiliation:
+      - ref: wlu
+  - name: Cheryl A. Johnson
+    affiliation:
+      - ref: eccc
+  - name: Mathieu Leblond
+    affiliation:
+      - ref: eccc
+  - name: Eric Neilson
+    affiliation:
+      - ref: nrcan
+  - name: Philip Weibe
+    affiliation:
+      - ref: nrcan
+  - name: Frances Stewart
+    orcid: 0000-0001-9344-8346
+    affiliation:
+      - ref: wlu
+  - name: Josie Hughes
+    affiliation:
+      - ref: eccc
+affiliations:
+  - id: wlu
     name: Wilfrid Laurier University
-  - eccc:
+  - id: eccc
     name: Environment & Climate Change Canada
-  - nrcan:
+  - id: nrcan
     name: Natural Resources Canada
@@
 format:
-  # docx:
-  #   reference-doc: "../templates/template_modified.docx" # Insert path for the DOCX file
-  pdf:
-    toc: true
-    number-sections: true
-    colorlinks: true
-    extra_dependencies: ["float"]
+  html: default
+  # pdf:
+  #   toc: true
+  #   number-sections: true
+  #   colorlinks: true
+  #   extra_dependencies: ["float"]
 execute:
@@
   fig-dpi: 600
-filters:
-  - ../templates/scholarly-metadata.lua
-  - ../templates/author-info-blocks.lua
-  - ../templates/pagebreak.lua
 bibliography: references.bib
-csl: "../templates/landscape-ecology.csl" # Insert path for the bib-style
-abstract: |
+abstract: >
   Northern Ontario holds some of the world’s last intact wild places, but it is under pressure from resource extraction and climate change.

### 2024-01-28 reports/GnC_report/paper/paper_test.qmd
add test version

---
title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North"
subtitle: "Prepared for: Canadian Wildlife Service, Canadian Forest Service, and Wilfrid Laurier University"
author:
  - name: Valentin Lucet
    email: valentin.lucet@gmail.com
    orcid: 0000-0003-0268-818X
    affiliation: 
      - ref: wlu
affiliations:
  - id: wlu
    name: Wilfrid Laurier University
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  # pdf:
  #   toc: true
  #   number-sections: true
  #   colorlinks: true
  #   extra_dependencies: ["float"]
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
bibliography: references.bib
abstract: >
  Test.
---

## Test

### 2024-01-28 R/crop.R
progress on package part

@@ -17,9 +17,10 @@ make_crop_name_label <- function(the_row,
 crop_from_annotations <- function(annotations,
                                   base_dir,
                                   out_dir,
-                                  overwrite = FALSE,
-                                  save_anns = TRUE,
-                                  save_labels = TRUE) {
+                                  overwrite = FALSE #,
+                                  # save_anns = TRUE,
+                                  # save_labels = TRUE
+                                  ) {
   browser()
 
   annotations <- annotations |>

### 2024-01-28 R/processing.R

@@ -1,0 +1,232 @@
+## Utilities for processing annotations and detections
+
+#' @export
+process_detections <- function(input_dir = "2_LabelStudio/0_inputs/") {
+
+  # Folders with Megadetector detections in coco format
+  deploy_dirs <- list.dirs(input_dir)
+
+  # List all the files with the proper file name format
+  file_list <- purrr::map(deploy_dirs, list.files, recursive = F,
+                   full.names = T, pattern = "output_coco.json") |>
+    purrr::list_c()
+
+  # Lapply did make it quite ressource heavy so we do a for loop
+  detections <- data.frame()
+  pb <- progress::progress_bar$new(total = length(file_list))
+
+  for (file in file_list) {
+    outls <- jsonlite::fromJSON(file, flatten = T)$annotations
+    detections <- rbind(detections, janitor::clean_names(outls))
+    pb$tick()
+  }
+
+  # if (save_dets) {
+  #   saveRDS(detections, file.path(out_dir, "detections.rds"))
+  #
+  # }
+
+  return(detections)
+}
+
+#' @export
+process_annotations <- function(input_dir = "2_LabelStudio/2_outputs_processed/") {
+
+  # List all the files with the proper file name format
+  file_list <- list.files(input_dir, recursive = F,
+                          full.names = T, pattern = "output.csv")
+
+  # Create an empty dataframe to be grown
+  annotations_no_dup <- data.frame()
+  annotations_dup <- data.frame()
+
+  # separate into dups and not dups. process not dups normally and
+  # process dups specifically
+  base_list <- basename(file_list)
+  base_list <- stringr::str_replace_all(base_list, "TrailCamStorage_2", "T2")
+  base_list <- stringr::str_replace_all(base_list, "TrailCamStorage", "T1")
+
+  projects <- unlist(lapply(stringr::str_split(base_list, "_"),
+                            function(x) paste0(x[1:2], collapse = "_")))
+  names(file_list) <- projects
+  dups_names <- names(which(table(projects) > 1))
+  non_dups_names <- names(which(!(table(projects) > 1)))
+
+  dups_names
+
+  file_list_no_dups <- file_list[names(file_list) %in% non_dups_names]
+  file_list_dups <- file_list[names(file_list) %in% dups_names]
+
+  # -------------------------------------------------------------------------
+
+  for (file in file_list_no_dups) {
+
+    outls <- read_ls_file(file)
+
[…]

### 2024-01-29 R/processing.R
update

@@ -1,14 +1,14 @@
 ## Utilities for processing annotations and detections
 
 #' @export
-process_detections <- function(input_dir = "2_LabelStudio/0_inputs/") {
+process_detections <- function(input_dir) {
 
   # Folders with Megadetector detections in coco format
   deploy_dirs <- list.dirs(input_dir)
 
   # List all the files with the proper file name format
   file_list <- purrr::map(deploy_dirs, list.files, recursive = F,
-                   full.names = T, pattern = "output_coco.json") |>
+                          full.names = T, pattern = "output_coco.json") |>
     purrr::list_c()
 
   # Lapply did make it quite ressource heavy so we do a for loop
@@ -30,7 +30,7 @@ process_detections <- function(input_dir = "2_LabelStudio/0_inputs/") {
 }
 
 #' @export
-process_annotations <- function(input_dir = "2_LabelStudio/2_outputs_processed/") {
+process_annotations <- function(input_dir) {
 
   # List all the files with the proper file name format
   file_list <- list.files(input_dir, recursive = F,
@@ -59,12 +59,11 @@ process_annotations <- function(input_dir = "2_LabelStudio/2_outputs_processed/"
 
   # -------------------------------------------------------------------------
 
+  pb <- progress::progress_bar$new(total = length(file_list_no_dups))
   for (file in file_list_no_dups) {
-
     outls <- read_ls_file(file)
-
     annotations_no_dup <- dplyr::bind_rows(annotations_no_dup, outls)
-
+    pb$tick()
   }
 
   annotations_no_dup <- annotations_no_dup |>
@@ -83,9 +82,11 @@ process_annotations <- function(input_dir = "2_LabelStudio/2_outputs_processed/"
 
   all_anns <- data.frame()
 
+  pb2 <- progress::progress_bar$new(total = length(dups_names))
   for (dup in dups_names) {
 
     print(dup)
+    pb2$tick()
 
     dup_files <- file_list_dups[names(file_list_dups) %in% dup]
 
@@ -184,20 +185,22 @@ process_annotations <- function(input_dir = "2_LabelStudio/2_outputs_processed/"
 
     }
 
-    all_anns <- dplyr::bind_rows(all_anns, anns) # |>
+    all_anns <- dplyr::bind_rows(all_anns, anns)
     # dplyr::select(-orig)
 
-    annotations_wide <- widen(annotations_no_dup) |>
-      dplyr::bind_rows(all_anns) |>
[…]

### 2024-01-29 README.Rmd
update readme

@@
 
+## Processing detections
+
+<!-- 1.  Process and tabulate the annotations. -->
+
+```{r, eval=FALSE}
+detections <- rofcamtrap::process_detections(
+  input_dir = "2_LabelStudio/0_inputs/")
+saveRDS(detections, "data/objects/detections.rds")
+```
+
 ## Processing annotations
 
-1.  Process and tabulate the annotations. TODO: first local then containerized
+1.  Process and tabulate the annotations.
 
-```{r}
+```{r, eval=FALSE}
+annotations <- rofcamtrap::process_annotations(
+   input_dir = "2_LabelStudio/2_outputs_processed/")
 
+saveRDS(annotations, "data/objects/annotations_wide_treated.rds")
+
+# Filter for non empty images
+annotations_noempty <- annotations |>
+  dplyr::filter(!is.na(to_name))
+
+saveRDS(annotations_noempty, "data/objects/annotations_wide_noempty.rds")
 ```

### 2024-01-29 README.Rmd
update readme

@@
 ```{bash eval=FALSE}
-mamba deactivate 
-cd mdtools
-poetry shell
-cd ../
 rofcamtrap/scripts/bash/camtrap.sh \
@@
 docker exec -it label-studio-app-1 bash
-curl -X GET http://localhost:8080/api/projects/?page_size=1000 -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d' -o files/outputs/project_counts.json
+curl -X GET http://localhost:8080/api/projects/?page_size=1000 -H 'Authorization: Token INSERT_TOKEN' -o files/outputs/project_counts.json
 ```
@@
 ```{bash eval=FALSE}
-ls data/outputs/ls # 81 Jan 11 2024
+ls data/outputs/ls # 82 Jan 29 2024
 ```

### 2024-01-29 reports/GnC_report/paper/paper.qmd
update docs website

@@
   html: default
-  # pdf:
-  #   toc: true
-  #   number-sections: true
-  #   colorlinks: true
-  #   extra_dependencies: ["float"]
+  pdf:
+    toc: true
+    number-sections: true
+    colorlinks: true
+    extra_dependencies: ["float"]
+format-links: false
 execute:
@@
 ```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '~/Documents/WILDLab/rofcamtrap/')
+root_dir <- here::here()
+paper_dir <- here::here("reports", "GnC_report", "paper")
+knitr::opts_knit$set(root.dir = root_dir)
 knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
@@
 options(scipen = 1, digits = 2)
-```
-
-<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored.
-
-With the following code you can access and display values from the yml header above.
-Keywords: `r rmarkdown::metadata$keywords`
-Highlights: `r rmarkdown::metadata$highlights`
-
-Here is a citation [@Marwick2017] -->
+do_eval <- FALSE
+```
 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: spatial-basics
-#| echo: false
 
@@
 
-# zones <- 
-#   sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
-#               quiet=TRUE, ) |> 
-#   sf::st_transform(crs = sf::st_crs(cams_sf))
-
 zones <- 
@@
 
-# regions_labels <- zones_filter |> 
-#   sf::st_drop_geometry() |> 
-# dplyr::bind_cols(data.frame(x = c(-88.5, -84, -88.5),
-#                             y = c(55, 52, 52.6)))
-
 regions_labels <- zones_filter |> 
@@
                               y = c(52, 52.6)))
+```
+
+```{r eval=do_eval, echo=FALSE}
+#| label: spatial-compute
 
 map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
-                            zoom = 9, maptype = "stamen_terrain")
-```
-
-```{r}
-#| label: fig-deployment
-#| echo: false
-#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
+                            zoom = 5, maptype = "stamen_terrain")
 
@@
   ggplot2::labs(x = "Longitude", y = "Latitude") +
-  # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
   ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)
+saveRDS(base, here::here(paper_dir, "base_map.rds"))
 
 p <- base +
-  # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
-  #                     size = 2.5, data = regions_labels) +
   ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA), inherit.aes = FALSE,
                       size = 2.5, data = regions_labels) +
-  ggplot2::geom_point(data = cams_sf_coords, #cex = 1, 
+  ggplot2::geom_point(data = cams_sf_coords,
                       shape = 24, col = "black", fill = "red",
-                      ggplot2::aes(x = X, y = Y, #fill=is_retrieved
-                                   ), 
-                      inherit.aes = FALSE) # +
-  # ggplot2::labs(fill = "Camera \nRetrieved")
-p
+                      ggplot2::aes(x = X, y = Y), 
+                      inherit.aes = FALSE)
 ggplot2::ggsave(plot = p, filename = "figures/camera_array.png", 
                 width = 6, height = 8, dpi = 300)
-ggplot2::ggsave(plot = p, filename = "reports/presentations/rofcamtrap-nov2023/images/camera_array.png", 
-                width = 6, height = 8, dpi = 300)
-```
-
-```{r}
+saveRDS(p, here::here(paper_dir, "camera_array.rds"))
+```
+
+```{r echo=FALSE}
+#| label: fig-deployment
+#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
+p <- readRDS(here::here(paper_dir, "camera_array.rds"))
+p
+```
+
+```{r eval=do_eval, echo=FALSE}
 #| label: fig-deployment-poster
-#| eval: false
-#| echo: false
 
@@
 
-```{r}
-#| label: load-data
-#| echo: false
-
-anns_wide <- readRDS("data/objects/annotations_wide_treated.rds")
-dets <- readRDS("data/objects/detections.rds")
-
-# Compute total amount of images
-nb_image <- length(unique(dets$image_id))
-nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
-
-# Hardcoded check
-stopifnot(nb_image == 3288178) # 2788375
+```{r eval=do_eval, echo=FALSE}
+#| label: process-exif
 
@@
   dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
-
   dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
   dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
-
-  dplyr::mutate(source_file = stringr::str_remove(source_file, 
-                                               "/workspace/storage/TrailCamStorage_2/")) |>
-  dplyr::mutate(source_file = stringr::str_remove(source_file, 
-                                               "/workspace/storage/TrailCamStorage/")) |>
-
+  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
+  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
   dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
   dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
-
-  dplyr::mutate(image_id = stringr::str_remove(image_id, 
-                                               "_workspace_storage_TrailCamStorage_2_")) |>
-  dplyr::mutate(image_id = stringr::str_remove(image_id, 
-                                               "_workspace_storage_TrailCamStorage_")) |>
-
+  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
+  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
   dplyr::mutate(image_id = strsplit(image_id, "_") |> 
@@
                   unlist()) |> 
-
-  dplyr::mutate(
-    source_file = 
-      stringr::str_remove(source_file, 
-                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
-  dplyr::mutate(
-    source_file = 
-      stringr::str_remove(source_file, 
-                          stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
-  # dplyr::mutate(image_id = stringr::str_sub(image_id, category6))
+  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
+  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
+
+readr::write_csv(exif, here::here("data", "tabular", "exif_processed.csv"))
+```
+
+```{r echo=FALSE}
+#| label: load-data
+
+anns_wide <- readRDS("data/objects/annotations_wide_treated.rds")
+anns_wide_noempty  <- readRDS("data/objects/annotations_wide_noempty.rds")
+dets <- readRDS("data/objects/detections.rds")
+exif <- readr::read_csv(here::here("data", "tabular", "exif_processed.csv"))
+
+# Compute total amount of images
+nb_image <- length(unique(dets$image_id))
+nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
+
+# Hardcoded check
+stopifnot(nb_image == 3288178) # 2788375
 ```
@@
 Images were retrieved from all cameras using a 32 GB SD card and copied onto a 10 TB hard drive (LaCie d2 Professional).
-The images amounted to 1.8TB of memory on this disk, for a total of __`r nb_image_fmt`__ images.
+The images amounted to 1.8TB of memory on this disk, for a total of **`r nb_image_fmt`** images.
 Images were stored with the following directory structure:
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: basics-stats-1
-#| echo: false
 
 # Select and distinct down to image-based infos
-dets_images <- dets |>  dplyr::select(image_id, max_confidence, isempty) |> 
+dets_images <- dets |> 
+  dplyr::select(image_id, max_confidence, isempty) |> 
   dplyr::distinct()
@@
 At this first step into the process, no filtering was applied beyond the model predictions which classifies images as either likely to contain an object of relevance or empty.
-About __`r dets_tb_percent[1]` %__ of the images, (__`r dets_tb_fmt[1]`__) images, were found by the model to possibly contain either a person, animal, or vehicle.
-The model did not produce detections for __`r dets_tb_fmt[2]`__ images, (__`r dets_tb_percent[2]` %__) of the total.
-
-```{r}
+About **`r dets_tb_percent[1]` %** of the images, (**`r dets_tb_fmt[1]`**) images, were found by the model to possibly contain either a person, animal, or vehicle.
+The model did not produce detections for **`r dets_tb_fmt[2]`** images, (**`r dets_tb_percent[2]` %**) of the total.
+
+```{r echo=FALSE}
 #| label: basics-stats-2
-#| echo: false
+
 dets_0 <- dets |> dplyr::filter(confidence > 0)
@@
 
-Because the model can produce more than one detection per image, this resulted in a total of __`r dets_tb_cat_unfil_sum_fmt`__ detections.
-The totals for each category are: __`r dets_tb_cat_unfil_fmt[1]`__ animals (or __`r dets_tb_cat_unfil_percent[1]` %__); __`r dets_tb_cat_unfil_fmt[2]`__ persons (or __`r dets_tb_cat_unfil_percent[2]` %__); and __`r dets_tb_cat_unfil_fmt[3]`__ vehicles (or __`r dets_tb_cat_unfil_percent[3]` %__).
+Because the model can produce more than one detection per image, this resulted in a total of **`r dets_tb_cat_unfil_sum_fmt`** detections.
+The totals for each category are: **`r dets_tb_cat_unfil_fmt[1]`** animals (or **`r dets_tb_cat_unfil_percent[1]` %**); **`r dets_tb_cat_unfil_fmt[2]`** persons (or **`r dets_tb_cat_unfil_percent[2]` %**); and **`r dets_tb_cat_unfil_fmt[3]`** vehicles (or **`r dets_tb_cat_unfil_percent[3]` %**).
 
@@
 Using a cut-off value for the detection confidence of 0.1 (,i.e.
-discarding all detections with confidence below 0.1), this left us with a total of __`r dets_tb_cat_sum_fmt`__ detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
-After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process  for each category are:  __`r dets_tb_cat_fmt[1]`__ animals (__`r dets_tb_cat_percent[1]` %__); __`r dets_tb_cat_fmt[2]`__ persons (__`r dets_tb_cat_percent[2]` %__); and __`r dets_tb_cat_fmt[3]`__ vehicles (__`r dets_tb_cat_percent[3]` %__).
-
-```{r}
+discarding all detections with confidence below 0.1), this left us with a total of **`r dets_tb_cat_sum_fmt`** detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process for each category are: **`r dets_tb_cat_fmt[1]`** animals (**`r dets_tb_cat_percent[1]` %**); **`r dets_tb_cat_fmt[2]`** persons (**`r dets_tb_cat_percent[2]` %**); and **`r dets_tb_cat_fmt[3]`** vehicles (**`r dets_tb_cat_percent[3]` %**).
+
+```{r echo=FALSE}
 #| label: fig-detections-histogram
-#| echo: false
 #| fig-cap: "Histogram of detection confidence scores produced for each bounding box predicted by Megadetector, showing a bimodal distribution.
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: basics-stats-3
-#| echo: false
 
@@
 
-# Filter for non empty images
-anns_wide_noempty <- anns_wide |>
-  dplyr::filter(!is.na(to_name))
-
-saveRDS(anns_wide_noempty, "data/objects/annotations_wide_noempty.rds")
-
 # Number of processed annotations
@@
 
-Out of the __`r dets_tb_cat_fmt[1]`__ animal detections, __`r processed_fmt`__ have been processed to date using our platform (or __`r processed_pct` %__) for a total of __`r processed_img_fmt`__ images (i.e.
-__`r processed_img_pct` %__ of the total amount of images, or __`r processed_img_up_pct` %__ of the total amount of images uploaded to the platform).
+Out of the **`r dets_tb_cat_fmt[1]`** animal detections, **`r processed_fmt`** have been processed to date using our platform (or **`r processed_pct` %**) for a total of **`r processed_img_fmt`** images (i.e.
+**`r processed_img_pct` %** of the total amount of images, or **`r processed_img_up_pct` %** of the total amount of images uploaded to the platform).
 
@@
 This include potentially empty images.
-If we add the empty images viewed as part of a burst, it brings the total amount to __`r viewed_fmt`__ images, (i.e.
-__`r viewed_pct` %__ of the total amount of images, or __`r viewed_pct_up` %__ of the total amount of images uploaded to the platform).
-
-```{r}
+If we add the empty images viewed as part of a burst, it brings the total amount to **`r viewed_fmt`** images, (i.e.
+**`r viewed_pct` %** of the total amount of images, or **`r viewed_pct_up` %** of the total amount of images uploaded to the platform).
+
+```{r echo=FALSE}
 #| label: mans
-#| echo: false
 
@@
 
-Out of the __`r processed_fmt`__ detections, __`r dels_nb_fmt`__ were false positives (or __`r dels_nb_pct` %__), where Megadetector identified something, but there was nothing.
-In addition, __`r non_mans_nb_fmt`__ were true positives (or __`r non_mans_nb_pct` %__), to which __`r mans_nb_fmt`__ false negatives were added (manually added bounding boxes, __`r mans_nb_pct` %__).
-Out of the __`r non_mans_nb_fmt`__ true positives, __`r orig_tb_fmt[3]`__ had to have their bounding box adjusted.
+Out of the **`r processed_fmt`** detections, **`r dels_nb_fmt`** were false positives (or **`r dels_nb_pct` %**), where Megadetector identified something, but there was nothing.
+In addition, **`r non_mans_nb_fmt`** were true positives (or **`r non_mans_nb_pct` %**), to which **`r mans_nb_fmt`** false negatives were added (manually added bounding boxes, **`r mans_nb_pct` %**).
+Out of the **`r non_mans_nb_fmt`** true positives, **`r orig_tb_fmt[3]`** had to have their bounding box adjusted.
 The table below summarises information for the tagging progress.
 
-```{r}
+```{r echo=FALSE}
 #| label: tab-summary-table
-#| echo: false
 
@@
 All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap).
-The two software tools used in our workflow are also available at online repositories:
-
-- [**mdtools**](https://github.com/StewartWILDlab/mdtools): a python-based command line tool for converting Megadetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
-
-- [**mdscripts**](https://github.com/StewartWILDlab/mdscripts): a bash-based command line tool for running Megadetector and mdtools.
+The repo makes use of [**mdtools**](https://github.com/StewartWILDlab/mdtools) module, a python-based command line tool for converting Megadetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: quality-control
-#| echo: false
 
 # Only annotations for animals
-# TODO !!!! investigate weird id of cut length 1
+# TODO investigate id of cut length 1
 anns_wide_animals <- anns_wide_noempty |>
@@
 
+# TODO streamline exif so you dont have to join
 anns_wide_animals_joined <- anns_wide_animals |>
   dplyr::left_join(exif, by = c("source_file", "image_id"))
+
+stopifnot(dim(anns_wide_animals)[1] ==  dim(anns_wide_animals_joined)[1])
+stopifnot((anns_wide_animals_joined$source_file |> is.na() |> sum()) == 0)
 
@@
 #   dplyr::right_join(dets_images, by = "image_id")
+
+saveRDS(anns_wide_animals_joined_mut, "data/objects/annotations_animals.rds")
 ```
@@
 
-```{r eval=FALSE}
+```{r eval=FALSE, echo=FALSE}
 # library(forcats)
@@
 
-```{r eval=FALSE}
+```{r eval=FALSE, echo=FALSE}
 # for_vic <- anns_wide_animals_joined_mut |> 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: species
-#| echo: false
+
 anns_wide_animals_table <- anns_wide_animals |>
@@
 
+base <- readRDS(here::here(paper_dir, "base_map.rds"))
 base +
@@
 
-```{r, eval = FALSE}
+```{r, eval=FALSE}
 #| label: heatmap
@@
 # raster::plot(rast)
-
+```
+
+```{r, eval=FALSE}
+#| label: heatmap2
+#| echo: false
+#| eval: false
 base_cut +
@@
 
-```{r}
+```{r eval=do_eval, echo=FALSE}
 #| label: bubbleplots
-#| echo: false
 
@@
 
-for (sp in unique(cams_sf_sp_sum$species)){
+plot_list <- list()
+
+for (sp in unique(cams_sf_sp_sum$species)) {
 
@@
   gg <- base_cut +
-    ggplot2::geom_point(data = dat, col="black",
-                        ggplot2::aes(x=X, y=Y, size = count), 
+    ggplot2::geom_point(data = dat, col = "black",
+                        ggplot2::aes(x = X, y = Y, size = count), 
                         inherit.aes = FALSE) +
     ggplot2::facet_wrap(~month) +
-    ggplot2::labs(size="Days \ndetected",
-                  title=sp)
-
-  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_bplot.png"))
+    ggplot2::labs(size = "Days \ndetected",
+                  title = sp)
+  if (sp %in% c("Caribou", "Sandhill crane", "Moose", "Wolf")) {
+    plot_list[[sp]] <- gg
   }
+
+  ggplot2::ggsave(gg, filename = paste0("figures/",sp,"_bplot.png"))
+}
+
+saveRDS(plot_list, here::here(paper_dir, "bbplot_list.rds"))
 ```
@@
 
-![](../figures/Caribou_bplot.png)
-\vspace{1pt}
-
-![](../figures/Sandhill crane_bplot.png)
-\vspace{1pt}
-
-![](../figures/Moose_bplot.png)
-\vspace{1pt}
-
-![](../figures/Wolf_bplot.png)
-\vspace{1pt}
+```{r echo=FALSE}
+plot_list <- readRDS(here::here(paper_dir, "bbplot_list.rds"))
+plot_list["Caribou"]
+```
+```{r echo=FALSE}
+plot_list["Sandhill crane"]
+```
+```{r echo=FALSE}
+plot_list["Moose"]
+```
+```{r echo=FALSE}
+plot_list["Wolf"]
+```
 

### 2024-01-29 R/processing.R
update processing

@@ -82,10 +82,22 @@ process_annotations <- function(input_dir) {
 
   all_anns <- data.frame()
 
-  pb2 <- progress::progress_bar$new(total = length(dups_names))
+  # for (dup in dups_names) {
+  #
+  #   dup_files <- file_list_dups[names(file_list_dups) %in% dup]
+  #
+  #   print(dup)
+  #
+  #   assertthat::assert_that(length(dup_files) == 2)
+  # }
+
+  p_length <- length(dups_names)
+  pb2 <- progress::progress_bar$new(total = p_length)
+
   for (dup in dups_names) {
 
     print(dup)
+
     pb2$tick()
 
     dup_files <- file_list_dups[names(file_list_dups) %in% dup]

### 2024-01-31 README.Rmd
update readme

@@
 
-## Processing detections
+## Processing detections & annotations
 
-<!-- 1.  Process and tabulate the annotations. -->
-
-```{r, eval=FALSE}
-detections <- rofcamtrap::process_detections(
-  input_dir = "2_LabelStudio/0_inputs/")
-saveRDS(detections, "data/objects/detections.rds")
-```
-
-## Processing annotations
-
-1.  Process and tabulate the annotations.
-
-```{r, eval=FALSE}
-annotations <- rofcamtrap::process_annotations(
-   input_dir = "2_LabelStudio/2_outputs_processed/")
-
-saveRDS(annotations, "data/objects/annotations_wide_treated.rds")
-
-# Filter for non empty images
-annotations_noempty <- annotations |>
-  dplyr::filter(!is.na(to_name))
-
-saveRDS(annotations_noempty, "data/objects/annotations_wide_noempty.rds")
-```
+See the quarto documents in the analysis directory.
 
@@
 
-We need to download the models... TBA
+We need to download the models... TBC.
 
+<!--
 ### False detections classifier
@@
 
-````{=html}
-<!--
 ### Licenses
@@
 > Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> 
+````
 -->
-````

### 2024-02-01 scripts/R/postprocess/3_ordination.R
change pipe

@@ -7,16 +7,16 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA,
     vegan::ordiArrowMul(vegan::scores(the_rda, display="bp", choices=c(1, 2),
                                       scaling=2))
 
-  sc_si <- vegan::scores(the_rda, display="sites", choices=c(1,2), scaling=2) %>%
-    as.data.frame() %>%
+  sc_si <- vegan::scores(the_rda, display="sites", choices=c(1,2), scaling=2) |>
+    as.data.frame() |>
     tibble::rownames_to_column("name")
-  sc_sp <- (vegan::scores(the_rda, display="species", choices=c(1,2), scaling=2) %>%
-              as.data.frame() * sp_scale) %>%
-    tibble::rownames_to_column("name") %>%
+  sc_sp <- (vegan::scores(the_rda, display="species", choices=c(1,2), scaling=2) |>
+              as.data.frame() * sp_scale) |>
+    tibble::rownames_to_column("name") |>
     dplyr::mutate(highlight =ifelse((abs(RDA1) > thres_plot | abs(RDA2) > thres_plot),
                                     name, ""))
-  sc_bp <- (vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2) %>%
-              as.data.frame() * mult) %>%
+  sc_bp <- (vegan::scores(the_rda, display="bp", choices=c(1, 2), scaling=2) |>
+              as.data.frame() * mult) |>
     tibble::rownames_to_column("name")
 
   base <-  ggplot2::ggplot() +
@@ -69,7 +69,7 @@ custom_rda_plot <- function(the_rda, sp_scale=2.3, clust=NA,
     grps <- data.frame(name = names(clust),
                        grp = as.factor(clust))
 
-    sc_si <- sc_si %>%
+    sc_si <- sc_si |>
       dplyr::left_join(grps, by = "name")
 
     base <- base +

### 2024-02-01 reports/GnC_report/paper/paper.qmd
update report itself

@@
 ```{r, setup, include=FALSE}
-root_dir <- here::here()
-paper_dir <- here::here("reports", "GnC_report", "paper")
-knitr::opts_knit$set(root.dir = root_dir)
 knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
@@
 do_eval <- FALSE
+source(here::here("scripts", "R", "postprocess", "3_ordination.R"))
 ```
@@
 
+```{r echo=FALSE}
+#| label: load-data
+anns_wide <- readRDS(here::here("data", "objects", "annotations_wide_treated.rds"))
+anns_wide_noempty  <- readRDS(here::here("data", "objects", "annotations_wide_noempty.rds"))
+dets <- readRDS(here::here("data", "objects", "detections.rds"))
+exif <- readr::read_csv(here::here("data", "tabular", "exif_processed.csv"))
+cams_sf_coords <- readRDS(here::here("data", "objects", "cams_sf_coords.rds"))
+zones_filtered  <- readRDS(here::here("data", "objects", "zones_filtered.rds"))
+zones_pols <- readRDS(here::here("data", "objects", "zones_pols.rds"))
+base_map <- readRDS(here::here("data", "objects", "base_map.rds"))
+camera_days <- readRDS(here::here("data", "objects" ,"camera_days.rds"))
+camera_regions <- readRDS(here::here("data", "objects" ,"camera_regions.rds"))
+extr_table_sub_tb <- readRDS(here::here("data", "objects" ,"extr_table_sub_tb.rds"))
+anns_wide_animals_joined_mut <- readRDS(here::here("data", "objects", "annotations_animals.rds"))
+anns_wide_animals_table_tb <- readRDS(here::here("data", "objects" ,"anns_wide_animals_table_tb.rds"))
+events_tb_filt_with_col <- readRDS(here::here("data", "objects" ,"events_tb_filt_with_col.rds"))
+events_cam_days <- readRDS(here::here("data", "objects" ,"events_cam_days.rds"))
+events_tb_cam <- readRDS(here::here("data", "objects" ,"events_tb_cam.rds"))
+sci_names_processed <- readRDS(here::here("data", "objects" ,"sci_names_processed.rds"))
+table_cat <- readRDS(here::here("data", "objects" ,"table_cat.rds"))
+```
+
 ### Camera deployment and setup
@@
 ```{r echo=FALSE}
-#| label: spatial-basics
-
-RoF_CameraDeploymentLog_WLU <- 
-  readxl::read_excel(paste0("data-raw/tabular/RoF_CameraDeploymentLog.xlsx")) |> 
-  janitor::clean_names()
-
-cams_sf <- RoF_CameraDeploymentLog_WLU |> 
-  dplyr::rename(longitude = x, latitude = y) |> 
-  dplyr::filter(!is.na(latitude)) |> 
-  dplyr::filter(!is.na(longitude)) |> 
-  sf::st_as_sf(coords = c("longitude","latitude"), crs = "EPSG:4326", remove = F) |> 
-  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))
-
-coords <- cams_sf |> 
-  sf::st_coordinates() |> 
-  as.data.frame()
-
-cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)
-
-zones <- 
-  sf::st_read("data-raw/spatial/ecozones/EcoZone/EcoZone.shp",
-              quiet = TRUE, ) |> 
-  sf::st_transform(crs = sf::st_crs(cams_sf))
-
-zones_filter <- zones |>
-  sf::st_filter(cams_sf)
-
-pols <- sf::st_coordinates(zones_filter) |> 
-  as.data.frame()
-
-regions_labels <- zones_filter |> 
+#| label: make-labels
+
+regions_labels <- zones_filtered |> 
   sf::st_drop_geometry() |> 
@@
 
-```{r eval=do_eval, echo=FALSE}
-#| label: spatial-compute
-
-map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
-                            zoom = 5, maptype = "stamen_terrain")
-
-base <- ggmap::ggmap(map) +
-  ggplot2::labs(x = "Longitude", y = "Latitude") +
-  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)
-saveRDS(base, here::here(paper_dir, "base_map.rds"))
-
-p <- base +
+```{r echo=FALSE}
+#| label: fig-deployment
+#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
+
+p <- base_map +
   ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA), inherit.aes = FALSE,
@@
                       inherit.aes = FALSE)
-ggplot2::ggsave(plot = p, filename = "figures/camera_array.png", 
+
+ggplot2::ggsave(plot = p, filename = here::here("figures", "camera_array.png"), 
                 width = 6, height = 8, dpi = 300)
-saveRDS(p, here::here(paper_dir, "camera_array.rds"))
-```
-
-```{r echo=FALSE}
-#| label: fig-deployment
-#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
-p <- readRDS(here::here(paper_dir, "camera_array.rds"))
+
+# saveRDS(p, here::here("data", "objects", "camera_array.rds"))
+# p <- readRDS(here::here("data", "objects", "camera_array.rds"))
+
 p
@@
 
-```{r eval=do_eval, echo=FALSE}
-#| label: fig-deployment-poster
-
-# base <- ggmap::ggmap(map) +
-#   ggplot2::labs(x = "Longitude", y = "Latitude") +
-#   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
-#   ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
-# 
-# p <- base +
-#   # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
-#   #                     size = 2.5, data = regions_labels) +
-#   ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
-#                       size = 2.5, data = regions_labels) +
-#   ggplot2::geom_point(data = cams_sf_coords |> 
-#                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
-#                       col="black", fill = "red",
-#                       ggplot2::aes(x=X, y=Y), 
-#                       inherit.aes = FALSE) +
-#   ggplot2::labs(fill = "Camera \nRetrieved")
-# p
-```
-
 ### Image detection
 
-```{r eval=do_eval, echo=FALSE}
-#| label: process-exif
-
-exif <- readr::read_csv("data/tabular/all_exif.csv") |>
-  janitor::clean_names() |>
-  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
-  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
-  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
-  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
-  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
-  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
-  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
-  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
-  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
-  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
-  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
-                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
-                  unlist()) |> 
-  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
-  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
-
-readr::write_csv(exif, here::here("data", "tabular", "exif_processed.csv"))
-```
-
-```{r echo=FALSE}
-#| label: load-data
-
-anns_wide <- readRDS("data/objects/annotations_wide_treated.rds")
-anns_wide_noempty  <- readRDS("data/objects/annotations_wide_noempty.rds")
-dets <- readRDS("data/objects/detections.rds")
-exif <- readr::read_csv(here::here("data", "tabular", "exif_processed.csv"))
+```{r echo=FALSE}
+#| label: basic-stats-zero
 
@@
 ```{r echo=FALSE}
-#| label: mans
+#| label: manual-annotations
 
@@
 
-We summarized our existing species data to date for analysis in two main ways:  (1) species presence/absence data for each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site.
-
-```{r echo=FALSE}
-#| label: quality-control
-
-# Only annotations for animals
-# TODO investigate id of cut length 1
-anns_wide_animals <- anns_wide_noempty |>
-  dplyr::filter(species != 'Human') |> 
-  dplyr::filter(species != 'Vehicle') |> 
-  dplyr::mutate(id_cut_length = purrr::map_dbl(id , \(x) {
-    x |> stringr::str_split("_") |> purrr::map_dbl(~length(.x))
-  })) |> 
-  dplyr::filter(id_cut_length > 1) |> 
-  dplyr::mutate(camera_id = purrr::map_chr(id , \(x) {
-    x |>
-      stringr::str_split("_") |>
-      purrr::map(~.x[[2]]) |>
-      unlist()
-  }))
-
-# Unknowns and bird breakdown
-# anns_wide_animals |> dplyr::filter(species == "Unknown") |> 
-#   dplyr::pull(id) |> 
-#   stringr::str_sub(1,4) |> table() |> sort() 
-# anns_wide_animals |> dplyr::filter(stringr::str_detect(species,"Bird")) |> 
-#   dplyr::pull(id) |> 
-#   stringr::str_sub(1,4) |> table() |> sort()
-
-# Joins
-
-# TODO streamline exif so you dont have to join
-anns_wide_animals_joined <- anns_wide_animals |>
-  dplyr::left_join(exif, by = c("source_file", "image_id"))
-
-stopifnot(dim(anns_wide_animals)[1] ==  dim(anns_wide_animals_joined)[1])
-stopifnot((anns_wide_animals_joined$source_file |> is.na() |> sum()) == 0)
-
-anns_wide_animals_joined_mut <- anns_wide_animals_joined |> 
-  dplyr::filter(species != "Unknown") |>
-  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) |> 
-  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
-                                               tz = "Canada/Eastern"),
-                year = lubridate::year(date_time),
-                month = lubridate::month(date_time, label = TRUE),
-                day = lubridate::day(date_time),
-                ymd = paste(year, month, day, sep = "_"))
-
-# anns_wide_animals_joined_dets <- anns_wide_animals |> 
-#   dplyr::right_join(dets_images, by = "image_id")
-
-saveRDS(anns_wide_animals_joined_mut, "data/objects/annotations_animals.rds")
-```
-
-<!-- Other camera plots -->
-
-```{r eval=FALSE, echo=FALSE}
-# library(forcats)
-library(ggplot2)
-# anns_wide_animals_joined_mut |> 
-#   ggplot(aes(x=date_time, y=species)) + geom_point()
-# anns_wide_animals_joined_mut |> 
-#   dplyr::filter(species == "Fisher") |> View()
-
-camera_days <- exif |> dplyr::group_by(maker_notes_user_label) |> 
-  dplyr::summarise(camera_first_time=min(exif_date_time_original)) |> 
-  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time,
-                                               tz = "Canada/Eastern"),
-                year = lubridate::year(date_time),
-                month = lubridate::month(date_time, label = TRUE),
-                day = lubridate::day(date_time),
-                ymd = paste(year, month, day, sep = "_")) |> 
-  dplyr::select(camera_id = maker_notes_user_label, 
-                camera_first_day = date_time)
-
-camera_regions <- cams_sf |> sf::st_join(zones) |> 
-  sf::st_drop_geometry() |> dplyr::select(camera_id, ECOZONE_NA)
-
-bp <- anns_wide_animals_joined_mut |>
-
-  dplyr::filter(species != "Fisher") |> 
+We summarized our existing species data to date for analysis in two main ways:  (1) species presence/absence data for each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site. The detection time for species is highly variable.
+
+```{r echo=FALSE}
+#| label: camera-plots
+#| fig-cap: "Box plot of detection time for each species in either ecozone."
+#| fig.width: 8
+#| fig.height: 4.5
+
+bp_data <-  anns_wide_animals_joined_mut |> 
   dplyr::left_join(camera_days, by = "camera_id") |>
@@
 
-  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
-                ECOZONE_NA = as.factor(ECOZONE_NA)) |>
-
-  ggplot(aes(x = diff_time, 
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA))
+
+bp <- bp_data |> 
+  ggplot2::ggplot(ggplot2::aes(x = diff_time, 
                                y = forcats::fct_reorder(species, total_events),
                                color = ECOZONE_NA)) +
-  geom_boxplot() +
-
-  scale_x_continuous()+
-  labs(x = "Detection time (in days)",
+  ggplot2::geom_boxplot() +
+  ggplot2::scale_x_continuous() +
+  ggplot2::labs(x = "Detection time (in days)",
                 y = "Species",
                 color = "Ecozone") + 
-  theme_bw()
-
-ggsave(plot = bp, filename = "figures/bp.png", width = 10, height = 15)
-
-pts <- anns_wide_animals_joined_mut |>
-
-  dplyr::filter(species != "Fisher") |> 
+  ggplot2::theme_bw()
+
+ggplot2::ggsave(plot = bp, filename = "figures/bp.png", width = 10, height = 15)
+```
+
+```{r echo=FALSE}
+#| label: camera-plots-2
+#| fig-cap: "Bubble plot of detection time for each species in either ecozone"
+#| fig.width: 8
+#| fig.height: 4.5
+
+pts_data <- anns_wide_animals_joined_mut |>
   dplyr::left_join(camera_days, by = "camera_id") |>
@@
 
-  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units="days"),
-                ECOZONE_NA = as.factor(ECOZONE_NA)) |>
-
-  ggplot(aes(x = diff_time, 
+  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
+                ECOZONE_NA = as.factor(ECOZONE_NA))
+
+pts <- pts_data |> 
+  ggplot2::ggplot(ggplot2::aes(x = diff_time, 
                                y = forcats::fct_reorder(species, total_events),
@@
                                color = ECOZONE_NA)) +
-
-  geom_point(aes(x = diff_time, 
+  ggplot2::geom_point(ggplot2::aes(x = diff_time, 
                                    y = forcats::fct_reorder(species, total_events),
@@
                                    group = ECOZONE_NA)) +
-
-  scale_x_continuous()+
-  labs(x = "Detection time (in days)",
+  ggplot2::scale_x_continuous() +
+  ggplot2::labs(x = "Detection time (in days)",
                 y = "Species",
@@
                 size = "Day counts") + 
-  theme_bw()
-ggsave(plot = pts, filename = "figures/pts.png", width = 10, height = 15)
+  ggplot2::theme_bw()
+
+ggplot2::ggsave(plot = pts, filename = "figures/pts.png", width = 10, height = 15)
+
 pts
-```
-
-```{r eval=FALSE, echo=FALSE}
-# for_vic <- anns_wide_animals_joined_mut |> 
-#   dplyr::select(id,species,date_time, camera_id) |> 
-#   dplyr::left_join(cams_sf |> dplyr::select(camera_id, plot_id) |> sf::st_drop_geometry(),
-#                    by = "camera_id")
-# tot_camera <- for_vic$camera_id |> unique() |> length()
-# hit_rates <- for_vic |> dplyr::select(species, camera_id) |> 
-#   dplyr::group_by(species) |> dplyr::summarise(n_cameras=length(unique(camera_id))) |> 
-#   dplyr::ungroup() |> dplyr::mutate(rate = n_cameras/tot_camera*100) |> 
-#   dplyr::arrange(dplyr::desc(rate))
-# # det_history_two_weeks <- 
-# 
-# readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
-# readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
-```
-
-```{r echo=FALSE}
-#| label: species
-
-anns_wide_animals_table <- anns_wide_animals |>
-  dplyr::select(camera_id, species) |> 
-  dplyr::filter(species != "Unknown") |>
-  table() 
-anns_wide_animals_table <- anns_wide_animals_table > 0
-anns_wide_animals_table <- anns_wide_animals_table + 0
-anns_wide_animals_table <- anns_wide_animals_table |> 
-  as.data.frame() |> 
-  tibble::rownames_to_column("camera_id") |> 
-  dplyr::arrange(camera_id)
-anns_wide_animals_table_tb <- anns_wide_animals_table |> 
-  dplyr::select(-camera_id)
-rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id
 ```
@@
 
-```{r}
-#| label: extract
-#| echo: false
-
-lu_16 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM16.tif")
-lu_17 <- raster::raster("data-raw/spatial/land_use/Class/FarNorth_LandCover_Class_UTM17.tif")
-
-lu_dat <- readr::read_csv("data-raw/spatial/land_use/attr_table_northen_ont_lc.txt") |> 
-  dplyr::mutate(cats = as.factor(code))
-
-cams_sf_buffers_16 <- cams_sf |> 
-  sf::st_transform(sf::st_crs(lu_16)) |> 
-  sf::st_buffer(dist = 1000)
-
-cams_sf_buffers_17 <- cams_sf |> 
-  sf::st_transform(sf::st_crs(lu_17)) |> 
-  sf::st_buffer(dist = 1000)
-
-extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
-                                        progress = FALSE)
-extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
-                                        progress = FALSE)
-
-ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) |> unlist())
-ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) |> unlist())
-
-stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))
-
-extr_16[ind_16] <- extr_17[ind_16]
-
-extr <- extr_16
-
-extr_list <- extr |> 
-  lapply(FUN = function(x){
-    the_tab <- table(x$value)
-    props <- the_tab/sum(the_tab)
-    cats <- names(the_tab)
-    df <- tibble::tibble(cats, props)
-  })
-names(extr_list) <- cams_sf_buffers_16$camera_id
-
-extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") |> 
-  dplyr::mutate(props = as.double(props)) |> 
-  dplyr::left_join(lu_dat, by="cats") |> 
-  dplyr::filter(!is.na(category_code)) |> 
-  dplyr::select(camera_id,category_code, props) |> 
-  tidyr::pivot_wider(names_from = "category_code", values_from = "props") |> 
-  as.data.frame()
-extr_table[is.na(extr_table)] <- 0
-
-# Combine the categories
-extr_table_simple <- extr_table |> 
-  dplyr::mutate(WAT = WAT + XWAT,
-                MixTRE = MixTRE + DecTRE,
-                Dist  = TrOrSHr + NSWood, 
-                SWA = ThSWA + ConSWA, 
-                Anth = URB + MIN + BED) |> 
-  dplyr::select(-XWAT,
-                -DecTRE,
-                -TrOrSHr, -NSWood, 
-                -ThSWA, - ConSWA, 
-                -URB, -MIN, -BED)
-
-extr_table_sub <- extr_table_simple |> 
-  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) |> 
-  dplyr::arrange(camera_id)
-extr_table_sub_tb <- extr_table_sub |> 
-  dplyr::select(-camera_id)
-rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
-extr_table_sub_tb <- extr_table_sub_tb[, colSums(extr_table_sub_tb) != 0]
-
-# lu_cropped <- raster::crop(lu_16, cams_sf_buffers_16)
-# raster::plot(lu_cropped)
-```
-
-```{r}
-#| label: extract-cats
-#| echo: false
-#| eval: false
-
-# sum(unlist(lapply(extr_list, function(x) sum(x$cats == "NA"))))
-# sum(unlist(lapply(extr_list, function(x) is.na(x$cats))))
-# #  [1] "1"   "2"   "8"   "9"   "12"  "15"  "16"  "17"  "18"  "11"  "13"  "14"  "19"  "20"  "23"  "157" "247" "21"  "22"
-# 
-# pairs(extr_table_sub_tb, col = "dodgerblue")
-# round(cor(extr_table_sub_tb), 2)
-```
-
 ## Data exploration and preliminary results
@@
 
-```{r}
-#| label: events-Plots
-#| echo: false
-
-events_plot_days <- anns_wide_animals_joined_mut |> 
-  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) |> 
-  dplyr::select(plot_id, species, ymd) |> 
-  dplyr::distinct()
-
-events_plot <- events_plot_days |> 
-  dplyr::group_by(plot_id, species) |>
-  dplyr::summarise(day_count=dplyr::n()) |> 
-  dplyr::ungroup()
-
-events_plot_tb <- events_plot |> 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
-  dplyr::arrange(plot_id) |>
-  tibble::column_to_rownames("plot_id")
-
-events_plot_tb[is.na(events_plot_tb)] <- 0
-events_plot_tb_filt <- events_plot_tb[,colSums(events_plot_tb)>50]
-tmp_col <- ncol(events_plot_tb_filt)
-
-events_plot_tb_filt_with_col <- events_plot_tb_filt |> 
-  tibble::rownames_to_column("plot_id")
-```
-
-```{r}
-#| label: events-Cams
-#| echo: false
-
-events_cam_days <- anns_wide_animals_joined_mut |> 
-  dplyr::select(camera_id, species, month, ymd) |> 
-  dplyr::distinct()
-
-events_cam <- events_cam_days |> 
-  dplyr::group_by(camera_id, species) |>
-  dplyr::summarise(day_count=dplyr::n()) |> 
-  dplyr::ungroup()
-
-events_tb_cam <- events_cam |> 
-  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
-  dplyr::arrange(camera_id) |>
-  tibble::column_to_rownames("camera_id")
-
-events_tb_cam[is.na(events_tb_cam)] <- 0
-# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
-# tmp_col <- ncol(events_tb_cam)
-```
-
-```{r}
+```{r echo=FALSE}
 #| label: tab-species-table
-#| echo: false
 
@@
                           sort(decreasing = TRUE) |> as.data.frame(), 
-                        days_detected=`sort(colSums(events_tb_cam), decreasing = TRUE)`) |> 
+                        days_detected = `sort(colSums(events_tb_cam), decreasing = TRUE)`) |> 
   tibble::rownames_to_column("species")
 
-# sci_names <- taxize::comm2sci(sp_tab$species)
-# saveRDS(sci_names, "analysis/data/derived_data/sci_names.rds")
-sci_names <- readRDS("data/objects/sci_names.rds") |> 
-  unlist() |> tibble::as_tibble(rownames = NA) |> 
-  tibble::rownames_to_column() |> 
-  dplyr::rename(Scientific_name = value,
-                species = rowname) |> 
-  tibble::add_row(Scientific_name = "Ursus americanus", species="Black bear") |> 
-  tibble::add_row(Scientific_name = "Perisoreus canadensis", species="Canada Jay")
-
 sp_tab |>
-  dplyr::left_join(sci_names, by="species") |> 
-  dplyr::relocate(Scientific_name, .after=species) |> 
+  dplyr::left_join(sci_names_processed, by = "species") |> 
+  dplyr::relocate(Scientific_name, .after = species) |> 
   dplyr::filter(days_detected >= 10) |>
@@
   gt::cols_label(
-    days_detected ="Days",
+    days_detected = "Days",
     species = "Species",
@@
   gt::tab_style(style = list(
-      gt::cell_text(style="italic")),
+      gt::cell_text(style = "italic")),
     locations = gt::cells_body(columns = Scientific_name)) |> 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-map-pies
-#| echo: false
 #| fig-cap: "Pie chart map of relative species abundances across our study area Detections have grouped by plot for better visibility, with a plot containing up to 4 cameras.
@@
   dplyr::group_by(plot_id) |> 
-  dplyr::mutate(X=mean(X), Y=mean(Y)) |>
+  dplyr::mutate(X = mean(X), Y = mean(Y)) |>
   dplyr::distinct() |> 
   dplyr::ungroup() |> 
-  dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")
-
-# ggplot2::ggplot() +
-#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
-#                               ggplot2::aes(x=X, y=Y, group=plot_id),
-#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
-#   ggplot2::coord_equal() #+ 
-#ggplot2::theme(legend.position = "none")
-
-# base2 <- ggplot2::ggplot() +
-#   ggplot2::theme_bw()
-# for (pol in 1:2){
-#   base2 <- base2 + 
-#     ggplot2::geom_polygon(data = dplyr::filter(pols, L2 == pol),
-#                           ggplot2::aes(x=X, y=Y),
-#                           col = 1, fill = "grey90")
-# }
-# 
-# base2 +
-#   ggplot2::coord_equal(xlim = c(-88.5, -84),
-#                        ylim = c(50, 54)) +
-#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
-#                               ggplot2::aes(x=X, y=Y, group=plot_id),
-#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
-#                               sorted_by_radius = TRUE,
-#                               legend_name = "Species",
-#                               pie_scale = 1.5) +
-#   ggplot2::labs(x="Lat", y="Long")
-
-regions_labels_bis <- zones_filter |> 
+  dplyr::left_join(events_tb_filt_with_col, by = "plot_id")
+
+regions_labels_bis <- zones_filtered |> 
   sf::st_drop_geometry() |> 
@@
 
-base <- readRDS(here::here(paper_dir, "base_map.rds"))
-base +
+base_map +
   ggplot2::coord_equal(xlim = c(-88.5, -84),
@@
   scatterpie::geom_scatterpie(data = cams_sf_with_props,
-                              ggplot2::aes(x=X, y=Y, group=plot_id),
-                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
+                              ggplot2::aes(x = X, y = Y, group = plot_id),
+                              cols = names(cams_sf_with_props)[4:(4 + ncol(cams_sf_with_props) - 4)],
                               sorted_by_radius = TRUE,
@@
                               pie_scale = 1.5) +
-  ggplot2::labs(x="Lat", y="Long") +
-    ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
+  ggplot2::labs(x = "Lat", y = "Long") +
+    ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA),inherit.aes = FALSE,
                       size = 2.5, data = regions_labels_bis)
@@
 ### Ordination - presence-absence
-
-```{r}
-#| label: rda-pa
-#| echo: false
-
-source("scripts/R/postprocess/3_ordination.R")
-
-# Ordination
-stopifnot(all(rownames(anns_wide_animals_table_tb)==
-                rownames(extr_table_sub_tb)))
-
-anns_wide_animals_table_tb_ch <- vegan::decostand(anns_wide_animals_table_tb, 
-                                                  "hellinger")
-
-pa_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
-                     extr_table_sub_tb)
-```
 
@@
 Ordinations also combine very well with clustering analyses to produce insights about associations between sites.
+
+```{r echo=FALSE}
+#| label: rda-pa
+#| echo: false
+
+anns_wide_animals_table_tb_ch <- vegan::decostand(anns_wide_animals_table_tb, 
+                                                  "hellinger")
+pa_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
+                     extr_table_sub_tb)
+```
 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-rda-pa
-#| echo: false
 #| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected in March through September of 2022 reveals that different species associate with different land cover classes.
 For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
-library(magrittr)
 custom_rda_plot(pa_rda)
@@
 
-```{r}
-#| label: jaccard-pcoa
-#| echo: false
-#| eval: false
-
-# anns_wide_animals_table_tb_jac <- vegan::vegdist(anns_wide_animals_table_tb, 
-#                                                  "jac", binary = TRUE)
-# 
-# pa_pcoa <- ape::pcoa(anns_wide_animals_table_tb_jac)
-# 
-# biplot(pa_pcoa, anns_wide_animals_table_tb)
-```
-
 ### Ordination - relative abundances
 
-```{r}
+```{r echo=FALSE}
 #| label: rda-events
-#| echo: false
 #| fig-cap: "Triplot of RDA on relative abundance data"
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-rda-events
-#| echo: false
 #| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data."
@@
 # vegan::RsquareAdj(events_rda)
-custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
+custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp = 0.4, scale_y_sp = 0.3,
                 thres_plot = 0.3)
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: cluster-base
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-rda-events-clust
-#| echo: false
 #| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
@@
 
-clust_grp <- cutree(clust, k=4)
-custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
+clust_grp <- cutree(clust, k = 4)
+custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp = 0.4, scale_y_sp = 0.3,
                 thres_plot = 0.2716, clust = clust_grp)
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-map-clust
-#| echo: false
 #| fig-cap: "Map of clustered camera sites. The two clusters were determined using the Ward clustering method."
@@
 
-base_cut <- base +
+base_cut <- base_map +
   ggplot2::coord_map(xlim = c(-89, -84),
@@
 base_cut +
-  ggplot2::geom_point(data = cams_sf_clust, cex = 1, shape=21, col="black",
-                      ggplot2::aes(x=X, y=Y, fill=grp), 
+  ggplot2::geom_point(data = cams_sf_clust, cex = 1, shape = 21, col = "black",
+                      ggplot2::aes(x = X, y = Y, fill = grp), 
                       inherit.aes = FALSE) +
-  ggplot2::labs(fill="Group")
+  ggplot2::labs(fill = "Group")
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: clust-geo
-#| echo: false
 
@@
                                  dist_geo, alpha = 0.1)
-clust_geo_grp <- cutree(clust_geo, k=K)
+clust_geo_grp <- cutree(clust_geo, k = K)
 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-map-clust-geo
-#| echo: false
 #| fig-cap: "Map of geographically constrained clustered camera sites, using geographically constrained clustering with a mixing parameter of 0.1."
@@
 base_cut +
-  ggplot2::geom_point(data = cams_sf_geo_clust, cex = 1, shape=21, col="black",
-                      ggplot2::aes(x=X, y=Y, fill=grp), 
+  ggplot2::geom_point(data = cams_sf_geo_clust, cex = 1, shape = 21, col = "black",
+                      ggplot2::aes(x = X, y = Y, fill = grp), 
                       inherit.aes = FALSE) +
-  ggplot2::labs(fill="Group")
+  ggplot2::labs(fill = "Group")
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: fig-rda-clust-geo
-#| echo: false
 #| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: join-sf
@@
   dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) # |> 
-  # dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))
+  # dplyr::filter(species %in% colnames(events_tb_filt_with_col[,-1]))
 
@@
   dplyr::group_by(camera_id, X, Y, geometry, species, month) |> 
-  dplyr::summarise(count=dplyr::n()) |> 
+  dplyr::summarise(count = dplyr::n()) |> 
   dplyr::ungroup() # |> 
@@
 # stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])
-```
-
-```{r, eval=FALSE}
-#| label: heatmap
-#| echo: false
-#| eval: false
-
-options(backup_options)
-
-cams_sf_sp_filt <- cams_sf_sp |> 
-  dplyr::filter(species == "Caribou", month == "Jun")
-
-cams_sf_sp_sum_filt <- cams_sf_sp_sum |> 
-  dplyr::filter(species == "Caribou", month == "Jun")
-
-mod <- mgcv::gam(count~s(X,Y), data = cams_sf_sp_sum_filt)
-pred_grid <- 
-  tidyr::expand_grid(X=seq(-89, -84, by=0.1),
-                     Y=seq(50, 55, by=0.1))
-preds <- pred_grid |> 
-  dplyr::mutate(count=predict(mod, pred_grid))
-
-# rast <- raster::rasterFromXYZ(preds)
-# raster::plot(rast)
-# rast[rast<0] <- NA
-# raster::plot(rast)
-```
-
-```{r, eval=FALSE}
-#| label: heatmap2
-#| echo: false
-#| eval: false
-base_cut +
-  ggplot2::geom_density2d_filled(
-    data = cams_sf_sp_filt, ggplot2::aes(X, Y),
-    binwidth = 0.05, alpha = 0.5)
-
-# base_cut +
-#   ggplot2::geom_tile(
-#     data = dplyr::filter(preds, count>=0), ggplot2::aes(X, Y, fill=count),
-#     binwidth = 0.05, alpha = 0.5, inherit.aes = FALSE) +
-#   ggplot2::scale_fill_viridis_c()
-```
-
-```{r}
-#| label: heatmap-facets
-#| echo: false
-#| eval: false
-
-cams_sf_sp_caribou <- cams_sf_sp |> 
-  dplyr::filter(species == "Caribou")
-
-base_cut +
-  ggplot2::geom_density2d_filled(
-    data = cams_sf_sp_caribou, ggplot2::aes(X, Y),
-    alpha = 0.5,breaks = c(seq(0.001,1,by=0.1),5, 10, 20, 50,100), n=20) +
-  ggplot2::facet_wrap(~month) +
-  ggplot2::coord_equal(xlim = c(-89, -84),
-                       ylim = c(50, 55)) +
-  ggplot2::theme(
-    legend.position='none'
-  )
 ```
@@
 
-saveRDS(plot_list, here::here(paper_dir, "bbplot_list.rds"))
+saveRDS(plot_list, here::here("data", "objects", "bbplot_list.rds"))
 ```
@@
 ```{r echo=FALSE}
-plot_list <- readRDS(here::here(paper_dir, "bbplot_list.rds"))
+#| label: bubbleplots-Caribou
+plot_list <- readRDS(here::here("data", "objects", "bbplot_list.rds"))
 plot_list["Caribou"]
@@
 ```{r echo=FALSE}
+#| label: bubbleplots-Sandhill
 plot_list["Sandhill crane"]
@@
 ```{r echo=FALSE}
+#| label: bubbleplots-Moose
 plot_list["Moose"]
@@
 ```{r echo=FALSE}
+#| label: bubbleplots-Wolf
 plot_list["Wolf"]
@@
 
-```{r, eval = FALSE}
-#| label: heatmaps
-#| eval: false
-#| echo: false
-
-for (sp in unique(cams_sf_sp_sum$species)){
-
-  cams_sf_sp_spe <- cams_sf_sp |>
-    dplyr::filter(species == sp)
-
-  gg <- base_cut +
-    ggplot2::geom_density2d_filled(
-      data = cams_sf_sp_spe, ggplot2::aes(X, Y),
-      alpha = 0.5,breaks = c(seq(0.001,1,by=0.1),5, 10, 20, 50,100), n=20) +
-    ggplot2::facet_wrap(~month) +
-    ggplot2::coord_equal(xlim = c(-89, -84),
-                         ylim = c(50, 55)) +
-    ggplot2::theme(
-      legend.position='none'
-    )
-
-  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_heatmap.png"))
-}
-```
-
 ### Take Home messages
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: tab-species-table-full
-#| echo: false
 
@@
 
-```{r}
+```{r echo=FALSE}
 #| label: tab-landcover
-#| echo: false
-
-lu_dat |> 
-  dplyr::select(category_code, label) |> 
-  dplyr::filter(category_code != "ND") |> 
-  dplyr::arrange(category_code) |> 
-  gt::gt() |>
-  gt::tab_header(
-    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
-    subtitle = "Adapted from the ontario Far North land cover dataset"
-  ) |>
-  gt::cols_label(
-    label ="Land cover class",
-    category_code = "Abbreviations",
-  ) |> 
-  gt::opt_stylize(style = 1, color = "gray")
+table_cat
 ```
@@
 <!-- The following line inserts a page break  -->
-
 \newpage
@@
 ```
+
+```{r}
+#| label: fig-deployment-poster
+
+# base <- ggmap::ggmap(map) +
+#   ggplot2::labs(x = "Longitude", y = "Latitude") +
+#   # ggplot2::geom_polygon(data = pols, ggplot2::aes(x=X, y=Y), col = 1, fill = NA)
+#   ggplot2::geom_path(data = pols, ggplot2::aes(x=X, y=Y), col = 1, linewidth=0.2)
+# 
+# p <- base +
+#   # ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION),inherit.aes = FALSE,
+#   #                     size = 2.5, data = regions_labels) +
+#   ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOZONE_NA),inherit.aes = FALSE,
+#                       size = 2.5, data = regions_labels) +
+#   ggplot2::geom_point(data = cams_sf_coords |> 
+#                         dplyr::filter(is_retrieved == "Yes"), cex = 1, shape=21, 
+#                       col="black", fill = "red",
+#                       ggplot2::aes(x=X, y=Y), 
+#                       inherit.aes = FALSE) +
+#   ggplot2::labs(fill = "Camera \nRetrieved")
+# p
+```
+
+```{r}
+# anns_wide_animals_joined_mut |> 
+#   ggplot(aes(x=date_time, y=species)) + geom_point()
+# anns_wide_animals_joined_mut |> 
+#   dplyr::filter(species == "Fisher") |> View()
+
+# for_vic <- anns_wide_animals_joined_mut |> 
+#   dplyr::select(id,species,date_time, camera_id) |> 
+#   dplyr::left_join(cams_sf_coords |> dplyr::select(camera_id, plot_id) |> sf::st_drop_geometry(),
+#                    by = "camera_id")
+# tot_camera <- for_vic$camera_id |> unique() |> length()
+# hit_rates <- for_vic |> dplyr::select(species, camera_id) |> 
+#   dplyr::group_by(species) |> dplyr::summarise(n_cameras=length(unique(camera_id))) |> 
+#   dplyr::ungroup() |> dplyr::mutate(rate = n_cameras/tot_camera*100) |> 
+#   dplyr::arrange(dplyr::desc(rate))
+# # det_history_two_weeks <- 
+# 
+# readr::write_csv(for_vic, "../data/derived_data/observations_by_camera_2023Aug15.csv")
+# readr::write_csv(hit_rates, "../data/derived_data/hit_rates_by_species_2023Aug15.csv")
+```
+
+```{r}
+#| label: extract-cats
+#| echo: false
+#| eval: false
+
+# sum(unlist(lapply(extr_list, function(x) sum(x$cats == "NA"))))
+# sum(unlist(lapply(extr_list, function(x) is.na(x$cats))))
+# #  [1] "1"   "2"   "8"   "9"   "12"  "15"  "16"  "17"  "18"  "11"  "13"  "14"  "19"  "20"  "23"  "157" "247" "21"  "22"
+# 
+# pairs(extr_table_sub_tb, col = "dodgerblue")
+# round(cor(extr_table_sub_tb), 2)
+```
+
+```{r}
+# ggplot2::ggplot() +
+#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
+#                               ggplot2::aes(x=X, y=Y, group=plot_id),
+#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
+#   ggplot2::coord_equal() #+ 
+#ggplot2::theme(legend.position = "none")
+
+# base2 <- ggplot2::ggplot() +
+#   ggplot2::theme_bw()
+# for (pol in 1:2){
+#   base2 <- base2 + 
+#     ggplot2::geom_polygon(data = dplyr::filter(pols, L2 == pol),
+#                           ggplot2::aes(x=X, y=Y),
+#                           col = 1, fill = "grey90")
+# }
+# 
+# base2 +
+#   ggplot2::coord_equal(xlim = c(-88.5, -84),
+#                        ylim = c(50, 54)) +
+#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
+#                               ggplot2::aes(x=X, y=Y, group=plot_id),
+#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
+#                               sorted_by_radius = TRUE,
+#                               legend_name = "Species",
+#                               pie_scale = 1.5) +
+#   ggplot2::labs(x="Lat", y="Long")
+```
+
+```{r}
+#| label: jaccard-pcoa
+#| echo: false
+#| eval: false
+
+# anns_wide_animals_table_tb_jac <- vegan::vegdist(anns_wide_animals_table_tb, 
+#                                                  "jac", binary = TRUE)
+# 
+# pa_pcoa <- ape::pcoa(anns_wide_animals_table_tb_jac)
+# 
+# biplot(pa_pcoa, anns_wide_animals_table_tb)
+```
+
+```{r}
+#| label: heatmap
+#| echo: false
+#| eval: false
+
+options(backup_options)
+
+cams_sf_sp_filt <- cams_sf_sp |> 
+  dplyr::filter(species == "Caribou", month == "Jun")
+
+cams_sf_sp_sum_filt <- cams_sf_sp_sum |> 
+  dplyr::filter(species == "Caribou", month == "Jun")
+
+mod <- mgcv::gam(count~s(X,Y), data = cams_sf_sp_sum_filt)
+pred_grid <- 
+  tidyr::expand_grid(X = seq(-89, -84, by = 0.1),
+                     Y = seq(50, 55, by = 0.1))
+preds <- pred_grid |> 
+  dplyr::mutate(count = predict(mod, pred_grid))
+
+# rast <- raster::rasterFromXYZ(preds)
+# raster::plot(rast)
+# rast[rast<0] <- NA
+# raster::plot(rast)
+```
+
+```{r}
+#| label: heatmap2
+#| echo: false
+#| eval: false
+base_cut +
+  ggplot2::geom_density2d_filled(
+    data = cams_sf_sp_filt, ggplot2::aes(X, Y),
+    binwidth = 0.05, alpha = 0.5)
+
+# base_cut +
+#   ggplot2::geom_tile(
+#     data = dplyr::filter(preds, count>=0), ggplot2::aes(X, Y, fill=count),
+#     binwidth = 0.05, alpha = 0.5, inherit.aes = FALSE) +
+#   ggplot2::scale_fill_viridis_c()
+```
+
+```{r}
+#| label: heatmap-facets
+#| echo: false
+#| eval: false
+
+cams_sf_sp_caribou <- cams_sf_sp |> 
+  dplyr::filter(species == "Caribou")
+
+base_cut +
+  ggplot2::geom_density2d_filled(
+    data = cams_sf_sp_caribou, ggplot2::aes(X, Y),
+    alpha = 0.5,breaks = c(seq(0.001,1, by = 0.1),5, 10, 20, 50,100), n = 20) +
+  ggplot2::facet_wrap(~month) +
+  ggplot2::coord_equal(xlim = c(-89, -84),
+                       ylim = c(50, 55)) +
+  ggplot2::theme(
+    legend.position = 'none'
+  )
+```
+
+```{r}
+#| label: heatmaps
+#| eval: false
+#| echo: false
+
+for (sp in unique(cams_sf_sp_sum$species)) {
+
+  cams_sf_sp_spe <- cams_sf_sp |>
+    dplyr::filter(species == sp)
+
+  gg <- base_cut +
+    ggplot2::geom_density2d_filled(
+      data = cams_sf_sp_spe, ggplot2::aes(X, Y),
+      alpha = 0.5,breaks = c(seq(0.001,1,by = 0.1),5, 10, 20, 50,100), n = 20) +
+    ggplot2::facet_wrap(~month) +
+    ggplot2::coord_equal(xlim = c(-89, -84),
+                         ylim = c(50, 55)) +
+    ggplot2::theme(
+      legend.position = 'none'
+    )
+
+  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_heatmap.png"))
+}
+```
 -->

### 2024-02-01 analysis/documents/1-data-preparation.qmd
add contents

---
title: "Data Processing"
author:
  - name: Valentin Lucet
    email: valentin.lucet@gmail.com
    orcid: 0000-0003-0268-818X
    affiliation: 
      - ref: wlu
affiliations:
  - id: wlu
    name: Wilfrid Laurier University
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    extra_dependencies: ["float"]
format-links: false
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-dpi: 600
# bibliography: references.bib
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
backup_options <- options()
options(scipen = 1, digits = 2)
do_eval <- FALSE
```

In this script, we format the data needed for further analyses.

#### Process detections

```{r eval=do_eval}
#| label: process-detections

detections <- rofcamtrap::process_detections(
  input_dir = here::here("2_LabelStudio", "0_inputs"))

saveRDS(detections,  here::here("data", "objects" ,"detections.rds"))
```

#### Process annotations

```{r eval=do_eval}
#| label: process-annotationa

anns_wide <- rofcamtrap::process_annotations(
   input_dir = here::here("2_LabelStudio", "2_outputs_processed/"))

saveRDS(anns_wide, here::here("data", "objects", "annotations_wide_treated.rds"))

# Filter for non empty images
anns_wide_noempty <- anns_wide |>
  dplyr::filter(!is.na(to_name))

saveRDS(anns_wide_noempty, here::here("data", "objects", "annotations_wide_noempty.rds"))
```

#### Process exif data

```{r eval=do_eval}
#| label: process-exif

exif <- readr::read_csv(here::here("data", "tabular", "all_exif.csv")) |>
  janitor::clean_names() |>
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
                  unlist()) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))

readr::write_csv(exif, here::here("data", "tabular", "exif_processed.csv"))
```

#### Load data

```{r eval=do_eval}
#| label: load-data
detections <- readRDS(here::here("data", "objects", "detections.rds"))
anns_wide <- readRDS(here::here("data", "objects", "annotations_wide_treated.rds"))
anns_wide_noempty <- readRDS(here::here("data", "objects", "annotations_wide_noempty.rds"))
exif <- readr::read_csv(here::here("data", "tabular", "exif_processed.csv"))
```

#### Process spatial information

```{r eval=do_eval}
#| label: process-cams-sf

RoF_CameraDeploymentLog_WLU <- 
  readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx")) |> 
  janitor::clean_names()

cams_sf <- RoF_CameraDeploymentLog_WLU |> 
  dplyr::rename(longitude = x, latitude = y) |> 
  dplyr::filter(!is.na(latitude)) |> 
  dplyr::filter(!is.na(longitude)) |> 
  sf::st_as_sf(coords = c("longitude","latitude"), crs = "EPSG:4326", remove = F) |> 
  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))

coords <- cams_sf |> 
  sf::st_coordinates() |> 
  as.data.frame()

cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)

saveRDS(cams_sf_coords, here::here("data", "objects", "cams_sf_coords.rds"))
```

```{r eval=do_eval}
#| label: process-zones

zones_filtered <- 
  sf::st_read(here::here("data-raw","spatial", "ecozones", "EcoZone" ,"EcoZone.shp"),
              quiet = TRUE) |> 
  sf::st_transform(crs = sf::st_crs(cams_sf)) |>
  sf::st_filter(cams_sf_coords)

saveRDS(zones_filtered, here::here("data", "objects", "zones_filtered.rds"))
```

```{r eval=do_eval}
#| label: process-pols

pols <- sf::st_coordinates(zones_filtered) |> 
  as.data.frame()

saveRDS(pols, here::here("data", "objects", "zones_pols.rds"))
```

```{r eval=do_eval}
#| label: process-map

map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
                            zoom = 5, maptype = "stamen_terrain")

base <- ggmap::ggmap(map) +
  ggplot2::labs(x = "Longitude", y = "Latitude") +
  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)

saveRDS(base, here::here("data", "objects", "base_map.rds"))
```

#### Process animal annotations

```{r eval=do_eval}
#| label: process-animals

# Only annotations for animals
# TODO investigate id of cut length 1

anns_wide_animals <- anns_wide_noempty |>
  dplyr::filter(species != 'Human') |> 
  dplyr::filter(species != 'Vehicle') |> 
  dplyr::mutate(id_cut_length = purrr::map_dbl(id , \(x) {
    x |> stringr::str_split("_") |> purrr::map_dbl(~length(.x))
  })) |> 
  dplyr::filter(id_cut_length > 1) |> 
  dplyr::mutate(camera_id = purrr::map_chr(id , \(x) {
    x |>
      stringr::str_split("_") |>
      purrr::map(~.x[[2]]) |>
      unlist()
  }))

# Join to EXIF
# TODO streamline exif so you dont have to join

anns_wide_animals_joined <- anns_wide_animals |>
  dplyr::left_join(exif, by = c("source_file", "image_id"))

stopifnot(dim(anns_wide_animals)[1] ==  dim(anns_wide_animals_joined)[1])
stopifnot((anns_wide_animals_joined$source_file |> is.na() |> sum()) == 0)

anns_wide_animals_joined_mut <- anns_wide_animals_joined |> 
  dplyr::filter(species != "Unknown") |>
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

saveRDS(anns_wide_animals_joined_mut, here::here("data", "objects", "annotations_animals.rds"))

## Spare code

# Unknowns and bird breakdown
# anns_wide_animals |> dplyr::filter(species == "Unknown") |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort() 
# anns_wide_animals |> dplyr::filter(stringr::str_detect(species,"Bird")) |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort()

# anns_wide_animals_joined_dets <- anns_wide_animals |> 
#   dplyr::right_join(dets_images, by = "image_id")

# anns_wide_animals_joined_mut$species |> unique() |> grepl(pattern='sp', fixed = T) |> sum()
```

#### Process deployments from exif data

```{r eval=do_eval}
#| label: process-deployments

camera_days <- exif |> 
  dplyr::group_by(maker_notes_user_label) |> 
  dplyr::summarise(camera_first_time = min(exif_date_time_original)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time, 
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) |> 
  dplyr::select(camera_id = maker_notes_user_label, 
                camera_first_day = date_time)

camera_regions <- cams_sf_coords |> 
  sf::st_join(zones_filtered) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(camera_id, ECOZONE_NA)

saveRDS(camera_days,  here::here("data", "objects" ,"camera_days.rds"))
saveRDS(camera_regions,  here::here("data", "objects" ,"camera_regions.rds"))
```

#### Process presence absence data

First, absolute occurrence.

```{r eval=do_eval}
#| label: process-pa

anns_wide_animals_table <- anns_wide_animals_joined_mut |>
  dplyr::select(camera_id, species) |> 
  dplyr::filter(species != "Unknown") |>
  table() 
anns_wide_animals_table <- anns_wide_animals_table > 0
anns_wide_animals_table <- anns_wide_animals_table + 0
anns_wide_animals_table <- anns_wide_animals_table |> 
  as.data.frame() |> 
  tibble::rownames_to_column("camera_id") |> 
  dplyr::arrange(camera_id)
anns_wide_animals_table_tb <- anns_wide_animals_table |> 
  dplyr::select(-camera_id)
rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id

saveRDS(anns_wide_animals_table_tb,  here::here("data", "objects" ,"anns_wide_animals_table_tb.rds"))
```

#### Process land cover data

1 km buffers

```{r eval=do_eval}
#| label: process-land-cover
lu_16 <- raster::raster(here::here("data-raw", "spatial", "land_use", "Class", "FarNorth_LandCover_Class_UTM16.tif"))
lu_17 <- raster::raster(here::here("data-raw", "spatial", "land_use", "Class", "FarNorth_LandCover_Class_UTM17.tif"))

lu_dat <- readr::read_csv(here::here("data-raw", "spatial", "land_use", "attr_table_northen_ont_lc.txt")) |> 
  dplyr::mutate(cats = as.factor(code))

cams_sf_buffers_16 <- cams_sf_coords |> 
  sf::st_transform(sf::st_crs(lu_16)) |> 
  sf::st_buffer(dist = 1000)

cams_sf_buffers_17 <- cams_sf_coords |> 
  sf::st_transform(sf::st_crs(lu_17)) |> 
  sf::st_buffer(dist = 1000)

extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
                                        progress = FALSE)
extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
                                        progress = FALSE)

ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) |> unlist())
ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) |> unlist())

stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))

extr_16[ind_16] <- extr_17[ind_16]

extr <- extr_16

extr_list <- extr |> 
  lapply(FUN = function(x){
    the_tab <- table(x$value)
    props <- the_tab/sum(the_tab)
    cats <- names(the_tab)
    df <- tibble::tibble(cats, props)
    return(df)
  })
names(extr_list) <- cams_sf_buffers_16$camera_id

extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") |> 
  dplyr::mutate(props = as.double(props)) |> 
  dplyr::left_join(lu_dat, by = "cats") |> 
  dplyr::filter(!is.na(category_code)) |> 
  dplyr::select(camera_id,category_code, props) |> 
  tidyr::pivot_wider(names_from = "category_code", values_from = "props") |> 
  as.data.frame()
extr_table[is.na(extr_table)] <- 0

# Combine the categories
extr_table_simple <- extr_table |> 
  dplyr::mutate(WAT = WAT + XWAT,
                MixTRE = MixTRE + DecTRE,
                Dist  = TrOrSHr + NSWood, 
                SWA = ThSWA + ConSWA, 
                Anth = URB + MIN + BED) |> 
  dplyr::select(-XWAT,
                -DecTRE,
                -TrOrSHr, -NSWood, 
                -ThSWA, -ConSWA, 
                -URB, -MIN, -BED)

extr_table_sub <- extr_table_simple |> 
  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) |> 
  dplyr::arrange(camera_id)
extr_table_sub_tb <- extr_table_sub |> 
  dplyr::select(-camera_id)
rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
extr_table_sub_tb <- extr_table_sub_tb[, colSums(extr_table_sub_tb) != 0]

stopifnot(all(rownames(anns_wide_animals_table_tb) == 
                rownames(extr_table_sub_tb)))

saveRDS(extr_table_sub_tb,  here::here("data", "objects" ,"extr_table_sub_tb.rds"))

table_cat <- lu_dat |> 
  dplyr::select(category_code, label) |> 
  dplyr::filter(category_code != "ND") |> 
  dplyr::arrange(category_code) |> 
  gt::gt() |>
  gt::tab_header(
    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
    subtitle = "Adapted from the ontario Far North land cover dataset"
  ) |>
  gt::cols_label(
    label = "Land cover class",
    category_code = "Abbreviations",
  ) |> 
  gt::opt_stylize(style = 1, color = "gray")

saveRDS(table_cat,  here::here("data", "objects" ,"table_cat.rds"))
```

Then, based on events, assuming day as an event window.

```{r eval=do_eval}
#| label: events-Plots

events_days <- anns_wide_animals_joined_mut |> 
  dplyr::select(plot_id, species, month, ymd) |> 
  dplyr::distinct()

events_summarised <- events_days |> 
  dplyr::group_by(plot_id, species) |>
  dplyr::summarise(day_count = dplyr::n()) |> 
  dplyr::ungroup()

events_tb <- events_summarised |> 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
  dplyr::arrange(plot_id) |>
  tibble::column_to_rownames("plot_id")

events_tb[is.na(events_tb)] <- 0
events_tb_filt <- events_tb[,colSums(events_tb) > 50]
tmp_col <- ncol(events_tb_filt)

events_tb_filt_with_col <- events_tb_filt |> 
  tibble::rownames_to_column("plot_id")

saveRDS(events_tb_filt_with_col,  here::here("data", "objects" ,"events_tb_filt_with_col.rds"))
```

Same, but using camera instead of plot id.

```{r eval=do_eval}
#| label: events-Cams

events_cam_days <- anns_wide_animals_joined_mut |> 
  dplyr::select(camera_id, species, month, ymd) |> 
  dplyr::distinct()

saveRDS(events_cam_days,  here::here("data", "objects" ,"events_cam_days.rds"))

events_cam <- events_cam_days |> 
  dplyr::group_by(camera_id, species) |>
  dplyr::summarise(day_count = dplyr::n()) |> 
  dplyr::ungroup()

events_tb_cam <- events_cam |> 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
  dplyr::arrange(camera_id) |>
  tibble::column_to_rownames("camera_id")

events_tb_cam[is.na(events_tb_cam)] <- 0
# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
# tmp_col <- ncol(events_tb_cam)

saveRDS(events_tb_cam,  here::here("data", "objects" ,"events_tb_cam.rds"))
```

#### Process scientific names

```{r eval=do_eval}
# sci_names <- taxize::comm2sci(sp_tab$species)
sci_names <- readRDS(here::here("data", "objects", "sci_names.rds")) |> 
  unlist() |> tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column() |> 
  dplyr::rename(Scientific_name = value,
                species = rowname) |> 
  tibble::add_row(Scientific_name = "Ursus americanus", species = "Black bear") |> 
  tibble::add_row(Scientific_name = "Perisoreus canadensis", species = "Canada Jay")

saveRDS(sci_names,here::here("data", "objects" , "sci_names_processed.rds"))
```

### 2024-02-01 analysis/documents/2-data-exploration.qmd
add contents

---
title: "Data Exploration"
author:
  - name: Valentin Lucet
    email: valentin.lucet@gmail.com
    orcid: 0000-0003-0268-818X
    affiliation: 
      - ref: wlu
affiliations:
  - id: wlu
    name: Wilfrid Laurier University
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    extra_dependencies: ["float"]
format-links: false
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-dpi: 600
# bibliography: references.bib
---

## Introduction

In this exploration of the data, we will make use of the code provided in "[Introduction to Camera Trap Data Management and Analysis in R](https://wildcolab.github.io/Introduction-to-Camera-Trap-Data-Management-and-Analysis-in-R/index.html)", hereafter refered to as ICTDMA. I adapt the code in this document into functions which should later serve as a basis for a R package.

```{r, include=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
backup_options <- options()
options(scipen = 1, digits = 2)
source("functions.R")
```

### Data loading and conversion

We load the annotation data and convert it into the proper format, which is the "Wildlife Insights" format. As described in ICTDMA:

" Their standard format is composed of four different elements:

-   Project data: a dataframe containing key information about the project itself, e.g. how the cameras were deployed and what the target features were.

-   Image data: a dataframe containing all of the information contained within each image. This information is typically added by humans, but increasing we are using artificial intelligence to speed up this process.

-   Deployment data: a dataframe listing the activity of the camera traps involved in your study, and any issues encountered during deployments which may influence their analysis

-   Camera data: a dataframe all the cameras deployed in the project "

Ideally, we would use the CamtrapDP format, but we shall circle back to that in the future.

#### Project data

We will let that aside for the moment as it is not relevant to the specific task as hand, but here are the fields this data format expects:

```{r, eval=FALSE}
 [1] "project_id"                          "project_name"                        "project_short_name"                 
 [4] "project_objectives"                  "project_species"                     "project_species_individual"         
 [7] "project_sensor_layout"               "project_sensor_layout_targeted_type" "project_bait_use"                   
[10] "project_bait_type"                   "project_stratification"              "project_stratification_type"        
[13] "project_sensor_method"               "project_individual_animals"          "project_blank_images"               
[16] "project_sensor_cluster"              "project_admin"                       "project_admin_email"                
[19] "project_admin_organization"          "country_code"                        "embargo"                            
[22] "initiative_id"                       "metadata_license"                    "image_license"                      
[25] "data_citation"                       "count_optional"                      "project_type
```

#### Image data

The image data is the annotations data, which we load:

```{r}
image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
  dplyr::select(id, source_file_with_dep,
                image_id, species:comment, camera_id,
                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
  dplyr::rename(file_path = source_file_with_dep)
head(image_data)
```

And then reformat and rename:

#### Deployment data

This is

#### Camera data

### 2024-02-01 reports/GnC_report/paper/paper.qmd
update all
correction author list

@@
   - name: Cheryl A. Johnson
-    affiliation:
-      - ref: eccc
-  - name: Mathieu Leblond
     affiliation:
@@
 
-  ggplot2::ggsave(gg, filename = paste0("figures/",sp,"_bplot.png"))
+  ggplot2::ggsave(gg, filename = here::here("figures",paste0(sp,"_bplot.png")))
 }
@@
 ```{r}
+#| label: git
+
 # what commit is this file at? 

### 2024-02-01 analysis/documents/1-data-preparation.qmd
update all
add plot

@@
 RoF_CameraDeploymentLog_WLU <- 
-  readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx")) |> 
-  janitor::clean_names()
+  suppressWarnings(readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx"),
+                                      guess_max = 10)) |>
+  janitor::clean_names() |> 
+  dplyr::mutate(date_retrieved = as.Date.character(date_retrieved))
 
@@
 map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
-                            zoom = 5, maptype = "stamen_terrain")
+                            zoom = 3, maptype = "stamen_terrain")
 

### 2024-02-01 analysis/documents/2-data-exploration.qmd
update
add plot

@@
 
-In this exploration of the data, we will make use of the code provided in "[Introduction to Camera Trap Data Management and Analysis in R](https://wildcolab.github.io/Introduction-to-Camera-Trap-Data-Management-and-Analysis-in-R/index.html)", hereafter refered to as ICTDMA.
-I adapt the code in this document into functions which should later serve as a basis for a R package.
+In this exploration of the data, we will loosely follow the protocol in "[Introduction to Camera Trap Data Management and Analysis in R](https://wildcolab.github.io/Introduction-to-Camera-Trap-Data-Management-and-Analysis-in-R/index.html)", hereafter referred to as ICTDMA.
 
@@
 ```{r, eval=FALSE}
+#| label: project
  [1] "project_id"                          "project_name"                        "project_short_name"                 
@@
 ```{r}
+#| label: image
 image_data <- readRDS(here::here("data", "objects", "annotations_animals.rds")) |> 
@@
 
-And then reformat and rename:
+#### Camera and deployment data
 
-#### Deployment data
+The camera and deployment data are currently together as each camera was only deployed once
 
-This is
+```{r}
+#| label: camdep
+cam_dep_data <- readRDS(here::here("data", "objects", "cams_sf_coords.rds")) |> 
+  sf::st_drop_geometry() |> 
+  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude)
+stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
+```
 
-#### Camera data
+#### Deployment plot
+
+```{r, fig.height=20, fig.width=8}
+#| label: camdep-plot
+cam_dep_data |> 
+  dplyr::mutate(date_deployed = as.Date(date_deployed),
+                date_retrieved = as.Date(date_retrieved)) |> 
+  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
+  dplyr::mutate(status = ifelse(is.na(status), "invalid", status)) |> 
+  ggplot2::ggplot() +
+  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = camera_id, 
+                                     xend = date_retrieved, yend = camera_id,
+                                     color = status)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id, 
+                                     color = status)) + 
+    ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id, 
+                                     color = status)) + 
+  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
+  ggplot2::theme_bw()
+```

### 2024-02-08 analysis/documents/2-data-exploration.qmd
update,fixing issues

@@
                 maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
-  dplyr::rename(file_path = source_file_with_dep)
+  dplyr::mutate(site_id = sapply(image_id, \(x) strsplit(x, "_")[[1]][1]),
+                date_time = lubridate::as_date(date_time)) |> 
+  dplyr::rename(file_path = source_file_with_dep) |> 
+  dplyr::filter(!is.na(date_time))
+
+# Fixing error
+# image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
+image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
+image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2024_Aug_19"
+
 head(image_data)
@@
   sf::st_drop_geometry() |> 
-  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude)
+  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude) |> 
+  dplyr::mutate(date_deployed = lubridate::as_date(date_deployed),
+                date_retrieved = lubridate::as_date(date_retrieved)) 
 stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
@@
 
+By cameras
+
 ```{r, fig.height=20, fig.width=8}
 #| label: camdep-plot
-cam_dep_data |> 
-  dplyr::mutate(date_deployed = as.Date(date_deployed),
-                date_retrieved = as.Date(date_retrieved)) |> 
+
+cam_dep_data_with_status <- cam_dep_data |> 
   dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
-  dplyr::mutate(status = ifelse(is.na(status), "invalid", status)) |> 
+  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))
+
+cam_dep_data_with_status |> 
   ggplot2::ggplot() +
@@
 ```
+
+By sites
+
+```{r, fig.height=40, fig.width=8}
+cam_dep_data_with_status |> 
+  ggplot2::ggplot() +
+  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = site_id,
+                                     xend = date_retrieved, yend = site_id,
+                                     color = status), alpha = 0.15, linewidth = 3) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
+                                   color = status)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
+                                   color = status)) +
+  ggplot2::geom_text(ggplot2::aes(x = date_deployed-40, y = site_id,
+                                  label = camera_id,
+                                  color = status), size = 3) +
+  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
+  ggplot2::theme_bw()
+```
+
+With image density
+
+```{r, fig.height=40, fig.width=8}
+cam_dep_data_with_status |> 
+  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(site_id, 
+                                                              date_deployed,
+                                                              decreasing = TRUE))) +
+  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = site_id,
+                                     color = status), alpha = 0.15, linewidth = 3) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
+                                   color = status)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
+                                   color = status)) +
+
+  ggplot2::geom_point(data = image_data,
+                      ggplot2::aes(x = date_time, y = site_id), 
+                      color = "firebrick", size = 0.5) +
+
+  ggplot2::geom_text(ggplot2::aes(x = date_deployed-60, y = site_id,
+                                  label = camera_id,
+                                  color = status), size = 3) +
+  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
+  ggplot2::theme_bw()
+```
+
+Sorted by deployment date
+
+```{r, fig.height=40, fig.width=8}
+cam_dep_data_with_status |> 
+  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id, 
+                                                              date_deployed,
+                                                              decreasing = TRUE))) +
+  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id,
+                                     color = status), alpha = 0.15, linewidth = 3) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id,
+                                   color = status)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id,
+                                   color = status)) +
+
+  ggplot2::geom_point(data = image_data,
+                      ggplot2::aes(x = date_time, y = camera_id), 
+                      color = "firebrick", size = 0.5) +
+
+  ggplot2::geom_text(ggplot2::aes(x = date_deployed-60, y = camera_id,
+                                  label = site_id,
+                                  color = status), size = 3) +
+  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
+  ggplot2::theme_bw()
+```

### 2024-02-08 analysis/documents/2-data-exploration.qmd
update

@@
 
-By cameras
-
-```{r, fig.height=20, fig.width=8}
-#| label: camdep-plot
-
-cam_dep_data_with_status <- cam_dep_data |> 
-  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |> 
-  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))
-
-cam_dep_data_with_status |> 
-  ggplot2::ggplot() +
-  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = camera_id, 
-                                     xend = date_retrieved, yend = camera_id,
-                                     color = status)) +
-  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id, 
-                                   color = status)) + 
-  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id, 
-                                   color = status)) + 
-  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
-  ggplot2::theme_bw()
-```
-
-By sites
-
-```{r, fig.height=40, fig.width=8}
-cam_dep_data_with_status |> 
-  ggplot2::ggplot() +
-  ggplot2::geom_segment(ggplot2::aes(x = date_deployed, y = site_id,
-                                     xend = date_retrieved, yend = site_id,
-                                     color = status), alpha = 0.15, linewidth = 3) +
-  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
-                                   color = status)) +
-  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
-                                   color = status)) +
-  ggplot2::geom_text(ggplot2::aes(x = date_deployed-40, y = site_id,
-                                  label = camera_id,
-                                  color = status), size = 3) +
-  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
-  ggplot2::theme_bw()
-```
-
-With image density
-
-```{r, fig.height=40, fig.width=8}
-cam_dep_data_with_status |> 
-  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(site_id, 
-                                                              date_deployed,
-                                                              decreasing = TRUE))) +
-  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = site_id,
-                                     color = status), alpha = 0.15, linewidth = 3) +
-  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = site_id,
-                                   color = status)) +
-  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = site_id,
-                                   color = status)) +
-
-  ggplot2::geom_point(data = image_data,
-                      ggplot2::aes(x = date_time, y = site_id), 
-                      color = "firebrick", size = 0.5) +
-
-  ggplot2::geom_text(ggplot2::aes(x = date_deployed-60, y = site_id,
-                                  label = camera_id,
-                                  color = status), size = 3) +
-  ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
-  ggplot2::theme_bw()
-```
-
-Sorted by deployment date
+Cameras, Sorted by deployment date, with associated site labelled.
 

### 2024-02-19 R/renaming.R
base code for renaming

@@ -1,0 +1,79 @@
+
+rename_images <- function(images_from,
+                          images_to,
+                          deployment_code,
+                          file_ext = "\\.(jpg|jpeg|JPG|JPEG)$") {
+
+  # Get all location folders in this deployment, some will have _repeat in them,
+  # due to the way things were done before
+  locations_all <- list.dirs(images_from, recursive = FALSE, full.names = FALSE)
+  # Remove those repeat folders from the list
+  locations <- locations_all[!grepl("repeat", locations_all)]
+
+  # For all locations, proceed to the copy and renaming
+  # TODO improve performance
+  for (loc in locations) {
+
+    message(paste("Proccessing folder", loc))
+
+    # Path to location dir
+    root_path <- file.path(images_from, loc)
+    # List all images
+    all_images <- list.files(root_path, recursive = TRUE, pattern = file_ext)
+
+    # For all images, make the appropriate file name
+    pb <- progress::progress_bar$new(total = length(all_images))
+    for (image in all_images) {
+
+      # Reconstruct image path
+      old_path <- file.path(root_path, image)
+
+      # Split at appropriate file separator
+      image_split <- strsplit(image, .Platform$file.sep) |>
+        unlist()
+
+      # Filter the path for useless sections, removing DCIM first
+      image_split_filtered <-  image_split[!(image_split %in% c("DCIM"))] |>
+          # Remove useless suffix for overflow folder and file name
+          gsub(pattern = "RECNX", replacement = "") |>
+          gsub(pattern = "RCNX", replacement = "") |>
+          gsub(pattern = file_ext, replacement = "")
+
+      # Get datetime from exif data
+      exif <- exifr::read_exif(old_path)
+      datetime <- exif$DateTimeOriginal |>
+        gsub(pattern = ":", replacement = "_") |>
+        gsub(pattern = " ", replacement = "_") |>
+        gsub(pattern = "Wildlife", replacement = "W") |>
+        gsub(pattern = "NonWildlife", replacement = "W")
+
+      # Construct the new file name...
+      new_name <- paste0(
+        paste0(c(deployment_code, loc,
+                 paste0(image_split_filtered, collapse = "_"), datetime),
+               collapse = "_"),
+        ".JPG")
+
+      # ...and new file path
+      new_path <- file.path(images_to, deployment_code,
+                            paste0(c(deployment_code, loc,image_split_filtered[1:2]), collapse = "_"),
[…]

### 2024-02-19 scripts/R/file_management/rename_images.R

@@ -1,0 +1,10 @@
+
+rename_images(images_from = "/media/vlucet/TrailCamST1/TrailCamStorage",
+              images_to = "/media/vlucet/TrailCamST1/renamed",
+              deployment_code = "TC1",
+              file_ext = "\\.(jpg|jpeg|JPG|JPEG)$")
+
+rename_images(images_from = "/media/vlucet/TrailCamST1/TrailCamStorage_2/",
+              images_to = "/media/vlucet/TrailCamST1/renamed",
+              deployment_code = "TC2",
+              file_ext = "\\.(jpg|jpeg|JPG|JPEG)$")

### 2024-02-19 analysis/documents/2-data-exploration.qmd
add lines

@@
 ---
-title: "Data Exploration"
+title: "Data Exploration and Correction"
 author:
@@
 # Fixing error 
+# TODO make it more explicit
 # image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
@@
 ```
+
+#### Adjust for daylight savings
+
+November 6th 2022 to March 12th 2023
+
+November 5th 2023 to March 10th 2024

### 2024-02-19 R/renaming.R
continue coding renaming

@@ -2,13 +2,21 @@
 rename_images <- function(images_from,
                           images_to,
                           deployment_code,
+                          locations_subset = c(),
                           file_ext = "\\.(jpg|jpeg|JPG|JPEG)$") {
 
-  # Get all location folders in this deployment, some will have _repeat in them,
-  # due to the way things were done before
+  # Get all location folders in this deployment, some will have _repeat or _crop
+  # in them, due to the way things were done before
   locations_all <- list.dirs(images_from, recursive = FALSE, full.names = FALSE)
   # Remove those repeat folders from the list
-  locations <- locations_all[!grepl("repeat", locations_all)]
+  locations <- locations_all[!(grepl("repeat", locations_all) |
+                                 grepl("crop", locations_all))]
+
+  # Subset if necessary
+  if (length(locations_subset) > 0) {
+    locations <- locations[locations %in% locations_subset]
+  }
+  print(locations)
 
   # For all locations, proceed to the copy and renaming
   # TODO improve performance
@@ -20,6 +28,9 @@ rename_images <- function(images_from,
     root_path <- file.path(images_from, loc)
     # List all images
     all_images <- list.files(root_path, recursive = TRUE, pattern = file_ext)
+
+    print(all_images)
+    stop()
 
     # For all images, make the appropriate file name
     pb <- progress::progress_bar$new(total = length(all_images))

### 2024-02-19 scripts/R/file_management/rename_images.R

@@ -1,9 +1,14 @@
 
+# First deployment
 rename_images(images_from = "/media/vlucet/TrailCamST1/TrailCamStorage",
               images_to = "/media/vlucet/TrailCamST1/renamed",
               deployment_code = "TC1",
+              locations_subset = c(
+                "Q647", "P216", "P100", "P087", "P058", "P045", "P293", "P080", "P224"
+                ),  # "P352", "P132", "P128", "P121", "P030"
               file_ext = "\\.(jpg|jpeg|JPG|JPEG)$")
 
+# Second deployment
 rename_images(images_from = "/media/vlucet/TrailCamST1/TrailCamStorage_2/",
               images_to = "/media/vlucet/TrailCamST1/renamed",
               deployment_code = "TC2",

### 2024-02-19 R/renaming.R
remove early stop

@@ -29,9 +29,6 @@ rename_images <- function(images_from,
     # List all images
     all_images <- list.files(root_path, recursive = TRUE, pattern = file_ext)
 
-    print(all_images)
-    stop()
-
     # For all images, make the appropriate file name
     pb <- progress::progress_bar$new(total = length(all_images))
     for (image in all_images) {

### 2024-04-08 analysis/documents/2-data-exploration.qmd
fix date

@@
 image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
-image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2024_Aug_19"
+image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"
 

### 2024-04-08 analysis/documents/1-data-preparation.qmd
fix base map zoom, add loading of data set

@@
 saveRDS(anns_wide_noempty, here::here("data", "objects", "annotations_wide_noempty.rds"))
+
+# anns_wide_noempty <- readRDS(here::here("data", "objects", "annotations_wide_noempty.rds"))
+readr::write_csv(anns_wide_noempty, here::here("data", "tabular", "annotations_wide_noempty.csv"))
 ```
@@
 map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
-                            zoom = 3, maptype = "stamen_terrain")
+                            zoom = 5, maptype = "stamen_terrain")
 

### 2024-04-08 reports/GnC_report/paper/paper.qmd
quick update
working through comments

@@
 ---
-title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North"
-subtitle: "Prepared for: Canadian Wildlife Service, Canadian Forest Service, and Wilfrid Laurier University"
+title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-occurrence Where the Hudsons's Bay Lowlands meets the Ontario Shield"
+subtitle: "Prepared for: Canadian Wildlife Service, Canadian Forest Service, and Wilfrid Laurier University - 2024 update"
 author:
@@
 format:
-  html: default
+  # html: default
   pdf:
@@
     number-sections: true
+    number-offset: 1
     colorlinks: true
@@
 The extent to which these methods perform for northern regions is not well established.
-Using an array of cameras deployed across an heterogeneous sampling area straddling two Northern Ontario ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
-We’ve recently developed an open-source workflow for camera trap image identification, access, and storage.
-We continue to use this workflow for ongoing image identification while investigating requirements to develop a machine learning image classifier.
+Using an array of cameras deployed across an heterogeneous sampling area straddling two Northern Ontario ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.In 2023 we developed an open source workflow for camera trap image identification, access, and storage.
+We have updated it in 2024 for efficiency.
+We also continue to investigate the requirements to develop a machine learning image classifier.
+{{< pagebreak >}}
 
@@
 
-## Introduction
+{{< pagebreak >}}
+
+# Introduction
 
@@
 We aim to evaluate the viability of deep neural networks for camera trap surveys currently taking place in Northern Ontario’s heterogeneous landscapes, including the Hudson Bay Lowlands and Ontario Shield ecozones, as variables like weather, seasonality, vegetation type, and species characteristics can impact the performance of these methods.
-The extent to which these methods perform for northern regions is not well established.
+The extent to which these methods perform for northern regions and northern specie is not well established.
 There are currently no available machine learning models for identifying species from images in northern regions and developing one would be beneficial as a monitoring and modeling tool for northern boreal ecosystems.
@@
 
-## Methods
+## Update for the 2024 progress report
+
+Since the last progress report (2023 report), we flag the following improvements and changes, which are detailed in the new sections of this reports.
+
+- The number of fully processed images was doubled from about 50000 to more than 100000, keeping with the rate of progress of the first year of tagging.
+- We made a first attempt at at using machine learning methods to train a species classifier using about 23000 images as the basis dataset, which produced encouraging results.
+- After reevaluation of the ability of our current worlflow and platform to scale to multiple years of data, we have transitioned the image tagging workflow to [WildTrax](https://wildtrax.ca/).
+- This report repeatedly mentions the AI model MegaDetector.
+Although this model now falls under the toolset known as PytorchWildlife, as of writing in April 2024, it is still commonly known as MegaDetector by the camera trapping community.
+
+### Outputs of note
+
+1. Poster submitted and presented at 2023 conference of the Ontario chapter of The Wildlife Society (OCTWS).
+2. Poster submitted and displayed at the Mushkegowuk council 2023 Research Meeting.
+3. Application to participate in the Workshop on Computer Vision Methods for Ecology at the Caltech Resnick Sustainability Institute. Rejectec in 2023, reapplied in 2024.
+
+# Methods
 
@@
 base_map <- readRDS(here::here("data", "objects", "base_map.rds"))
+ont_base_map <- readRDS(here::here("data", "objects", "ont_base_map.rds"))
 camera_days <- readRDS(here::here("data", "objects" ,"camera_days.rds"))
@@
 
-### Camera deployment and setup
+## Camera deployment methods
 
@@
 
+TODO: add more from Sam
+
+## Deployment summary
+
 Briefly, 5 ARUs are deployed per site, with 100 m between ARUs.
-Up to 4 cameras were paired with ARUs at each of these sites, with at least one of these cameras placed at a random ARU site and one other prioritized to focus on a game trail or open area to increase detection probability (in total, we have 43 plots with 4 cameras, 5 plots with 3 cameras and 3 plots with 2 cameras). @fig-deployment shows the location of the clustered cameras and their current retrieval status.
-170 cameras were retrieved in September 2022, totaling a deployment spanning 5-7 months.
-5 additional cameras were only retrieved in March 2023 due to logistical and inaccessibility issues.
-There remains 18 cameras in the field to be retrieved in 2023.
-
-The work conducted under this project focuses on quantifying and developing an open-source workflow to quantifying the wildlife species detected in those images that have been retrieved to date.
+Up to 4 cameras were paired with ARUs at each of these sites, with at least one of these cameras placed at a random ARU site and one other prioritized to focus on a game trail or open area to increase detection probability.
+Weather restrictions and field conditions limited access to some sites and resulting in some variation in this design.
+In total we have 43 plots with 4 cameras, 5 plots with 3 cameras and 3 plots with 2 cameras. @fig-deployment shows the location of the clustered cameras and their current retrieval status.
+54 cameras were retrieved in May 2022 and 116 in September 2022.
+6 additional cameras were retrieved in March 2023 due to logistical and accessibility issues.
+Another 77 cameras were retrieved in Mar 2023.
+The last 18 cameras were retrieved in September 2023 and are now being processed.
+
+The work conducted under this project focuses on developing an open-source workflow to quantify the wildlife species detected in those images that have been retrieved to date. This workflow is also being designed to include anticipated additional images from future deployments on this, and potentially other, projects.
 
@@
 regions_labels <- zones_filtered |> 
+  dplyr::mutate(ECOZONE_NA = c("Hudson Bay\nLowlands", "Ontario\nShield")) |> 
   sf::st_drop_geometry() |> 
-  dplyr::bind_cols(data.frame(x = c(-84.5, -88.5),
+  dplyr::bind_cols(data.frame(x = c(-84.5, -88.75),
                               y = c(52, 52.6)))
@@
 #| label: fig-deployment
-#| fig-cap: "Camera deployment status for the 193 cameras, 170 of which were retrieved in September 2022. Labels for the 3 ecoregions spanned by the sampling area are included."
+#| fig-cap: "Left: Map of the province of Ontario with ecoregion and inset map delimitation.
+Right: Inset map of camera deployment locations as wellas deployment and retrieval dates for the 227 sites.
+Labels for the 2 ecoregions spanned by the sampling area are included."
+#| fig.width: 8
+#| fig.height: 6
 
@@
   ggplot2::geom_point(data = cams_sf_coords,
-                      shape = 24, col = "black", fill = "red",
-                      ggplot2::aes(x = X, y = Y), 
-                      inherit.aes = FALSE)
+                      shape = 24, col = "black", #fill = "red",
+                      ggplot2::aes(x = X, y = Y, fill = deployment_group), 
+                      inherit.aes = FALSE) +
+  ggplot2::labs(fill = "Deployment - retrieval") #+
+  # ggplot2::theme_void() +
+  # ggplot2::theme(
+  #   # legend.justification defines the edge of the legend that the legend.position coordinates refer to
+  #   legend.justification = c(0, 1),
+  #   # Set the legend flush with the left side of the plot, and just slightly below the top of the plot
+  #   legend.position = c(0, .95)
+  #   ) +
+  # ggplot2::theme(
+  #   text = ggplot2::element_text(family = "Futura-Medium"),
+  #   legend.title = ggplot2::element_text(family = "Futura-Bold", size = 10),
+  #   legend.text = ggplot2::element_text(family = "Futura-Medium", size = 10)
+  #   ) +
 
@@
 
-p
-```
-
-### Image detection
-
-```{r echo=FALSE}
-#| label: basic-stats-zero
-
-# Compute total amount of images
-nb_image <- length(unique(dets$image_id))
-nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
-
-# Hardcoded check
-stopifnot(nb_image == 3288178) # 2788375
-```
+cowplot::plot_grid(ont_base_map, p, ncol = 2, rel_widths = c(1,2))
+```
+
+## Data access and storage
 
 Images were retrieved from all cameras using a 32 GB SD card and copied onto a 10 TB hard drive (LaCie d2 Professional).
-The images amounted to 1.8TB of memory on this disk, for a total of **`r nb_image_fmt`** images.
-Images were stored with the following directory structure:
+The images amounted to 2+TB of memory on this disk, for a total of **`r nb_image_fmt`** images.
+They were stored with the following directory structure:
 
@@
 
-The images were processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
-The model processes each image and produces “detections” (bounding boxes) at a variable confidence score.
+We have made a backup of the these images on a compute Canada cloud instance , accessible via SSH terminal. We also plan to keep an additional copy of the images on a separate harddrive.
+
+## Image processing
+
+In this section, we separate the workflow we developed and used between September 2022 and December 2023, from the new workflow we are experimenting with and hoping to perfect for 2024 and onward.
+In 2022 we explored alternative tagging methods as no platform was at that time equipped for our needs.
+We developed a custom and reproducible platform for image tagging through lable studio and R scripts.
+We deployed our own LabelStudio platform and used custom (yet reproducible) scripts for image analysis.
+This has proven to be a workable yet flawed method for processing our large amount of images.
+The main issue we ran into is due to the fact that MegaDetector (which is described in more details in the next section), produced large amount of false positives in our region.
+This may have resulted from Megadetector not being calibrated to highly seasonal landscapes, resulting in very slow processing of images through Labelstudio.
+This made data processing in LabelStudio slower and more cumbersome than we initially anticipated, as LabelStudio is not well equipped for group tagging of false positives.
+However, in the absence of a better alternative, we persevered in using the platform, and attempted to circumvent these issues by trying to reduce the number of false positives in the MegaDetector output.
+We attempted to find patterns in the false positives rates and to train a separate classifier but we were unsuccessful.
+As we continued, it slowly became clear that LabelStudio could not scale appropriately to multi-year data.
+Concurrently, the wildtrax platform improved significantly.
+Wildtrax now supports MegaDetector and image filtering based on score; something we had been conducting manually through our LabelStudio protocol.
+It started being able to display the bounding boxes created by MegaDetector and exposed these results in its API, allowing for further processing.
+Compared to 2022, it now has much better batch tagging and species verification abilities.
+Currently, it is still missing the ability to assign specific tags to bounding boxes within the app.
+As of March 2024 we are actively shifting out workflow to WildTrax, reanalyzing all previous work conducted in LabelStudio.
+For transparency we outline both LabelStudio and wildtrax protocols below.
+
+### Previous workflow (LabelStudio)
+
+#### Image detection
+
+```{r echo=FALSE}
+#| label: basic-stats-zero
+
+# Compute total amount of images
+nb_image <- length(unique(dets$image_id))
+nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
+
+# Hardcoded check
+stopifnot(nb_image == 3288178) # 2788375
+```
+
+For the LabelStudio workflow, the images were first processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
+MegaDetector processes each image and produces “detections” (bounding boxes) at a variable confidence score.
 
@@
 At this first step into the process, no filtering was applied beyond the model predictions which classifies images as either likely to contain an object of relevance or empty.
-About **`r dets_tb_percent[1]` %** of the images, (**`r dets_tb_fmt[1]`**) images, were found by the model to possibly contain either a person, animal, or vehicle.
+About **`r dets_tb_percent[1]` %** of the images, (**`r dets_tb_fmt[1]`** images), were found by the model to possibly contain either a person, animal, or vehicle.
 The model did not produce detections for **`r dets_tb_fmt[2]`** images, (**`r dets_tb_percent[2]` %**) of the total.
@@
 
-Megadetector is an AI model which produces confidence scores with each of its predictive detections.
+MegaDetector is an AI model which produces confidence scores with each of its predictive detections.
 With no filtering, the overwhelming majority of the detections are of very low confidence (85% of detections are below a score of 0.1).
 This explains the high number of "person" and "vehicle" detections which are of course unrealistic.
-It is recommended to filter low probability detections, and cut-off values of 0.1 or 0.2 are usually suggested in the Megadetector documentation.
-
-Using a cut-off value for the detection confidence of 0.1 (,i.e.
-discarding all detections with confidence below 0.1), this left us with a total of **`r dets_tb_cat_sum_fmt`** detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
-After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process for each category are: **`r dets_tb_cat_fmt[1]`** animals (**`r dets_tb_cat_percent[1]` %**); **`r dets_tb_cat_fmt[2]`** persons (**`r dets_tb_cat_percent[2]` %**); and **`r dets_tb_cat_fmt[3]`** vehicles (**`r dets_tb_cat_percent[3]` %**).
+It is recommended to filter low probability detections, and cut-off values of 0.1 or 0.2 were suggested in the [MegaDetector documentation](https://github.com/microsoft/CameraTraps) at the time model was ran.
+
+As the NSERC Alliance Wildlife Working Group, we decided to use a cut-off value of 0.1 for detection confidence of 0.1 (,i.e.
+discarding all detections with confidence below 0.1), leaving us with a total of **`r dets_tb_cat_sum_fmt`** detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process for each category were: **`r dets_tb_cat_fmt[1]`** animals (**`r dets_tb_cat_percent[1]` %**); **`r dets_tb_cat_fmt[2]`** persons (**`r dets_tb_cat_percent[2]` %**); and **`r dets_tb_cat_fmt[3]`** vehicles (**`r dets_tb_cat_percent[3]` %**).
 
@@
 #| label: fig-detections-histogram
-#| fig-cap: "Histogram of detection confidence scores produced for each bounding box predicted by Megadetector, showing a bimodal distribution.
-The colors represent the three categories of object identified by Megadetector: animals, persons and vehicles.
+#| fig-cap: "Histogram of detection confidence scores produced for each bounding box predicted by MegaDetector on camera trap images collected from the Northern Ontario array between March 2022 and March 2023, showing a bimodal distribution.
+The colors represent the three categories of object identified by MegaDetector: animals, persons and vehicles.
 Note that the hsitogram is truncated at our selected confidence threshold of 0.1."
@@
 Finally, it is important to note that not only pictures with detections were uploaded to the LabelStudio platform.
-As stated before, camera traps produce images in bursts (in our case, 5 images per trigger).
-It is possible for Megadetector to miss a detection within a burst, or to produce variable confidence scores within that burst, which could lead to missing true detections if not all pictures from a given burst are filtered in.
-Therefore, we included all the images from a given burst if any of the images in the burst include a detection with certainty above the threshold, including potentially empty images from that burst.
-
-### Tagging platform selection
+Camera traps produce images in bursts (in our case, 5 images per trigger).
+It is possible for MegaDetector to miss a detection within a burst, or to produce variable confidence scores within that burst, which could lead to missing true detections if not all pictures from a given burst are included.
+Therefore, if any image from a burst exceeded the selected MegaDetector threshold, we included it in our analysis .
+
+#### Tagging platform selection
 
@@
 -   *Optionally, open source,* to allow for maximum reproducibility of the process.
-
-To guide our choice, we used these criteria to informally compare 51 different platforms and tools currently available for processing and managing camera trap images.
+-   *Ideally,* stored on a cloud system hosted in Canada.
+
+To guide our choice, we used these criteria to informally compare 51 different platforms and tools currently available for processing and managing camera trap images (a list of the tools compared in given in Appendix III).
 We first looked at platforms traditionally used for tagging camera trap images, but found them often lacking in two main aspects: flexibility of the tagging interface, and ability to manipulate bounding box information.
@@
 These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners.
-Indeed, machine learning datasets outside can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features.
+Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features.
 Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.
@@
 We found two platforms that matched all of our criteria: LabelStudio (@LabelStudio) and CVAT (@boris_sekachev_2020_4009388).
-After testing both platforms, LabelStudio was chosen because the data format it uses is more easily compatible with Megadetector's COCO format output, making it easier to setup.
+After testing both platforms, LabelStudio was chosen because the data format it uses is more easily compatible with MegaDetector's COCO format output, making it easier to setup.
 We deployed the Community edition LabelStudio (version 1.6.0rc5) on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and a NGINX reverse proxy, which is a standard approach for custom hosting of websites.
 
-### Tagging progress
+### Updated workflow (Wildtrax)
+
+#### Image renaming and organization
+
+In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which all images were manually sorted into "wildlife" and "nonwildlife" folders inside each subfolders.
+In addition, because wildtrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images.
+Therefore, we developed a simple and very efficient script in the Julia programming language to paralellize the renaming of each file into a standardized manner.
+We had originally refrained from using this method of uniquely identifying files and instead relied on tracking the unique path of each file as we wanted to reduce the impact on our storage capacity (as it essentially required to double the required storage space).
+
+The standard format we used for file renaming follows this convention:
+
+- "\textcolor{red}{retrievalID}\_\textcolor{blue}{siteID}\_\textcolor{green}{locationID}\_\textcolor{purple}{cameraID}\_\textcolor{orange}{subfolderID}\_\textcolor{PineGreen}{fileID}\_\textcolor{cyan}{dateTime}.JPG",
+
+Giving, for example, a file with the name:
+
+- "\textcolor{red}{TC1}\_\textcolor{blue}{P028}\_\textcolor{green}{P028-1}\_\textcolor{purple}{WLU-10}\_\textcolor{orange}{100}\_\textcolor{PineGreen}{0001}\_\textcolor{cyan}{2022\_03\_18\_09\_44\_10}.JPG".
+
+We also added a "W" if the image had been sorted into a "wildlife" folder, ie: "TC2_INTC02_INTC02-2_CWS-ON46_*W*_0099_2022_12_17_13_41_00.JPG".
+
+#### Current tagging workflow and postprocessing
+
+Wildtrax keeps track of image batches using "location folders".
+We created our own location folders but combining images within the same retrieval, site and location IDs.
+For example the image mentioned above would be placed in the "TC1_P028_P028-1_WLU-10" location folders.
+For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.
+
+As a first step in testing the wildtrax workflow, we are uploading the folders from the first retrieval that showed the highest rates of false positives, retagging and evaluating the level of overlap with the LabelStudio workflow and also allowing us to complete these folders.
+As a second step, we are uploading the folders from the very last retrieval, to speed up the tagging progress.
+What remains to be determined is how we will properly assign tags to bounding boxes.
+We are currently considering two main avenues: 1.
+re-purposing the previous LabelStudio workflow for bounding box assignment, and 2.
+using a non networked platform, for example the EcoAssist software.
+
+## Species Classifier
+
+We made a first attempt at developing a species classifier using the 20000+ cropped bounding boxes resulting from the tagging process.
+We used the Pytorch framework, training a resnet100 model for 100 epochs, and modifying the last layer to include a regularization (i.e.
+dropout) layer in order to avoid overfitting.
+We tested the species classifier training via both transfer learning (frozen weights) and fresh learning (unfrozen weights).
+We used a batch size of 32.
+
+## Tagging progress - LabelStudio platform
 
@@
 
-Out of the **`r dets_tb_cat_fmt[1]`** animal detections, **`r processed_fmt`** have been processed to date using our platform (or **`r processed_pct` %**) for a total of **`r processed_img_fmt`** images (i.e.
+Out of the **`r dets_tb_cat_fmt[1]`** animal detections, **`r processed_fmt`** have been processed to date using the LabelStudio platform (or **`r processed_pct` %**) for a total of **`r processed_img_fmt`** images (i.e.
 **`r processed_img_pct` %** of the total amount of images, or **`r processed_img_up_pct` %** of the total amount of images uploaded to the platform).
 
-As stated above, all images of a given burst are viewed if at least one image in the burst contains a detections with confidence above the threshold.
-This include potentially empty images.
-If we add the empty images viewed as part of a burst, it brings the total amount to **`r viewed_fmt`** images, (i.e.
+As stated above, all images of a given burst are viewed if at least one image in the burst contains at least one detection with confidence above the MegaDetector confidence threshold.
+This includes potentially empty images.
+If we add the empty images viewed as part of a burst, it brings down the total amount of viewed images to **`r viewed_fmt`** images, (i.e.
 **`r viewed_pct` %** of the total amount of images, or **`r viewed_pct_up` %** of the total amount of images uploaded to the platform).
@@
 
-Out of the **`r processed_fmt`** detections, **`r dels_nb_fmt`** were false positives (or **`r dels_nb_pct` %**), where Megadetector identified something, but there was nothing.
+Out of the **`r processed_fmt`** detections, **`r dels_nb_fmt`** were false positives (or **`r dels_nb_pct` %**).
 In addition, **`r non_mans_nb_fmt`** were true positives (or **`r non_mans_nb_pct` %**), to which **`r mans_nb_fmt`** false negatives were added (manually added bounding boxes, **`r mans_nb_pct` %**).
-Out of the **`r non_mans_nb_fmt`** true positives, **`r orig_tb_fmt[3]`** had to have their bounding box adjusted.
+Out of the **`r non_mans_nb_fmt`** true positives, **`r orig_tb_fmt[3]`** had to have their bounding box adjusted: the bounding box drawn by MegaDetector was at least partially innacurate and had to be corrected.
 The table below summarises information for the tagging progress.
@@
 knitr::kable(sum_tab,
-             caption = "Summary table on tagging progress")
+             caption = "Summary table on tagging progress. The 0.1 filter value corresponds to the megadetector cinfidence threshold.")
 ```
@@
 The relatively high rate of false positives is attributable to a combination of factors.
-First, the high seasonality of Ontario's far north makes it so that it is difficult to predict whether a camera that was set up in the winter won't be overgrown by vegetation in the spring.
-Locations where vegetation has grown in front of the sensor, combined with windy events, creates a perfect storm for false cameras triggers when vegetation is moving in front of the sensors.
-In addition, Megadetector is a likely to mistake a certain proportion of those vegetation pictures for actual animals.
-We are currently exploring avenues to tackle this roadblock by identifying repeating bounding images and by using a model to sort false from true positives.
-
-#### Reproducibility
+First, the high seasonality of Ontario's far north makes it difficult to predict whether a camera that was set up in the winter won't be overgrown by vegetation in the spring.
+This isn't a unique problem to this project, but an issue in many ecosystems.
+Locations where vegetation has grown in front of the sensor, combined with windy events, increases the number of photos triggered by vegetation, rather than wildlife.
+MegaDetector is likely to mistake many of the vegetation pictures for animals.
+We are currently exploring avenues to ways to address this by identifying repeating bounding images and by using a model to sort false from true positives.
+
+## Tagging progress - WildTrax
+
+TODO: update once data is harmonized
+
+## Reproducibility
 
 We are committed to producing a reproducible workflow.
-All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap).
-The repo makes use of [**mdtools**](https://github.com/StewartWILDlab/mdtools) module, a python-based command line tool for converting Megadetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
-
-### Analysis
-
-#### Species detections
-
-We summarized our existing species data to date for analysis in two main ways:  (1) species presence/absence data for each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site. The detection time for species is highly variable.
+All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap), which is currently private but will be opened upon submission of the final version of this report.
+The repo makes use of the [**mdtools**](https://github.com/StewartWILDlab/mdtools) module, a python-based command line tool for converting MegaDetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
+
+## Analysis
+
+### Species detections
+
+We summarized our existing species data in two ways:  (1) species presence/absence at each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site. The detection time for species is highly variable.
 
@@
 
-#### Ordination and clustering
-
-We analysed the relative abundance data using well known numerical ecology community analysis methods (Legendre and Legendre 2013).
+### Ordination and clustering
+
+We analysed the relative abundance data using published and standardized numerical ecology community analysis methods (Legendre and Legendre 2013).
 As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
@@
 
-## Data exploration and preliminary results
-
-### Wildlife community compposition
-
-We obtained repeat detections of at least 31 mammal and bird species across 193 sites and 5-7 months of detections The most common species detected (detected more than 10 days in the season) are listed in Table 2 below (the full table is available in the appendix at the end of this report).
-Note the large amount detections in the “bird” and “duck” species categories which could be further distinguished into specific species.
+# Data exploration and preliminary results
+
+## Wildlife community composition
+
+We obtained repeat detections of at least 31 mammal and bird species across 193 sites and 5-7 months of deployment.
+The most common species detected (detected more than 50 days in the season) are listed in Table 2 below (the full table is available in Appendix I).
+Note the large amount detections in the “bird” and “duck” species categories which could be further distinguished into specific species; eight of the top 17 detected species categories were avian.
 
@@
   dplyr::relocate(Scientific_name, .after = species) |> 
-  dplyr::filter(days_detected >= 10) |>
+  dplyr::filter(days_detected >= 50) |>
   gt::gt() |>
   gt::tab_header(
-    title = "Table 2: Relative abundances for species with more than 10 detection days",
-    subtitle = "Computed in number of days detected in the season"
+    title = "Table 2: Relative abundances for species with more than 50 detection days, computed in number of days detected in the season."
   ) |>
@@
 
-As for the spatial distribution of those relative abundances, @fig-map-pies shows relative species abundances across the landscape, for species detected more than 50 days in the season (this number is roughly equal to the standard deviation of the distribution of detection days, which has a mean of 32 days).
+As for the spatial distribution of species relative abundances, @fig-map-pies shows relative species abundances across the landscape, for species detected more than 50 days in the season (this number is roughly equal to the standard deviation of the distribution of detection days, which has a mean of 32 days).
 Note that for this figure, the detections have grouped by plot number for better visibility, with a plot containing up to 4 cameras (see deployment methods).
@@
 
-### Ordination - presence-absence
+## Ordination - presence-absence
 
 To start exploring general species-habitat associations we completed an redundancy analysis (RDA, using the `vegan` R package, @vegan).
-RDAs allow to study the relationships between two matrices of data, and is often used in community studies, using a constraining matrix of environmental variables for each site and a corresponding matrix of species observations (either presence absence or relative abundances).
-Ordinations also combine very well with clustering analyses to produce insights about associations between sites.
+RDAs allow to study the relationships between two matrices of data, and is often used in community ecology.
 
@@
 
-We used the Ontario Far North Land Cover dataset to produce proportions of land covers at each site (@hogg2014far, a full legend of the land cover abbreviations are to be found in the appendix of this report).
-@fig-rda-pa shows the triplot for the RDA on presence-absence data, which yielded a low $R^2$ of `r vegan::RsquareAdj(pa_rda)[2]`, but which is still relevant for exploring species associations.
+We used the Ontario Far North Land Cover dataset to produce proportions of land covers at each site (@hogg2014far, a full legend of the land cover abbreviations are to be found in Appendix II).
+@fig-rda-pa shows the triplot for the RDA on presence-absence data, which yielded a low $R^2$ of `r round(as.numeric(vegan::RsquareAdj(pa_rda)[2]), 2)`, but which is still relevant for exploring species associations.
 Land cover might not be able to explain community structure, but it can still show which species are found together and around what kind of land cover class.
-Here, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
-Canada geese are, as is to be expected, associated with water (WAT) while smaller species like squirrels and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
+Here, as expected, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
+Canada geese are associated with water (WAT) while smaller species like snowshoe Hares and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
 
@@
 #| label: fig-rda-pa
-#| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected in March through September of 2022 reveals that different species associate with different land cover classes.
+#| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected in March through September of 2022 reveals that different species associate with different Far North land cover classes.
 For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
@@
 
-### Ordination - relative abundances
+## Ordination - relative abundances
 
@@
 @fig-rda-events shows the triplot for the RDA on relative abundance data.
-This analysis yielded a slightly larger $R^2$ of `r vegan::RsquareAdj(events_rda)[2]`, than the analysis in @fig-rda-pa presenting presence-absence data, but reduces the axis scores of the rarer species.
+This analysis yielded a slightly larger $R^2$ of `r round(as.numeric(vegan::RsquareAdj(events_rda)[2]), 2)`, than the analysis in @fig-rda-pa presenting presence-absence data, but reduces the axis scores of the rarer species.
 We observe similar patterns for caribou, crane, moose, black bear and marten.
@@
 #| label: fig-rda-events
-#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data."
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns as the triplot using presence-absence data."
 
@@
 custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp = 0.4, scale_y_sp = 0.3,
-                thres_plot = 0.3)
-```
-
-### Clustering
+                thres_plot = 0.2)
+```
+
+## Clustering
 
@@
 
-### Spatial clustering
+## Spatial clustering
 
@@
 
-Spatial clustering allows to add distance between sites as a soft constraint in the clustering algorithm.
+Spatial clustering allows to add Euclidean distance between sites as a soft constraint in the clustering algorithm.
 The extent to which these distances influence the result in balance with the community data can be adjusted, by setting the strength of the mixing parameter alpha.
@@
 
-We found that the optimal value for alpha be `r 0.1`, and performed mixed clustering using the `hclustgeo` function.
+We found an alpha level of 0.1 to be optimal for this data set, and performed mixed clustering using the `hclustgeo` function.
 @fig-map-clust-geo shows the resulting map of geographically constrained clustered camera sites.
-This reveals a North/South pattern in the spatial clustering of sites.
 
@@
 
-We can plot those geographically constrained clusters on top of the abundance-based RDA of @fig-rda-events.
-@fig-rda-clust-geo shows that those geographic clusters do not map well on top of the ordination, with one cluster seeming to represent a subset of a larger trend experienced in the larger part of the data.
-
-```{r echo=FALSE}
-#| label: fig-rda-clust-geo
-#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
-Two clusters determined using geographically constrained clustering with a mixing parameter of 0.1."
-
-custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
-```
-
-### Species heatmaps and bubbleplots
+## Species heatmaps and bubbleplots
 
@@
 
-In order to visualize relative abundances through time and space, we can use bubbleplots.
+We use a series bubbleplots to visualize species relative abundance through space and time.
 The following plots are for caribou, sandhill crane, moose, and wolf.
@@
 
-As more of the camera images are labeled and our dataset grows, we can transform the above bubble plots into heat maps of species presence/absence or relative abundance.
-Below is an example using the existing caribou data to date.
-It is important to recognize that not all cameras had been deployed until June, leaving April and Ma.
-y maps with large portions of the northern study area where cameras were not yet present.
-
-### Take Home messages
+## Take Home messages
 
@@
 
--   Using an object detection model allowed us to filter out large quantities of empty images, but performed somewhat poorly in certain conditions (namely, windy sites with drastic changes in vegetation).
-
--   According to our analysis, species tend to associate with land use changes classes that generally corresponds to their expected habitat, such as wetlands for caribou and cranes, or forested areas for moose.
-
--   Clustering reveals two large groups among sites, but spatial patterns are difficult to tease apart at this stage.
-
-## Next steps
-
-Our existing image identification workflow and tagging process has allowed us to produce an open and reproducible data set of spatially explicit wildlife species counts in Ontario’s Far North.
-Using this data set we have started to explore spatial patterns of species richness, occurrence, and abundance.
-All of this work is conducted through opensource platforms, and data exploration and analysis code hosted as a [GitHub repository](https://github.com/StewartWILDlab/rofcamtrap) for ease of access and reproducibilty.
-The information necessay to reproduce this report is in the colophon in the appendix.
-
-Over the coming months we will continue to process the retrieved data and anticipate adding to this data set over the summer, as data from 18 cameras has yet to be retrieved.
+-   Using an object detection model allowed us to filter out large quantities of empty images, but performed somewhat poorly in certain conditions (ie, windy sites with drastic changes in vegetation).
+
+-   Associative patterns: species associate with land cover classes that corresponds to expected habitat, such as wetlands for caribou and cranes, or forested areas for moose.
+
+-   Spatial clustering and considerations: Latitudional relationships may dominate this data set and drive species-habitat associations.
+
+-   Species classifier: we obtained better accuracy (94%) with fresh learning than transfer learning (91%).
+We are still exploring avenues for improvement, but it is likely that the small sample size is the main limiting factor.
+
+# Next steps
+
+Over the coming months we will continue to process the retrieved data and anticipate adding to our data set.
 The next steps for this project are to complete the tagging process and identify all the detections tagged as unknowns, bird, duck and mustelid species.
 We will also further explore the spatial community composition patterns at each camera site.
-Given a high enough data density for certain species future possibilities for this project include modelling species occupancy (@MacKenzie2002) and spatially explicit density estimation using spatially explicit capture-recapture models (@Chandler2013).
-
-After data processing and exploration is complete the next major step in this project will be to use the bounding boxes information we have generated to train an image classifier.
-This classifier will be based on Megaclassifier (of type EfficientNet architecture, @EfficientNet), an classifier already trained to pair with Megadetector's output.
-We will retrain this model to fit our taxa, possibly leveraging mentoring opportunities offered by the CV4E workshop ("Computer Vision for Ecology"), if availability allows.
-
-## Acknowledgements
+
+Given a high enough data density for certain species future possibilities for this project include modelling species occupancy (e.g.
+@MacKenzie2002) and spatially explicit density estimation using spatially explicit capture-recapture models (e.g.@Chandler2013). We will perfect the new tagging workflow and continue testing classifier methods.
+
+# Acknowledgements
 
@@
 We also thank members of the WILDlab for feedback throughout this process: Claudia Haas, Eric Jolin, Charlotte Rentmeister, and Sheyda Zand.
-This work was supported by a G&C to Frances Stewart.
-
-## References
+This work was supported by a G&C to Frances Stewart and the NSERC Alliance full project team.
+
+{{< pagebreak >}}
+
+# References
 
@@
 
-## Appendix
+# Appendix I: Species table
 
@@
 
+# Appendix II: Land cover classes table
+
 ```{r echo=FALSE}
@@
 
+# Appendix III: List of camera trapping tools compared
+
 <!-- The following line inserts a page break  -->
@@
 
-### Colophon
+## Colophon
 
@@
 
+As more of the camera images are labeled and our dataset grows, we can transform the above bubble plots into heat maps of species presence/absence or relative abundance.
+Below is an example using the existing caribou data to date.
+It is important to recognize that not all cameras had been deployed until June, leaving April and May.
+y maps with large portions of the northern study area where cameras were not yet present.
+
 ```{r}
@@
 ```
+We can plot those geographically constrained clusters on top of the abundance-based RDA of @fig-rda-events.
+@fig-rda-clust-geo shows that those geographic clusters do not map well on top of the ordination, with one cluster seeming to represent a subset of a larger trend experienced in the larger part of the data.
+
+```{r echo=FALSE}
+#| label: fig-rda-clust-geo
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
+Two clusters determined using geographically constrained clustering with a mixing parameter of 0.1."
+
+custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
+```
 -->

### 2024-04-08 scripts/R/file_management/rename_images.R
playing with wildtrax code

@@ -16,3 +16,42 @@ rename_images(images_from = "/media/vlucet/TrailCamST1/TrailCamStorage_2/",
 
 # [1] "P045" "P058" "P080!!!" "P087" "P100"
 # [6] "P216" "P224" "P293" "Q647"
+
+library(wildRtrax)
+
+Sys.setenv(WT_USERNAME = 'valentin.lucet@gmail.com',
+           WT_PASSWORD = 'FF3DhjsN9D$MMDn#')
+
+# debugonce(wt_auth)
+wt_auth(force = T)
+
+# debugonce(wt_get_download_summary)
+my_projects <- wt_get_download_summary(sensor_id = 'CAM')
+
+# client <- httr2::oauth_client(
+#   id = "Eg2MPVtqkf3SuKS5uXzP97nxU13Z2K1i",
+#   token_url = "https://abmi.auth0.com/oauth/token",
+#   name = "wildtrax"
+# )
+
+# debugonce(httr2::oauth_flow_password)
+# tok <- httr2::oauth_flow_password(client = client, username = Sys.getenv("WT_USERNAME"),
+#                            password = Sys.getenv("WT_PASSWORD"))
+
+# -------------------------------------------------------------------------
+
+# .wt_api_pr(path = "/bis/get-download-summary", sensorId = sensor_id,
+#            sort = "fullNm", order = "asc")
+
+debugonce(wt_download_report)
+my_report <- wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "image_report", weather_cols = F) |>
+  tibble::as_tibble()
+
+my_report <- wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "main", weather_cols = F) |>
+  tibble::as_tibble()
+img_report <- wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "image_report", weather_cols = F) |>
+  tibble::as_tibble()
+md_report <- wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "megadetector", weather_cols = F) |>
+  tibble::as_tibble()
+
+saveRDS(my_report, "my_report.rds")

### 2024-04-08 analysis/documents/1-data-preparation.qmd
add Ontario map

@@
 
-cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)
+cams_sf_coords <- dplyr::bind_cols(cams_sf, coords) |>
+  dplyr::mutate(
+    month_year_deployed = paste(lubridate::month(date_deployed, label = T, abbr = T), 
+                                lubridate::year(date_deployed)), 
+    month_year_retrieved = paste(lubridate::month(date_retrieved, label = T, abbr = T), 
+                                 lubridate::year(date_retrieved)), 
+    deployment_group = paste0(month_year_deployed, ifelse(
+      month_year_retrieved == "NA NA", "", paste(" -", month_year_retrieved))))
 
@@
 #| label: process-zones
-
-zones_filtered <- 
-  sf::st_read(here::here("data-raw","spatial", "ecozones", "EcoZone" ,"EcoZone.shp"),
+zones <- sf::st_read(here::here("data-raw","spatial", "ecozones", "EcoZone" ,"EcoZone.shp"),
               quiet = TRUE) |> 
-  sf::st_transform(crs = sf::st_crs(cams_sf)) |>
+  sf::st_transform(crs = sf::st_crs(cams_sf))
+zones_filtered <- zones |>
   sf::st_filter(cams_sf_coords)
@@
 saveRDS(base, here::here("data", "objects", "base_map.rds"))
+
+ont_map <- ggmap::get_stadiamap(bbox = c(left = as.numeric(sf::st_bbox(zones)[1])-1, 
+                                         bottom = as.numeric(sf::st_bbox(zones)[2])+1, 
+                                         right = as.numeric(sf::st_bbox(zones)[3])-1,
+                                         top = as.numeric(sf::st_bbox(zones)[4])+1), 
+                            zoom = 5, maptype = "stamen_terrain")
+
+ont_base <- ggmap::ggmap(ont_map) +
+  ggplot2::labs(x = "Longitude", y = "Latitude") +
+  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2) +
+  ggplot2::geom_rect(
+    xmin = -90,
+    ymin = 50,
+    xmax = -83,
+    ymax = 55.5,
+    fill = NA, 
+    colour = "black",
+    size = 0.6
+  )
+
+saveRDS(ont_base, here::here("data", "objects", "ont_base_map.rds"))
 ```

### 2024-04-09 reports/GnC_report/paper/paper.qmd
ptogress

@@
 
+```{r echo=FALSE}
+#| label: basic-stats-zero
+
+# Compute total amount of images
+nb_image <- length(unique(dets$image_id))
+nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
+
+# Hardcoded check
+stopifnot(nb_image == 3288178) # 2788375
+```
+
 Images were retrieved from all cameras using a 32 GB SD card and copied onto a 10 TB hard drive (LaCie d2 Professional).
@@
 
-```{r echo=FALSE}
-#| label: basic-stats-zero
-
-# Compute total amount of images
-nb_image <- length(unique(dets$image_id))
-nb_image_fmt <- prettyNum(nb_image, big.mark = ",")
-
-# Hardcoded check
-stopifnot(nb_image == 3288178) # 2788375
-```
-
 For the LabelStudio workflow, the images were first processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
@@
 These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners.
-Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features.
+Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features (for example, see @Schneider2021).
 Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.
@@
 
-In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which all images were manually sorted into "wildlife" and "nonwildlife" folders inside each subfolders.
+Wildtrax keeps track of image batches using "location folders".
+We created our own location folders by combining images within the same retrieval, site and location IDs.
+The standard format we used for location folders follows this convention:
+
+- "\textcolor{red}{retrievalID}\_\textcolor{blue}{siteID}\_\textcolor{green}{locationID}\_\textcolor{purple}{cameraID}
+
+Giving, for example, a folder with the name:
+
+- "\textcolor{red}{TC1}\_\textcolor{blue}{P028}\_\textcolor{green}{P028-1}\_\textcolor{purple}{WLU-10}"
+
+For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.
+
+In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which most of the images were manually sorted into "wildlife" and "nonwildlife" folders inside each of the camera subfolders.
 In addition, because wildtrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images.
@@
 #### Current tagging workflow and postprocessing
-
-Wildtrax keeps track of image batches using "location folders".
-We created our own location folders but combining images within the same retrieval, site and location IDs.
-For example the image mentioned above would be placed in the "TC1_P028_P028-1_WLU-10" location folders.
-For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.
 
@@
 sum_tab <- tibble::tribble(
-  ~Status,                ~Images,           ~Detections,
-  "Total",                nb_image_fmt,      dets_tb_cat_unfil_sum_fmt,
-  "With detections",      dets_tb_fmt[1],    "N/A",
-  "Empty",                dets_tb_fmt[2],    "N/A",
-  "Filtered (> 0.1)",     uploaded_fmt,      dets_tb_cat_sum_fmt,
-  "Processed",            processed_img_fmt, processed_fmt
+  ~Status,                ~Images,           ~Detections,               ~`False Positives`,
+  "Total",                nb_image_fmt,      dets_tb_cat_unfil_sum_fmt, "N/A",
+  "With detections",      dets_tb_fmt[1],    "N/A",                     "N/A",
+  "Empty",                dets_tb_fmt[2],    "N/A",                     "N/A",
+  "Filtered (> 0.1)",     uploaded_fmt,      dets_tb_cat_sum_fmt,       "N/A",
+  "Processed",            processed_img_fmt, processed_fmt,             dels_nb_fmt
 )
@@
 
-\newpage
-
-# Appendix I: Species table
+{{< pagebreak >}}
+
+# Appendices
+
+## Appendix I: Species table
 
@@
 
-# Appendix II: Land cover classes table
+{{< pagebreak >}}
+
+## Appendix II: Land cover classes table
 
@@
 
-# Appendix III: List of camera trapping tools compared
-
-<!-- The following line inserts a page break  -->
-\newpage
+{{< pagebreak >}}
+
+## Appendix III: List of camera trapping tools compared
+
+-  Aardwolf
+-  Agouti
+-  AIDE
+-  animl
+-  animl (R)
+-  CAIMAN
+-  Camelot
+-  Camera Base
+-  Camera Trap Manager
+-  CameraTrapDetectorR
+-  CameraTraps
+-  Camtrap DP
+-  camtrapR
+-  Conservation AI
+-  Cos4Cloud FASTCAT
+-  CPW Photo Warehouse
+-  DeepFaune
+-  digiKam
+-  EcoAssist
+-  eMammal
+-  Event Finder Suite
+-  IBEIS
+-  IL2BB BBoxEE
+-  Mapview
+-  mbaza
+-  MD Fast API
+-  MLWIC MLWIC2 
+-  PhotoSpread
+-  rapidPop
+-  Renamer & CamTrap
+-  Snoopy
+-  sparcD
+-  SpeedyMouse
+-  Timelapse
+-  TrailCamData
+-  TRAPPER
+-  U-Infuse
+-  ViXeN
+-  Wild ID
+-  WILDCAM
+-  wildcam
+-  Wildepod
+-  WildEye TrapTagger
+-  Wildlife Insights
+-  Wildlife observer network
+-  Wildme Wildbook
+-  WildTrax
+-  WTB Where's The Bear
+-  zamba
+-  Zooniverse
+
+{{< pagebreak >}}
 

### 2024-04-09 reports/GnC_report/paper/paper.qmd
progress

@@
 re-purposing the previous LabelStudio workflow for bounding box assignment, and 2.
-using a non networked platform, for example the EcoAssist software.
+using a different, non networked platform.
 
@@
 
-We made a first attempt at developing a species classifier using the 20000+ cropped bounding boxes resulting from the tagging process.
-We used the Pytorch framework, training a resnet100 model for 100 epochs, and modifying the last layer to include a regularization (i.e.
-dropout) layer in order to avoid overfitting.
-We tested the species classifier training via both transfer learning (frozen weights) and fresh learning (unfrozen weights).
-We used a batch size of 32.
+We made a first attempt at developing a species classifier using the tagged data from all the images retrieved in 2022.
+Following the work of @beery2019efficient, we chose to use the bounding box of each verified detection instead of the entire image as the basis for our data.
+This resulted in 23730 cropped bounding boxes to be used for training and validation.
+We used the Pytorch framework (@Paszke_PyTorch_An_Imperative_2019), a widely used state of the art library for developing neural networks models in Python.
+As a first attempt, we settled on a model of the ResNet 100 architecture (@he2015deep), as it is widely used basal architecture for training species classifiers (for example, see @Khan2020, @Ukwuoma2022, and @BintaIslam2023), although the field is progressing and innovating on new methods and architectures rapidly (for example, see @Elhamod2021 or @radford2021learning).
+
+We chose to train the model for 100 epochs with a batch size of 32 as an arbitrary starting point for training parameters.
+We encountered issues in overfitting, which we tackled by modifying the last layer of our network to include a regularization (i.e.
+dropout) layer (as recommended by @Srivastava2014).
+We tested the species classifier training via both transfer learning (frozen weights of a pretrained ResNet model) and classic learning (unfrozen weights, learning from the data only), as this is currently a comparison of interest in the field of AI applied to camera trap data (@schneider2018deep, @BintaIslam2023).
 
@@
   dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
-                ECOZONE_NA = as.factor(ECOZONE_NA))
+                ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
+  dplyr::filter(diff_time < 200) # For mustelid error currently
 
@@
 For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
+#| fig.width: 8
+#| fig.height: 6
+
 custom_rda_plot(pa_rda)
@@
 #| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns as the triplot using presence-absence data."
+#| fig.width: 8
+#| fig.height: 6
 
@@
 Two clusters determined using Ward clustering are shown as ellipses."
+#| fig.width: 8
+#| fig.height: 6
 
@@
 #| fig-cap: "Map of clustered camera sites. The two clusters were determined using the Ward clustering method."
+#| fig.width: 8
+#| fig.height: 6
 
@@
 #| fig-cap: "Map of geographically constrained clustered camera sites, using geographically constrained clustering with a mixing parameter of 0.1."
+#| fig.width: 8
+#| fig.height: 6
 
@@
 #| label: bubbleplots-Caribou
+#| fig.width: 8
 plot_list <- readRDS(here::here("data", "objects", "bbplot_list.rds"))
@@
 #| label: bubbleplots-Sandhill
+#| fig.width: 8
 plot_list["Sandhill crane"]
@@
 #| label: bubbleplots-Moose
+#| fig.width: 8
 plot_list["Moose"]
@@
 #| label: bubbleplots-Wolf
+#| fig.width: 8
 plot_list["Wolf"]

### 2024-04-22 reports/GnC_report/paper/paper.qmd
another round of review

@@
 We aim to evaluate the viability of deep neural networks for camera trap surveys currently taking place in Northern Ontario’s heterogeneous landscapes, including the Hudson Bay Lowlands and Ontario Shield ecozones, as variables like weather, seasonality, vegetation type, and species characteristics can impact the performance of these methods.
-The extent to which these methods perform for northern regions and northern specie is not well established.
+The extent to which these methods perform for northern regions and northern species is not well established.
 There are currently no available machine learning models for identifying species from images in northern regions and developing one would be beneficial as a monitoring and modeling tool for northern boreal ecosystems.
 
-Using an array of cameras deployed in an heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
-We’ve recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with next steps involving the development of an image classifier.
+Using an array of cameras deployed in a heterogeneous sampling area straddling two of Northern Ontario’s ecozones (Hudson Bay Lowlands and Ontario Shield), we first look to quantify current species diversity and co-occurrence.
+We’ve recently developed an open-source workflow for camera trap image identification, access, and storage, which we are using to identify images, with the next steps involving the development of an image classifier.
 
@@
 
-Since the last progress report (2023 report), we flag the following improvements and changes, which are detailed in the new sections of this reports.
-
-- The number of fully processed images was doubled from about 50000 to more than 100000, keeping with the rate of progress of the first year of tagging.
-- We made a first attempt at at using machine learning methods to train a species classifier using about 23000 images as the basis dataset, which produced encouraging results.
-- After reevaluation of the ability of our current worlflow and platform to scale to multiple years of data, we have transitioned the image tagging workflow to [WildTrax](https://wildtrax.ca/).
+Since the last progress report (2023 report), we have flagged the following improvements and changes, which are detailed in the new sections of this report.
+
+- The number of fully processed images doubled from about 50000 to more than 100000, keeping with the rate of progress of the first year of tagging.
+- We made a first attempt at using machine learning methods to train a species classifier using about 23000 images as the basis dataset, which produced encouraging results.
+- After reevaluation of the ability of our current worlflow and platform to scale to multiple years of data, we have transitioned the image tagging workflow to [WildTrax](https://WildTrax.ca/).
 - This report repeatedly mentions the AI model MegaDetector.
@@
 
-1. Poster submitted and presented at 2023 conference of the Ontario chapter of The Wildlife Society (OCTWS).
-2. Poster submitted and displayed at the Mushkegowuk council 2023 Research Meeting.
+1. Poster submitted and presented at the 2023 conference of the Ontario chapter of The Wildlife Society (OCTWS).
+2. Poster submitted and displayed at the Mushkegowuk Council 2023 Research Meeting.
 3. Application to participate in the Workshop on Computer Vision Methods for Ecology at the Caltech Resnick Sustainability Institute. Rejectec in 2023, reapplied in 2024.
@@
 In March through June of 2022 193 unbaited and unlured wildlife camera traps ([Reconyx Hyper Fire II](https://reconyx.com/product/hyperfire-2-covert-ir-camera)) were added to this study design.
-All cameras were all programmed to take bursts of 5 images (one per second) at each trigger of their motion sensor.
-
-TODO: add more from Sam
+All cameras were programmed to take bursts of 5 images (one per second) at each trigger of their motion sensor.
+
+TODO: more detailed summary of deployment methods from Sam's report to be added before submission
 
@@
 Up to 4 cameras were paired with ARUs at each of these sites, with at least one of these cameras placed at a random ARU site and one other prioritized to focus on a game trail or open area to increase detection probability.
-Weather restrictions and field conditions limited access to some sites and resulting in some variation in this design.
-In total we have 43 plots with 4 cameras, 5 plots with 3 cameras and 3 plots with 2 cameras. @fig-deployment shows the location of the clustered cameras and their current retrieval status.
-54 cameras were retrieved in May 2022 and 116 in September 2022.
-6 additional cameras were retrieved in March 2023 due to logistical and accessibility issues.
-Another 77 cameras were retrieved in Mar 2023.
-The last 18 cameras were retrieved in September 2023 and are now being processed.
+Weather restrictions and field conditions limited access to some sites resulting in some variation in this design.
+In total, we have 43 plots with 4 cameras, 5 plots with 3 cameras, and 3 plots with 2 cameras. @fig-deployment shows the location of the clustered cameras and their current retrieval status: 54 cameras were retrieved in May 2022 and 116 in September 2022.
+Six additional cameras meant to be retrieved then were instead retrieved in March 2023 due to logistical and accessibility issues, along with another 77 cameras that were meant to be retrieved at that date.
+The last 18 cameras were retrieved in September 2023.
 
@@
 p <- base_map +
-  ggplot2::geom_label(ggplot2::aes(x = x, y = y, label = ECOZONE_NA), inherit.aes = FALSE,
+  ggplot2::geom_label(ggplot2::aes(x = x, y = y,label = ECOZONE_NA), inherit.aes = FALSE,
                       size = 2.5, data = regions_labels) +
@@
 
-We have made a backup of the these images on a compute Canada cloud instance , accessible via SSH terminal. We also plan to keep an additional copy of the images on a separate harddrive.
+We have made a backup of these images on a Compute Canada cloud instance, accessible via SSH terminal. We also plan to keep an additional copy of the images on a separate hard drive.
 
@@
 In 2022 we explored alternative tagging methods as no platform was at that time equipped for our needs.
-We developed a custom and reproducible platform for image tagging through lable studio and R scripts.
+We developed a custom and reproducible platform for image tagging through LabelStudio and R scripts.
 We deployed our own LabelStudio platform and used custom (yet reproducible) scripts for image analysis.
 This has proven to be a workable yet flawed method for processing our large amount of images.
-The main issue we ran into is due to the fact that MegaDetector (which is described in more details in the next section), produced large amount of false positives in our region.
-This may have resulted from Megadetector not being calibrated to highly seasonal landscapes, resulting in very slow processing of images through Labelstudio.
-This made data processing in LabelStudio slower and more cumbersome than we initially anticipated, as LabelStudio is not well equipped for group tagging of false positives.
-However, in the absence of a better alternative, we persevered in using the platform, and attempted to circumvent these issues by trying to reduce the number of false positives in the MegaDetector output.
-We attempted to find patterns in the false positives rates and to train a separate classifier but we were unsuccessful.
+The main issue we ran into is because MegaDetector (which is described in more detail in the next section), produced a large amount of false positives in our region.
+This may have resulted from MegaDetector not being calibrated to highly seasonal landscapes, resulting in very slow processing of images through LabelStudio This made data processing in LabelStudio slower and more cumbersome than we initially anticipated, as LabelStudio is not well equipped for group tagging of false positives.
+However, in the absence of a better alternative, we persevered in using the platform and attempted to circumvent these issues by trying to reduce the number of false positives in the MegaDetector output.
+We attempted to find patterns in the false positive rates and to train a separate classifier but we were unsuccessful.
 As we continued, it slowly became clear that LabelStudio could not scale appropriately to multi-year data.
-Concurrently, the wildtrax platform improved significantly.
-Wildtrax now supports MegaDetector and image filtering based on score; something we had been conducting manually through our LabelStudio protocol.
+Concurrently, the WildTrax platform improved significantly.
+WildTrax now supports MegaDetector and image filtering based on score; something we had been conducting manually through our LabelStudio protocol.
 It started being able to display the bounding boxes created by MegaDetector and exposed these results in its API, allowing for further processing.
@@
 Currently, it is still missing the ability to assign specific tags to bounding boxes within the app.
-As of March 2024 we are actively shifting out workflow to WildTrax, reanalyzing all previous work conducted in LabelStudio.
-For transparency we outline both LabelStudio and wildtrax protocols below.
+As of March 2024, we are actively shifting our workflow to WildTrax, reanalyzing all previous work conducted in LabelStudio.
+For transparency, we outline both LabelStudio and WildTrax protocols below.
 
@@
 
-#### Image detection
-
-For the LabelStudio workflow, the images were first processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with a NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
-MegaDetector processes each image and produces “detections” (bounding boxes) at a variable confidence score.
+#### Image Detection
+
+For the LabelStudio workflow, the images were first processed through the animal detection model MegaDetector (@beery2019efficient), version 5.0, using a Lenovo X1 laptop with an NVIDIA GeForce GTX 1650 Max-Q graphics card, which took about 2 weeks of quasi-continuous processing to complete.
 
@@
 
-At this first step into the process, no filtering was applied beyond the model predictions which classifies images as either likely to contain an object of relevance or empty.
-About **`r dets_tb_percent[1]` %** of the images, (**`r dets_tb_fmt[1]`** images), were found by the model to possibly contain either a person, animal, or vehicle.
+MegaDetector is an AI model that predicts areas of an image that are likely to contain an animal, a person, or a vehicle, producing rectangular bounding boxes that cover such areas.
+It processes each image and produces “detections” at a variable confidence score (between 0 and 1, excluding 0 as this would mean the model predicts nothing on the image).
+About **`r dets_tb_percent[1]` %** of the images, (**`r dets_tb_fmt[1]`** images), were found by the model to possibly contain either an animal, person, or vehicle.
 The model did not produce detections for **`r dets_tb_fmt[2]`** images, (**`r dets_tb_percent[2]` %**) of the total.
@@
 
-MegaDetector is an AI model which produces confidence scores with each of its predictive detections.
-With no filtering, the overwhelming majority of the detections are of very low confidence (85% of detections are below a score of 0.1).
+The overwhelming majority of the detections were of very low confidence (85% of detections are below a score of 0.1).
 This explains the high number of "person" and "vehicle" detections which are of course unrealistic.
-It is recommended to filter low probability detections, and cut-off values of 0.1 or 0.2 were suggested in the [MegaDetector documentation](https://github.com/microsoft/CameraTraps) at the time model was ran.
-
-As the NSERC Alliance Wildlife Working Group, we decided to use a cut-off value of 0.1 for detection confidence of 0.1 (,i.e.
-discarding all detections with confidence below 0.1), leaving us with a total of **`r dets_tb_cat_sum_fmt`** detections to sort through (see @fig-detections-histogram for an histogram of the detection scores, grouped by detection types).
+It is recommended to filter low probability detections and cut-off values of 0.1 or 0.2 were suggested in the [MegaDetector documentation](https://github.com/microsoft/CameraTraps) at the time the model was run.
+
+As the NSERC Alliance Wildlife Working Group, we decided to use a cut-off value of 0.1 for detection confidence (i.e.
+discarding all detections with confidence below 0.1), leaving us with a total of **`r dets_tb_cat_sum_fmt`** detections to sort through (see @fig-detections-histogram for a histogram of the detection scores, grouped by detection types).
 After applying this confidence threshold (decided on by the Wildlife Working Group) the total images to process for each category were: **`r dets_tb_cat_fmt[1]`** animals (**`r dets_tb_cat_percent[1]` %**); **`r dets_tb_cat_fmt[2]`** persons (**`r dets_tb_cat_percent[2]` %**); and **`r dets_tb_cat_fmt[3]`** vehicles (**`r dets_tb_cat_percent[3]` %**).
@@
 Note that the hsitogram is truncated at our selected confidence threshold of 0.1."
-#| fig.width: 8
-#| fig.height: 4.5
+#| fig.width: 6
+#| fig.height: 3
 
@@
 It is possible for MegaDetector to miss a detection within a burst, or to produce variable confidence scores within that burst, which could lead to missing true detections if not all pictures from a given burst are included.
-Therefore, if any image from a burst exceeded the selected MegaDetector threshold, we included it in our analysis .
+Therefore, if any image from a burst exceeded the selected MegaDetector threshold, we included it in our analysis.
 
@@
 bounding boxes of detections produced by MegaDetector).
-This is necessary for when we will crop our images to train our neural network model.
+This is necessary for when we crop our images to train our neural network model.
 -   *Optionally, open source,* to allow for maximum reproducibility of the process.
@@
 
-To guide our choice, we used these criteria to informally compare 51 different platforms and tools currently available for processing and managing camera trap images (a list of the tools compared in given in Appendix III).
+To guide our choice, we used these criteria to informally compare 51 different platforms and tools currently available for processing and managing camera trap images (a list of the tools compared is given in Appendix III).
 We first looked at platforms traditionally used for tagging camera trap images, but found them often lacking in two main aspects: flexibility of the tagging interface, and ability to manipulate bounding box information.
 
-These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners.
+These two features are often better addressed by tools not traditionally used for tagging camera trap images but used by the larger community of machine learning practitioners.
 Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features (for example, see @Schneider2021).
-Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.
+Tagging is a common step in machine learning research and industry and is often referred to as labeling.
 
 We found two platforms that matched all of our criteria: LabelStudio (@LabelStudio) and CVAT (@boris_sekachev_2020_4009388).
-After testing both platforms, LabelStudio was chosen because the data format it uses is more easily compatible with MegaDetector's COCO format output, making it easier to setup.
-We deployed the Community edition LabelStudio (version 1.6.0rc5) on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and a NGINX reverse proxy, which is a standard approach for custom hosting of websites.
-
-### Updated workflow (Wildtrax)
+After testing both platforms, LabelStudio was chosen because the data format it uses is more easily compatible with MegaDetector's COCO format output, making it easier to set up.
+We deployed the Community edition LabelStudio (version 1.6.0rc5) on an Ubuntu virtual machine hosted by Compute Canada, using an Apache server and an NGINX reverse proxy, which is a standard approach for custom hosting of websites.
+
+### Updated workflow (WildTrax)
 
@@
 
-Wildtrax keeps track of image batches using "location folders".
-We created our own location folders by combining images within the same retrieval, site and location IDs.
+WildTrax keeps track of image batches using "location folders".
+We created our location folders by combining images within the same retrieval, site, and location IDs.
 The standard format we used for location folders follows this convention:
@@
 
-For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.
-
-In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which most of the images were manually sorted into "wildlife" and "nonwildlife" folders inside each of the camera subfolders.
-In addition, because wildtrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images.
-Therefore, we developed a simple and very efficient script in the Julia programming language to paralellize the renaming of each file into a standardized manner.
-We had originally refrained from using this method of uniquely identifying files and instead relied on tracking the unique path of each file as we wanted to reduce the impact on our storage capacity (as it essentially required to double the required storage space).
+For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folder into "_W" and "_NW" variants and uploaded as such to WildTrax.
+
+To reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which most of the images were manually sorted into "wildlife" and "nonwildlife" folders inside each of the camera subfolders.
+In addition, because WildTrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images.
+Therefore, we developed a simple and very efficient script in the Julia programming language to parallelize the renaming of each file in a standardized manner.
+We had originally refrained from using this method of uniquely identifying files and instead relied on tracking the unique path of each file as we wanted to reduce the impact on our storage capacity (as it essentially required doubling the required storage space).
 
@@
 
-As a first step in testing the wildtrax workflow, we are uploading the folders from the first retrieval that showed the highest rates of false positives, retagging and evaluating the level of overlap with the LabelStudio workflow and also allowing us to complete these folders.
+As a first step in testing the WildTrax workflow, we are uploading the folders from the first retrieval that showed the highest rates of false positives, retagging and evaluating the level of overlap with the LabelStudio workflow and also allowing us to complete these folders.
 As a second step, we are uploading the folders from the very last retrieval, to speed up the tagging progress.
@@
 re-purposing the previous LabelStudio workflow for bounding box assignment, and 2.
-using a different, non networked platform.
+using a different, nonnetworked platform.
 
@@
 This resulted in 23730 cropped bounding boxes to be used for training and validation.
-We used the Pytorch framework (@Paszke_PyTorch_An_Imperative_2019), a widely used state of the art library for developing neural networks models in Python.
+We used the Pytorch framework (@Paszke_PyTorch_An_Imperative_2019), a widely used state-of-the-art library for developing neural network models in Python.
 As a first attempt, we settled on a model of the ResNet 100 architecture (@he2015deep), as it is widely used basal architecture for training species classifiers (for example, see @Khan2020, @Ukwuoma2022, and @BintaIslam2023), although the field is progressing and innovating on new methods and architectures rapidly (for example, see @Elhamod2021 or @radford2021learning).
@@
 dropout) layer (as recommended by @Srivastava2014).
-We tested the species classifier training via both transfer learning (frozen weights of a pretrained ResNet model) and classic learning (unfrozen weights, learning from the data only), as this is currently a comparison of interest in the field of AI applied to camera trap data (@schneider2018deep, @BintaIslam2023).
+We tested the species classifier training via both transfer learning (frozen weights of a pre-trained ResNet model) and classic learning (unfrozen weights, learning from the data only), as this is currently a comparison of interest in the field of AI applied to camera trap data (@schneider2018deep, @BintaIslam2023).
 
@@
 In addition, **`r non_mans_nb_fmt`** were true positives (or **`r non_mans_nb_pct` %**), to which **`r mans_nb_fmt`** false negatives were added (manually added bounding boxes, **`r mans_nb_pct` %**).
-Out of the **`r non_mans_nb_fmt`** true positives, **`r orig_tb_fmt[3]`** had to have their bounding box adjusted: the bounding box drawn by MegaDetector was at least partially innacurate and had to be corrected.
-The table below summarises information for the tagging progress.
+Out of the **`r non_mans_nb_fmt`** true positives, **`r orig_tb_fmt[3]`** had to have their bounding box adjusted: the bounding box drawn by MegaDetector was at least partially inaccurate and had to be corrected.
+The table below summarises information on the tagging progress.
 
@@
 knitr::kable(sum_tab,
-             caption = "Summary table on tagging progress. The 0.1 filter value corresponds to the megadetector cinfidence threshold.")
+             caption = "Summary table on tagging progress. The 0.1 filter value corresponds to the MegaDetector cinfidence threshold.")
 ```
@@
 This isn't a unique problem to this project, but an issue in many ecosystems.
-Locations where vegetation has grown in front of the sensor, combined with windy events, increases the number of photos triggered by vegetation, rather than wildlife.
+Locations where vegetation has grown in front of the sensor, combined with windy events, increase the number of photos triggered by vegetation, rather than wildlife.
 MegaDetector is likely to mistake many of the vegetation pictures for animals.
@@
 
-TODO: update once data is harmonized
+TODO: update on the tagging progress on WildTrax at the end of April
+
+## Ordination and clustering
+
+We summarized our existing species data in two ways:  (1) species presence/absence at each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site.
+We analyzed the presence/absence and relative abundance data using published and standardized numerical ecology community analysis methods (Legendre and Legendre 2013).
+
+As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
+We used Ward clustering on the relative abundances matrix.
+We extracted land cover information in a 1-km radius around each camera site, using the Far North Land Cover data set from Ontario's Ministry of Natural Resources and Forestry.
+We computed the frequencies of land cover classes and used these as constraining variables in a redundancy analysis, a method well suited for multidimensional data.
 
@@
 We are committed to producing a reproducible workflow.
-All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap), which is currently private but will be opened upon submission of the final version of this report.
-The repo makes use of the [**mdtools**](https://github.com/StewartWILDlab/mdtools) module, a python-based command line tool for converting MegaDetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
-
-## Analysis
-
-### Species detections
-
-We summarized our existing species data in two ways:  (1) species presence/absence at each camera and, (2) relative species abundance, measured as the number of days, within each month that each species was detected at a camera site. The detection time for species is highly variable.
-
-```{r echo=FALSE}
+All code necessary to reproduce this report is available at an online [repository](https://github.com/StewartWILDlab/rofcamtrap), which is currently private but will be opened upon publication.
+The repository makes use of the [**mdtools**](https://github.com/StewartWILDlab/mdtools) module, a Python-based command line tool for converting MegaDetector outputs into LabelStudio inputs, and LabelStudio outputs into CSV tables.
+
+# Preliminary results
+
+## Species detections
+
+```{r eval=FALSE, echo=FALSE}
 #| label: camera-plots
@@
 #| fig.width: 8
-#| fig.height: 4.5
+#| fig.height: 6
 
@@
                 ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
-  dplyr::filter(diff_time < 200) # For mustelid error currently
+  dplyr::filter(diff_time < 200) # For mustelid error currentlycams_sf_coords TODO
 
@@
 ggplot2::ggsave(plot = bp, filename = "figures/bp.png", width = 10, height = 15)
-```
-
-```{r echo=FALSE}
-#| label: camera-plots-2
-#| fig-cap: "Bubble plot of detection time for each species in either ecozone"
-#| fig.width: 8
-#| fig.height: 4.5
-
+
+# bp
+```
+
+```{r echo=FALSE}
 pts_data <- anns_wide_animals_joined_mut |>
@@
   dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
-                ECOZONE_NA = as.factor(ECOZONE_NA))
+                ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
+  dplyr::filter(diff_time < 200) # For mustelid error currentlycams_sf_coords TODO
 
@@
 ggplot2::ggsave(plot = pts, filename = "figures/pts.png", width = 10, height = 15)
-
+```
+
+To better visualize patterns in species detection, we computed the detection time (number of days from the start of a camera deployment until the species is detected at that camera) for each species in either ecozone, for data from cameras in the field between March 2022 and March 2023.
+The detection time for species is highly variable, with a mean of `r pts_data$diff_time |> mean() |> as.numeric()` days and a standard deviation of `r pts_data$diff_time |> sd() |> as.numeric()` days.
+
+We also produced a camera deployment graph, which is too large to be included in the main text but can be found in Appendix IV.
+The graph shows the deployment period for all cameras (with camera ID on the y-axis, date on the x-axis, and site ID as graph labels).
+The graphs show deployment and retrieval dates as black dots and also include days with at least one animal detection as red dots.
+
+```{r echo=FALSE}
+#| label: camera-plots-2
+#| fig-cap: "Bubble plot of detection time (number of days from the start of a camera deployment until the species is detection at that camera) for each species in either ecozone, for data between March 2022 and March 2023."
+#| fig.width: 8
+#| fig.height: 6
 pts
 ```
-
-### Ordination and clustering
-
-We analysed the relative abundance data using published and standardized numerical ecology community analysis methods (Legendre and Legendre 2013).
-As a first step, we combined clustering and ordination to visualize associations between species, camera sites, and land cover in the vicinity of the detection site.
-We used Ward clustering on the relative abundances matrix.
-We extracted land cover information in a 1-km radius around each camera site, using the Far North Land Cover data set from Ontario's Ministry of Natural Resources and Forestry.
-We computed the frequencies of land cover classes and used these as constraining variable in a redundancy analysis, a method well suited for multidimensional data.
-
-# Data exploration and preliminary results
 
@@
   gt::tab_header(
-    title = "Table 2: Relative abundances for species with more than 50 detection days, computed in number of days detected in the season."
+    title = "Table 2: Relative abundances for species with more than 50 detection days, computed in number of days detected in the season, for data between March 2022 and March 2023."
   ) |>
@@
 
-As for the spatial distribution of species relative abundances, @fig-map-pies shows relative species abundances across the landscape, for species detected more than 50 days in the season (this number is roughly equal to the standard deviation of the distribution of detection days, which has a mean of 32 days).
-Note that for this figure, the detections have grouped by plot number for better visibility, with a plot containing up to 4 cameras (see deployment methods).
-There are no clear spatial patterns, but some generalities can be noted: Moose detections are predominant in the northwestern and southeastern zones of the sampling array, while caribou and sandhill crane detections share the most of the detections for the central part of the array.
+As for the spatial distribution of species relative abundances, @fig-map-pies shows relative species abundances across the landscape, for species detected more than 50 days in the season (the distribution of detection days, has a mean of `r sp_tab$days_detected |> mean()` days and a standard deviation of `r `sp_tab$days_detected |> sd()`).
+Note that for this figure, the detections have been grouped by plot number for better visibility, with a plot containing up to 4 cameras (see deployment methods).
+There are no clear spatial patterns, but some generalities can be noted: Moose detections are predominant in the northwestern and southeastern zones of the sampling array, while caribou and sandhill crane detections share most of the detections for the central part of the array.
 
@@
 #| label: fig-map-pies
-#| fig-cap: "Pie chart map of relative species abundances across our study area Detections have grouped by plot for better visibility, with a plot containing up to 4 cameras.
+#| fig-cap: "Pie chart map of relative species abundances across our study area, for data between March 2022 and March 2023.
+Detections have been aggreagted by plot for better visibility, with a plot containing up to 4 cameras.
 Only species detected more than 50 days in the season are shown for clarity."
@@
 
-To start exploring general species-habitat associations we completed an redundancy analysis (RDA, using the `vegan` R package, @vegan).
-RDAs allow to study the relationships between two matrices of data, and is often used in community ecology.
+To start exploring general species-habitat associations we completed a redundancy analysis (RDA, using the `vegan` R package, @vegan).
+RDAs allow to study of the relationships between two matrices of data and is often used in community ecology.
 
@@
 
-We used the Ontario Far North Land Cover dataset to produce proportions of land covers at each site (@hogg2014far, a full legend of the land cover abbreviations are to be found in Appendix II).
+We used the Ontario Far North Land Cover dataset to produce the proportions of land covers at each site (@hogg2014far, a full legend of the land cover abbreviations are to be found in Appendix II).
 @fig-rda-pa shows the triplot for the RDA on presence-absence data, which yielded a low $R^2$ of `r round(as.numeric(vegan::RsquareAdj(pa_rda)[2]), 2)`, but which is still relevant for exploring species associations.
 Land cover might not be able to explain community structure, but it can still show which species are found together and around what kind of land cover class.
-Here, as expected, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
+Here, as expected, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bears and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
 Canada geese are associated with water (WAT) while smaller species like snowshoe Hares and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
-
 ```{r echo=FALSE}
 #| label: fig-rda-pa
-#| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected in March through September of 2022 reveals that different species associate with different Far North land cover classes.
+#| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected between March 2022 and March 2023 reveals that different species associate with different Far North land cover classes.
 For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
@@
 
-To continue exploring abundance-habitat relationships as a proxy for species relative habitat use, we also conducted an RDA using relative abundances instead of presence absence.
+To continue exploring abundance-habitat relationships as a proxy for species' relative habitat use, we also conducted an RDA using relative abundances instead of presence-absence.
 @fig-rda-events shows the triplot for the RDA on relative abundance data.
 This analysis yielded a slightly larger $R^2$ of `r round(as.numeric(vegan::RsquareAdj(events_rda)[2]), 2)`, than the analysis in @fig-rda-pa presenting presence-absence data, but reduces the axis scores of the rarer species.
-We observe similar patterns for caribou, crane, moose, black bear and marten.
-
+We observe similar patterns for caribou, cranes, moose, black bears, and marten.
 ```{r echo=FALSE}
 #| label: fig-rda-events
-#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns as the triplot using presence-absence data."
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected between March 2022 and March 2023, showing similar patterns as the triplot using presence-absence data."
 #| fig.width: 8
@@
 
-To start inferring patters from the above ordination analyses we used a clustering algorithm on the relative abundance data presented in @fig-rda-events.
+To start inferring patterns from the above ordination analyses we used a clustering algorithm on the relative abundance data presented in @fig-rda-events.
 @fig-rda-events-clust shows the triplot for the RDA on relative abundance data, with Ward clusters as ellipses (@Legendre2012).
 We divided our ordination plot into 2 clusters along the first RDA axis.
-One cluster for sites primarily found with caribou and cranes, and a second cluster for the rest of the cameras and species.This clustering generally reflects the spatial patterns in species abundance that we observed in @fig-map-pies.
+One cluster for sites primarily found with caribou and cranes, and a second cluster for the rest of the cameras and species.
+This clustering generally reflects the spatial patterns in species abundance that we observed in @fig-map-pies.
 
@@
 #| label: fig-rda-events-clust
-#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected in March through September of 2022, showing similar patterns than the triplot using presence-absence data.
+#| fig-cap: "Triplot of wildlife relative abundance data from Northern Ontario camera trapping images collected between March 2022 and March 2023, showing similar patterns than the triplot using presence-absence data.
 Two clusters determined using Ward clustering are shown as ellipses."
@@
 
-clust_grp <- cutree(clust, k = 4)
+clust_grp <- cutree(clust, k = 2)
 custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp = 0.4, scale_y_sp = 0.3,
@@
 #| label: fig-map-clust
-#| fig-cap: "Map of clustered camera sites. The two clusters were determined using the Ward clustering method."
+#| fig-cap: "Map of camera sites clustered by Ward Clustering for data collected between March 2022 and March 2023."
 #| fig.width: 8
@@
 The extent to which these distances influence the result in balance with the community data can be adjusted, by setting the strength of the mixing parameter alpha.
-The `ClustGeo` package (@Chavent2018) allows to test different values of that parameter to determine optimal mixing of the clustering forces, with the function `choicealpha`.
-
-We found an alpha level of 0.1 to be optimal for this data set, and performed mixed clustering using the `hclustgeo` function.
+The `ClustGeo` package (@Chavent2018) allows testing different values of that parameter to determine optimal mixing of the clustering forces, with the function `choicealpha`.
+
+We found an alpha level of 0.1 to be optimal for this data set and performed mixed clustering using the `hclustgeo` function.
 @fig-map-clust-geo shows the resulting map of geographically constrained clustered camera sites.
@@
 #| label: fig-map-clust-geo
-#| fig-cap: "Map of geographically constrained clustered camera sites, using geographically constrained clustering with a mixing parameter of 0.1."
+#| fig-cap: "Map of geographically constrained clustered camera sites, for data collected between March 2022 and March 2023, using geographically constrained clustering with a mixing parameter of 0.1."
 #| fig.width: 8
@@
 
-We use a series bubbleplots to visualize species relative abundance through space and time.
+We use a series of bubble plots to visualize species' relative abundance through space and time.
 The following plots are for caribou, sandhill crane, moose, and wolf.
@@
 
-We can derive a few take home messages from our initial exploration of the data:
-
--   Using an object detection model allowed us to filter out large quantities of empty images, but performed somewhat poorly in certain conditions (ie, windy sites with drastic changes in vegetation).
-
--   Associative patterns: species associate with land cover classes that corresponds to expected habitat, such as wetlands for caribou and cranes, or forested areas for moose.
-
--   Spatial clustering and considerations: Latitudional relationships may dominate this data set and drive species-habitat associations.
-
--   Species classifier: we obtained better accuracy (94%) with fresh learning than transfer learning (91%).
-We are still exploring avenues for improvement, but it is likely that the small sample size is the main limiting factor.
+We can derive a few take-home messages from our initial exploration of the data:
+
+-   Using an established object detection model for camera trap images (i.e., MegaDetector) allowed us to filter out large quantities of empty images, but performed somewhat poorly in certain conditions (ie, windy sites with drastic changes in vegetation).
+
+-   Associative patterns: species associate with land cover classes that correspond to expected habitat, such as wetlands for caribou and cranes, or forested areas for moose.
+
+-   Spatial clustering and considerations: latitudinal relationships may dominate this data set and drive species-habitat associations.
+
+-   Species classifier: we obtained better accuracy (94%) with classic learning than transfer learning (91%).
+We are still exploring avenues for improvement, but the small sample size is likely the main limiting factor.
 
@@
 
-Over the coming months we will continue to process the retrieved data and anticipate adding to our data set.
-The next steps for this project are to complete the tagging process and identify all the detections tagged as unknowns, bird, duck and mustelid species.
+Over the coming months, we will continue to process the retrieved data and anticipate adding to this data set as images from the 2024 deployments are retrieved.
+The next steps for this project are to complete the tagging process and refine to the species level all detections currently categorized as "unknown", "bird", "Duck", and "mustelid".
+Dr.
+Stewart will host an undergraduate research student in the summer of 2024 to help with the "mustelid" category if funding is approved through the NSERC USRA program.
 We will also further explore the spatial community composition patterns at each camera site.
-
-Given a high enough data density for certain species future possibilities for this project include modelling species occupancy (e.g.
-@MacKenzie2002) and spatially explicit density estimation using spatially explicit capture-recapture models (e.g.@Chandler2013). We will perfect the new tagging workflow and continue testing classifier methods.
+Given a high enough data density for certain species future possibilities for this project include modeling species occupancy (e.g.
+@MacKenzie2002) and spatially explicit density estimation using spatially explicit capture-recapture models (e.g.
+@Chandler2013). We will perfect the new tagging workflow and continue testing classifier methods.
 
@@
 
-We thank the student volunteers at Wilfrid Laurier University for their help with tagging: Teea Curlew, Bridget Matthews, Meghna Pal, Rafay Siraj, and Dietrich Westberg.
+We thank our research assistant, Meghna Pal, as well as our student volunteers at Wilfrid Laurier University for their help with camera tagging: Teea Curlew, Bridget Matthews, Rafay Siraj, and Dietrich Westberg.
 We also thank members of the WILDlab for feedback throughout this process: Claudia Haas, Eric Jolin, Charlotte Rentmeister, and Sheyda Zand.
-This work was supported by a G&C to Frances Stewart and the NSERC Alliance full project team.
+This work was supported by an ECCC G&C, an NSERC Canada Research Chair (Tier II), an NSERC Alliance grant, and Canadian Foundation for Innovation funds to Dr.
+Frances Stewart.
 
@@
   gt::tab_header(
-    title = "Table 3: Relative species abundances",
+    title = "Table 3: Relative abundances for species for all species, computed in number of days detected in the season, for data between March 2022 and March 2023.
+values represent the number of days each species was detected at all available cameras",
     subtitle = "Computed in number of days detected in the season"
@@
 #| label: tab-landcover
-table_cat
+table_cat|>
+  gt::tab_header(
+    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset (adapted from dataset's metadata), along with the percentage land cover within the 1km buffer areas used for the the multivariate analyses in this report."
+  )
 ```
@@
 
+## Appendix IV: Camera Deployment History
+
+```{r, echo=FALSE}
+#| fig.width:  8
+#| fig.height: 12
+
+image_data <- anns_wide_animals_joined_mut |> 
+  dplyr::select(id, source_file_with_dep,
+                image_id, species:comment, camera_id,
+                maker_notes_sequence:maker_notes_user_label, plot_id:ymd) |> 
+  dplyr::mutate(site_id = sapply(image_id, \(x) strsplit(x, "_")[[1]][1]),
+                date_time = lubridate::as_date(date_time)) |> 
+  dplyr::rename(file_path = source_file_with_dep) |> 
+  dplyr::filter(!is.na(date_time))
+
+cam_dep_data <- cams_sf_coords |> 
+  sf::st_drop_geometry() |> 
+  dplyr::select(dplyr::ends_with("_id"), date_deployed, date_retrieved, longitude, latitude,
+                deployment_group) |> 
+  dplyr::mutate(date_deployed = lubridate::as_date(date_deployed),
+                date_retrieved = lubridate::as_date(date_retrieved))
+
+stopifnot(sum(table(cam_dep_data$camera_id) != 1) == 0)
+
+cam_dep_data_with_status <- cam_dep_data |>
+  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |>
+  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))
+
+cam_dep_data_with_status_1 <- 
+  cam_dep_data_with_status[1:(nrow(cam_dep_data_with_status)/2),]
+cam_dep_data_with_status_2 <- 
+  cam_dep_data_with_status[((nrow(cam_dep_data_with_status)/2) + 1):nrow(cam_dep_data_with_status),]
+
+cam_dep_data_with_status_1 |>
+  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id,
+                                                              date_deployed,
+                                                              decreasing = TRUE))) +
+  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id), 
+                        alpha = 0.15, linewidth = 3) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id)) +
+
+  ggplot2::geom_point(data = image_data |> 
+                        dplyr::filter(camera_id %in% cam_dep_data_with_status_1$camera_id),
+                      ggplot2::aes(x = date_time, y = camera_id),
+                      color = "firebrick", size = 0.5) +
+
+  ggplot2::geom_text(ggplot2::aes(x = date_deployed - 60, y = camera_id,
+                                  label = site_id), size = 3) +
+  # ggplot2::scale_color_discrete(type = c("#E69F00","#56B4E9")) +
+  ggplot2::theme_bw() +
+  ggplot2::labs(x = "Date", y = "Camera ID")
+```
+
+```{r, echo=FALSE}
+#| fig.width:  8
+#| fig.height: 12
+cam_dep_data_with_status_2 |>
+  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id,
+                                                              date_deployed,
+                                                              decreasing = TRUE))) +
+  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id),
+                        alpha = 0.15, linewidth = 3) +
+  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id)) +
+  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id)) +
+
+  ggplot2::geom_point(data = image_data |> 
+                        dplyr::filter(camera_id %in% cam_dep_data_with_status_2$camera_id),
+                      ggplot2::aes(x = date_time, y = camera_id),
+                      color = "firebrick", size = 0.5) +
+
+  ggplot2::geom_text(ggplot2::aes(x = date_deployed - 60, y = camera_id,
+                                  label = site_id), size = 3) +
+  ggplot2::theme_bw() +
+  ggplot2::labs(x = "Date", y = "Camera ID")
+```
+
 {{< pagebreak >}}
@@
     )
-
   ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,"_heatmap.png"))

### 2024-04-29 reports/GnC_report/paper/paper.qmd
fix typos

@@
       - ref: eccc
+  - name: Meghna Pal
+    affiliation:
+      - ref: wlu
   - name: Jennifer Baltzer
@@
   - name: Frances Stewart
-    orcid: 0000-0001-9344-8346
     affiliation:
@@
 - We made a first attempt at using machine learning methods to train a species classifier using about 23000 images as the basis dataset, which produced encouraging results.
-- After reevaluation of the ability of our current worlflow and platform to scale to multiple years of data, we have transitioned the image tagging workflow to [WildTrax](https://WildTrax.ca/).
+- After reevaluation of the ability of our current workflow and platform to scale to multiple years of data, we have transitioned the image tagging workflow to [WildTrax](https://WildTrax.ca/).
 - This report repeatedly mentions the AI model MegaDetector.
@@
 2. Poster submitted and displayed at the Mushkegowuk Council 2023 Research Meeting.
-3. Application to participate in the Workshop on Computer Vision Methods for Ecology at the Caltech Resnick Sustainability Institute. Rejectec in 2023, reapplied in 2024.
+3. Application to participate in the Workshop on Computer Vision Methods for Ecology at the Caltech Resnick Sustainability Institute.
 
@@
 The network was established by CWS in 2020 and seasonally equipped in a rolling design with 500 autonomous recording units (ARUs) to monitor audible species diversity paired with site-specific on-ground vegetation imagery.
-In March through June of 2022 193 unbaited and unlured wildlife camera traps ([Reconyx Hyper Fire II](https://reconyx.com/product/hyperfire-2-covert-ir-camera)) were added to this study design.
+
+In March through Sepember of 2022, 272 unbaited and unlured wildlife camera traps ([Reconyx Hyper Fire II](https://reconyx.com/product/hyperfire-2-covert-ir-camera)) were added to this study design, spread over 60 plots and 227 sites.
+The cameras were deployed by CWS Ontario staff and contractors along linear features such as game trails, along streams and wetlands, on shorelines of small lakes, or facing beaver lodges and dams.
 All cameras were programmed to take bursts of 5 images (one per second) at each trigger of their motion sensor.
 
-TODO: more detailed summary of deployment methods from Sam's report to be added before submission
+Instructions for determining where to set up the camera included looking for signs of animal presence, such as scat, scratch marks, tracks, and ungulate carcasses; as well as targeting landscape features likely to attract wildlife, such as game trails, riparian corridors, ridgelines, or vegetation.
+The cameras were enclosed in a cage (Security Enclosure) and attached to a tree or post overlooking the area, making sure that the attachment point could support the camera’s weight during strong winds.
+The cameras were placed about 1.2 m above the ground and angled slightly downwards, with the target detection zone about 5 m away.
+If the snow depth was more than 1.2 m at the time of deployment, the camera was raised to a height of 1.5 m instead.
+The cameras were set up to have an unobstructed view, at least 5 m wide and 10 m long, of the detection zone.
+Ideally, they were placed facing either north, northwest, or northeast to reduce glare from direct sunlight.
+After the activation of the camera, the lens was checked for cleanliness, and test photos were taken to ensure the camera was working correctly.
 
@@
 
-Briefly, 5 ARUs are deployed per site, with 100 m between ARUs.
+Briefly, 5 ARUs were deployed per site.
 Up to 4 cameras were paired with ARUs at each of these sites, with at least one of these cameras placed at a random ARU site and one other prioritized to focus on a game trail or open area to increase detection probability.
 Weather restrictions and field conditions limited access to some sites resulting in some variation in this design.
-In total, we have 43 plots with 4 cameras, 5 plots with 3 cameras, and 3 plots with 2 cameras. @fig-deployment shows the location of the clustered cameras and their current retrieval status: 54 cameras were retrieved in May 2022 and 116 in September 2022.
-Six additional cameras meant to be retrieved then were instead retrieved in March 2023 due to logistical and accessibility issues, along with another 77 cameras that were meant to be retrieved at that date.
-The last 18 cameras were retrieved in September 2023.
+In total, we have 43 plots with 4 cameras, 5 plots with 3 cameras, and 3 plots with 2 cameras. @fig-deployment shows the location of the clustered cameras and their deployment and retrieval dates.
+
+  - 59 cameras were deployed in March 2022, 53 of which were retrieved in September 2022.
+The remaining 6 cameras meant to be retrieved then were instead retrieved in March 2023 due to logistical and accessibility issues.
+
+  - 116 cameras were deployed in May 2022, and retrieved in September 2022.
+
+  - 18 cameras were deployed in June 2022, and retrieved in September 2023
+
+  - 79 cameras were deployed in September 2022, and 77 were retrieved in March 2023. The remaining 3 have not been retrieved yet due to weather and logistical constraints.
 
@@
 #| fig-cap: "Left: Map of the province of Ontario with ecoregion and inset map delimitation.
-Right: Inset map of camera deployment locations as wellas deployment and retrieval dates for the 227 sites.
+Right: Inset map of camera deployment locations as well as deployment and retrieval dates for the 227 sites.
 Labels for the 2 ecoregions spanned by the sampling area are included."
@@
 The main issue we ran into is because MegaDetector (which is described in more detail in the next section), produced a large amount of false positives in our region.
-This may have resulted from MegaDetector not being calibrated to highly seasonal landscapes, resulting in very slow processing of images through LabelStudio This made data processing in LabelStudio slower and more cumbersome than we initially anticipated, as LabelStudio is not well equipped for group tagging of false positives.
+This may have resulted from MegaDetector not being calibrated to highly seasonal landscapes, resulting in very slow processing of images through LabelStudio.
+This made data processing in LabelStudio slower and more cumbersome than we initially anticipated, as LabelStudio is not well equipped for group tagging of false positives.
 However, in the absence of a better alternative, we persevered in using the platform and attempted to circumvent these issues by trying to reduce the number of false positives in the MegaDetector output.
@@
 The colors represent the three categories of object identified by MegaDetector: animals, persons and vehicles.
-Note that the hsitogram is truncated at our selected confidence threshold of 0.1."
+Note that the histogram is truncated at our selected confidence threshold of 0.1."
 #| fig.width: 6
@@
 
-For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folder into "_W" and "_NW" variants and uploaded as such to WildTrax.
+For locations in which images were previously sorted into "wildlife" and "nonwildlife", we divided each location folder into "_W" and "_NW" variants and uploaded as such to WildTrax.
 
@@
 #| fig-cap: "Pie chart map of relative species abundances across our study area, for data between March 2022 and March 2023.
-Detections have been aggreagted by plot for better visibility, with a plot containing up to 4 cameras.
+Detections have been aggregted by plot for better visibility, with a plot containing up to 4 cameras.
 Only species detected more than 50 days in the season are shown for clarity."
@@
 Here, as expected, we see that caribou and sandhill cranes are associated with open and treed wetlands such as open and treed bogs and fens (OBOG, OFEN, TrBOG, and TrFEN) while black bears and moose are found together in more forested areas of mixed trees (MixTRE), sparse Trees (SpTRE) and coniferous trees (ConTRE).
-Canada geese are associated with water (WAT) while smaller species like snowshoe Hares and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
+Canada geese are associated with water (WAT) while smaller species like snowshoe hares and martens score away from the larger mammals in a mix of habitats, including swamps (SWA).
 ```{r echo=FALSE}
@@
 #| fig-cap: "Triplot of wildlife presence-absence data from Northern Ontario camera trapping images collected between March 2022 and March 2023 reveals that different species associate with different Far North land cover classes.
-For example, caribou cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
+For example, caribou and cranes are more likely to be detected near open and treed wetlands (OBOG, OFEN, TrBOG, and TrFEN) while black bear and moose near forested areas (MixTRE, SpTRE, ConTRE)."
 #| fig.width: 8

### 2024-04-29 scripts/julia/rename-images.jl
updates

@@ -18,7 +18,7 @@ using CSV
     # Open the file to read a single tag
     io = open(image, "r")
     tags =  read_tags(io, read_all = false, tags=["EXIF_TAG_DATE_TIME_ORIGINAL"])
-    
+
     # If the tag is empty, throw an error, else format it
     if isempty(tags)
         throw("Image $image has no datetime")
@@ -35,18 +35,18 @@ end
 
 # Construct a new path
 @everywhere function make_new_path(image, datetime ; images_to, deployment_code, loc, loc_base, file_ext)
-    
+
     # If the datetime is empty, throw an error, else format it
-    if isnothing(datetime) 
+    if isnothing(datetime)
         throw("Image $image has no datetime")
     else
         # Filter the path for useless sections, removing DCIM first
         image_split_filtered = replace.(replace.(replace.(replace.(replace.(
-            filter(x -> x != "DCIM", 
+            filter(x -> x != "DCIM",
                 # Split at appropriate file separator
-                splitpath(replace(image, loc=>""))[2:end]), 
+                splitpath(replace(image, loc=>""))[2:end]),
             "RECNX"=>""), "RCNX"=>""), file_ext=>""), "NonWildlife"=>"NW"), "Wildlife"=>"W")
-        
+
         # Construct the new file name...
         new_name =  join([deployment_code, loc_base,
             join(vcat(image_split_filtered, [datetime]), "_") * ".JPG"], "_")
@@ -60,7 +60,7 @@ end
             sub = join(vcat([deployment_code], [loc_base], image_split_filtered[1:2]), "_")
         end
 
-        new_path = joinpath(images_to, deployment_code, 
+        new_path = joinpath(images_to, deployment_code,
                             join([deployment_code, loc_base], "_"),
                             sub,
                             new_name)
@@ -75,9 +75,9 @@ end
 @everywhere function process_image(image, new_path; force = true)
 
     if !isfile(new_path)
-        if !isdir(dirname(new_path)) 
+        if !isdir(dirname(new_path))
             @info("Creating path $(dirname(new_path))")
-            mkpath(dirname(new_path)) 
+            mkpath(dirname(new_path))
         end
         # @info("Creating file $(basename(new_path))")
         cp(image, new_path, force = force)
@@ -95,14 +95,14 @@ function rename_images(
 
[…]

### 2024-04-29 analysis/documents/1-data-preparation.qmd
updates

@@
   dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
-  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))
+  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/"))) #|> 
+#   dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
+#                                                tz = "Canada/Eastern"),
+#                 year = lubridate::year(date_time),
+#                 month = lubridate::month(date_time, label = TRUE),
+#                 day = lubridate::day(date_time),
+#                 ymd = paste(year, month, day, sep = "_"))
+# 
+# exif[
+#   exif$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
+# exif[
+#   exif$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"
 
@@
 
+anns_wide_animals_joined_mut[
+  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
+anns_wide_animals_joined_mut[
+  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"
+
 saveRDS(anns_wide_animals_joined_mut, here::here("data", "objects", "annotations_animals.rds"))
@@
                 camera_first_day = date_time)
+
+camera_days[camera_days$camera_first_day == "2024-05-23 13:31:46",]$camera_first_day <- as.Date("2022-05-23 13:31:46")
 
@@
 
+all_table <- dplyr::bind_rows(extr) |> 
+  dplyr::group_by(value) |> 
+  dplyr::summarise(total = sum(coverage_fraction)) |> 
+  dplyr::mutate(value = as.factor(value), 
+                Percentage = round(total/sum(total)*100, 2))
+
 extr_list <- extr |> 
@@
 table_cat <- lu_dat |> 
-  dplyr::select(category_code, label) |> 
+  dplyr::left_join(all_table, by = dplyr::join_by(cats == value)) |> 
+  dplyr::mutate(Percentage = ifelse(is.na(Percentage), 0, Percentage)) |>  
+  dplyr::select(category_code, label, Percentage) |> 
   dplyr::filter(category_code != "ND") |> 
@@
   gt::gt() |> 
-  gt::tab_header(
-    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
-    subtitle = "Adapted from the ontario Far North land cover dataset"
-  ) |>
   gt::cols_label(
@@
     category_code = "Abbreviations",
+    Percentage = "% Cover in 1km buffers"
   ) |> 
@@
 ```
+
+#### Process WildTrax data
+
+```{r}
+# debugonce(wt_auth)
+wildRtrax::wt_auth(force = T)
+
+# debugonce(wt_get_download_summary)
+my_projects <- wildRtrax::wt_get_download_summary(sensor_id = 'CAM')
+
+# debugonce(wt_download_report)
+my_report <- wildRtrax::wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "image_report", weather_cols = F) |>
+  tibble::as_tibble()
+```

### 2024-04-29 analysis/documents/2-data-exploration.qmd
updates

@@
 # image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
-image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
-image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"
+# image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
+# image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"
 
@@
 
-```{r, fig.height=40, fig.width=8}
+```{r, eval: FALSE fig.height=35, fig.width=7}
+#| fig.width: 8
+#| fig.height: 35
+
+cam_dep_data_with_status <- cam_dep_data |>
+  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |>
+  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))
+
 cam_dep_data_with_status |>
@@
 November 5th 2023 to March 10th 2024
+
+#### Compare platform
+
+```{r}
+wt <- readRDS("../../my_report.rds")
+```
+
+```{r}
+image_data |> dplyr::filter(plot_id == "P045") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage_2/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()
+
+wt |> dplyr::filter(grepl(.data$location, pattern = "TC2_P045", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
+```
+
+```{r}
+image_data |> dplyr::filter(plot_id == "P045") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()
+
+wt |> dplyr::filter(grepl(.data$location, pattern = "TC1_P045", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
+```
+
+```{r}
+image_data |> dplyr::filter(plot_id == "P080") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()
+
+wt |> dplyr::filter(grepl(.data$location, pattern = "TC1_P080", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
+```
+
+#### AI results explore
+
+```{r}
+library(tidyverse)
+
+labels <- read.csv("../../3_Classifiers/1_species_classifier/labels.csv") |> 
+  pivot_longer(OtherBird:Bear, names_to = "Category_true", values_to = "val") |> 
+  filter(val != 0) |> select(-val)
+preds <- read.csv("../../3_Classifiers/1_species_classifier/predictions.csv") |> 
+  pivot_longer(!X, names_to = "Category_pred", values_to = "conf") |> 
+  group_by(X) |> 
+  arrange(desc(conf)) |> 
+  slice(1) |> ungroup()
+
+joined <- labels |> 
+  left_join(preds, by=c("id"="X")) |> 
+  drop_na() |> 
+  mutate(correct = Category_true == Category_pred)
+
+errors <- joined |> filter(!correct)
+```
+
+```{r}
+ex <- errors |> filter(Category_true == "Moose", Category_pred == "Caribou") |> 
+  pull(filepath) |> 
+  str_replace("/media/vlucet/TrailCamST/", "/media/vlucet/TrailCamST1/_old/")
+
+for (thefile in ex) {
+  # print(file.path("/media/vlucet/TrailCamST1/_old/errors_ex", basename(thefile)))
+  file.copy(thefile, file.path("/media/vlucet/TrailCamST1/_old/errors_ex/", basename(thefile)))
+}
+```

### 2024-04-29 reports/GnC_report/paper/paper.qmd
updates

@@
 
-TODO: update on the tagging progress on WildTrax at the end of April
+About 97% of all images previously tagged in LabelStudio have been uploaded to the WildTrax platform (3,192,316 images), in addition to all the images from the cameras most recently retrieved.
+The remaining images are still being uploaded.
+These images are being re-tagged at the species level, and are in various stages of progress, with about half of those folders fully tagged and verified, illustrating the superior speed of tagging on the platform.
+A more detailed report of this progress and tagging method will be present in the final version of this report, once the revision of the ragging workflow is finalized.
 
@@
                 ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
-  dplyr::filter(diff_time < 200) # For mustelid error currentlycams_sf_coords TODO
+  dplyr::filter(diff_time < 200) # For mustelid error currently TODO
 
